### 2014-09-09 draft_lm1.tex

\documentclass[12pt]{article}              
\usepackage[margin=1.5cm]{geometry}
\usepackage{setspace,relsize}               
\usepackage{moreverb}                        
\usepackage{url}
\usepackage{hyperref}
\hypersetup{colorlinks=true,citecolor=blue}
\usepackage{amsmath}
\usepackage{mathtools} 
\usepackage{amssymb}
\usepackage{indentfirst}
\usepackage[authoryear,round]{natbib}
\bibliographystyle{apalike}
\usepackage[pdftex]{lscape}
% \usepackage{float}
% \usepackage{longtable}
% \usetikzlibrary{arrows}
%%%%
\title{
Estimating the Initial Susceptible Fraction in Dengue Dynamics
}           
\author{Flavio Code\c{c}o Coelho \\
School of Applied Mathematics, Get\'ulio Vargas Foundation (FGV) \\
\& \\
Luiz Max de Carvalho \\
Program for Scientific Computing (PROCC), Oswaldo Cruz Foundation}

\date{\today}
% Notation defs
\def \rr {$R_{t}\ $}

\usepackage{Sweave}
\begin{document}                                  
\input{draft_lm1-concordance}

%
\maketitle
\begin{abstract}
We derive confidence/credibility bounds for the uncertainty about the estimate of the effective reproductive number $R_t$ under the assumption that reported disease counts are Poisson distributed.
It is shown that one can construct a conservative confidence interval for \rr using the induced probability distribution that results from the approach of \citet{mantel}.
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Background}
\label{sec:intro}

In monitoring of infectious diseases, it is important to assess whether the incidence of a a particular disease is increasing significantly, in order to decide to take preventive measures.
The effective reproductive number at time $t$, \rr, can be understood as a real-time estimate of the basic reproductive number ($R_{0}$) and is defined as the average number of secondary cases per primary case at time $t$.
Let $Y_t$ be the number of reported disease cases for a particular time $t \in (0, T)$.
\citet{nishiura} extend the theory developed by \citet{stallybrass} and propose to estimate \rr as
\begin{equation}
\label{eq:Rtestimate}
R_t = \left( \frac{Y_{t+1}}{Y_t}\right)^{1/n}
\end{equation}
where $n$ is taken to be the ratio between the length of reporting interval and the mean generation time of the disease.
Here we are interested in the simpler case $n=1$.
If \rr is to be used as a decision tool, one needs to be able to quantify the uncertainty about estimate in equation~\ref{eq:Rtestimate}. 
Here we detail how to obtain confidence intervals for \rr, exploring two approaches under the assumption that the counts $Y_t$ are Poisson distributed for all $t$.

\newpage
\section{Approaches}

\subsection{Ederer \& Mantel (1974)}
\label{sec:mantel}

First we explore the approach of \citet{mantel}, whose objective is to obtain confidence intervals for the ratio of two Poisson counts. 
Let $Y_{t} \sim Poisson(\lambda_t)$ and $Y_{t+1} \sim Poisson(\lambda_{t+1})$ and define $S = Y_{t} + Y_{t+1}$.
The authors then proceed to note that by conditioning on the sum $S$
\begin{align}
\label{eq:binlike}
Y_{t+1} | S &\sim Binomial(S, \theta_t) \\
\theta_t &= \frac{\lambda_{t+1}}{\lambda_{t} + \lambda_{t+1}}
\end{align}
Let $c_{\alpha}(\theta_t) = \{\theta_t^{(L)} , \theta_t^{(U)} \}$ be such that $Pr(\theta_t^{(L)}<\theta_t <\theta_t^{(U)}) = \alpha$.
Analogously, define $c_{\alpha}(R_t) = \{R_t^{(L)} , R_t^{(U)} \}$ such that $Pr(R_t^{(L)}<R_t<R_t^{(U)}) = \alpha$.
\citet{mantel} show that one can construct a $100\alpha \%$ confidence interval for \rr by noting that
\begin{align}
 R_t^{(L)} &= \frac{\theta_t^{(L)}}{(1-\theta_t^{(L)})}\\
 R_t^{(U)} &= \frac{\theta_t^{(U)}}{(1-\theta_t^{(U)})}
\end{align}

Many authors have derived $c_{\alpha}(\theta_t)$ following orthodox approaches such as those of \citet{wilson} and \citet{clopper}, mainly for simplicity reasons.
We choose instead to take a Bayesian approach and use the  $100\alpha \%$ posterior credibility interval for $\theta_t$ as $c_{\alpha}(\theta_t)$.
If we choose the conjugate beta prior with parameters $a_0$ and $b_0$ for the binomial likelihood in (\ref{eq:binlike}), the posterior distribution for $\theta_t$ is
\begin{equation}
\label{eq:thetapost}
p(\theta_t| Y_{t+1}, S) \sim Beta(Y_{t+1} + a_0, Y_t + b_0)
\end{equation}
Equation~(\ref{eq:thetapost}) tells us that the induced distribution of $R_t$ is a beta prime with parameters $ a_1 = Y_{t+1} + a_0$ and $b_1 =  Y_t + b_0$.
The density of this induced distribution  $f_P(R_t)$ is then 
\begin{equation}
\label{eq:densityMantel}
f_P(R_t| a_1, b_1) = \frac{\Gamma(a_1 + b_1)}{\Gamma(a_1)\Gamma(b_1)} R_t^{a_1 - 1} (1 + R_t)^{-(a_1 + b_1)}
\end{equation}
Thus, the expectation of \rr under $f_P$ is $a_1/(b_1 - 1)$ and its variance is $a_1(a_1 + b_1 - 1)/\left((b_1 - 2)(b_1 - 1)^2 \right) $.

Since when $R_t > 1$ indicates sustained transmission, one may be interested in computing the probability of this event.
By noting that
\begin{align}
\label{cumprobMantel}
Pr(R_t > 1) &= 1 - \int_0^1 f_P(r)dr \\
            &= 1- Pr(\theta_t < \frac{1}{2})
\end{align}
one can avoid dealing with the density in~(\ref{eq:densityMantel}) directly.

% \subsection{Nishiura et al. (2010)}
% \label{sec:nishiura}
% 
% An alternative approach is to use the conditional distribution of \rr on $Y_{t+1}$ and $Y_t$ as defined in equation A7 of \citet{nishiura}:
% \begin{equation}
% \label{eq:unorm}
% f_{R}(R_{t}) = (Y_tR_{t})^{Y_{t+1}} e^{-Y_tR_{t}}
% \end{equation}
% Noticing the kernel of (\ref{eq:unorm}) is that of a gamma distribution with $a_2 = Y_{t+1}+1$ and $b_2 = Y_t$, we obtain a proper density from which to construct $c_{\alpha}(R_t)$, simply by computing the appropriate quantiles of said distribution.
%  This density is
% \begin{equation}
% \label{eq:densityNishiura}
% f_N(R_t| a_2, b_2) =  \frac{b_2^{a_2}}{\Gamma(a_2)} R_t^{a_2-1} e^{-b_2 R_t}
% \end{equation}
% Once again $Pr(R_t>1)$ can be easily computed using
% \begin{equation}
% \label{eq:cdfNishiura}
% P(R_t > 1) = 1-F_N(1) = 1- \frac{1}{\Gamma(a_2)} \gamma(a_2, b_2)
% \end{equation}
% where $F_N(R_t)$ is the distribution function of $R_t$ and $\gamma(\cdot)$ is the incomplete gamma function.

\subsection{A remark on prior distributions and tail behaviour}
\label{sec:tails}

In order to decide which approach to take, it may be of use analysing the variance and tail behaviour of the derived distributions for \rr. 
Consider the case of using a flat $Uniform(0, 1)$ prior for $\theta_t$ in Section~\ref{sec:mantel}.
With $a_0 = b_0 = 1$, $a_1 = a_2$ and $b_1 = b_2 + 1$.
The beta prime (inverse beta distribution) will have heavier tails compared to the conditional distribution in (\ref{eq:densityNishiura}), thus providing more conservative confidence/credibility intervals.  
As a side note, the Bayesian approach presented in Section~\ref{sec:mantel} will give similar results to those of \citet{wilson} and \citet{wilson} for $Y_{t+1}$ and $Y_t >> 1$.
Under the flat  uniform prior for $\theta_t$, the Bayesian posterior credibility interval is nearly indistinguishable from the confidence interval proposed by \citet{clopper} for $Y_{t+1}, Y_t > 20$.
Note that the $Beta(1, 1)$ uniform prior for $\theta_t$ constitutes a poor prior choice mainly because the induced distribution for \rr is only well-defined for $b_0 > 2$.

The main advange of the Bayesian approach is that one can devise prior distributions for $\theta_t$ taking advantage of the intuitive parametrisation and flexibility of the beta family of distributions.
Prior elicitation can also be done for \rr and the hyperparameters directly plugged into the prior for $\theta_t$. 
One can, for example, choose a priori mean and variance for \rr and find $a_0$ and $b_0$ that satisfy those conditions.
Let $m_0$ and $v_0$ be the prior expectation and variance for $R_t$. 
After some tedious algebra one finds
\begin{align}
\label{eq:elicitation}
a_0 &= \frac{m_0v_0 + m_0^3 + m_0^2}{v_0} \\
b_0 &= \frac{2v_0 + m_0^2 + m_0}{v_0}
\end{align}
If one wants only to specify $m_0$ and a coefficient of variation $c$ \footnote{$c = \sqrt{v_0}/ m_0$.} for $R_t$ \textit{a priori}, some less boring algebra gives:
\begin{align}
\label{eq:elicitationcv}
a_0 &= \frac{m_0^3c^2 + m_0^3 + m_0^2}{m_0^2c^2} \\
b_0 &= \frac{2m_0^2c^2 + m^2 + m}{m_0^2c^2}
\end{align}

\newpage
\section{Preliminary results -- Dengue notifications in Rio de Janeiro}
\label{sec:dataanalysis}

We now show the application of the methods described above to dengue notification data in the city of Rio de Janeiro.
The data consists of weekly dengue notifications for 10 Health Planning Areas (APS, in Portuguese) for the period 2010-2014.
Figure~\ref{fig:casesperAPS} shows the data.
We compute $c_{\alpha}(R_t)$ for the total (sum) number of notifications using the two approaches described, using a generation time of $3$ weeks and choose $a_0$ and $b_0$ such that $E[R_t] = 1/2$ and $Var(R_t) = 2$.
The resulting estimates and $95\%$ confidence bounds are presented in Figure~\ref{fig:confidencetotals}.
Finally, we compute $Pr(R_t > 1)$ using the two distributions discussed here (Figure~\ref{fig:pRt1}).

\section{Final Remarks}
\label{sec:remarks}

Both strategies presented here seem appropriate to the quantification of uncertainty about $R_t$.
The distribution obtained by combining the approach of \citet{mantel} with Bayesian inference for $\theta_t$, however, possesses heavier tails and thus offers more conservative bounds.
Additionally, the ability to specify prior information on $R_t$ based on expert knowledge may prove useful.

\section{Acknowlegements}
LMC is grateful to Dr. Leonardo Bastos for useful discussions.
\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!h]
\setkeys{Gin}{scale=.7}
\begin{center}
\includegraphics{FIGURES/figure-002}
\end{center}
\caption{Total number of dengue notifications per HPA (APS)}
\label{fig:casesperAPS}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!h]
\setkeys{Gin}{scale=.7}
\begin{center}
\includegraphics{FIGURES/figure-003}
\end{center}
\caption{\rr estimates for the total notifications and confidence intervals obtained using the two proposed methods.}
\label{fig:confidencetotals}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!h]
\setkeys{Gin}{scale=.7}
\begin{center}
\includegraphics{FIGURES/figure-004}
\end{center}
\caption{Probability that $R_t \geq 1$.
The interrupted gray line marks the arbitrary value of $0.80$.  
}
\label{fig:pRt1}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\bibliography{lm1}
\end{document}

### 2014-09-09 fit_dengue_my.py
more changes, still not ready...

@@ -2,7 +2,8 @@
 # Multi-year fit of influenza epidemics
 #
 import pyximport;
-pyximport.install(pyimport=True)
+
+pyximport.install(pyimport=False)
 from BIP.Bayes.Melding import FitModel
 from scipy.integrate import odeint
 import scipy.stats as st
@@ -14,41 +15,36 @@ import datetime
 import pandas as pd
 
 
-
-beta = 1.0  #Transmission coefficient
+beta = 1.0  # Transmission coefficient
 b1 = 19.9
-b0 = 1.5  #low beta during winters
-eta = .0  #infectivity of asymptomatic infections relative to clinical ones. FIXED
-epsilon = 2.8  #latency rate
-mu = 0  #nat/mortality rate
+b0 = 1.5  # low beta during winters
+eta = .0  # infectivity of asymptomatic infections relative to clinical ones. FIXED
+epsilon = 2.8  # latency rate
+mu = 0  # nat/mortality rate
 
+tau = 0.2  # recovery rate. FIXED
 
-tau = 1.4  #infectious rate. FIXED
-nflu = 0  #.00142
-s0 = .241  #fraction of susceptibles at the beginning of first epidemic
+s0 = .241  # fraction of susceptibles at the beginning of first epidemic
 s1 = 0.699;
 s2 = 0.699;
-s3 = 0.692;
-s4 = 0.691;
-s5 = 0.655;
-s6 = .233
 
-Ss = {0: s0, 1: s1, 2: s2, 3: s3, 4: s4, 5: s5, 6: s6}  #Multiple Ss map
+
+Ss = {0: s0, 1: s1, 2: s2}  # Multiple Ss map
 r0 = 1.2
 # Initial conditions
-inits = [.9999, .0001, 0]  #initial values for state variables.
+inits = [.9999, .0001, 0]  # initial values for state variables.
 
 
 def model(theta):
-    #setting parameters
-    s0, s1, s2, s3, s4, s5, s6, r0 = theta
-    Ss = {0: s0, 1: s1, 2: s2, 3: s3, 4: s4, 5: s5, 6: s6}  #Multiple Ss map
+    # setting parameters
+    s0, s1, s2, r0 = theta
+    Ss = {0: s0, 1: s1, 2: s2}  #Multiple Ss map
 
     #~ b1=19.9
     def sir(y, t):
         '''ODE model'''
         S, I, R = y
-        b1 = r0 * tau / (As[astep[int(t)]] * Ss[astep[int(t)]])
[…]

### 2014-09-09 fit_dengue_my.py
more changes, still not ready...

@@ -64,8 +64,6 @@ def model(theta):
         inits[0] = Ss[i];
         inits[-1] = 1 - sum(inits[:1])
         Y[t0:tf, :] = odeint(sir, inits, np.arange(t0, tf, 1))  #,tcrit=tcrit)
-        Y[t0:tf, 1] = Y[t0:tf, 1] * As[
-            ycode[int(t0)]]  #Adjusting the output to just the symptomatic I to compare with data
 
     return Y
 

### 2014-09-09 fit_dengue_my.py
more changes, still not ready...

@@ -143,14 +143,14 @@ def year_code(bstep):
 
     return yc
 
-
+# TODO: Fix this based on the actual dates
 def beta_step(timeline):
     """
     Returns the date ranges where beta should be low, i.e., during summers. 
     """
     bstep = np.zeros(len(timeline) + 200)
     for i, d in enumerate(timeline):
-        if d.year == 2009:  # In 209 h1n1 started in the summer
+        if d.year == 2009:  # In 2009 h1n1 started in the summer
             if d.month != 6:
                 bstep[i] = 1
         else:
@@ -172,8 +172,8 @@ def s_index(bstep):
 
 def calc_R0s():
     r0s = []
-    for s in Ss.itervalues():
-        r0s.append(b1 / tau * s)
+    for i in range(len(Ss)):
+        r0s.append(b1 / tau * SS[i])
     return np.array(r0s)
 
 
@@ -186,7 +186,7 @@ def constrain_Re(theta):
 # # running the analysys
 if __name__ == "__main__":
 
-    dt = prepdata('data_Rt_dengue.csv', 0, 243, 1);
+    dt = prepdata('data_Rt_dengue.csv', 0, 243, 1)
     modname = "Dengue_S0"
     #print dt['I'][:,1]
 
@@ -227,27 +227,15 @@ if __name__ == "__main__":
     #~ P.legend(['d','e']+[pnames[1]])
     #~ P.show()
     #Priors and limits for all countries
-    tpars_be = [(0, 1), (0, 1), (0, 1), (0, 1), (0, 1), (0, 1), (0, 1)] + \
-               [(0, 1), (0, 1), (0, 1), (0, 1), (0, 1), (0, 1), (0, 1)] + [(1, .4), (0, 4e-6)]
-    tlims_be = [(0, 1), (0, 1), (0, 1), (0, 1), (0, 1), (0, 1), (0, 1)] + \
-               [(0.1, 1), (0.2, 1), (0.2, 1), (0.2, 1), (0.2, 1), (0.2, 1), (0.1, 1)] + [(1.01, 1.4),
-                                                                                         (0, .00004)]  #beta and k
-    tpars_nl = [(0, .6), (0, .4), (0, .4), (0, .4), (0, .4), (0, .4), (0, .6)] + \
-               [(0, 1), (0, 1), (0, 1), (0, 1), (0, 1), (0, 1), (0, 1)] + [(1, .4), (0, 4e-6)]  #beta and k
[…]

### 2014-09-11 aux/make_Rt_data.R
Adding SEs

@@ -5,11 +5,11 @@
 source("Rt_aux.R")
 source("elicit_Rt_prior.R")
 
-dat <- read.csv("dadojuntos_201433.csv", header=TRUE)
-
+dat <- read.csv("dadojuntos_201433.csv", header = TRUE)
 N <- length(unique(dat$SE))
 cases <- rowSums(matrix(dat$casos, nrow = N))
 pars <- elicit.Rt.hyperpars(m0 = 1/2, v0 = 2)  
 Rtb <- Rt.beta(cases, a0 = pars[1], b0 = pars[2])
-dadoexp <- data.frame(cases, Rtb)
+dadoexp <- data.frame(SE=unique(dat$SE), cases, Rtb)
+
 write.csv(dadoexp, file = "data_Rt_dengue.csv", row.names = FALSE)

### 2014-09-15 draft_lm1.tex
Changing to PLOS template

@@
-\documentclass[12pt]{article}              
-\usepackage[margin=1.5cm]{geometry}
-\usepackage{setspace,relsize}               
-\usepackage{moreverb}                        
-\usepackage{url}
+% Template for PLoS
+% Version 2.0 July 2014
+%
+% To compile to pdf, run:
+% latex plos.template
+% bibtex plos.template
+% latex plos.template
+% latex plos.template
+% dvipdf plos.template
+%
+% % % % % % % % % % % % % % % % % % % % % %
+%
+% -- IMPORTANT NOTE
+%
+% Be advised that this is merely a template 
+% designed to facilitate accurate translation of manuscript content 
+% into our production files. 
+%
+% This template contains extensive comments intended 
+% to minimize problems and delays during our production 
+% process. Please follow the template 
+% whenever possible.
+%
+% % % % % % % % % % % % % % % % % % % % % % % 
+%
+% Once your paper is accepted for publication and enters production, 
+% PLEASE REMOVE ALL TRACKED CHANGES in this file and leave only
+% the final text of your manuscript.
+%
+% DO NOT ADD EXTRA PACKAGES TO THIS TEMPLATE unless absolutely necessary.
+% Packages included in this template are intentionally
+% limited and basic in order to reduce the possibility
+% of issues during our production process.
+%
+% % % % % % % % % % % % % % % % % % % % % % %
+%
+% -- FIGURES AND TABLES
+%
+% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
+% - Figures should be uploaded separately from your manuscript file. 
+% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
+% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
+% See http://www.plosone.org/static/figureGuidelines for PLOS figure guidelines.
+%
+% Tables should be cell-based and may not contain:
+% - tabs/spacing/line breaks within cells to alter layout
+% - vertically-merged cells (no tabular environments within tabular environments, do not use \multirow)
+% - colors, shading, or graphic objects
+% See http://www.plosone.org/static/figureGuidelines#tables for table guidelines.
+%
+% For sideways tables, use the {rotating} package and use \begin{sidewaystable} instead of \begin{table} in the appropriate section. PLOS guidelines do not accomodate sideways figures.
+%
+% % % % % % % % % % % % % % % % % % % % % % % %
+%
+% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
+%
+% IMPORTANT
+% Below are a few tips to help format your equations and other special characters according to our specifications.
+For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://www.plosone.org/static/latexGuidelines
+%
+% Please be sure to include all portions of an equation in the math environment, and for any superscripts or subscripts also include the base number/text.
+For example, use $mathrm{mm}^2$ instead of mm$^2$ (do not use \textsuperscript command).
+%
+% DO NOT USE the \rm command to render mathmode characters in roman font, instead use $\mathrm{}$
+% For bolding characters in mathmode, please use $\mathbf{}$ 
+%
+% Please add line breaks to long equations when possible in order to fit our 2-column layout. 
+%
+% For inline equations, please do not include punctuation within the math environment unless this is part of the equation.
+%
+% For spaces within the math environment please use the \; or \: commands, even within \text{} (do not use smaller spacing as this does not convert well).
+%
+%
+% % % % % % % % % % % % % % % % % % % % % % % %
+
+\documentclass[10pt]{article}
+% amsmath package, useful for mathematical formulas
+\usepackage{amsmath}
+% amssymb package, useful for mathematical symbols
+\usepackage{amssymb}
+% cite package, to clean up citations in the main text. Do not remove.
+\usepackage{cite}
 \usepackage{hyperref}
-\hypersetup{colorlinks=true,citecolor=blue}
-\usepackage{amsmath}
-\usepackage{mathtools} 
-\usepackage{amssymb}
-\usepackage{indentfirst}
-\usepackage[authoryear,round]{natbib}
-\bibliographystyle{apalike}
-\usepackage[pdftex]{lscape}
-% \usepackage{float}
-% \usepackage{longtable}
-% \usetikzlibrary{arrows}
-%%%%
-\title{
-Estimating the Initial Susceptible Fraction in Dengue Dynamics
+% line numbers
+\usepackage{lineno}
+% ligatures disabled
+\usepackage{microtype}
+\DisableLigatures[f]{encoding = *, family = * }
+% rotating package for sideways tables
+%\usepackage{rotating}
+% If you wish to include algorithms, please use one of the packages below.
+Also, please see the algorithm section of our LaTeX guidelines (http://www.plosone.org/static/latexGuidelines) for important information about required formatting.
+%\usepackage{algorithmic}
+%\usepackage{algorithmicx}
+% Use doublespacing - comment out for single spacing
+%\usepackage{setspace} 
+%\doublespacing
+% Text layout
+\topmargin 0.0cm
+\oddsidemargin 0.5cm
+\evensidemargin 0.5cm
+\textwidth 16cm 
+\textheight 21cm
+% Bold the 'Figure #' in the caption and separate it with a period
+% Captions will be left justified
+\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}
+% Use the PLoS provided BiBTeX style
+\bibliographystyle{plos2009}
+% Remove brackets from numbering in List of References
+\makeatletter
+\renewcommand{\@biblabel}[1]{\quad#1.}
+\makeatother
+
+% Leave date blank
+\date{}
+
+\pagestyle{myheadings}
+
+%% Include all macros below. Please limit the use of macros.
+
+%% END MACROS SECTION
+
+\begin{document}
+
+% Title must be 150 characters or less
+\begin{flushleft}
+{\Large
+\textbf{Estimating the Initial Susceptible Fraction in Dengue Dynamics}
 }
-\author{Flavio Code\c{c}o Coelho \\
-School of Applied Mathematics, Get\'ulio Vargas Foundation (FGV) \\
-\& \\
-Luiz Max de Carvalho \\
-Program for Scientific Computing (PROCC), Oswaldo Cruz Foundation}
+% Insert Author names, affiliations and corresponding author email.
+\\
+Flavio Code\c{c}o Coelho$^{1}$, 
+Luiz Max de Carvalho$^{2}$, 
+Author3$^{3,\ast}$
+\\
+\bf{1} School of Applied Mathematics, Get\'ulio Vargas Foundation (FGV), Rio de Janeiro -- RJ, Brazil.
+\\
+\bf{2} Program for Scientific Computing (PROCC), Oswaldo Cruz Foundation, Rio de Janeiro -- RJ, Brazil.
+\\
+$\ast$ E-mail: fccoelho@fgv.br
+\end{flushleft}
 
-\date{\today}
-% Notation defs
-\def \rr {$R_{t}\ $}
+% Please keep the abstract between 250 and 300 words
+\section*{Abstract}
 
-\usepackage{Sweave}
-\begin{document}                                  
-\input{draft_lm1-concordance}
+% Please keep the Author Summary between 150 and 200 words
+% Use first person. PLOS ONE authors please skip this step. 
+% Author Summary not valid for PLOS ONE submissions.   
+\section*{Author Summary}
 
+\section*{Introduction}
+% You may title this section "Methods" or "Models". 
+% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
+% authors may use "Analysis" 
+\section*{Methods}
+\subsection*{The model}
+\subsection*{Parameter estimation using Bayesian melding}
+% Results and Discussion can be combined.
+\section*{Results}
+
+% We only support three levels of headings, please do not create a heading level below \subsubsection.
+\subsection*{Subsection 1}
+
+\section*{Discussion}
+
+% Do NOT remove this, even if you are not including acknowledgments.
+
+\section*{Acknowledgments}
+LMC is grateful to Dr. Leonardo Bastos for useful discussions.
+
+\section*{References}
+\bibliography{lm1}
+\section*{Figure Legends}
+% This section is for figure legends only, do not include
+% graphics in your manuscript file.
 %
-\maketitle
-\begin{abstract}
-We derive confidence/credibility bounds for the uncertainty about the estimate of the effective reproductive number $R_t$ under the assumption that reported disease counts are Poisson distributed.
-It is shown that one can construct a conservative confidence interval for \rr using the induced probability distribution that results from the approach of \citet{mantel}.
-\end{abstract}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Background}
-\label{sec:intro}
+%\begin{figure}
+%\caption{
+%{\bf Bold the first sentence.}  Rest of figure caption.  
+%}
+%\label{Figure_label}
+%\end{figure}
 
-In monitoring of infectious diseases, it is important to assess whether the incidence of a a particular disease is increasing significantly, in order to decide to take preventive measures.
-The effective reproductive number at time $t$, \rr, can be understood as a real-time estimate of the basic reproductive number ($R_{0}$) and is defined as the average number of secondary cases per primary case at time $t$.
-Let $Y_t$ be the number of reported disease cases for a particular time $t \in (0, T)$.
-\citet{nishiura} extend the theory developed by \citet{stallybrass} and propose to estimate \rr as
-\begin{equation}
-\label{eq:Rtestimate}
-R_t = \left( \frac{Y_{t+1}}{Y_t}\right)^{1/n}
-\end{equation}
-where $n$ is taken to be the ratio between the length of reporting interval and the mean generation time of the disease.
-Here we are interested in the simpler case $n=1$.
-If \rr is to be used as a decision tool, one needs to be able to quantify the uncertainty about estimate in equation~\ref{eq:Rtestimate}. 
-Here we detail how to obtain confidence intervals for \rr, exploring two approaches under the assumption that the counts $Y_t$ are Poisson distributed for all $t$.
+\section*{Tables}
+% 
+% See introductory notes if you wish to include sideways tables.
+%
+% NOTE: Please look over our table guidelines at http://www.plosone.org/static/figureGuidelines#tables to make sure that your tables meet our requirements.
+Certain types of spacing, cell merging, and other formatting tricks may have unintended results and will be returned for revision.
+%
+%\begin{table}[!ht]
+%\caption{
+%\bf{Table title}}
+%\begin{tabular}{|c|c|c|}
+%table information
+%\end{tabular}
+%\begin{flushleft}Table caption
+%\end{flushleft}
+%\label{tab:label}
+% \end{table}
 
-\newpage
-\section{Approaches}
+\section*{Supporting Information Legends}
+%
+% Please enter your Supporting Information captions below in the following format:
+%\item{\bf Figure SX. Enter mandatory title here.} Enter optional descriptive information here.
+% 
+%\begin{description}
+%\item {\bf}
+%\item {\bf}
+%\end{description}
 
-\subsection{Ederer \& Mantel (1974)}
-\label{sec:mantel}
-
-First we explore the approach of \citet{mantel}, whose objective is to obtain confidence intervals for the ratio of two Poisson counts. 
-Let $Y_{t} \sim Poisson(\lambda_t)$ and $Y_{t+1} \sim Poisson(\lambda_{t+1})$ and define $S = Y_{t} + Y_{t+1}$.
-The authors then proceed to note that by conditioning on the sum $S$
-\begin{align}
-\label{eq:binlike}
-Y_{t+1} | S &\sim Binomial(S, \theta_t) \\
-\theta_t &= \frac{\lambda_{t+1}}{\lambda_{t} + \lambda_{t+1}}
-\end{align}
-Let $c_{\alpha}(\theta_t) = \{\theta_t^{(L)} , \theta_t^{(U)} \}$ be such that $Pr(\theta_t^{(L)}<\theta_t <\theta_t^{(U)}) = \alpha$.
-Analogously, define $c_{\alpha}(R_t) = \{R_t^{(L)} , R_t^{(U)} \}$ such that $Pr(R_t^{(L)}<R_t<R_t^{(U)}) = \alpha$.
-\citet{mantel} show that one can construct a $100\alpha \%$ confidence interval for \rr by noting that
-\begin{align}
- R_t^{(L)} &= \frac{\theta_t^{(L)}}{(1-\theta_t^{(L)})}\\
- R_t^{(U)} &= \frac{\theta_t^{(U)}}{(1-\theta_t^{(U)})}
-\end{align}
-
-Many authors have derived $c_{\alpha}(\theta_t)$ following orthodox approaches such as those of \citet{wilson} and \citet{clopper}, mainly for simplicity reasons.
-We choose instead to take a Bayesian approach and use the  $100\alpha \%$ posterior credibility interval for $\theta_t$ as $c_{\alpha}(\theta_t)$.
-If we choose the conjugate beta prior with parameters $a_0$ and $b_0$ for the binomial likelihood in (\ref{eq:binlike}), the posterior distribution for $\theta_t$ is
-\begin{equation}
-\label{eq:thetapost}
-p(\theta_t| Y_{t+1}, S) \sim Beta(Y_{t+1} + a_0, Y_t + b_0)
-\end{equation}
-Equation~(\ref{eq:thetapost}) tells us that the induced distribution of $R_t$ is a beta prime with parameters $ a_1 = Y_{t+1} + a_0$ and $b_1 =  Y_t + b_0$.
-The density of this induced distribution  $f_P(R_t)$ is then 
-\begin{equation}
-\label{eq:densityMantel}
-f_P(R_t| a_1, b_1) = \frac{\Gamma(a_1 + b_1)}{\Gamma(a_1)\Gamma(b_1)} R_t^{a_1 - 1} (1 + R_t)^{-(a_1 + b_1)}
-\end{equation}
-Thus, the expectation of \rr under $f_P$ is $a_1/(b_1 - 1)$ and its variance is $a_1(a_1 + b_1 - 1)/\left((b_1 - 2)(b_1 - 1)^2 \right) $.
-
-Since when $R_t > 1$ indicates sustained transmission, one may be interested in computing the probability of this event.
-By noting that
-\begin{align}
-\label{cumprobMantel}
-Pr(R_t > 1) &= 1 - \int_0^1 f_P(r)dr \\
-            &= 1- Pr(\theta_t < \frac{1}{2})
-\end{align}
-one can avoid dealing with the density in~(\ref{eq:densityMantel}) directly.
-
-% \subsection{Nishiura et al. (2010)}
-% \label{sec:nishiura}
-% 
-% An alternative approach is to use the conditional distribution of \rr on $Y_{t+1}$ and $Y_t$ as defined in equation A7 of \citet{nishiura}:
-% \begin{equation}
-% \label{eq:unorm}
-% f_{R}(R_{t}) = (Y_tR_{t})^{Y_{t+1}} e^{-Y_tR_{t}}
-% \end{equation}
-% Noticing the kernel of (\ref{eq:unorm}) is that of a gamma distribution with $a_2 = Y_{t+1}+1$ and $b_2 = Y_t$, we obtain a proper density from which to construct $c_{\alpha}(R_t)$, simply by computing the appropriate quantiles of said distribution.
-%  This density is
-% \begin{equation}
-% \label{eq:densityNishiura}
-% f_N(R_t| a_2, b_2) =  \frac{b_2^{a_2}}{\Gamma(a_2)} R_t^{a_2-1} e^{-b_2 R_t}
-% \end{equation}
-% Once again $Pr(R_t>1)$ can be easily computed using
-% \begin{equation}
-% \label{eq:cdfNishiura}
-% P(R_t > 1) = 1-F_N(1) = 1- \frac{1}{\Gamma(a_2)} \gamma(a_2, b_2)
-% \end{equation}
-% where $F_N(R_t)$ is the distribution function of $R_t$ and $\gamma(\cdot)$ is the incomplete gamma function.
-
-\subsection{A remark on prior distributions and tail behaviour}
-\label{sec:tails}
-
-In order to decide which approach to take, it may be of use analysing the variance and tail behaviour of the derived distributions for \rr. 
-Consider the case of using a flat $Uniform(0, 1)$ prior for $\theta_t$ in Section~\ref{sec:mantel}.
-With $a_0 = b_0 = 1$, $a_1 = a_2$ and $b_1 = b_2 + 1$.
-The beta prime (inverse beta distribution) will have heavier tails compared to the conditional distribution in (\ref{eq:densityNishiura}), thus providing more conservative confidence/credibility intervals.
-As a side note, the Bayesian approach presented in Section~\ref{sec:mantel} will give similar results to those of \citet{wilson} and \citet{wilson} for $Y_{t+1}$ and $Y_t >> 1$.
-Under the flat  uniform prior for $\theta_t$, the Bayesian posterior credibility interval is nearly indistinguishable from the confidence interval proposed by \citet{clopper} for $Y_{t+1}, Y_t > 20$.
-Note that the $Beta(1, 1)$ uniform prior for $\theta_t$ constitutes a poor prior choice mainly because the induced distribution for \rr is only well-defined for $b_0 > 2$.
-
-The main advange of the Bayesian approach is that one can devise prior distributions for $\theta_t$ taking advantage of the intuitive parametrisation and flexibility of the beta family of distributions.
-Prior elicitation can also be done for \rr and the hyperparameters directly plugged into the prior for $\theta_t$. 
-One can, for example, choose a priori mean and variance for \rr and find $a_0$ and $b_0$ that satisfy those conditions.
-Let $m_0$ and $v_0$ be the prior expectation and variance for $R_t$. 
-After some tedious algebra one finds
-\begin{align}
-\label{eq:elicitation}
-a_0 &= \frac{m_0v_0 + m_0^3 + m_0^2}{v_0} \\
-b_0 &= \frac{2v_0 + m_0^2 + m_0}{v_0}
-\end{align}
-If one wants only to specify $m_0$ and a coefficient of variation $c$ \footnote{$c = \sqrt{v_0}/ m_0$.} for $R_t$ \textit{a priori}, some less boring algebra gives:
-\begin{align}
-\label{eq:elicitationcv}
-a_0 &= \frac{m_0^3c^2 + m_0^3 + m_0^2}{m_0^2c^2} \\
-b_0 &= \frac{2m_0^2c^2 + m^2 + m}{m_0^2c^2}
-\end{align}
-
-\newpage
-\section{Preliminary results -- Dengue notifications in Rio de Janeiro}
-\label{sec:dataanalysis}
-
-We now show the application of the methods described above to dengue notification data in the city of Rio de Janeiro.
-The data consists of weekly dengue notifications for 10 Health Planning Areas (APS, in Portuguese) for the period 2010-2014.
-Figure~\ref{fig:casesperAPS} shows the data.
-We compute $c_{\alpha}(R_t)$ for the total (sum) number of notifications using the two approaches described, using a generation time of $3$ weeks and choose $a_0$ and $b_0$ such that $E[R_t] = 1/2$ and $Var(R_t) = 2$.
-The resulting estimates and $95\%$ confidence bounds are presented in Figure~\ref{fig:confidencetotals}.
-Finally, we compute $Pr(R_t > 1)$ using the two distributions discussed here (Figure~\ref{fig:pRt1}).
-
-\section{Final Remarks}
-\label{sec:remarks}
-
-Both strategies presented here seem appropriate to the quantification of uncertainty about $R_t$.
-The distribution obtained by combining the approach of \citet{mantel} with Bayesian inference for $\theta_t$, however, possesses heavier tails and thus offers more conservative bounds.
-Additionally, the ability to specify prior information on $R_t$ based on expert knowledge may prove useful.
-
-\section{Acknowlegements}
-LMC is grateful to Dr. Leonardo Bastos for useful discussions.
-\newpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[!h]
-\setkeys{Gin}{scale=.7}
-\begin{center}
-\includegraphics{FIGURES/figure-002}
-\end{center}
-\caption{Total number of dengue notifications per HPA (APS)}
-\label{fig:casesperAPS}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[!h]
-\setkeys{Gin}{scale=.7}
-\begin{center}
-\includegraphics{FIGURES/figure-003}
-\end{center}
-\caption{\rr estimates for the total notifications and confidence intervals obtained using the two proposed methods.}
-\label{fig:confidencetotals}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[!h]
-\setkeys{Gin}{scale=.7}
-\begin{center}
-\includegraphics{FIGURES/figure-004}
-\end{center}
-\caption{Probability that $R_t \geq 1$.
-The interrupted gray line marks the arbitrary value of $0.80$.  
-}
-\label{fig:pRt1}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-\bibliography{lm1}
 \end{document}

### 2014-09-15 S1_lm1.tex
More additions for PLOS template

\documentclass[12pt]{article}              
\usepackage[margin=1.5cm]{geometry}
\usepackage{setspace,relsize}               
\usepackage{moreverb}                        
\usepackage{url}
\usepackage{hyperref}
\hypersetup{colorlinks=true,citecolor=blue}
\usepackage{amsmath}
\usepackage{mathtools} 
\usepackage{amssymb}
\usepackage{indentfirst}
\usepackage[authoryear,round]{natbib}
\bibliographystyle{apalike}
\usepackage[pdftex]{lscape}
\usepackage[toc,page]{appendix}
% \usepackage{float}
% \usepackage{longtable}
% \usetikzlibrary{arrows}
%%%%
\title{
Supporting Information to ¨Estimating the Initial Susceptible Fraction in Dengue Dynamics"
}           
\author{Flavio Code\c{c}o Coelho \\
\\
\& \\
Luiz Max de Carvalho \\
}

\date{\today}
% Notation defs
\def \rr {$R_{t}\ $}

%\usepackage{Sweave}
\begin{document}                                  
% \input{draft_lm1-concordance}

%
\maketitle

In monitoring of infectious diseases, it is important to assess whether the incidence of a a particular disease is increasing significantly, in order to decide to take preventive measures.
The effective reproductive number at time $t$, \rr, can be understood as a real-time estimate of the basic reproductive number ($R_{0}$) and is defined as the average number of secondary cases per primary case at time $t$.
Here we derive confidence/credibility bounds for the uncertainty about the estimate of the effective reproductive number $R_t$ under the assumption that reported disease counts are Poisson distributed.
It is shown that one can construct a conservative confidence interval for \rr using the induced probability distribution that results from the approach of \citet{mantel}.

Let $Y_t$ be the number of reported disease cases for a particular time $t \in (0, T)$.
\citet{nishiura} extend the theory developed by \citet{stallybrass} and propose to estimate \rr as
\begin{equation}
\label{eq:Rtestimate}
R_t = \left( \frac{Y_{t+1}}{Y_t}\right)^{1/n}
\end{equation}
where $n$ is taken to be the ratio between the length of reporting interval and the mean generation time of the disease.
Here we are interested in the simpler case $n=1$.
If \rr is to be used as a decision tool, one needs to be able to quantify the uncertainty about estimate in equation~\ref{eq:Rtestimate}. 
Here we detail how to obtain confidence intervals for \rr, exploring two approaches under the assumption that the counts $Y_t$ are Poisson distributed for all $t$.

\section*{The approach Ederer \& Mantel (1974)}

\label{sec:mantel}
First we explore the approach of \citet{mantel}, whose objective is to obtain confidence intervals for the ratio of two Poisson counts. 
Let $Y_{t} \sim Poisson(\lambda_t)$ and $Y_{t+1} \sim Poisson(\lambda_{t+1})$ and define $S = Y_{t} + Y_{t+1}$.
The authors then proceed to note that by conditioning on the sum $S$
\begin{align}
\label{eq:binlike}
Y_{t+1} | S &\sim Binomial(S, \theta_t) \\
\theta_t &= \frac{\lambda_{t+1}}{\lambda_{t} + \lambda_{t+1}}
\end{align}
Let $c_{\alpha}(\theta_t) = \{\theta_t^{(L)} , \theta_t^{(U)} \}$ be such that $Pr(\theta_t^{(L)}<\theta_t <\theta_t^{(U)}) = \alpha$.
Analogously, define $c_{\alpha}(R_t) = \{R_t^{(L)} , R_t^{(U)} \}$ such that $Pr(R_t^{(L)}<R_t<R_t^{(U)}) = \alpha$.
\citet{mantel} show that one can construct a $100\alpha \%$ confidence interval for \rr by noting that
\begin{align}
 R_t^{(L)} &= \frac{\theta_t^{(L)}}{(1-\theta_t^{(L)})}\\
 R_t^{(U)} &= \frac{\theta_t^{(U)}}{(1-\theta_t^{(U)})}
\end{align}

Many authors have derived $c_{\alpha}(\theta_t)$ following orthodox approaches such as those of \citet{wilson} and \citet{clopper}, mainly for simplicity reasons.
We choose instead to take a Bayesian approach and use the  $100\alpha \%$ posterior credibility interval for $\theta_t$ as $c_{\alpha}(\theta_t)$.
If we choose the conjugate beta prior with parameters $a_0$ and $b_0$ for the binomial likelihood in (\ref{eq:binlike}), the posterior distribution for $\theta_t$ is
\begin{equation}
\label{eq:thetapost}
p(\theta_t| Y_{t+1}, S) \sim Beta(Y_{t+1} + a_0, Y_t + b_0)
\end{equation}
Equation~(\ref{eq:thetapost}) tells us that the induced distribution of $R_t$ is a beta prime with parameters $ a_1 = Y_{t+1} + a_0$ and $b_1 =  Y_t + b_0$.
The density of this induced distribution  $f_P(R_t)$ is then 
\begin{equation}
\label{eq:densityMantel}
f_P(R_t| a_1, b_1) = \frac{\Gamma(a_1 + b_1)}{\Gamma(a_1)\Gamma(b_1)} R_t^{a_1 - 1} (1 + R_t)^{-(a_1 + b_1)}
\end{equation}
Thus, the expectation of \rr under $f_P$ is $a_1/(b_1 - 1)$ and its variance is $a_1(a_1 + b_1 - 1)/\left((b_1 - 2)(b_1 - 1)^2 \right) $.

Since when $R_t > 1$ indicates sustained transmission, one may be interested in computing the probability of this event.
By noting that
\begin{align}
\label{cumprobMantel}
Pr(R_t > 1) &= 1 - \int_0^1 f_P(r)dr \\
            &= 1- Pr(\theta_t < \frac{1}{2})
\end{align}
one can avoid dealing with the density in~(\ref{eq:densityMantel}) directly.

\subsection*{A remark on prior distributions and tail behaviour}
\label{sec:tails}

In order to decide which approach to take, it may be of use analysing the variance and tail behaviour of the derived distributions for \rr. 
Consider the case of using a flat $Uniform(0, 1)$ prior for $\theta_t$ in Section~\ref{sec:mantel}.
With $a_0 = b_0 = 1$, $a_1 = a_2$ and $b_1 = b_2 + 1$.
The beta prime (inverse beta distribution) will have heavier tails compared to the conditional distribution in (\ref{eq:densityNishiura}), thus providing more conservative confidence/credibility intervals.  
As a side note, the Bayesian approach presented in Section~\ref{sec:mantel} will give similar results to those of \citet{wilson} and \citet{wilson} for $Y_{t+1}$ and $Y_t >> 1$.
Under the flat  uniform prior for $\theta_t$, the Bayesian posterior credibility interval is nearly indistinguishable from the confidence interval proposed by \citet{clopper} for $Y_{t+1}, Y_t > 20$.
Note that the $Beta(1, 1)$ uniform prior for $\theta_t$ constitutes a poor prior choice mainly because the induced distribution for \rr is only well-defined for $b_0 > 2$.

The main advange of the Bayesian approach is that one can devise prior distributions for $\theta_t$ taking advantage of the intuitive parametrisation and flexibility of the beta family of distributions.
Prior elicitation can also be done for \rr and the hyperparameters directly plugged into the prior for $\theta_t$. 
One can, for example, choose a priori mean and variance for \rr and find $a_0$ and $b_0$ that satisfy those conditions.
Let $m_0$ and $v_0$ be the prior expectation and variance for $R_t$. 
After some tedious algebra one finds
\begin{align}
\label{eq:elicitation}
a_0 &= \frac{m_0v_0 + m_0^3 + m_0^2}{v_0} \\
b_0 &= \frac{2v_0 + m_0^2 + m_0}{v_0}
\end{align}
If one wants only to specify $m_0$ and a coefficient of variation $c$ \footnote{$c = \sqrt{v_0}/ m_0$.} for $R_t$ \textit{a priori}, some less boring algebra gives:
\begin{align}
\label{eq:elicitationcv}
a_0 &= \frac{m_0^3c^2 + m_0^3 + m_0^2}{m_0^2c^2} \\
b_0 &= \frac{2m_0^2c^2 + m^2 + m}{m_0^2c^2}
\end{align}
\newpage
\bibliography{lm1}
\end{document}

### 2014-09-19 fit_dengue_my.py
script adaptado e funcionando. fit perfeito.

@@ -6,12 +6,14 @@ import pyximport;
 pyximport.install(pyimport=False)
 from BIP.Bayes.Melding import FitModel
 from scipy.integrate import odeint
+from scipy.interpolate import interp1d
 import scipy.stats as st
 import numpy as np
 import pylab as P
 import copy
 from collections import defaultdict
 import datetime
+import sys
 import pandas as pd
 
 
@@ -22,48 +24,55 @@ eta = .0  # infectivity of asymptomatic infections relative to clinical ones. FI
 epsilon = 2.8  # latency rate
 mu = 0  # nat/mortality rate
 
-tau = 0.2  # recovery rate. FIXED
+tau = 1  # recovery rate. FIXED
 
-s0 = .241  # fraction of susceptibles at the beginning of first epidemic
-s1 = 0.699;
-s2 = 0.699;
-
+N = 1  # Population of Rio
+s0 = 0.9999*N  # fraction of susceptibles at the beginning of first epidemic
+s1 = 0.999*N;
+s2 = 0.999*N;
 
 Ss = {0: s0, 1: s1, 2: s2}  # Multiple Ss map
-r0 = 1.2
-# Initial conditions
-inits = [.9999, .0001, 0]  # initial values for state variables.
 
+
+# Initial conditions
+inits = np.array([s0, 100, 0.0])  # initial values for state variables.
 
 def model(theta):
     # setting parameters
-    s0, s1, s2, r0 = theta
-    Ss = {0: s0, 1: s1, 2: s2}  #Multiple Ss map
+    s0, s1, s2 = theta
+    Ss = {0: s0, 1: s1, 2: s2}  # Multiple Ss map
 
-    #~ b1=19.9
+    # ~ b1=19.9
     def sir(y, t):
         '''ODE model'''
         S, I, R = y
-        b1 = r0 * tau / (Ss[ycode[int(t)]])
-        beta = b1 if bstep[int(t)] else b0
+        
+        beta = 0 if t > 242 else  iRt(t) * tau/ N#(Ss[ycode[int(t)]])
 
         #~ if (calc_R0s()<1.1).any():
         #~ print "flat"
-        lamb = beta * I * S
+        lamb = (beta * I * S)
         return [-lamb,  #dS/dt
                 lamb - tau * I,  #dI/dt
                 tau * I,  #dR/dt
         ]
 
     Y = np.zeros((wl, 3))
+    # Initial conditions
+
[…]

### 2014-09-19 draft_lm1.tex
minor changes to d manuscript

@@
 \\
-Flavio Code\c{c}o Coelho$^{1}$, 
+Flavio Code\c{c}o Coelho$^{1\ast}$, 
 Luiz Max de Carvalho$^{2}$, 
-Author3$^{3,\ast}$
 \\
-\bf{1} School of Applied Mathematics, Get\'ulio Vargas Foundation (FGV), Rio de Janeiro -- RJ, Brazil.
+\bf{1} Escola de Matem\'atica Aplicada , Funda\c{c}\~ao Get\'ulio Varga (FGV), Rio de Janeiro -- RJ, Brazil.
 \\
-\bf{2} Program for Scientific Computing (PROCC), Oswaldo Cruz Foundation, Rio de Janeiro -- RJ, Brazil.
+\bf{2} Programa de Computa\c{c}\~ao Cient\'ifica (PROCC), Funda\c{c}\~ao Oswaldo Cruz, Rio de Janeiro -- RJ, Brazil.
 \\
@@
 \section*{Abstract}
-
+Our paper is just awesome.
 % Please keep the Author Summary between 150 and 200 words
@@
 \section*{Results}
-
 % We only support three levels of headings, please do not create a heading level below \subsubsection.
-\subsection*{Subsection 1}
-
+%\subsection*{Subsection 1}
 \section*{Discussion}
-
 % Do NOT remove this, even if you are not including acknowledgments.
-
 \section*{Acknowledgments}
-LMC is grateful to Dr. Leonardo Bastos for useful discussions.
-
+LMC is grateful to Dr. Leonardo Bastos for useful discussions on the estimation of $R(t)$.
 \section*{References}

### 2014-09-22 fit_dengue_my.py
Atualizei o script de plotagem

@@ -247,8 +247,8 @@ if __name__ == "__main__":
     tpars = [(9, 1), (9, 1), (9, 1)]
     tlims = [(0, 1), (0, 1), (0, 1)]
 
-    F = FitModel(2000, model, inits, tf, tnames, pnames,
-                 wl, nw, verbose=1, burnin=200, constraints=[])
+    F = FitModel(10000, model, inits, tf, tnames, pnames,
+                 wl, nw, verbose=1, burnin=2000, constraints=[])
     F.set_priors(tdists=nt * [st.beta],
                  tpars=tpars,
                  tlims=tlims,
@@ -257,4 +257,4 @@ if __name__ == "__main__":
     F.run(dt, 'DREAM', likvar=1e-7, pool=False, ew=0, adjinits=False, dbname=modname, monitor=['I'])
     #~ print F.AIC, F.BIC, F.DIC
     #print F.optimize(data=dt,p0=[s0,s1,s2], optimizer='scipy',tol=1e-55, verbose=1, plot=1)
-    F.plot_results(['I'], dbname=modname, savefigs=1)
+    F.plot_results(['S', 'I'], dbname=modname, savefigs=1)

### 2014-09-22 fit_dengue_my_plot.py

@@ -15,6 +15,9 @@ import Gnuplot
 from scipy.stats import gaussian_kde
 import datetime
 import matplotlib.collections as collections
+import seaborn as sns
+import pandas as pd
+from sqlalchemy import create_engine
 
 from BIP.Bayes.PlotMeld import violin_plot, pred_new_cases, peakdet, plot_par_violin
 
@@ -34,7 +37,20 @@ def _read_results(nam):
         pt.append(a)
         series.append(b)
         predseries.append(pred)
+
     return pt, series, predseries, obs, nf
+
+def read_data(dbname):
+    """
+    Returns dataframes from tables of the sqlite database
+    :param dbname: sqlite3 database name
+    :return: list of pandas dataframes
+    """
+    eng = create_engine('sqlite:///{}.sqlite'.format(dbname))
+    df_data = pd.read_sql_table('data', eng, index_col=['pk', 'time'], parse_dates={'time': '%Y-%m-%d %H:%M:%S'})
+    df_pt = pd.read_sql_table('post_theta', eng, index_col=['pk', 'time'], parse_dates={'time': '%Y-%m-%d %H:%M:%S'})
+    df_series = pd.read_sql_table('series', eng, index_col=['pk', 'time'], parse_dates={'time': '%Y-%m-%d %H:%M:%S'})
+    return df_data, df_pt, df_series
 
 
 def create_tex_table(nms):
@@ -81,16 +97,16 @@ def plot_series2(tim, obs, series, names=[], title='Simulated vs Observed series
     else:
         lag = int(lag) * wl
     for i, n in enumerate(names):
-        ax = P.gca()  #fig.add_subplot(len(names), 1, i+1)
-        #~ if isinstance(tim[0], datetime.date):
-        #~ ax.xaxis_date()
+        fig = P.figure()
+        ax = P.gca()
         co = c.next()
         labs = ['Influenzanet', 'EISN']
         if n in obs:
-            #print obs[n].shape
-            for i in range(obs[n].shape[1]):
-                ax.plot(tim, obs[n][:len(tim), i], c.next() + s.next(), alpha=0.7, label=labs[i])
-                #print len(tim),  ls
+            if len(obs[n].shape) > 1:
+                for i in range(obs[n].shape[1]):
[…]

### 2014-09-22 draft_lm1.tex
updated the introduction

@@
 \usepackage{microtype}
+\usepackage{todonotes}
 \DisableLigatures[f]{encoding = *, family = * }
@@
 \\
-\bf{1} Escola de Matem\'atica Aplicada , Funda\c{c}\~ao Get\'ulio Varga (FGV), Rio de Janeiro -- RJ, Brazil.
+\bf{1} Escola de Matem\'atica Aplicada , Funda\c{c}\~ao Getulio Vargas (FGV), 
+Rio de Janeiro -- RJ, Brazil.
 \\
@@
 \section*{Introduction}
-% You may title this section "Methods" or "Models". 
-% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
-% authors may use "Analysis" 
+When modeling infecctious disease spread using dynamic models, the initial 
+conditions of the system, although key to match the simulated dynamics to 
+observed data, are frequently overlooked. If the model is about a disease 
+invading a new population one can simply assume the vast majority of the 
+population to be susceptible. However for a disease which is well in its way to 
+endemicity but not yet stable, it can be very hard to determine the 
+immunological structure of the population without resorting to expensive large 
+scale prevalence tests.
+
+In the particular case of Dengue fever in Brasil, with four circulating 
+serotypes, it can be very challenging to parameterize a simulation starting at 
+any time other than when the date the disease was reintroduced in the country 
+in the late 1980s. For SIR-like models, the relevant initial values are the 
+number of infectious and susceptible  individuals. The number of infectious 
+individuals are less challenging to determine, because the get very small 
+during the winters and can be derived from incidence data  with is monitored 
+continuously. The real challenge is to determine the number of susceptibles to 
+a particular strain right at the beginning of the dengue season.
+
+Multiple factors contribute to the dificulty in estimating the number of 
+susceptibles in the population, first and foremost is the fact that dengue 
+confers permanente immunity within serotypes but with only short-lived 
+cross-immunity. Secondly, there are a larger number of asymptomatic cases in 
+any epidemic, which nevertheless acquire immunity. Lastly Demography and 
+migrations affect the number of susceptible in ways which are not easy to 
+determine.
+
+The initial number of susceptibles -- $S_0$ -- is a a very powerful parameter 
+with the regards to epidemic risk. Its value affects directly the local $R_0$, 
+thus being intrumental in predicting epidemics, but it also affects the speed 
+and size of the epidemic.
+
+In this paper, we estimate the initial number of susceptibles right before 
+every epidemic in the last 18 years. For this we use a simplified model of 
+dengue transmission in which the vector is left out. We also model the mix of 
+circulating serotypes without discriminating between them. We then fit the 
+model output to observed data to estimate the $S_0$ before every 
+recorded epidemic in the Brazilian city of Rio de Janeiro. Rio de Janeiro has 
+bee through many Dengue epidemics of various serotypes and sizes.
+
 \section*{Methods}
+\subsection*{The data}
+The Data used to fit the model consists of weekly incidence from 1996 to 2014 
+in the city of Rio de Janeiro, Brazil. From this data, we calculated series 
+effective reproduction number, $R_t$, time series using the method described by 
+Nishiura et al.\cite{nishiura}.\todo[inline]{Add here more information about 
+the calculation of Rt such as generation time used, moving windows, etc}
 \subsection*{The model}
@@
 LMC is grateful to Dr. Leonardo Bastos for useful discussions on the estimation of $R(t)$.
-\section*{References}
+
 \bibliography{lm1}

### 2014-09-22 draft_lm1.tex
added more refs

@@
 cross-immunity. Secondly, there are a larger number of asymptomatic cases in 
-any epidemic, which nevertheless acquire immunity. Lastly Demography and 
-migrations affect the number of susceptible in ways which are not easy to 
+any epidemic, which nevertheless acquire immunity, Thirdly, under-reporting is 
+a serious issue in Brazil. Duarte and Franca\cite{duarte_data_2006}, estimated 
+the sensitivity of dengue reporting in Belo-Horizonte, Brazil to be of 63\%, 
+meaning that aproximately 37\% of the cases go unreported. Lastly, demography 
+and migrations affect the number of susceptible in ways which are not easy to 
 determine.

### 2014-09-22 aux/plot_serotypes.R
added some data and plotting code for serotypes 1-3 in the period 1986-2009

@@ -1,0 +1,3 @@
+dt <- read.table("..//denv_serotypes_Rio.csv", sep = "\t", header = TRUE)
+matplot(dt, type = "l", lwd = 2, col = 1:3, lty =1)
+legend(x="topleft", legend = c("DENV-1", "DENV-2", "DENV-3"), col = 1:3, lty = 1, bty = "n")

### 2014-09-22 draft_lm1.tex
working on the model section

--- 2014-09-22 draft_lm1.tex
@@
 conditions of the system, although key to match the simulated dynamics to 
-observed data, are frequently overlooked. If the model is about a disease 
+observed data, are frequently overlooked.
+If the model is about a disease 
 invading a new population one can simply assume the vast majority of the 
-population to be susceptible. However for a disease which is well in its way to 
+population to be susceptible.
+However for a disease which is well in its way to 
 endemicity but not yet stable, it can be very hard to determine the 
@@
 
-In the particular case of Dengue fever in Brasil, with four circulating 
+In the particular case of Dengue fever in Brazil, with four circulating 
 serotypes, it can be very challenging to parameterize a simulation starting at 
 any time other than when the date the disease was reintroduced in the country 
-in the late 1980s. For SIR-like models, the relevant initial values are the 
-number of infectious and susceptible  individuals. The number of infectious 
+in the late 1980s.
+For SIR-like models, the relevant initial values are the 
+number of infectious and susceptible  individuals. 
+The number of infectious 
 individuals are less challenging to determine, because the get very small 
 during the winters and can be derived from incidence data  with is monitored 
-continuously. The real challenge is to determine the number of susceptibles to 
+continuously.
+The real challenge is to determine the number of susceptibles to 
 a particular strain right at the beginning of the dengue season.
@@
 susceptibles in the population, first and foremost is the fact that dengue 
-confers permanente immunity within serotypes but with only short-lived 
-cross-immunity. Secondly, there are a larger number of asymptomatic cases in 
-any epidemic, which nevertheless acquire immunity. Lastly Demography and 
+confers permanent immunity within serotypes but with only short-lived 
+cross-immunity.
+Secondly, there are a larger number of asymptomatic cases in 
+any epidemic, which nevertheless acquire immunity.
+Lastly Demography and 
 migrations affect the number of susceptible in ways which are not easy to 
@@
 The initial number of susceptibles -- $S_0$ -- is a a very powerful parameter 
-with the regards to epidemic risk. Its value affects directly the local $R_0$, 
+with regard to epidemic risk.
+Its value affects directly the local $R_0$, 
 thus being intrumental in predicting epidemics, but it also affects the speed 
@@
 In this paper, we estimate the initial number of susceptibles right before 
-every epidemic in the last 18 years. For this we use a simplified model of 
-dengue transmission in which the vector is left out. We also model the mix of 
-circulating serotypes without discriminating between them. We then fit the 
+every epidemic in the last 18 years.
+For this we use a simplified model of 
+dengue transmission based on a Susceptible-infecctious-Removed (SIR) model.
+We then fit the 
 model output to observed data to estimate the $S_0$ before every 
@@
 \subsection*{The data}
-The Data used to fit the model consists of weekly incidence from 1996 to 2014 
-in the city of Rio de Janeiro, Brazil. From this data, we calculated series 
-effective reproduction number, $R_t$, time series using the method described by 
-Nishiura et al.\cite{nishiura}.\todo[inline]{Add here more information about 
-the calculation of Rt such as generation time used, moving windows, etc}
+
+The data used to fit the model consists of weekly incidence from 1996 to 2014 
+in the city of Rio de Janeiro, Brazil.
+From this series, we calculated the effective reproduction number, $R_t$, according to the expression given by Nishiura et al.~\cite{nishiura}.
+The method assumes the disease counts $Y(t)$ are Poisson random variables such that $R_t$ is 
+\[ R_t = \left(\frac{Y(t + 1)}{Y(t)}\right)^{1/n}\]
+where $n$ is the ratio between the length of reporting interval and the mean generation time of the disease.
+In this paper we obtain the estimates of $R_t$ using a moving window of three weeks, which roughly corresponds to the generation time of dengue, and thus $n = 1$.
+We provide more detail on the estimation and confidence/credibility intervals in Text S1.
+
 \subsection*{The model}
-\subsection*{Parameter estimation using Bayesian melding}
-% Results and Discussion can be combined.
+
+A Susceptible-Infectious-Removed (SIR) model is proposed to model dengue dynamics.
+In the traditional formulation of the model, transmission is governed by a transmission rate $\beta$ and recovery happens at a rate $\tau$.
+As the epidemic progresses, the effective transmission  rate changes and is
+\begin{equation}
+ \label{eq:effbeta}
+ \beta(t) = R_t	\cdot\tau
+\end{equation}
+
+We parameterize our model with the varying transmission using the system of ordinary differential equations
+\begin{align}
+ \frac{dS}{dt} &= -\beta(t)SI \\
+ \frac{dI}{dt} &= \beta(t)SI - \tau I&\\
+ \frac{dR}{dt} &= \tau I&
+\end{align}
+where $S + I + R = 1 \: \forall\: t$. % \in [T_0, T_1]$. 
+This is a rather simplified model, in which, for instance, the vector is left out.
+Also, although there probably are multiple circulating serotypes, our approach does not discriminate between them.
+
+\subsection*{Bayesian parameter estimation}
+For the fitting procedure the time series were normalized to lie on the $[0,1]$ interval, therefore avoiding scaling issues.
+% Results and Discussion can be combined.xx
 \section*{Results}

### 2014-09-22 aux/Rt_aux.R
New implementation of R_t to also use the posterior mean

@@ -28,8 +28,9 @@ Rt.beta <- function(d, gt = 3, alpha = .95, a0 = 2 , b0 = 3){
   K <- length(ac)
   jk1 <- ac[2:K] # J[k+1] 
   jk  <- ac[1:(K-1)]# J[k]  
-  Rt <- NA
+  Rt <- Rt2 <- NA
   Rt[(gt + 1):N] <- jk1/jk
+  Rt2[(gt + 1):N] <- (jk1 + a0)/(jk + b0 - 1)
   CIs <- data.frame (p1 = rep(NA, N), lwr = rep(NA, N), upr = rep(NA, N) )
   ## p1 = Pr(R>1) 
   for( k in 1: (N-gt)){
@@ -37,7 +38,7 @@ Rt.beta <- function(d, gt = 3, alpha = .95, a0 = 2 , b0 = 3){
     CIs[k+gt, 2:3] <- ll(betaconf(alpha = alpha, x = jk1[k], 
                                   n = jk1[k] + jk[k], a = a0, b = b0 ))
   } 
-  return(data.frame(Rt, CIs))
+  return(data.frame(Rt, Rt2, CIs))
 }
 ###################################################
 Rt.gamma <- function(d, gt = 3, alpha = .95){

### 2014-09-22 aux/plot_serotypes.R
new serotype plot

@@ -1,3 +1,4 @@
 dt <- read.table("..//denv_serotypes_Rio.csv", sep = "\t", header = TRUE)
-matplot(dt, type = "l", lwd = 2, col = 1:3, lty =1)
+matplot(1993:2010, dt, type = "l", lwd = 2, col = 1:3,
+        lty = 1, xlab = "Time (years)", ylab = "Viral isolation")
 legend(x="topleft", legend = c("DENV-1", "DENV-2", "DENV-3"), col = 1:3, lty = 1, bty = "n")

### 2014-09-24 draft_lm1.tex
updated the introduction and references

@@
 susceptibles in the population, first and foremost is the fact that dengue
-confers permanent immunity within serotypes but with only short-lived 
-cross-immunity.
-Secondly, there are a larger number of asymptomatic cases in 
-any epidemic, which nevertheless acquire immunity.
-Lastly Demography and 
-migrations affect the number of susceptible in ways which are not easy to 
+
+confers permanente immunity within serotypes but with only short-lived 
+cross-immunity. Secondly, there are a larger number of asymptomatic cases in 
+any epidemic, which nevertheless acquire immunity, Thirdly, under-reporting is 
+a serious issue in Brazil. Duarte and Franca\cite{duarte_data_2006}, estimated 
+the sensitivity of dengue reporting in Belo-Horizonte, Brazil to be of 63\%, 
+meaning that aproximately 37\% of the cases go unreported. Lastly, demography 
+and migrations affect the number of susceptible in ways which are not easy to
 determine.
@@
 
-In this paper, we estimate the initial number of susceptibles right before 
-every epidemic in the last 18 years.
-For this we use a simplified model of 
+In this paper, we estimate the number of susceptibles, allowing for 
+discontinuities right before every epidemic in the last 18 years. THis way we 
+can account for replacement of the predominant serotype, which can redefine 
+susceptibility for the following season.
+
+To accomplish this we use a simplified model of 
 dengue transmission based on a Susceptible-infecctious-Removed (SIR) model. 
-We then fit the 
-model output to observed data to estimate the $S_0$ before every 
-recorded epidemic in the Brazilian city of Rio de Janeiro. Rio de Janeiro has 
-bee through many Dengue epidemics of various serotypes and sizes.
+Since Dengue is not caused by a single pathogen, but by four separate strains 
+with limited cross-immunity, when we talk about number of susceptibles, in the 
+context of a single serotype SIR model, we are actually referring to the sum of 
+the susceptibles to the currently circulating serotypes at any given time.
+
+Some information is available about the predominant serotypes during the 
+period of study\cite{macedo_virological_2013}.
+
+The estimation of the  number of susceptibles have been attempted 
+before, for other diseases\cite{bjornstad_dynamics_2002, 
+wallinga_reconstruction_2003}. These methods try to reconstruct the entire 
+series of infectious and susceptibles for measles 
+outbreaks from case data.\todo[inline]{anyone else?} In the case of Dengue the 
+series of susceptibles to all possible serotypes, cannot be reconstructed based 
+solely on a deterministic transmission model, since the arrival/re-emergenge of 
+new serotypes (which are a stochastic events) can change drastically the pool 
+of susceptibles throwing off any sequential estimation of the dynamics. 
+Adopting a multi-strain transmission model, does not help because of 
+insufficient information about the proportions of each serotype in case data.
+
+The authors have applied similar methods as used here, to estimate the number 
+of susceptibles for the predominant strain of influenza at the beginning of  
+influenza seasons in europe\cite{pone2011}.
 

### 2014-09-24 fit_dengue_my_big.py

@@ -27,27 +27,25 @@ mu = 0  # nat/mortality rate
 tau = 1  # recovery rate. FIXED
 
 N = 1  # Population of Rio
-s0 = 0.9999*N  # fraction of susceptibles at the beginning of first epidemic
-s1 = 0.999*N;
-s2 = 0.999*N;
 
-Ss = {0: s0, 1: s1, 2: s2}  # Multiple Ss map
+
+#~ Ss = {0: s0, 1: s1, 2: s2}  # Multiple Ss map
 
 
 # Initial conditions
-inits = np.array([s0, 100, 0.0])  # initial values for state variables.
+inits = np.array([1, 0.001, 0.0])  # initial values for state variables.
 
 def model(theta):
     # setting parameters
-    s0, s1, s2 = theta
-    Ss = {0: s0, 1: s1, 2: s2}  # Multiple Ss map
+    s0, s1, s2, s3, s4, s5, s6, s7, s8, s9 = theta
+    Ss = {0: s0, 1: s1, 2: s2, 3: s3, 4: s4, 5: s5, 6: s6, 7: s7, 8: s8, 9: s9}  # Multiple Ss map
 
     # ~ b1=19.9
     def sir(y, t):
         '''ODE model'''
         S, I, R = y
         
-        beta = 0 if t > 242 else  iRt(t) * tau/ N#(Ss[ycode[int(t)]])
+        beta = 0 if t > 728 else iRt(t) * tau#(Ss[ycode[int(t)]])
 
         #~ if (calc_R0s()<1.1).any():
         #~ print "flat"
@@ -68,7 +66,6 @@ def model(theta):
             #~ inits[1] = Y[t0-1,1]
             #~ print inits
             inits[1] = dt['I'][t0-1] if N-Ss[i] > dt['I'][t0-1] else N-Ss[i]
-            #~ print inits
         inits[0] = Ss[i];  # Define S0
         inits[-1] = N - sum(inits[:2])  # Define R(0)
         Y[t0:tf, :] = odeint(sir, inits, np.arange(t0, tf, 1))  #,tcrit=tcrit)
@@ -87,7 +84,7 @@ def prepdata(fname, sday=0, eday=None, mao=7):
     eday: final day
     mao: Moving average's Order
     """
-    data = pd.read_csv(fname, header=0, delimiter=',', skiprows=[1, 2 , 3], parse_dates=True)
[…]

### 2014-09-29 draft_lm1.tex
Improvements of Sep 29th 2014

@@
 any epidemic, which nevertheless acquire immunity, Thirdly, under-reporting is 
-a serious issue in Brazil. Duarte and Franca\cite{duarte_data_2006}, estimated 
+a serious issue in Brazil. Duarte and Fran\c{c}a\cite{duarte_data_2006}, estimated 
 the sensitivity of dengue reporting in Belo-Horizonte, Brazil to be of 63\%, 
@@
 \section*{Methods}
+
 \subsection*{The data}
@@
 in the city of Rio de Janeiro, Brazil.
-From this series, we calculated the effective reproduction number, $R_t$, according to the expression given by Nishiura et al.~\cite{nishiura}.
-The method assumes the disease counts $Y(t)$ are Poisson random variables such that $R_t$ is 
+From this series, we calculated the effective reproduction number, $R_t$, 
+according to the expression given by Nishiura et al.~\cite{nishiura}.
+The method assumes the disease counts $Y(t)$ are Poisson random variables such 
+that $R_t$ is 
 \[ R_t = \left(\frac{Y(t + 1)}{Y(t)}\right)^{1/n}\]
-where $n$ is the ratio between the length of reporting interval and the mean generation time of the disease.
-In this paper we obtain the estimates of $R_t$ using a moving window of three weeks, which roughly corresponds to the generation time of dengue, and thus $n = 1$.
-We provide more detail on the estimation and confidence/credibility intervals in Text S1.
+where $n$ is the ratio between the length of reporting interval and the mean 
+generation time of the disease.
+Here we obtain the estimates of $R_t$ using a moving window of three weeks, 
+which roughly corresponds to the generation time of dengue, and thus $n = 1$.
+
+We provide more detail on drawing posterior inference for $R_t$ in Text S1.
 
@@
 A Susceptible-Infectious-Removed (SIR) model is proposed to model dengue dynamics.
-In the traditional formulation of the model, transmission is governed by a transmission rate $\beta$ and recovery happens at a rate $\tau$.
+In the traditional formulation of the model, transmission is governed by a 
+transmission rate $\beta$ and recovery happens at a rate $\tau$.
+
 As the epidemic progresses, the effective transmission  rate changes and is
@@
  \label{eq:effbeta}
- \beta(t) = R_t	\cdot\tau
+ \beta(t) = \frac{R_t\cdot\tau}{S_0}
 \end{equation}
 
-We parameterize our model with the varying transmission using the system of ordinary differential equations
+We parameterize our model with the varying transmission using the system of 
+ordinary differential equations
 \begin{align}
@@
 where $S + I + R = 1 \: \forall\: t$. % \in [T_0, T_1]$. 
-This is a rather simplified model, in which, for instance, the vector is left out.
-Also, although there probably are multiple circulating serotypes, our approach does not discriminate between them.
+This is a rather simplified model, in which, for instance, the vector is left 
+out.
+Also, although there probably are multiple circulating serotypes, our approach
+does not discriminate between them.
 
 \subsection*{Bayesian parameter estimation}
-For the fitting procedure the time series were normalized to lie on the $[0,1]$ interval, therefore avoiding scaling issues.
-% Results and Discussion can be combined.xx
+
+For the fitting procedure the time series were divided into epidemic windows 
+that corresponded to significant raises in incidence and normalized to lie on 
+the $[0,1]$ interval.
+
+Also, as the numerical integration routine needs $\beta(t)$ to be continuous, we 
+used linear interpolation to obtain values of $R_t$ for any time point.
+
+We take a Bayesian approach to the estimation of $S_0$ for each window.
+For a given interval $j = \{ t_j^{\text{start}}, t_j^{\text{end}} \} $ we 
+observe an incidence time series $\mathbf{Y_{j}}$.
+We are thus interested in the posterior distribution
+\begin{equation}
+ \label{eq:S0post}
+ p(S_{0j}|\mathbf{Y_{j}}) \propto l(\mathbf{Y_{j}}|S_{0j}, R_t, m,  \tau 
+)\pi(S_{0j}) 
+\end{equation}
+The likelihood $l(\mathbf{Y_{j}}|\cdot)$ is assumed to be a Normal distribution 
+with fixed variance $\sigma^2$.
+In this estimation procedure we kept $R_t$ fixed at fixed at the posterior mean 
+obtained as described in TextS1 and fixed $\tau = 1/7\: \text{days}^{-1}$.
+% \subsubsection*{Priors}
+To complete the inference, we need to specify prior distributions for the 
+parameters of interest.
+We place a flat $\text{Beta}(1, 1)$ prior on $S_{0j}\:\forall j$
+% 
+% \begin{align}
+%  S_{0j} &\sim Beta(1, 1) \\
+%        m &\sim \\
+% \end{align}
+
+% \subsubsection*{Posterior approximation}
+
+To approximate the posterior in~(\ref{eq:S0post}) we use Markov chain Monte 
+Carlo techiniques implemented in the Bayesian inference with Python 
+(BIP)~\cite{pone2011} available at 
+\url{http://code.google.com/p/bayesian-inference/}.
+BIP uses a Differential Evolution Adaptive Metropolis (DREAM)~\cite{vrugt2008} 
+scheme that efficiently samples from high-dimensional joint distributions using 
+multiple adaptive chains running in parallel with delayed rejection.
+The chains were run until $5000$ samples were obtained after discarding $500$ 
+burn-in samples.
+Convergence of the parallel chains was verified at every 100 iterations by the 
+calculation of the Gelman-Rubins' R convergence diagnostic~\cite{brooks1998}.
+% Results and Discussion can be combined.
 \section*{Results}
@@
 \section*{Acknowledgments}
-LMC is grateful to Dr. Leonardo Bastos for useful discussions on the estimation of $R(t)$.
+LMC is grateful to Dr. Leonardo Bastos for useful discussions on the 
+posterior inference for $R(t)$.
 

### 2014-09-29 draft_lm1.tex
Flavio's corrections

@@
 that $R_t$ is 
-\[ R_t = \left(\frac{Y(t + 1)}{Y(t)}\right)^{1/n}\]
+\begin{equation}
+ \label{eq:Rt}
+ R_t = \left(\frac{Y(t + 1)}{Y(t)}\right)^{1/n}
+\end{equation}
+
 where $n$ is the ratio between the length of reporting interval and the mean 
@@
 \end{equation}
-
+where $R_t$ is the effective reproductive number as defined in~\ref{eq:Rt}.
 We parameterize our model with the varying transmission using the system of 
@@
 Also, although there probably are multiple circulating serotypes, our approach
-does not discriminate between them.
+does not discriminate between them due to the lack of serotype-specific data.
 
@@
 
-For the fitting procedure the time series were divided into epidemic windows 
-that corresponded to significant raises in incidence and normalized to lie on 
-the $[0,1]$ interval.
+For the fitting procedure the time series was divided into $J=10$ epidemic 
+windows that corresponded to significant raises in incidence and normalized to 
+lie on the $[0,1]$ interval.
 
-Also, as the numerical integration routine needs $\beta(t)$ to be continuous, we 
-used linear interpolation to obtain values of $R_t$ for any time point.
+Also, as the numerical integration routine needs to treat $\beta(t)$ as a
+continuous function of time because of the variable step size, we used linear 
+interpolation to obtain values of $R_t$ for any time point.
 
@@
 multiple adaptive chains running in parallel with delayed rejection.
+In this study we used one chain per parameter, i.e, 3 chains for each run.
 The chains were run until $5000$ samples were obtained after discarding $500$ 

### 2014-10-01 fit_dengue_my_big_plot.py
added figs; working on concatenated plots

@@ -1,0 +1,229 @@
+""" 
+Script to generate the plots for the
+multi-year simulations
+"""
+import cPickle as CP
+from itertools import cycle
+from scipy import stats
+import glob
+from numpy import *
+from matplotlib import font_manager
+from matplotlib.ticker import FormatStrFormatter
+import matplotlib.dates as mdates
+import matplotlib.pyplot as P
+import Gnuplot
+from scipy.stats import gaussian_kde
+import datetime
+import matplotlib.collections as collections
+import seaborn as sns
+import pandas as pd
+from sqlalchemy import create_engine
+
+from BIP.Bayes.PlotMeld import violin_plot, pred_new_cases, peakdet, plot_par_violin
+
+
+def _read_results(nam):
+    """
+    read results from disk
+    """
+    pt, series, predseries = [], [], []
+    nf = len(glob.glob('%s*.pickle' % nam))
+    for w in range(nf):
+        fn = "%s_%s.pickle" % (nam, w)
+        print fn
+        with open(fn, 'r') as f:
+            a, b, obs, pred, samples = CP.load(f)
+        # f.close()
+        pt.append(a)
+        series.append(b)
+        predseries.append(pred)
+
+    return pt, series, predseries, obs, nf
+
+def read_data(dbname):
+    """
+    Returns dataframes from tables of the sqlite database
+    :param dbname: sqlite3 database name
+    :return: list of pandas dataframes
+    """
+    eng = create_engine('sqlite:///{}.sqlite'.format(dbname))
+    df_data = pd.read_sql_table('data', eng, index_col=['pk', 'time'], parse_dates={'time': '%Y-%m-%d %H:%M:%S'})
+    df_pt = pd.read_sql_table('post_theta', eng, index_col=['pk', 'time'], parse_dates={'time': '%Y-%m-%d %H:%M:%S'})
+    df_series = pd.read_sql_table('series', eng, index_col=['pk', 'time'], parse_dates={'time': '%Y-%m-%d %H:%M:%S'})
+    return df_data, df_pt, df_series
+
+
+def create_tex_table(nms):
+    pts = {}
+    for nm in nms:
+        pt, series, predseries, obs, weeks = _read_results(nm)
+        pts[nm.split("-")[0]] = pt
[…]

### 2014-10-01 fit_dengue_split_big.py

@@ -30,59 +30,77 @@ tau = 1  # recovery rate. FIXED
 
 N = 1  # Population of Rio
 
-
+inicio=0
 #~ Ss = {0: s0, 1: s1, 2: s2}  # Multiple Ss map
 
 
 # Initial conditions
-inits = np.array([1, 0.001, 0.0])  # initial values for state variables.
+inits = np.array([0.999, 0.001, 0.0])  # initial values for state variables.
 
+
+#@jit#('f8[:](f8[:],f8,f8,f8)')
+def sir(y, t, *pars):
+    '''ODE model'''
+    S, I, R = y
+    s0, m, tau = pars
+    beta = 0 if t > 728 else iRt(t) * tau/s0
+    # print S, I, beta
+    
+    lamb = beta * (I+m) * S
+
+    return np.array([-lamb,  #dS/dt
+            lamb - tau * I,  #dI/dt
+            tau * I,  #dR/dt
+    ])
+
+#@jit#('f8[:,:](f8[:],f8,f8,f8)')
+def jac(y, t, *pars):
+    S, I, R = y
+    s0, m, tau = pars
+    beta = 0 if t > 728 else iRt(t) * tau/s0
+    return np.array([[-(I + m)*beta, -S*beta, 0],
+           [(I + m)*beta, S*beta - tau, 0],
+           [0, tau, 0]])
 
 def model(theta):
     # setting parameters
-    s0, s1, s2, s3, s4, s5, s6, s7, s8, s9 = theta
-    Ss = {0: s0, 1: s1, 2: s2, 3: s3, 4: s4, 5: s5, 6: s6, 7: s7, 8: s8, 9: s9}  # Multiple Ss map
+    s0, m, tau = theta
 
-    @jit('f8[:](f8[:],f8)')
-    def sir(y, t):
-        '''ODE model'''
-        S, I, R = y
-        
-        beta = 0 if t > 728 else iRt(t) * tau#(Ss[ycode[int(t)]])
+    t0 = 0
+    tf = fim
+    Y = np.zeros((tf-t0, 3))
 
-        #~ if (calc_R0s()<1.1).any():
-        #~ print "flat"
-        lamb = (beta * (I+m) * S)
-        
-        return np.array([-lamb,  #dS/dt
-                lamb - tau * I,  #dI/dt
-                tau * I,  #dR/dt
-        ])
-    @jit('f8[:,:](f8[:],f8)')
-    def jac(y,t):
[…]

### 2014-10-01 fit_dengue_split_big_plot.py

@@ -1,0 +1,249 @@
+""" 
+Script to generate the plots for the
+multi-year simulations
+"""
+import cPickle as CP
+from itertools import cycle
+from scipy import stats
+import glob
+from numpy import *
+from matplotlib import font_manager
+from matplotlib.ticker import FormatStrFormatter
+import matplotlib.dates as mdates
+import matplotlib.pyplot as P
+import Gnuplot
+from scipy.stats import gaussian_kde
+import datetime
+import matplotlib.collections as collections
+import seaborn as sns
+import pandas as pd
+from sqlalchemy import create_engine
+
+from BIP.Bayes.PlotMeld import violin_plot, pred_new_cases, peakdet, plot_par_violin
+
+
+def _read_results(nam):
+    """
+    read results from disk
+    """
+    pt, series, predseries = [], [], []
+    nf = len(glob.glob('%s*.pickle' % nam))
+    for w in range(nf):
+        fn = "%s_%s.pickle" % (nam, w)
+        print fn
+        with open(fn, 'r') as f:
+            a, b, obs, pred, samples = CP.load(f)
+        # f.close()
+        pt.append(a)
+        series.append(b)
+        predseries.append(pred)
+
+    return pt, series, predseries, obs, nf
+
+def read_data(dbname):
+    """
+    Returns dataframes from tables of the sqlite database
+    :param dbname: sqlite3 database name
+    :return: list of pandas dataframes
+    """
+    eng = create_engine('sqlite:///{}'.format(dbname))
+    df_data = pd.read_sql_table('data', eng, index_col=['pk', 'time'], parse_dates={'time': '%Y-%m-%d %H:%M:%S'})
+    df_pt = pd.read_sql_table('post_theta', eng, index_col=['pk', 'time'], parse_dates={'time': '%Y-%m-%d %H:%M:%S'})
+    df_series = pd.read_sql_table('series', eng, index_col=['pk', 'time'], parse_dates={'time': '%Y-%m-%d %H:%M:%S'})
+    return df_data, df_pt, df_series
+
+
+def create_tex_table(nms):
+    pts = {}
+    for nm in nms:
+        pt, series, predseries, obs, weeks = _read_results(nm)
+        pts[nm.split("-")[0]] = pt
[…]

### 2014-10-01 fit_dengue_split_big_plot.py
adicionei primeira versao da figura concatenada

@@ -217,6 +217,7 @@ def plot_concat_series(dbs):
         s_lwr = srs.S.groupby(level='time').aggregate(lwr)
         # P.legend(loc=0)
         P.fill_between(s_median.index, s_lwr, s_upr, facecolor=co, alpha=.2)
+    P.savefig('concat.svg')
 
 
 

### 2014-10-03 fit_dengue_split_big_plot.py
atualizei gráfico com posteriores das séries de S e I

@@ -47,9 +47,9 @@ def read_data(dbname):
     :return: list of pandas dataframes
     """
     eng = create_engine('sqlite:///{}'.format(dbname))
-    df_data = pd.read_sql_table('data', eng, index_col=['pk', 'time'], parse_dates={'time': '%Y-%m-%d %H:%M:%S'})
-    df_pt = pd.read_sql_table('post_theta', eng, index_col=['pk', 'time'], parse_dates={'time': '%Y-%m-%d %H:%M:%S'})
-    df_series = pd.read_sql_table('series', eng, index_col=['pk', 'time'], parse_dates={'time': '%Y-%m-%d %H:%M:%S'})
+    df_data = pd.read_sql_table('data', eng, index_col=['time'], parse_dates={'time': '%Y-%m-%d %H:%M:%S'})
+    df_pt = pd.read_sql_table('post_theta', eng, index_col=['time'], parse_dates={'time': '%Y-%m-%d %H:%M:%S'})
+    df_series = pd.read_sql_table('series', eng, index_col=['time'], parse_dates={'time': '%Y-%m-%d %H:%M:%S'})
     return df_data, df_pt, df_series
 
 
@@ -204,20 +204,42 @@ def series(nam='Dengue_S0'):
         #print P.setp(ax2)
 
 def plot_concat_series(dbs):
+    """
 
+    :param dbs:
+    """
+    series = []
+    obs = []
+    for db in dbs:
+        data, theta, srs = read_data(db)
+        series.append(srs)
+        obs.append(data)
     upr = lambda x: stats.scoreatpercentile(x, 97.5)
     lwr = lambda x: stats.scoreatpercentile(x, 2.5)
     c = cycle(['b', 'g', 'r', 'c', 'm', 'y', 'k'])
-    for db in dbs:
+    ax1 = P.subplot(211)
+    for srs in series:
         co = c.next()
-        data, theta, srs = read_data(db)
         s_median = srs.S.groupby(level='time').median()
         s_median.plot(style='k-', label='Median')
         s_upr = srs.S.groupby(level='time').aggregate(upr)
         s_lwr = srs.S.groupby(level='time').aggregate(lwr)
         # P.legend(loc=0)
         P.fill_between(s_median.index, s_lwr, s_upr, facecolor=co, alpha=.2)
-    P.savefig('concat.svg')
+
+    ax2 = P.subplot(212, sharex=ax1)
+    for srs, da in zip(series, obs):
[…]

### 2014-10-03 fit_dengue_split_big_plot.py
melhorei script para gerar figuras

@@ -18,27 +18,11 @@ import matplotlib.collections as collections
 import seaborn as sns
 import pandas as pd
 from sqlalchemy import create_engine
+from collections import OrderedDict
 
 from BIP.Bayes.PlotMeld import violin_plot, pred_new_cases, peakdet, plot_par_violin
 
 
-def _read_results(nam):
-    """
-    read results from disk
-    """
-    pt, series, predseries = [], [], []
-    nf = len(glob.glob('%s*.pickle' % nam))
-    for w in range(nf):
-        fn = "%s_%s.pickle" % (nam, w)
-        print fn
-        with open(fn, 'r') as f:
-            a, b, obs, pred, samples = CP.load(f)
-        # f.close()
-        pt.append(a)
-        series.append(b)
-        predseries.append(pred)
-
-    return pt, series, predseries, obs, nf
 
 def read_data(dbname):
     """
@@ -53,16 +37,26 @@ def read_data(dbname):
     return df_data, df_pt, df_series
 
 
-def create_tex_table(nms):
-    pts = {}
-    for nm in nms:
-        pt, series, predseries, obs, weeks = _read_results(nm)
-        pts[nm.split("-")[0]] = pt
+def create_tex_table(dbs):
+    """
+    Create Latex table with the ratios of susceptibles per year
+    :return:
+    """
+    pts = OrderedDict()
+    series = OrderedDict()
+    obs = OrderedDict()
+    for db in dbs:
+        y = db.split('_')[0][-4:]
+        data, theta, srs = read_data(db)
+        series[y] = srs
+        obs[y] = data
+        pts[y] = theta
+
     head = r"""\begin{center}
-            \begin{tabular}{c|c c c}
+            \begin{tabular}{c|c}
             \hline
             """
-    head += r"""Name & Belgium & Netherlands & Portugal\\
+    head += r"""Year & Ratio \\
             \hline
             """
     bot = r"""
@@ -72,10 +66,13 @@ def create_tex_table(nms):
         """
     body = r""
     st = []
-    for n in pts['be'][0].dtype.names:
-        body += n + r" & %1.3G, " % mean(pts['be'][0][n]) + "%1.3G " % var(pts['be'][0][n]) + \
-                r" & %1.3G, " % mean(pts['nl'][0][n]) + "%1.3G " % var(pts['nl'][0][n]) + \
[…]

### 2014-10-03 fit_dengue_split_big_plot.py
fixed table generating code

@@ -71,8 +71,9 @@ def create_tex_table(dbs):
             continue
         last_week_of_last_epi = series.values()[i-1].index[-1]
         first_week = series.values()[i].index[0]
-        ratio = series[Y].S.ix[first_week] / series.values()[i-1].S.ix[last_week_of_last_epi]
-        body += Y + r" & {}\pm{}".format(mean(ratio), std(ratio))
+        ratio = array(series[Y].S.ix[first_week]) / array(series.values()[i-1].S.ix[last_week_of_last_epi])
+        body += Y + r" & {}({}-{})\\".format(nanmedian(ratio), stats.scoreatpercentile(ratio, 2.5), stats.scoreatpercentile(ratio, 97.5))
+        body += "\n"
 
     return head + body + bot
 

### 2014-10-03 fit_dengue_split_big_plot.py
added more figures

@@ -72,7 +72,7 @@ def create_tex_table(dbs):
         last_week_of_last_epi = series.values()[i-1].index[-1]
         first_week = series.values()[i].index[0]
         ratio = array(series[Y].S.ix[first_week]) / array(series.values()[i-1].S.ix[last_week_of_last_epi])
-        body += Y + r" & {}({}-{})\\".format(nanmedian(ratio), stats.scoreatpercentile(ratio, 2.5), stats.scoreatpercentile(ratio, 97.5))
+        body += Y + r" & {:.2%}({:.2%}-{:.2%})\\".format(nanmedian(ratio), stats.scoreatpercentile(ratio, 2.5), stats.scoreatpercentile(ratio, 97.5))
         body += "\n"
 
     return head + body + bot

### 2014-10-06 draft_lm1.tex
added more figures
fixed bad link for PNG fig
worked on the introduction

@@
 observed data, are frequently overlooked.
-If the model is about a disease 
-invading a new population one can simply assume the vast majority of the 
-population to be susceptible.
-However for a disease which is well in its way to 
-endemicity but not yet stable, it can be very hard to determine the 
-immunological structure of the population without resorting to expensive large 
-scale prevalence tests.
+
+Dengue is a disease which is caused by 4 different types of viruses 
+which co-circulate between the vector and human populations. immunity is 
+considered permanent after the exposure of humans to each of the viruses but 
+cross immunity is considered very limited[ref].
+Thus, the proportions of viral types co-circulating at any point in time is 
+strongly dependent on previous incidence of the disease, which determines the 
+number of susceptibles for each viral type at any point in time.
+
+Dengue transmission is conditioned on temperature, due to its effects on the 
+vector reproduction.
+In places with sufficient temperature variations dengue is a summer disease 
+mainly.
+
+Determining the viral types in circulation during an epidemic is 
+expensive, as it requires virus isolatio and reverse-transcriptase-PCR. 
+Therefore only a very limited number of samples are examined on any epidemic. 
+Besides, mild cases are rarely detected let alone tested for dengue.
+
+Dengue in Brazil is a disease which is well in its way to 
+endemicity but not yet stable, therefore co-circulation patterns have not yet 
+reached a steady state as has been demonstrated for certain countries in 
+south-east asia which have had the disease much longer.
 
@@
 any time other than when the date the disease was reintroduced in the country 
-in the late 1980s.
-For SIR-like models, the relevant initial values are the 
-number of infectious and susceptible  individuals. 
-The number of infectious 
-individuals are less challenging to determine, because the get very small 
-during the winters and can be derived from incidence data  with is monitored 
-continuously.
+in the late 1980s and even so, without virological survey data from time to 
+time, to inform on the mixing of viruses in circulation, it is virtually 
+impossible to predict, by simulation alone, the imunological state of the 
+population at any point in the future.
+
+For SIR-like models, the relevant initial values are the number of infectious 
+and susceptible  individuals. 
+The number of infectious individuals are less challenging to determine, because 
+the get very small during the winters and can be derived from incidence data  
+with is monitored continuously.
 The real challenge is to determine the number of susceptibles to 
@@
 susceptibles in the population, first and foremost is the fact that dengue
-
-confers permanente immunity within serotypes but with only short-lived 
-cross-immunity. Secondly, there are a larger number of asymptomatic cases in 
-any epidemic, which nevertheless acquire immunity, Thirdly, under-reporting is 
-a serious issue in Brazil. Duarte and Fran\c{c}a\cite{duarte_data_2006}, estimated 
-the sensitivity of dengue reporting in Belo-Horizonte, Brazil to be of 63\%, 
-meaning that aproximately 37\% of the cases go unreported. Lastly, demography 
-and migrations affect the number of susceptible in ways which are not easy to
-determine.
+confers permanent immunity within serotypes but with only short-lived 
+cross-immunity. 
+Secondly, there are a larger number of asymptomatic cases in 
+any epidemic, which nevertheless acquire immunity. It is estimated that for 
+every case reported, 10-20 are not seen by health authorities (personnal 
+communication).
+Thirdly, under-reporting is a serious issue in Brazil. 
+Duarte and Fran\c{c}a\cite{duarte_data_2006}, estimated 
+the sensitivity of dengue reporting for hospitalized patients in 
+Belo-Horizonte, Brazil to be of 63\%, 
+meaning that aproximately 37\% of the suspected dengue cases go unreported.  
+Lastly, demography and migrations affect the number of susceptible in ways which 
+are not easy to determine.
 
 The initial number of susceptibles -- $S_0$ -- is a a very powerful parameter 
-with regard to epidemic risk.
-Its value affects directly the local $R_0$, 
-thus being intrumental in predicting epidemics, but it also affects the speed 
-and size of the epidemic.
+with regards to epidemic risk.
+Its value affects directly the local $R_0$, thus being intrumental in predicting 
+epidemics, but it also affects the speed and size of the epidemic.
 
 In this paper, we estimate the number of susceptibles, allowing for 
-discontinuities right before every epidemic in the last 18 years. THis way we 
-can account for replacement of the predominant serotype, which can redefine 
-susceptibility for the following season.
+discontinuities right before every epidemic in the last 18 years. 
+Through this method we hope to detect drastic changes in the mixture of 
+susceptibles cause by the replacement of the predominant serotype, which 
+can redefine susceptibility for the following season.
 
@@
 %\end{figure}
+\begin{figure}
+\begin{center}
+\includegraphics[width=.9\textwidth]{./plots/concat_SI.png}
+\end{center}
+\caption{
+{\bf Bold the first sentence.}  Rest of figure caption.  
+}
+\label{Fig:S0}
+\end{figure}
 
@@
 
+\begin{table}
+\caption{\bf{Table title}}
+ \begin{center}
+            \begin{tabular}{c|c}
+            \hline
+            Year & Ratio \\
+            \hline
+            1997 & 204.26(67.45-706.68)\\
+1998 & 207.45 (170.45-261.63)\\
+2000 & 2300.63 (1741.30-3156.85)\\
+2001 & 317.62 (226.47-526.45)\\
+2005 & 100.28 (48.95-171.58)\\
+2006 & 198.26 (114.59-317.88)\\
+2007 & 567.78 (388.55-834.58)\\
+2008 & 7.82 (2.97-34.43)\\
+1999 & 26.70 (5.97-87.36)\\
+
+        \hline
+        \end{tabular}
+        \end{center}
+\end{table}
+
 \section*{Supporting Information Legends}

### 2014-10-08 draft_lm1.tex
minor changes

@@
 %\subsection*{Subsection 1}
+\subsection{Serotype proportion}
+The susceptible population estimated in this paper is defined as individuals 
+susceptible to at least one of the circulating serotypes in any given year. 
+This is due to the fact that we use case data to estimate $S_0$ prior to the 
+epidemic. In general, susceptibility to dengue may be represented by 
+4 binomially distributed random variables. Each representing the 
+fraction of susceptibles to each virus type -- $S_i \sim Bin(p_i), \, i \in  
+\{1, 2, 3, 4\}$. In the absence of 
+cross-immunity, the suceptibility can be manifested, in individuals, as one of 
+the following forms:
+\begin{equation}
+ S = \{S_{1}, S_{2}, S_{3}, 
+S_{4},S_{12},S_{123},S_{1234},S_{13},S_{14},S_{23},S_{234},S_{24},S_{34}\}
+\end{equation}
+
+Even though we don't observe the incidences for each serotype independently, we 
+have information about the proportion of Serotypes, based on the lab confirmed 
+cases. We also know that the drop in susceptibles estimated by the model, 
+corresponds to the drop in the number of the one or more serotypes. For 
+example: in an epidemic of just DENV1, the debt of susceptibles would have to 
+be distributed between the following types: 
+$\{S_{1},S_{12},S_{123},S_{1234},S_{13},S_{14}\}$.
+
 \section*{Discussion}

### 2014-10-08 draft_lm1.tex
minor changes
added Poisson-binomial idea

--- 2014-10-06 draft_lm1.tex
@@
 %\subsection*{Subsection 1}
+\subsection*{Serotype proportion}
+The susceptible population estimated in this paper is defined as individuals 
+susceptible to at least one of the circulating serotypes in any given year. 
+Let this Total number of susceptibles be denoted by $S_T$.
+
+Let the number of susceptibles to a given dengue virus ($DENV_i$), $S_i$ be 
+represented by a Binomial random variables $S_i \sim Bin(n, p_i)$ where $n$ is 
+population size and p is the true proportion of susceptibles to $DENV_i$ in 
+the population, with $i \in  \{1, 2, 3, 4\}$. In the absence of 
+cross-immunity, i.e., the exposure to serotypes are independent, the 
+total number of susceptibles $S_T=\sum_i S_i$  should follow a Poisson-Binomial 
+distribution, $PoisBin(p_i,n)$, where
+\begin{equation}
+ P(S_T=k) = \sum \prod_i p_i \prod_i  (1-p_i)
+\end{equation}
+
+To elicit priors for the $p_i \sim Be(a,b)$, we used the proportion of 
+serotyped cases at the end of each epidemic as the expected value for the $p_i$.
+
+Let $S_{0T}$ and $S_{fT}$, be the total number of susceptibles before and 
+after an epidemic, respectively
+
+% the suceptibility can be manifested, in individuals, as one of 
+% the following forms:
+% \begin{equation}
+%  S = \{S_{1}, S_{2}, S_{3}, 
+% S_{4},S_{12},S_{123},S_{1234},S_{13},S_{14},S_{23},S_{234},S_{24},S_{34}\}
+% \end{equation}
+
 \section*{Discussion}

### 2014-10-08 draft_lm1.tex
minor improvements

@@
 
-To elicit priors for the $p_i \sim Be(a,b)$, we used the proportion of 
-serotyped cases at the end of each epidemic as the expected value for the $p_i$.
+Initially, we have no information about the distribution of the $p_i$ but they 
+can be learned gradually with each epidemic.
 
 Let $S_{0T}$ and $S_{fT}$, be the total number of susceptibles before and 
-after an epidemic, respectively
+after an epidemic, respectively. The total number of cases observed in an 
+epidemic, $I_T = \sum_i I_i$ can be expressed as $S_{0T}-S_{fT}$, thus
+\begin{align}
+ S_{fT} &= S_{0T} - I_T\\
+ &=\sum S_i - \sum I_i\\
+ &=\sum_i S_i-I_i
+\end{align}
+
+If $S_i/n \approx p_i$ and $I_i$ is observed, we can adjust the initial 
+$p_i$ proportionately to $I_i$.
+\todo[inline]{sera que da para fazer inferencia bayesiana aqui?}
 

### 2014-10-08 draft_lm1.tex
minor changes

@@
 \end{equation}
+\todo[inline]{this sum is over all types of susceptibles, I think: 
+$\{S_{1}, S_{2}, 
+S_{3},S_{4},S_{12},S_{123},S_{1234},S_{13},S_{14},S_{23},S_{234},S_{24},S_{34}\}
+$}
 
@@
  S_{fT} &= S_{0T} - I_T\\
- &=\sum S_i - \sum I_i\\
- &=\sum_i S_i-I_i
+ &=\sum S_{0i} - \sum I_i\\
+ &=\sum_i S_{0i}-I_i
 \end{align}
@@
 If $S_i/n \approx p_i$ and $I_i$ is observed, we can adjust the initial 
-$p_i$ proportionately to $I_i$.
+$p_i$ according to $I_i$.
 \todo[inline]{sera que da para fazer inferencia bayesiana aqui?}

### 2014-10-13 draft_lm1.tex
moving Rt estimation to main text

@@
 \makeatother
+
+% Our defs 
+\def \rr {$R_{t}\ $}
 
@@
 
-\subsection*{The data}
+\subsection*{Estimating the reproductive number}
+
+In monitoring of infectious diseases, it is important to assess whether the 
+incidence of a a particular disease is increasing significantly, in order to 
+decide to take preventive measures.
+The effective reproductive number at time $t$, \rr, can be understood as a 
+real-time estimate of the basic reproductive number ($R_{0}$) and is defined as 
+the average number of secondary cases per primary case at time $t$.
+Here we derive confidence/credibility bounds for the uncertainty about the 
+estimate of the effective reproductive number $R_t$ under the assumption that 
+reported disease counts are Poisson distributed.
+It is shown that one can construct a conservative confidence interval for \rr 
+using the induced probability distribution that results from the approach of 
+\cite{mantel}.
+
+Let $Y_t$ be the number of reported disease cases for a particular time $t \in 
+(0, T)$.
+\cite{nishiura} extend the theory developed by \cite{stallybrass} and propose 
+to estimate \rr as
+\begin{equation}
+\label{eq:Rtestimate}
+R_t = \left( \frac{Y_{t+1}}{Y_t}\right)^{1/n}
+\end{equation}
+where $n$ is taken to be the ratio between the length of reporting interval and 
+the mean generation time of the disease.
+Here we are interested in the simpler case $n=1$.
+If \rr is to be used as a decision tool, one needs to be able to quantify the 
+uncertainty about estimate in equation~\ref{eq:Rtestimate}. 
+Here we detail how to obtain confidence intervals for \rr, exploring two 
+approaches under the assumption that the counts $Y_t$ are Poisson distributed 
+for all $t$.
+
+First we explore the approach of \cite{mantel}, whose objective is to obtain 
+confidence intervals for the ratio of two Poisson counts. 
+Let $Y_{t} \sim Poisson(\lambda_t)$ and $Y_{t+1} \sim Poisson(\lambda_{t+1})$ 
+and define $S = Y_{t} + Y_{t+1}$.
+The authors then proceed to note that by conditioning on the sum $S$
+\begin{align}
+\label{eq:binlike}
+Y_{t+1} | S &\sim Binomial(S, \theta_t) \\
+\theta_t &= \frac{\lambda_{t+1}}{\lambda_{t} + \lambda_{t+1}}
+\end{align}
+Let $c_{\alpha}(\theta_t) = \{\theta_t^{(L)} , \theta_t^{(U)} \}$ be such that 
+$Pr(\theta_t^{(L)}<\theta_t <\theta_t^{(U)}) = \alpha$.
+Analogously, define $c_{\alpha}(R_t) = \{R_t^{(L)} , R_t^{(U)} \}$ such that 
+$Pr(R_t^{(L)}<R_t<R_t^{(U)}) = \alpha$.
+\cite{mantel} show that one can construct a $100\alpha \%$ confidence interval 
+for \rr by noting that
+\begin{align}
+ R_t^{(L)} &= \frac{\theta_t^{(L)}}{(1-\theta_t^{(L)})}\\
+ R_t^{(U)} &= \frac{\theta_t^{(U)}}{(1-\theta_t^{(U)})}
+\end{align}
+
+Many authors have derived $c_{\alpha}(\theta_t)$ following orthodox approaches 
+such as those of \cite{wilson} and \cite{clopper}, mainly for simplicity 
+reasons.
+We choose instead to take a Bayesian approach and use the  $100\alpha \%$ 
+posterior credibility interval for $\theta_t$ as $c_{\alpha}(\theta_t)$.
+If we choose the conjugate beta prior with parameters $a_0$ and $b_0$ for the 
+binomial likelihood in (\ref{eq:binlike}), the posterior distribution for 
+$\theta_t$ is
+\begin{equation}
+\label{eq:thetapost}
+p(\theta_t| Y_{t+1}, S) \sim Beta(Y_{t+1} + a_0, Y_t + b_0)
+\end{equation}
+Equation~(\ref{eq:thetapost}) tells us that the induced distribution of $R_t$ is 
+a beta prime with parameters $ a_1 = Y_{t+1} + a_0$ and $b_1 =  Y_t + b_0$.
+The density of this induced distribution  $f_P(R_t)$ is then 
+\begin{equation}
+\label{eq:densityMantel}
+f_P(R_t| a_1, b_1) = \frac{\Gamma(a_1 + b_1)}{\Gamma(a_1)\Gamma(b_1)} R_t^{a_1 - 
+1} (1 + R_t)^{-(a_1 + b_1)}
+\end{equation}
+Thus, the expectation of \rr under $f_P$ is $a_1/(b_1 - 1)$ and its variance is 
+$a_1(a_1 + b_1 - 1)/\left((b_1 - 2)(b_1 - 1)^2 \right) $.
+
+Since when $R_t > 1$ indicates sustained transmission, one may be interested in 
+computing the probability of this event.
+By noting that
+\begin{align}
+\label{cumprobMantel}
+Pr(R_t > 1) &= 1 - \int_0^1 f_P(r)dr \\
+            &= 1- Pr(\theta_t < \frac{1}{2})
+\end{align}
+one can avoid dealing with the density in~(\ref{eq:densityMantel}) directly.
 
@@
 in the city of Rio de Janeiro, Brazil.
-From this series, we calculated the effective reproduction number, $R_t$, 
-according to the expression given by Nishiura et al.~\cite{nishiura}.
-The method assumes the disease counts $Y(t)$ are Poisson random variables such 
-that $R_t$ is 
-\begin{equation}
- \label{eq:Rt}
- R_t = \left(\frac{Y(t + 1)}{Y(t)}\right)^{1/n}
-\end{equation}
-
-where $n$ is the ratio between the length of reporting interval and the mean 
-generation time of the disease.
-Here we obtain the estimates of $R_t$ using a moving window of three weeks, 
-which roughly corresponds to the generation time of dengue, and thus $n = 1$.
-
-We provide more detail on drawing posterior inference for $R_t$ in Text S1.
 
@@
 \end{equation}
-where $R_t$ is the effective reproductive number as defined in~\ref{eq:Rt}.
+where $R_t$ is the effective reproductive number as defined 
+in~\ref{eq:Rtestimate}.
 We parameterize our model with the varying transmission using the system of 
@@
 lie on the $[0,1]$ interval.
-
 Also, as the numerical integration routine needs to treat $\beta(t)$ as a
@@
 calculation of the Gelman-Rubins' R convergence diagnostic~\cite{brooks1998}.
+
 % Results and Discussion can be combined.

### 2014-10-13 S1_lm1.tex
moving Rt estimation to main text

@@
 \documentclass[12pt]{article}              
-\usepackage[margin=1.5cm]{geometry}
+\usepackage[margin=2.0cm]{geometry}
 \usepackage{setspace,relsize}               
@@
 \title{
-Supporting Information to ¨Estimating the Initial Susceptible Fraction in Dengue Dynamics"
+Supporting Information to Estimating the Initial Susceptible Fraction in Dengue Dynamics"
 }           
-\author{Flavio Code\c{c}o Coelho \\
+\author{
+Flavio Code\c{c}o Coelho \\
 \\
@@
 
-\date{\today}
+\date{}
 % Notation defs
@@
 
-In monitoring of infectious diseases, it is important to assess whether the incidence of a a particular disease is increasing significantly, in order to decide to take preventive measures.
-The effective reproductive number at time $t$, \rr, can be understood as a real-time estimate of the basic reproductive number ($R_{0}$) and is defined as the average number of secondary cases per primary case at time $t$.
-Here we derive confidence/credibility bounds for the uncertainty about the estimate of the effective reproductive number $R_t$ under the assumption that reported disease counts are Poisson distributed.
-It is shown that one can construct a conservative confidence interval for \rr using the induced probability distribution that results from the approach of \citet{mantel}.
-
-Let $Y_t$ be the number of reported disease cases for a particular time $t \in (0, T)$.
-\citet{nishiura} extend the theory developed by \citet{stallybrass} and propose to estimate \rr as
-\begin{equation}
-\label{eq:Rtestimate}
-R_t = \left( \frac{Y_{t+1}}{Y_t}\right)^{1/n}
-\end{equation}
-where $n$ is taken to be the ratio between the length of reporting interval and the mean generation time of the disease.
-Here we are interested in the simpler case $n=1$.
-If \rr is to be used as a decision tool, one needs to be able to quantify the uncertainty about estimate in equation~\ref{eq:Rtestimate}. 
-Here we detail how to obtain confidence intervals for \rr, exploring two approaches under the assumption that the counts $Y_t$ are Poisson distributed for all $t$.
-
-\section*{The approach Ederer \& Mantel (1974)}
-
-\label{sec:mantel}
-First we explore the approach of \citet{mantel}, whose objective is to obtain confidence intervals for the ratio of two Poisson counts. 
-Let $Y_{t} \sim Poisson(\lambda_t)$ and $Y_{t+1} \sim Poisson(\lambda_{t+1})$ and define $S = Y_{t} + Y_{t+1}$.
-The authors then proceed to note that by conditioning on the sum $S$
-\begin{align}
-\label{eq:binlike}
-Y_{t+1} | S &\sim Binomial(S, \theta_t) \\
-\theta_t &= \frac{\lambda_{t+1}}{\lambda_{t} + \lambda_{t+1}}
-\end{align}
-Let $c_{\alpha}(\theta_t) = \{\theta_t^{(L)} , \theta_t^{(U)} \}$ be such that $Pr(\theta_t^{(L)}<\theta_t <\theta_t^{(U)}) = \alpha$.
-Analogously, define $c_{\alpha}(R_t) = \{R_t^{(L)} , R_t^{(U)} \}$ such that $Pr(R_t^{(L)}<R_t<R_t^{(U)}) = \alpha$.
-\citet{mantel} show that one can construct a $100\alpha \%$ confidence interval for \rr by noting that
-\begin{align}
- R_t^{(L)} &= \frac{\theta_t^{(L)}}{(1-\theta_t^{(L)})}\\
- R_t^{(U)} &= \frac{\theta_t^{(U)}}{(1-\theta_t^{(U)})}
-\end{align}
-
-Many authors have derived $c_{\alpha}(\theta_t)$ following orthodox approaches such as those of \citet{wilson} and \citet{clopper}, mainly for simplicity reasons.
-We choose instead to take a Bayesian approach and use the  $100\alpha \%$ posterior credibility interval for $\theta_t$ as $c_{\alpha}(\theta_t)$.
-If we choose the conjugate beta prior with parameters $a_0$ and $b_0$ for the binomial likelihood in (\ref{eq:binlike}), the posterior distribution for $\theta_t$ is
-\begin{equation}
-\label{eq:thetapost}
-p(\theta_t| Y_{t+1}, S) \sim Beta(Y_{t+1} + a_0, Y_t + b_0)
-\end{equation}
-Equation~(\ref{eq:thetapost}) tells us that the induced distribution of $R_t$ is a beta prime with parameters $ a_1 = Y_{t+1} + a_0$ and $b_1 =  Y_t + b_0$.
-The density of this induced distribution  $f_P(R_t)$ is then 
-\begin{equation}
-\label{eq:densityMantel}
-f_P(R_t| a_1, b_1) = \frac{\Gamma(a_1 + b_1)}{\Gamma(a_1)\Gamma(b_1)} R_t^{a_1 - 1} (1 + R_t)^{-(a_1 + b_1)}
-\end{equation}
-Thus, the expectation of \rr under $f_P$ is $a_1/(b_1 - 1)$ and its variance is $a_1(a_1 + b_1 - 1)/\left((b_1 - 2)(b_1 - 1)^2 \right) $.
-
-Since when $R_t > 1$ indicates sustained transmission, one may be interested in computing the probability of this event.
-By noting that
-\begin{align}
-\label{cumprobMantel}
-Pr(R_t > 1) &= 1 - \int_0^1 f_P(r)dr \\
-            &= 1- Pr(\theta_t < \frac{1}{2})
-\end{align}
-one can avoid dealing with the density in~(\ref{eq:densityMantel}) directly.
-
-\subsection*{A remark on prior distributions and tail behaviour}
+\subsection*{A remark on prior distributions and tail behavior of the 
+distribution of $R_t$}
 \label{sec:tails}
@@
 In order to decide which approach to take, it may be of use analysing the variance and tail behaviour of the derived distributions for \rr. 
-Consider the case of using a flat $Uniform(0, 1)$ prior for $\theta_t$ in Section~\ref{sec:mantel}.
+Consider the case of using a flat $Uniform(0, 1)$ prior for $\theta_t$.
 With $a_0 = b_0 = 1$, $a_1 = a_2$ and $b_1 = b_2 + 1$.

### 2014-10-13 draft_lm1.tex
No more serotype differences; no more table of S0 ratios

@@
 %\subsection*{Subsection 1}
-\subsection*{Serotype proportion}
-The susceptible population estimated in this paper is defined as individuals 
-susceptible to at least one of the circulating serotypes in any given year. 
-Let this Total number of susceptibles be denoted by $S_T$.
-
-Let the number of susceptibles to a given dengue virus ($DENV_i$), $S_i$ be 
-represented by a Binomial random variables $S_i \sim Bin(n, p_i)$ where $n$ is 
-population size and p is the true proportion of susceptibles to $DENV_i$ in 
-the population, with $i \in  \{1, 2, 3, 4\}$. In the absence of 
-cross-immunity, i.e., the exposure to serotypes are independent, the 
-total number of susceptibles $S_T=\sum_i S_i$  should follow a Poisson-Binomial 
-distribution, $PoisBin(p_i,n)$, where
-\begin{equation}
- P(S_T=k) = \sum \prod_i p_i \prod_i  (1-p_i)
-\end{equation}
-\todo[inline]{this sum is over all types of susceptibles, I think: 
-$\{S_{1}, S_{2}, 
-S_{3},S_{4},S_{12},S_{123},S_{1234},S_{13},S_{14},S_{23},S_{234},S_{24},S_{34}\}
-$}
-
-Initially, we have no information about the distribution of the $p_i$ but they 
-can be learned gradually with each epidemic.
-
-Let $S_{0T}$ and $S_{fT}$, be the total number of susceptibles before and 
-after an epidemic, respectively. The total number of cases observed in an 
-epidemic, $I_T = \sum_i I_i$ can be expressed as $S_{0T}-S_{fT}$, thus
-\begin{align}
- S_{fT} &= S_{0T} - I_T\\
- &=\sum S_{0i} - \sum I_i\\
- &=\sum_i S_{0i}-I_i
-\end{align}
-
-If $S_i/n \approx p_i$ and $I_i$ is observed, we can adjust the initial 
-$p_i$ according to $I_i$.
-\todo[inline]{sera que da para fazer inferencia bayesiana aqui?}
-
-% the suceptibility can be manifested, in individuals, as one of 
-% the following forms:
-% \begin{equation}
-%  S = \{S_{1}, S_{2}, S_{3}, 
-% S_{4},S_{12},S_{123},S_{1234},S_{13},S_{14},S_{23},S_{234},S_{24},S_{34}\}
-% \end{equation}
-
 \section*{Discussion}
@@
 
-\begin{table}
-\caption{\bf{Table title}}
- \begin{center}
-            \begin{tabular}{c|c}
-            \hline
-            Year & Ratio \\
-            \hline
-            1997 & 204.26(67.45-706.68)\\
-1998 & 207.45 (170.45-261.63)\\
-2000 & 2300.63 (1741.30-3156.85)\\
-2001 & 317.62 (226.47-526.45)\\
-2005 & 100.28 (48.95-171.58)\\
-2006 & 198.26 (114.59-317.88)\\
-2007 & 567.78 (388.55-834.58)\\
-2008 & 7.82 (2.97-34.43)\\
-1999 & 26.70 (5.97-87.36)\\
-
-        \hline
-        \end{tabular}
-        \end{center}
-\end{table}
-
 \section*{Supporting Information Legends}

### 2014-10-13 Modelo Estocástico/dengue_full_SDE.py
adicionei script para modelo estocastico

@@ -1,0 +1,73 @@
+CC By-SA 3.0
+
+Licença
+
+O TRABALHO (CONFORME ABAIXO DEFINIDO) É DISPONIBILIZADO DE ACORDO COM OS TERMOS DESTA LICENÇA PÚBLICA CREATIVE COMMONS ("LPCC" OU "LICENÇA"). O TRABALHO ESTÁ PROTEGIDO PELO CÓDIGO DO DIREITO DE AUTOR E DOS DIREITOS CONEXOS E/OU POR OUTRAS LEIS DE PROPRIEDADE INTELECTUAL E/OU POR OUTRAS LEIS (“LEGISLAÇÃO APLICÁVEL”). QUALQUER USO DO TRABALHO QUE NÃO O AUTORIZADO POR ESTA LICENÇA OU PELA LEGISLAÇÃO APLICÁVEL É PROIBIDO.
+
+AO EXERCER QUALQUER UM DOS DIREITOS DE EXPLORAÇÃO DO TRABALHO PREVISTOS NA PRESENTE LICENÇA, O UTILIZADOR ESTARÁ A CONCORDAR COM OS TERMOS DESTA LICENÇA E A ACEITAR VINCULAR-SE AOS MESMOS. NA MEDIDA EM QUE A PRESENTE LICENÇA SEJA CONSIDERADA UM CONTRATO, O LICENCIANTE CONCEDE AO UTILIZADOR OS DIREITOS AQUI PREVISTOS, EM CONTRAPARTIDA PELA SUA ACEITAÇÃO DOS TERMOS E CONDIÇÕES AQUI CONTIDOS.
+
+1. Definições
+
+    "Apresentar ao Público" significa, para os propósitos desta Licença, tanto (i) comunicar ao público o Trabalho ou Trabalho Derivado, o que for apropriado, por meio de qualquer ato que torne o seu gozo acessível ao público independentemente da posse de uma cópia do mesmo, nomeadamente através da representação, recitação, declamação, execução, exibição, exposição, transmissão ou retransmissão por fio ou sem fio, do Trabalho ou Trabalho Derivado, como (ii) colocar o Trabalho ou Trabalho Derivado, o que for apropriado, à disposição do público, por fio ou sem fio, por forma a torná-lo acessível a qualquer pessoa a partir do local e no momento por ela escolhido.
[…]

### 2014-10-15 Modelo Estocastico/dengue_full_SDE.py
finished implementation of the stochastic model.

@@ -1,0 +1,95 @@
+# coding:utf8
+
+from BIP.SDE.gillespie import Model
+import time
+from numpy import array, genfromtxt
+import pylab as P
+
+vnames = ['S', 'I1', 'I2', 'I3', 'I4', 'R1', 'R2', 'R3', 'R4', 'I11',
+          'I12', 'I13', 'I14', 'I21', 'I22', 'I23', 'I24', 'I31', 'I32',
+          'I33', 'I34', 'I41', 'I42', 'I43', 'I44', 'R']
+
+# transitions
+# S->I1	S->I2	S->I3	S->I4	I1=>R1	I2=>R2	I3=>R3	I4=>R4	R1=>I11	R1=>I12	R1=>I13	R1=>I14	R2=>I21	R2=>I22	R2=>I23	R2=>I24	R3=>I31	R3=>I32	R3=>I33	R3=>I34	R4=>I41	R4=>I42	R4=>I43	R4=>I44	I11=>R	I12=R	I13=>R	I14=R	I21=>R	I22=R	I23=>R	I24=R	I31=>R	I32=R	I33=>R	I34=R	I41=>R	I42=R	I43=>R	I44=R
+
+tm = genfromtxt('tmat.csv', delimiter=',', skip_header=2, usecols=range(1, 67), dtype=int)
+
+# Initial Conditions
+# Initial conditions for each of the four Serotype model
+
+N = 1000
+
+ini = [N - 4, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0]
+
+pars = (1 / 70.,  # mu
+        400,  # beta
+        1,  # phi
+        1 / 3.65,  # sigma
+        1)  # gamma
+
+# Propensity functions
+def gen_prop_functs():
+    nat_code = "def fnat(r, ini): return r[0]*N\n"
+    mu_code = "def fmu{}(r, ini): return r[0]*ini[{}]\n"
+    inf_code = """def fS(r, ini): return ini[0]*sum([lamb(i) for i in range(4)])\n"""
+    inf2_code = """def f{}(r, ini): return ini[0]* lamb({})\n"""
+    rec_code = """def f{}(r, ini): return r[-2]*ini[{}]\n"""
+    inf3_code = """def f{}(r, ini): return r[-1]*ini[{}]*lamb({})\n"""
+
+    nt = tm.shape[1]
+    with open('propfun.py', 'w') as f:
[…]

### 2014-10-15 Modelo Estocastico/propfun.py

@@ -1,0 +1,87 @@
+pars = (0.014285714285714285, 400, 1, 0.273972602739726, 1)
+ini = [996, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0]
+N = 1000
+
+lamb = lambda i: 400*(ini[i+1] + sum(ini[9+i : 9+12+i+1 : 4])-ini[9+4*i+1])
+def fS(r, ini): return ini[0]*sum([lamb(i) for i in range(4)])
+
+#Infection rates
+
+def fS_I1(r, ini): return ini[0]* lamb(0)
+def fS_I2(r, ini): return ini[0]* lamb(1)
+def fS_I3(r, ini): return ini[0]* lamb(2)
+def fS_I4(r, ini): return ini[0]* lamb(3)
+
+#Primary recovery rates
+
+def fI1_R1(r, ini): return r[-2]*ini[5]
+def fI2_R2(r, ini): return r[-2]*ini[6]
+def fI3_R3(r, ini): return r[-2]*ini[7]
+def fI4_R4(r, ini): return r[-2]*ini[8]
+
+#Secondary Infection rates
+
+def fR1_I11(r, ini): return r[-1]*ini[5]*lamb(0)
+def fR1_I12(r, ini): return r[-1]*ini[5]*lamb(1)
+def fR1_I13(r, ini): return r[-1]*ini[5]*lamb(2)
+def fR1_I14(r, ini): return r[-1]*ini[5]*lamb(3)
+def fR2_I21(r, ini): return r[-1]*ini[6]*lamb(0)
+def fR2_I22(r, ini): return r[-1]*ini[6]*lamb(1)
+def fR2_I23(r, ini): return r[-1]*ini[6]*lamb(2)
+def fR2_I24(r, ini): return r[-1]*ini[6]*lamb(3)
+def fR3_I31(r, ini): return r[-1]*ini[7]*lamb(0)
+def fR3_I32(r, ini): return r[-1]*ini[7]*lamb(1)
+def fR3_I33(r, ini): return r[-1]*ini[7]*lamb(2)
+def fR3_I34(r, ini): return r[-1]*ini[7]*lamb(3)
+def fR4_I41(r, ini): return r[-1]*ini[8]*lamb(0)
+def fR4_I42(r, ini): return r[-1]*ini[8]*lamb(1)
[…]

### 2014-10-15 Modelo Estocástico/dengue_full_SDE.py

@@ -1,73 +1,0 @@
-CC By-SA 3.0
-
-Licença
-
-O TRABALHO (CONFORME ABAIXO DEFINIDO) É DISPONIBILIZADO DE ACORDO COM OS TERMOS DESTA LICENÇA PÚBLICA CREATIVE COMMONS ("LPCC" OU "LICENÇA"). O TRABALHO ESTÁ PROTEGIDO PELO CÓDIGO DO DIREITO DE AUTOR E DOS DIREITOS CONEXOS E/OU POR OUTRAS LEIS DE PROPRIEDADE INTELECTUAL E/OU POR OUTRAS LEIS (“LEGISLAÇÃO APLICÁVEL”). QUALQUER USO DO TRABALHO QUE NÃO O AUTORIZADO POR ESTA LICENÇA OU PELA LEGISLAÇÃO APLICÁVEL É PROIBIDO.
-
-AO EXERCER QUALQUER UM DOS DIREITOS DE EXPLORAÇÃO DO TRABALHO PREVISTOS NA PRESENTE LICENÇA, O UTILIZADOR ESTARÁ A CONCORDAR COM OS TERMOS DESTA LICENÇA E A ACEITAR VINCULAR-SE AOS MESMOS. NA MEDIDA EM QUE A PRESENTE LICENÇA SEJA CONSIDERADA UM CONTRATO, O LICENCIANTE CONCEDE AO UTILIZADOR OS DIREITOS AQUI PREVISTOS, EM CONTRAPARTIDA PELA SUA ACEITAÇÃO DOS TERMOS E CONDIÇÕES AQUI CONTIDOS.
-
-1. Definições
-
-    "Apresentar ao Público" significa, para os propósitos desta Licença, tanto (i) comunicar ao público o Trabalho ou Trabalho Derivado, o que for apropriado, por meio de qualquer ato que torne o seu gozo acessível ao público independentemente da posse de uma cópia do mesmo, nomeadamente através da representação, recitação, declamação, execução, exibição, exposição, transmissão ou retransmissão por fio ou sem fio, do Trabalho ou Trabalho Derivado, como (ii) colocar o Trabalho ou Trabalho Derivado, o que for apropriado, à disposição do público, por fio ou sem fio, por forma a torná-lo acessível a qualquer pessoa a partir do local e no momento por ela escolhido.
[…]

### 2014-10-15 Modelo Estocastico/dengue_full_SDE.py
fixed value of sigma

@@ -19,12 +19,12 @@ tm = genfromtxt('tmat.csv', delimiter=',', skip_header=2, usecols=range(1, 67), 
 
 N = 1000
 
-ini = [N - 4, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0]
+ini = [N - 40, 10, 10, 10, 10, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0]
 
 pars = (1 / 70.,  # mu
         400,  # beta
         1,  # phi
-        1 / 3.65,  # sigma
+        100,  # sigma
         1)  # gamma
 
 # Propensity functions

### 2014-10-15 Modelo Estocastico/propfun.py

@@ -1,5 +1,5 @@
-pars = (0.014285714285714285, 400, 1, 0.273972602739726, 1)
-ini = [996, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0]
+pars = (0.014285714285714285, 400, 1, 100, 1)
+ini = [960, 10, 10, 10, 10, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0]
 N = 1000
 
 lamb = lambda i: 400*(ini[i+1] + sum(ini[9+i : 9+12+i+1 : 4])-ini[9+4*i+1])

### 2014-10-16 draft_lm1.tex
fixed bugs in the model; it is still not parameterized correctly...

@@
 \section*{Introduction}
-When modeling infecctious disease spread using dynamic models, the initial 
-conditions of the system, although key to match the simulated dynamics to 
-observed data, are frequently overlooked.
 
@@
 Determining the viral types in circulation during an epidemic is 
-expensive, as it requires virus isolatio and reverse-transcriptase-PCR. 
+expensive, as it requires virus isolation and reverse-transcriptase-PCR. 
 Therefore only a very limited number of samples are examined on any epidemic. 

### 2014-10-16 Modelo Estocastico/dengue_full_SDE.py

@@ -1,5 +1,5 @@
 # coding:utf8
-
+from __future__ import division
 from BIP.SDE.gillespie import Model
 import time
 from numpy import array, genfromtxt
@@ -12,35 +12,38 @@ vnames = ['S', 'I1', 'I2', 'I3', 'I4', 'R1', 'R2', 'R3', 'R4', 'I11',
 # transitions
 # S->I1	S->I2	S->I3	S->I4	I1=>R1	I2=>R2	I3=>R3	I4=>R4	R1=>I11	R1=>I12	R1=>I13	R1=>I14	R2=>I21	R2=>I22	R2=>I23	R2=>I24	R3=>I31	R3=>I32	R3=>I33	R3=>I34	R4=>I41	R4=>I42	R4=>I43	R4=>I44	I11=>R	I12=R	I13=>R	I14=R	I21=>R	I22=R	I23=>R	I24=R	I31=>R	I32=R	I33=>R	I34=R	I41=>R	I42=R	I43=>R	I44=R
 
-tm = genfromtxt('tmat.csv', delimiter=',', skip_header=2, usecols=range(1, 67), dtype=int)
+tm = genfromtxt('tmat.csv', delimiter=',', skip_header=2, usecols=range(1, 68), dtype=int)
 
 # Initial Conditions
 # Initial conditions for each of the four Serotype model
 
-N = 1000
+N = 5000
 
-ini = [N - 40, 10, 10, 10, 10, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0]
+ini = [N, 150, 150, 150, 150, 150, 150, 150, 150, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0]
 
 pars = (1 / 70.,  # mu
         400,  # beta
         1,  # phi
         100,  # sigma
-        1)  # gamma
+        1.7)  # gamma
 
 # Propensity functions
 def gen_prop_functs():
-    nat_code = "def fnat(r, ini): return r[0]*N\n"
+    nat_code = "def fnat(r, ini): return r[0]*sum(ini)\n"
     mu_code = "def fmu{}(r, ini): return r[0]*ini[{}]\n"
-    inf_code = """def fS(r, ini): return ini[0]*sum([lamb(i) for i in range(4)])\n"""
+    # inf_code = """def fS(r, ini): return ini[0]*sum([lamb(i) for i in range(4)])\n"""
[…]

### 2014-10-16 Modelo Estocastico/propfun.py

@@ -1,9 +1,12 @@
-pars = (0.014285714285714285, 400, 1, 100, 1)
-ini = [960, 10, 10, 10, 10, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0]
-N = 1000
+pars = (0.014285714285714285, 400, 1, 10, 1.7)
+ini = [5000, 150, 150, 150, 150, 150, 150, 150, 150, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0]
+N = 5000
 
 lamb = lambda i: 400*(ini[i+1] + sum(ini[9+i : 9+12+i+1 : 4])-ini[9+4*i+1])
-def fS(r, ini): return ini[0]*sum([lamb(i) for i in range(4)])
+
+#Natality
+
+def fnat(r, ini): return r[0]*sum(ini)
 
 #Infection rates
 
@@ -14,48 +17,48 @@ def fS_I4(r, ini): return ini[0]* lamb(3)
 
 #Primary recovery rates
 
-def fI1_R1(r, ini): return r[-2]*ini[5]
-def fI2_R2(r, ini): return r[-2]*ini[6]
-def fI3_R3(r, ini): return r[-2]*ini[7]
-def fI4_R4(r, ini): return r[-2]*ini[8]
+def fI1_R1(r, ini): return r[3]*ini[1]
+def fI2_R2(r, ini): return r[3]*ini[2]
+def fI3_R3(r, ini): return r[3]*ini[3]
+def fI4_R4(r, ini): return r[3]*ini[4]
 
 #Secondary Infection rates
 
-def fR1_I11(r, ini): return r[-1]*ini[5]*lamb(0)
-def fR1_I12(r, ini): return r[-1]*ini[5]*lamb(1)
-def fR1_I13(r, ini): return r[-1]*ini[5]*lamb(2)
-def fR1_I14(r, ini): return r[-1]*ini[5]*lamb(3)
-def fR2_I21(r, ini): return r[-1]*ini[6]*lamb(0)
-def fR2_I22(r, ini): return r[-1]*ini[6]*lamb(1)
-def fR2_I23(r, ini): return r[-1]*ini[6]*lamb(2)
-def fR2_I24(r, ini): return r[-1]*ini[6]*lamb(3)
-def fR3_I31(r, ini): return r[-1]*ini[7]*lamb(0)
[…]

### 2014-10-16 Modelo Estocastico/dengue_full_SDE.py
changed plotting to show prevalence of the types

@@ -4,6 +4,7 @@ from BIP.SDE.gillespie import Model
 import time
 from numpy import array, genfromtxt
 import pylab as P
+import numpy as np
 
 vnames = ['S', 'I1', 'I2', 'I3', 'I4', 'R1', 'R2', 'R3', 'R4', 'I11',
           'I12', 'I13', 'I14', 'I21', 'I22', 'I23', 'I24', 'I31', 'I32',
@@ -19,13 +20,13 @@ tm = genfromtxt('tmat.csv', delimiter=',', skip_header=2, usecols=range(1, 68), 
 
 N = 5000
 
-ini = [N, 150, 150, 150, 150, 150, 150, 150, 150, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0]
+ini = [N, 150, 100, 75, 50, 15, 15, 15, 15, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0]
 
-pars = (1 / 70.,  # mu
-        400,  # beta
-        1,  # phi
-        100,  # sigma
-        1.7)  # gamma
+pars = (1 / 50.,  # mu
+        0.0004,  # beta
+        1.1,  # phi
+        .1,  # sigma
+        1.1)  # gamma
 
 # Propensity functions
 def gen_prop_functs():
@@ -38,8 +39,9 @@ def gen_prop_functs():
 
     nt = tm.shape[1]
     with open('propfun.py', 'w') as f:
+        f.write("import numpy as np\n\n")
         f.write("pars = {}\nini = {}\nN = {}\n\n".format(pars, ini, N))
-        f.write("lamb = lambda i: {}*(ini[i+1] + sum(ini[9+i : 9+12+i+1 : 4])-ini[9+4*i+1])\n".format(pars[1]))
+        f.write("lamb = lambda i: {}*(ini[i+1] + sum(np.array(ini[9+i : 9+12+i+1 : 4])*{})-ini[9+4*i+1])\n".format(pars[1], pars[4]))
         # f.write(inf_code)
 
         f.write("\n#Natality\n\n")
@@ -99,14 +101,25 @@ assert len(propensity) == tm.shape[1]
 
[…]

### 2014-10-16 Modelo Estocastico/propfun.py

@@ -1,8 +1,10 @@
-pars = (0.014285714285714285, 400, 1, 10, 1.7)
-ini = [5000, 150, 150, 150, 150, 150, 150, 150, 150, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0]
+import numpy as np
+
+pars = (0.02, 0.0004, 1.1, 0.1, 1.1)
+ini = [5000, 150, 100, 75, 50, 15, 15, 15, 15, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0]
 N = 5000
 
-lamb = lambda i: 400*(ini[i+1] + sum(ini[9+i : 9+12+i+1 : 4])-ini[9+4*i+1])
+lamb = lambda i: 0.0004*(ini[i+1] + sum(np.array(ini[9+i : 9+12+i+1 : 4])*1.1)-ini[9+4*i+1])
 
 #Natality
 

### 2014-11-04 fit_dengue_split_big.py
Added dates to estimate S_0 up to 2014

@@ -38,7 +38,7 @@ inicio=0
 inits = np.array([0.999, 0.001, 0.0])  # initial values for state variables.
 
 
-#@jit#('f8[:](f8[:],f8,f8,f8)')
+@jit
 def sir(y, t, *pars):
     '''ODE model'''
     S, I, R = y
@@ -53,7 +53,7 @@ def sir(y, t, *pars):
             tau * I,  #dR/dt
     ])
 
-#@jit#('f8[:,:](f8[:],f8,f8,f8)')
+@jit
 def jac(y, t, *pars):
     S, I, R = y
     s0, m, tau = pars
@@ -61,7 +61,7 @@ def jac(y, t, *pars):
     return np.array([[-(I + m)*beta, -S*beta, 0],
            [(I + m)*beta, S*beta - tau, 0],
            [0, tau, 0]])
-
+@jit
 def model(theta):
     # setting parameters
     s0, m, tau = theta
@@ -160,7 +160,8 @@ def fix_rt(rt):
 
 # # running the analysys
 if __name__ == "__main__":
-    dt = prepdata('aux/data_Rt_dengue_big.csv', 0, 728, 1)
+    # dt = prepdata('aux/data_Rt_dengue_big.csv', 0, 728, 1)
+    dt = prepdata('data_Rt_dengue_complete.csv', 0, 971, 1)
 
     
     # Defining start and end of the simulations
@@ -174,6 +175,10 @@ if __name__ == "__main__":
            dt['time'].index(datetime.datetime(2006, 9, 25)),  # Start of the 2007 epidemic
            dt['time'].index(datetime.datetime(2007, 8, 27)),  # Start of the 2008 epidemic
            dt['time'].index(datetime.datetime(2008, 9, 1)),  # Start of the 2009 epidemic
+
+           dt['time'].index(datetime.datetime(2010, 10, 17)),  # Start of the 2011 epidemic
+           dt['time'].index(datetime.datetime(2011, 8, 29)),  # Start of the 2012 epidemic
+           dt['time'].index(datetime.datetime(2012, 11, 12)),  # Start of the 2013 epidemic
     ]
     tfs = t0s[1:] + [len(dt['time'])]
     tfs = [ dt['time'].index(datetime.datetime(1996, 7, 29)),  # end of the 1996 epidemic
@@ -186,6 +191,10 @@ if __name__ == "__main__":
             dt['time'].index(datetime.datetime(2007, 8, 27)),  # end of the 2007 epidemic
[…]

### 2014-11-04 S1_lm1.tex
changes to S1, long forgotten, sorrey :-(

@@
 \label{sec:tails}
-
-In order to decide which approach to take, it may be of use analysing the variance and tail behaviour of the derived distributions for \rr. 
+There are a number of approaches to deriving the distribution of \rr.
+One could for example  use the distribution in equation A6 of~\citet{nishiura} 
+or estimate \rr using the approach described in the main text~\citep{mantel}.
+In order to decide which approach to take, it may be of use analyzing the 
+variance and tail behavior of the derived distributions for \rr. 
 Consider the case of using a flat $Uniform(0, 1)$ prior for $\theta_t$.
 With $a_0 = b_0 = 1$, $a_1 = a_2$ and $b_1 = b_2 + 1$.
-The beta prime (inverse beta distribution) will have heavier tails compared to the conditional distribution in (\ref{eq:densityNishiura}), thus providing more conservative confidence/credibility intervals.
-As a side note, the Bayesian approach presented in Section~\ref{sec:mantel} will give similar results to those of \citet{wilson} and \citet{wilson} for $Y_{t+1}$ and $Y_t >> 1$.
+The beta prime (inverse beta distribution) will have heavier tails compared to 
+the conditional distribution proposed by~\citet{nishiura}, thus providing more 
+conservative confidence/credibility intervals.  
+As a side note, the Bayesian approach presented in this 
+paper will give similar results to those of 
+\citet{wilson} and \citet{wilson} for $Y_{t+1}$ and $Y_t >> 1$.
 Under the flat  uniform prior for $\theta_t$, the Bayesian posterior credibility interval is nearly indistinguishable from the confidence interval proposed by \citet{clopper} for $Y_{t+1}, Y_t > 20$.
@@
 
-The main advange of the Bayesian approach is that one can devise prior distributions for $\theta_t$ taking advantage of the intuitive parametrisation and flexibility of the beta family of distributions.
+An advantage of the Bayesian approach is that one can devise prior 
+distributions for $\theta_t$ taking advantage of the intuitive parameterization 
+and flexibility of the beta family of distributions.
 Prior elicitation can also be done for \rr and the hyperparameters directly plugged into the prior for $\theta_t$. 
@@
 \end{align}
+
+This approach thus makes possible to incorporate epidemiological knowledge 
+about disease biology (e.g. the magnitude of $R_0$) into the computation of \rr.
+This may prove particularly important when disease counts are low and/or close 
+to the detection threshold.
 \newpage

### 2014-11-04 draft_lm1.tex
changed title. fixes issue #1
started rewriting the introduction, and added afew extra references to the bibliography

@@
 {\Large
-\textbf{Estimating the Initial Susceptible Fraction in Dengue Dynamics}
+\textbf{Estimating the Attack ratio of Dengue Epidemics without serotype 
+information}
 }
@@
 Dengue is a disease which is caused by 4 different types of viruses 
-which co-circulate between the vector and human populations. immunity is 
-considered permanent after the exposure of humans to each of the viruses but 
+which co-circulate between the vector and human populations. Immunity is 
+considered permanent after the exposure of humans to each of the viruses and 
 cross immunity is considered very limited[ref].
@@
 
-Dengue transmission is conditioned on temperature, due to its effects on the 
-vector reproduction.
-In places with sufficient temperature variations dengue is a summer disease 
-mainly.
+Dengue transmission is modulated by environmental conditions, of which 
+temperature, due to its effects on the 
+vector reproduction, stands out as a strong predictor of 
+incidence\cite{honorio_temporal_2009,wu_higher_2009}.
+In places with sufficient temperature variation dengue is predominantly a 
+summer disease.
 
-Determining the viral types in circulation during an epidemic is 
-expensive, as it requires virus isolation and reverse-transcriptase-PCR. 
-Therefore only a very limited number of samples are examined on any epidemic. 
-Besides, mild cases are rarely detected let alone tested for dengue.
+The long term dynamics of dengue is also modulated by the alternation of virus 
+types in circulation Demographics also plays a role in replenishing the 
+population of susceptibles.
 
-Dengue in Brazil is a disease which is well in its way to 
-endemicity but not yet stable, therefore co-circulation patterns have not yet 
-reached a steady state as has been demonstrated for certain countries in 
-south-east asia which have had the disease much longer.
+The Attack ratio of disease is a mesure of mobidity defined as the number of new
+cases divided byt the population at risk. For Dengue outbreaks and epidemics 
+this can be difficult to calculated due to the lack of information about the 
+population at risk. The population at risk in the case of dengue, is the number 
+of susceptibles to the circulating virus type(s) befor a given epidemic.
 
-In the particular case of Dengue fever in Brazil, with four circulating 
-serotypes, it can be very challenging to parameterize a simulation starting at 
-any time other than when the date the disease was reintroduced in the country 
-in the late 1980s and even so, without virological survey data from time to 
-time, to inform on the mixing of viruses in circulation, it is virtually 
-impossible to predict, by simulation alone, the imunological state of the 
-population at any point in the future.
+The main source of information about Dengue dynamics are the incidence 
+timeseries, which represent all the cases of Dengue reported by health 
+professionals with no discrimination of virus types. Only a very small sample 
+of the cases are ever examined to determined the virus type.
 
-For SIR-like models, the relevant initial values are the number of infectious 
-and susceptible  individuals. 
-The number of infectious individuals are less challenging to determine, because 
-the get very small during the winters and can be derived from incidence data  
-with is monitored continuously.
-The real challenge is to determine the number of susceptibles to 
-a particular strain right at the beginning of the dengue season.
+In order to calculate the attack ratio, we need to determine the number of 
+susceptibles to the circulating virus types right before the epidemic. Without 
+regular virological surveys, it is virtually impossible to determine the 
+population at risk.
 
-Multiple factors contribute to the dificulty in estimating the number of 
-susceptibles in the population, first and foremost is the fact that dengue
-confers permanent immunity within serotypes but with only short-lived 
-cross-immunity. 
+The attack ratio is also influenced by the reproductive number of the disease 
+\cite{bacaer_final_2009, katriel_attack_2012}. With Dengue, since transmission 
+is affected by environmental conditions as discussed above, the reproductive 
+number is a function of time.
+
+%%%% Revised up to here
+
 Secondly, there are a larger number of asymptomatic cases in 

### 2014-11-04 fit_dengue_split_big.py

@@ -38,7 +38,7 @@ inicio=0
 inits = np.array([0.999, 0.001, 0.0])  # initial values for state variables.
 
 
-@jit
+#@jit
 def sir(y, t, *pars):
     '''ODE model'''
     S, I, R = y
@@ -53,7 +53,7 @@ def sir(y, t, *pars):
             tau * I,  #dR/dt
     ])
 
-@jit
+#@jit
 def jac(y, t, *pars):
     S, I, R = y
     s0, m, tau = pars
@@ -61,7 +61,7 @@ def jac(y, t, *pars):
     return np.array([[-(I + m)*beta, -S*beta, 0],
            [(I + m)*beta, S*beta - tau, 0],
            [0, tau, 0]])
-@jit
+#@jit
 def model(theta):
     # setting parameters
     s0, m, tau = theta
@@ -177,8 +177,8 @@ if __name__ == "__main__":
            dt['time'].index(datetime.datetime(2008, 9, 1)),  # Start of the 2009 epidemic
 
            dt['time'].index(datetime.datetime(2010, 10, 17)),  # Start of the 2011 epidemic
-           dt['time'].index(datetime.datetime(2011, 8, 29)),  # Start of the 2012 epidemic
-           dt['time'].index(datetime.datetime(2012, 11, 12)),  # Start of the 2013 epidemic
+           dt['time'].index(datetime.datetime(2011, 8, 28)),  # Start of the 2012 epidemic
+           dt['time'].index(datetime.datetime(2012, 11, 11)),  # Start of the 2013 epidemic
     ]
     tfs = t0s[1:] + [len(dt['time'])]
     tfs = [ dt['time'].index(datetime.datetime(1996, 7, 29)),  # end of the 1996 epidemic
@@ -230,8 +230,8 @@ if __name__ == "__main__":
 
 
 
-    for inicio, fim in zip(t0s, tfs)[3:4]: # Slice to force start from a different point
-        dt = prepdata('data_Rt_dengue_big.csv', inicio, fim, 1)
+    for inicio, fim in zip(t0s, tfs):#[3:4]: # Slice to force start from a different point
+        dt = prepdata('data_Rt_dengue_complete.csv', inicio, fim, 1)
         # Interpolated Rt
         iRt = interp1d(np.arange(dt['Rt'].size), np.array(dt['Rt']), kind='linear', bounds_error=False, fill_value=0)
         ano = dt['time'][0].year
[…]

### 2014-11-06 draft_lm1.tex
finished rewriting the introduction

@@
 is affected by environmental conditions as discussed above, the reproductive 
-number is a function of time.
+number is also expected to vary with time.
 
@@
 
-The initial number of susceptibles -- $S_0$ -- is a a very powerful parameter 
-with regards to epidemic risk.
-Its value affects directly the local $R_0$, thus being intrumental in predicting 
-epidemics, but it also affects the speed and size of the epidemic.
+In this paper, we estimate the number of susceptibles before every major 
+outbreak ($S_{0, year}$) in the last 18 years. Due to the alternation of 
+ciculating serotypes, we detected fluctuating levels of susceptibility the 
+circulating virus(es). To Estimate the susceptibles we have we have fixed the 
+tramissibility based on the $R_t$ series derived from data\cite{nishiura}. Then, 
+from the incidence series and the population at risk, we could then calculate 
+the attack ratio for each epidemic.
 
-In this paper, we estimate the number of susceptibles, allowing for 
-discontinuities right before every epidemic in the last 18 years. 
-Through this method we hope to detect drastic changes in the mixture of 
-susceptibles cause by the replacement of the predominant serotype, which 
-can redefine susceptibility for the following season.
-
-To accomplish this we use a simplified model of 
-dengue transmission based on a Susceptible-infecctious-Removed (SIR) model. 
-Since Dengue is not caused by a single pathogen, but by four separate strains 
-with limited cross-immunity, when we talk about number of susceptibles, in the 
-context of a single serotype SIR model, we are actually referring to the sum of 
-the susceptibles to the currently circulating serotypes at any given time.
-
-Some information is available about the predominant serotypes during the 
-period of study\cite{macedo_virological_2013}.
+To accomplish the estimation of the $S_0$, we use a simplified model of 
+dengue transmission based on a single-strain Susceptible-infecctious-Removed 
+(SIR) model. The reason for the choice of a single strain model, is the lack of 
+accurate information regarding the actual proportions of each virus in 
+circulation. However, some information is available about the predominant 
+serotypes for some epidemics in the period 
+of study\cite{macedo_virological_2013}.
 
@@
 new serotypes (which are a stochastic events) can change drastically the pool 
-of susceptibles throwing off any sequential estimation of the dynamics. 
-Adopting a multi-strain transmission model, does not help because of 
-insufficient information about the proportions of each serotype in case data.
+of susceptibles throwing off any sequential estimation based on the incidence
+dynamics.
 

### 2014-11-06 fit_dengue_split_big_plot.py

@@ -230,7 +230,7 @@ def plot_concat_series(dbs):
         co = c.next()
         i_median = srs.I.groupby(level='time').median()
         i_median.plot(style='k-', label='Median')
-        da.I.groupby(level='time').plot(style='r*', label='cases', alpha=.5)
+        da.I.groupby(level='time').plot(style='r.', label='cases')
         i_upr = srs.I.groupby(level='time').aggregate(upr)
         i_lwr = srs.I.groupby(level='time').aggregate(lwr)
         P.fill_between(i_median.index, i_lwr, i_upr, facecolor=co, alpha=.2)
@@ -256,12 +256,12 @@ if __name__ == "__main__":
     #~ series('Dengue_S0_big')
 
     dbs = glob.glob("DengueS*.sqlite")
-    #plot_concat_series(dbs)
-    print create_tex_table(dbs)
+    plot_concat_series(dbs)
+    #print create_tex_table(dbs)
 
 
     # sns.tsplot(series.S, time=series.time)
 
     # series.groupby(level='time').plot()
     P.show()
-    # print create_tex_table(['be-multiyear', 'nl-multiyear', 'pt-multiyear'])
+

### 2014-11-06 draft_lm1.tex
improved introduction

@@
 
-%%%% Revised up to here
-
-Secondly, there are a larger number of asymptomatic cases in 
-any epidemic, which nevertheless acquire immunity. It is estimated that for 
+Various challenges exist to assess the attack rate of a dengue epidemic. One of 
+them is that there are a large number of asymptomatic cases in any epidemic, 
+which nevertheless acquire immunity. It is estimated that for 
 every case reported, 10-20 are not seen by health authorities (personnal 

### 2014-11-07 draft_lm1.tex
fixed stochastic model

@@
 In monitoring of infectious diseases, it is important to assess whether the 
-incidence of a a particular disease is increasing significantly, in order to 
+incidence of a  particular disease is increasing significantly, in order to 
 decide to take preventive measures.

### 2014-11-07 Modelo Estocastico/dengue_full_SDE.py

@@ -6,27 +6,36 @@ from numpy import array, genfromtxt
 import pylab as P
 import numpy as np
 
-vnames = ['S', 'I1', 'I2', 'I3', 'I4', 'R1', 'R2', 'R3', 'R4', 'I11',
-          'I12', 'I13', 'I14', 'I21', 'I22', 'I23', 'I24', 'I31', 'I32',
-          'I33', 'I34', 'I41', 'I42', 'I43', 'I44', 'R']
+from itertools import cycle
+
+vnames = ['S', 'I1', 'I2', 'I3', 'I4', 'R1', 'R2', 'R3', 'R4',
+          'I12', 'I13', 'I14', 'I21', 'I23', 'I24', 'I31', 'I32',
+            'I34', 'I41', 'I42', 'I43', 'R']
 
 # transitions
-# S->I1	S->I2	S->I3	S->I4	I1=>R1	I2=>R2	I3=>R3	I4=>R4	R1=>I11	R1=>I12	R1=>I13	R1=>I14	R2=>I21	R2=>I22	R2=>I23	R2=>I24	R3=>I31	R3=>I32	R3=>I33	R3=>I34	R4=>I41	R4=>I42	R4=>I43	R4=>I44	I11=>R	I12=R	I13=>R	I14=R	I21=>R	I22=R	I23=>R	I24=R	I31=>R	I32=R	I33=>R	I34=R	I41=>R	I42=R	I43=>R	I44=R
+# S->I1	S->I2	S->I3	S->I4	I1=>R1	I2=>R2	I3=>R3	I4=>R4	R1=>I12	R1=>I13	R1=>I14	R2=>I21	R2=>I23	R2=>I24	R3=>I31	R3=>I32	R3=>I34	R4=>I41	R4=>I42	R4=>I43	I12=R	I13=>R	I14=R	I21=>R	I23=>R	I24=R	I31=>R	I32=R	I34=R	I41=>R	I42=R	I43=>R
 
-tm = genfromtxt('tmat.csv', delimiter=',', skip_header=2, usecols=range(1, 68), dtype=int)
+tm = genfromtxt('tmat.csv', delimiter=',', skip_header=2, usecols=range(1, 56), dtype=int)
 
 # Initial Conditions
 # Initial conditions for each of the four Serotype model
 
 N = 5000
 
[…]

### 2014-11-07 Modelo Estocastico/propfun.py

@@ -1,10 +1,18 @@
 import numpy as np
 
-pars = (0.02, 0.0004, 1.1, 0.1, 1.1)
-ini = [5000, 150, 100, 75, 50, 15, 15, 15, 15, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0]
+pars = (0.014285714285714285, 400, 2.4, 0.1, 1.8)
+ini = [5000, 150, 100, 75, 50, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0]
 N = 5000
 
-lamb = lambda i: 0.0004*(ini[i+1] + sum(np.array(ini[9+i : 9+12+i+1 : 4])*1.1)-ini[9+4*i+1])
+# infectives for each serotype
+
+inf_types = {1: [ini[1], ini[12], ini[15], ini[18]],
+             2: [ini[2], ini[9], ini[16], ini[19]],
+             3: [ini[3], ini[10], ini[13], ini[20]],
+             4: [ini[4], ini[11], ini[14], ini[17]]
+            }
+
+lamb = lambda i: 400*(ini[i+1] + sum(np.array(inf_types[i+1])*2.4))
 
 #Natality
 
@@ -26,41 +34,33 @@ def fI4_R4(r, ini): return r[3]*ini[4]
 
 #Secondary Infection rates
 
-def fR1_I11(r, ini): return r[4]*ini[5]*lamb(0)
 def fR1_I12(r, ini): return r[4]*ini[5]*lamb(1)
 def fR1_I13(r, ini): return r[4]*ini[5]*lamb(2)
 def fR1_I14(r, ini): return r[4]*ini[5]*lamb(3)
 def fR2_I21(r, ini): return r[4]*ini[6]*lamb(0)
-def fR2_I22(r, ini): return r[4]*ini[6]*lamb(1)
 def fR2_I23(r, ini): return r[4]*ini[6]*lamb(2)
 def fR2_I24(r, ini): return r[4]*ini[6]*lamb(3)
 def fR3_I31(r, ini): return r[4]*ini[7]*lamb(0)
 def fR3_I32(r, ini): return r[4]*ini[7]*lamb(1)
-def fR3_I33(r, ini): return r[4]*ini[7]*lamb(2)
 def fR3_I34(r, ini): return r[4]*ini[7]*lamb(3)
 def fR4_I41(r, ini): return r[4]*ini[8]*lamb(0)
[…]

### 2014-11-16 Modelo Estocastico/dengue_full_SDE.py
improved methods section; started results

@@ -121,7 +121,7 @@ assert len(propensity) == tm.shape[1]
 
 M = Model(vnames=vnames, rates=pars, inits=ini, tmat=tm, propensity=propensity)
 t0 = time.time()
-M.run(tmax=10000, reps=1)
+M.run(tmax=1000, reps=1)
 print 'total time: {} seconds'.format(time.time() - t0)
 t, series, steps = M.getStats()
 

### 2014-11-17 Modelo Estocastico/dengue_full_SDE.py
updated model

@@ -22,7 +22,7 @@ tm = genfromtxt('tmat.csv', delimiter=',', skip_header=2, usecols=range(1, 56), 
 
 N = 5000
 
-ini = [N, 50, 40, 30, 20, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0]
+ini = (N, 50, 30, 0, 0, 0, 0, 0, 0, 0, 0, 5, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
 
 pars = (1 / 70.,  # mu
         400,  # beta
@@ -43,9 +43,9 @@ inf_types = {1: [ini[1], ini[12], ini[15], ini[18]],
 # Propensity functions
 def gen_prop_functs():
     nat_code = "def fnat(r, ini): return r[0]*sum(ini)\n"
-    mu_code = "def fmu{}(r, ini): return r[0]*ini[{}]\n"
+    mu_code = "def fmu{}(r, ini): return r[0]*max(0,ini[{}])\n"
     inf_code = """def f{}(r, ini): return ini[0]* lamb({})\n"""
-    rec_code = """def f{}(r, ini): return r[3]*ini[{}]\n"""
+    rec_code = """def f{}(r, ini): return r[3]*max(0,ini[{}])\n"""
     inf2_code = """def f{}(r, ini): return r[4]*r[5]*ini[{}]*lamb({})\n"""
 
     nt = tm.shape[1]
@@ -97,7 +97,7 @@ def gen_prop_functs():
 gen_prop_functs()
 from propfun import *
 
-propensity = [fnat,  # Natality
+propensity = [fnat,                             # Natality
               fS_I1, fS_I2, fS_I3, fS_I4,       # Primary infections
               fI1_R1, fI2_R2, fI3_R3, fI4_R4,   # primary recovery
               fR1_I12, fR1_I13, fR1_I14,        #Secondary infections
@@ -121,11 +121,12 @@ assert len(propensity) == tm.shape[1]
 
 M = Model(vnames=vnames, rates=pars, inits=ini, tmat=tm, propensity=propensity)
 t0 = time.time()
-M.run(tmax=10000, reps=1)
+M.run(tmax=500, reps=1)
 print 'total time: {} seconds'.format(time.time() - t0)
 t, series, steps = M.getStats()
 
 ser = series.mean(axis=2)
+print ser[:,11]
 
 #  Calculatin prevalence by serotype
 

### 2014-11-17 Modelo Estocastico/propfun.py

@@ -1,7 +1,7 @@
 import numpy as np
 
 pars = (0.014285714285714285, 400, 1.0, 0.1, 1.0, 0.001)
-ini = [5000, 50, 40, 30, 20, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0]
+ini = (5000, 50, 30, 0, 0, 0, 0, 0, 0, 0, 0, 5, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
 N = 5000
 
 # infectives for each serotype
@@ -27,10 +27,10 @@ def fS_I4(r, ini): return ini[0]* lamb(3)
 
 #Primary recovery rates
 
-def fI1_R1(r, ini): return r[3]*ini[1]
-def fI2_R2(r, ini): return r[3]*ini[2]
-def fI3_R3(r, ini): return r[3]*ini[3]
-def fI4_R4(r, ini): return r[3]*ini[4]
+def fI1_R1(r, ini): return r[3]*max(0,ini[1])
+def fI2_R2(r, ini): return r[3]*max(0,ini[2])
+def fI3_R3(r, ini): return r[3]*max(0,ini[3])
+def fI4_R4(r, ini): return r[3]*max(0,ini[4])
 
 #Secondary Infection rates
 
@@ -49,40 +49,40 @@ def fR4_I43(r, ini): return r[4]*r[5]*ini[8]*lamb(2)
 
 #Secondary Recovery rates
 
-def fI12_R(r, ini): return r[3]*ini[9]
-def fI13_R(r, ini): return r[3]*ini[10]
-def fI14_R(r, ini): return r[3]*ini[11]
-def fI21_R(r, ini): return r[3]*ini[12]
-def fI23_R(r, ini): return r[3]*ini[13]
-def fI24_R(r, ini): return r[3]*ini[14]
-def fI31_R(r, ini): return r[3]*ini[15]
-def fI32_R(r, ini): return r[3]*ini[16]
-def fI34_R(r, ini): return r[3]*ini[17]
-def fI41_R(r, ini): return r[3]*ini[18]
-def fI42_R(r, ini): return r[3]*ini[19]
-def fI43_R(r, ini): return r[3]*ini[20]
+def fI12_R(r, ini): return r[3]*max(0,ini[9])
+def fI13_R(r, ini): return r[3]*max(0,ini[10])
+def fI14_R(r, ini): return r[3]*max(0,ini[11])
[…]

### 2014-11-17 draft_lm1.tex

@@
 
-The estimation of the  number of susceptibles have been attempted 
+Methods for estimating the  number of susceptibles have been proposed 
 before, for other diseases\cite{bjornstad_dynamics_2002, 
@@
 
-\subsection*{Estimating the reproductive number}
+In this section we will start by describing how we estimated the effective 
+reproductive number, $R_t$, relative to the global dengue cases, and it's 
+posterior probability distribution. 
+We then proceed to describe the SIR model 
+used to represent the global incidence and how the $R_t$ is integrated into the 
+model to allow for time varying transmission. 
+Then we describe the Bayesian 
+estimation methodology applied to find the posterior probability distributions 
+of the susceptibles to the main circulating Dengue viruses for each epidemic, 
+using Dengue incidence time series for Rio de Janeiro from 1996 to 2014.
+In the last subsection we describe how we estimated the Attack ratio of each 
+epidemic from the estimated susceptibility and observed incidence.
+
+\subsection*{Estimating the effective reproductive number, $R_t$}
 
@@
 out.
-Also, although there probably are multiple circulating serotypes, our approach
+The rationale for this simplification is based on the ability of the 
+empirically derived $R_t$  to incorporate the effects of the fluctuating vector 
+populations.
+Also, although there are multiple circulating serotypes, our approach
 does not discriminate between them due to the lack of serotype-specific data.
@@
 
-For the fitting procedure the time series was divided into $J=10$ epidemic 
+For the fitting procedure the incidence time series was divided into $J=13$ 
+epidemic 
 windows that corresponded to significant raises in incidence and normalized to 
@@
 
+\subsection*{Calculating the attack ratio}
+
+The attack ratio of an epidemic is define by the number of infections divided 
+by the size of the population at risk.
+
+\begin{equation}
+\label{eq:AR}
+A=\frac{cases}{Pop.\, at\, risk} 
+\end{equation}
+
+Based on the what have been discussed so far, we can rewrite (\ref{eq:AR}) for 
+each epidemic $j$ as
+\begin{equation}
+\label{eq:AR2}
+ A_{j}=\frac{\sum Y_j}{S_{0j}}
+\end{equation}
+where $S_{0j}$ is the number of susceptibles before each epidemic $j$, 
+estimated as described above.
+
 % Results and Discussion can be combined.
-\section*{Results}
+\section*{Results and Discussion}
+
+The time series used for all the analyses consist of officially notified dengue 
+cases agregated weekly.
 % We only support three levels of headings, please do not create a heading level below \subsubsection.
@@
 LMC is grateful to Dr. Leonardo Bastos for useful discussions on the 
-posterior inference for $R(t)$.
+posterior inference for $R(t)$. The authors are also grateful to Claudia T. 
+Codeco for helpful discussions about the manuscript.
 

### 2014-11-17 draft_lm1.tex
added table for attack ratios; updated the figure for the estimation; started writing the results;

@@
 \begin{align}
- \frac{dS}{dt} &= -\beta(t)SI \\
- \frac{dI}{dt} &= \beta(t)SI - \tau I&\\
+   \label{eq:model}
+ \frac{dS}{dt} &= -\beta(t)SI \\     \nonumber
+ \frac{dI}{dt} &= \beta(t)SI - \tau I&\\      \nonumber
  \frac{dR}{dt} &= \tau I&
@@
 Convergence of the parallel chains was verified at every 100 iterations by the 
-calculation of the Gelman-Rubins' R convergence diagnostic~\cite{brooks1998}.
+calculation of the Gelman-Rubin's R convergence diagnostic~\cite{brooks1998}.
 
@@
 
-The time series used for all the analyses consist of officially notified dengue 
-cases agregated weekly.
+The time series used for all the analyses consists of officially notified 
+dengue cases aggregated weekly. 
+Figure XX  shows the $R_t$ series, according to 
+(\ref{eq:Rtestimate}). 
+It can be seen that the inter-epidemic periods are characterized by $R_t$ being 
+indistinguishable from 1.
+
+Figure \ref{Fig:S0}, shows the model from (\ref{eq:model}) fitted to the data.
+
+Table \ref{tab:AR} contains the attack ratios calculated for each 
+epidemic/outbreak.
 % We only support three levels of headings, please do not create a heading level below \subsubsection.
@@
 \begin{center}
-\includegraphics[width=.9\textwidth]{./plots/concat_SI.png}
+\includegraphics[width=.9\textwidth]{./concat_SI.png}
 \end{center}
 \caption{
-{\bf Bold the first sentence.}  Rest of figure caption.  
+{\bf Susceptibles and Infectious posterior curves.}  The curves were estimated 
+only for the periods where $R_t> 1$.  The susceptible curves in the top panel 
+reflect the prevalence of fraction of susceptibles to circulating strain(s) for 
+each epidemic/outbreak. In the lower panel see the posterior distribution 
+of infectious curves, represented by its Median and 95\% credible interval. 
+Scatter plot shows the observed cases.
 }
@@
 
+\begin{table}[!ht]
+\caption{
+\bf{Median Attack ration calculated according to (\ref{eq:AR2})}}
+\begin{center}
+\begin{tabular}{l|r}
+\hline
+Year & Median Attack Ratio \\
+\hline
+1996 & \textbf{38.64\%} (17.18\%-54.19\%)\\
+1997 & \textbf{87.31\%} (73.83\%-87.31\%)\\
+1998 & \textbf{49.86\%} (49.19\%-49.86\%)\\
+1999 & \textbf{10.76\%} (3.72\%-20.24\%)\\
+2000 & \textbf{25.49\%} (24.47\%-26.98\%)\\
+2001 & \textbf{48.12\%} (46.57\%-49.15\%)\\
+2005 & \textbf{14.74\%} (10.23\%-21.38\%)\\
+2006 & \textbf{11.22\%} (8.00\%-14.41\%)\\
+2007 & \textbf{15.08\%} (14.76\%-15.38\%)\\
+2008 & \textbf{13.73\%} (3.10\%-30.84\%)\\
+2010 & \textbf{18.30\%} (17.34\%-19.33\%)\\
+2011 & \textbf{8.59\%} (8.20\%-9.36\%)\\
+2012 & \textbf{13.92\%} (12.71\%-14.92\%)\\
+\hline
+\end{tabular}
+\begin{flushleft}Medians are presented with 95\% credible intervals.
+\end{flushleft}
+
+\end{center}
+
+\label{tab:AR}
+\end{table}
+
 \section*{Supporting Information Legends}

### 2014-11-17 fit_dengue_split_big.py

@@ -230,7 +230,7 @@ if __name__ == "__main__":
 
 
 
-    for inicio, fim in zip(t0s, tfs):#[3:4]: # Slice to force start from a different point
+    for inicio, fim in zip(t0s, tfs)[3:4]: # Slice to force start from a different point
         dt = prepdata('data_Rt_dengue_complete.csv', inicio, fim, 1)
         # Interpolated Rt
         iRt = interp1d(np.arange(dt['Rt'].size), np.array(dt['Rt']), kind='linear', bounds_error=False, fill_value=0)
@@ -261,7 +261,7 @@ if __name__ == "__main__":
                      pdists=[st.beta] * nph, ppars=[(1, 1)] * nph, plims=[(0, 1)] * nph)
         
         
-        F.run(dt, 'DREAM', likvar=1e-9, pool=False, ew=0, adjinits=True, dbname=modname, monitor=['I', 'S'])
+        F.run(dt, 'DREAM', likvar=1e-10, pool=False, ew=0, adjinits=True, dbname=modname, monitor=['I', 'S'])
         #~ print F.AIC, F.BIC, F.DIC
         #print F.optimize(data=dt,p0=[s0,s1,s2], optimizer='scipy',tol=1e-55, verbose=1, plot=1)
         F.plot_results(['S', 'I'], dbname=modname, savefigs=1)

### 2014-11-17 fit_dengue_split_big_plot.py

@@ -48,81 +48,40 @@ def create_tex_table(dbs):
     for db in dbs:
         y = db.split('_')[0][-4:]
         data, theta, srs = read_data(db)
+        #print data
         series[y] = srs
-        obs[y] = data
+        obs[y] = data.I
         pts[y] = theta
 
     head = r"""\begin{center}
-            \begin{tabular}{c|c}
-            \hline
-            """
+\begin{tabular}{c|c}
+\hline
+"""
     head += r"""Year & median Attack Ratio \\
-            \hline
-            """
+\hline
+"""
     bot = r"""
-        \hline
-        \end{tabular}
-        \end{center}
+\hline
+\end{tabular}
+\end{center}
         """
     body = r""
     st = []
+    # years = sorted(list(series.keys()))
     for i, Y in enumerate(series.keys()):
         cases = obs[Y].sum()
         first_week = series.values()[i].index[0]
-        ratio = 1.0*cases/array(series[Y].S.ix[first_week]) 
-        body += Y + r" & {:.2%}({:.2%}-{:.2%})\\".format(nanmedian(ratio), stats.scoreatpercentile(ratio, 2.5), stats.scoreatpercentile(ratio, 97.5))
-        body += "\n"
+        try:
+            ratio = 1.0*cases/array(series[Y].S.ix[first_week])
+            body += Y + r" & {:.2%} ({:.2%}-{:.2%})\\".format(nanmedian(ratio), stats.scoreatpercentile(ratio, 2.5), stats.scoreatpercentile(ratio, 97.5))
+            body += "\n"
+        except KeyError as e:
+            print Y, first_week, e
+        
 
     return head + body + bot
 
 
-def plot_series2(tim, obs, series, names=[], title='Simulated vs Observed series', wl=7, lag=False):
-    ser2 = {}
-    for n in series[0].dtype.names:
-        ser2[n] = concatenate([s[n] for s in series], axis=1)
-    ls = ser2[n].shape[1]
-    tim = tim[:ls]
-    # print type (series)#.I.shape
-    #fig =P.gcf()
-    if not names:
-        names = series[0].dtype.names
-    c = cycle(['b', 'g', 'r', 'c', 'm', 'y', 'k'])
-    s = cycle(['o', '^', '>', '<', 's', '*', '+', '1'])
-    if isinstance(tim[0], datetime.date):
-        lag = datetime.timedelta(int(lag) * wl)
-    else:
[…]

### 2014-11-18 draft_lm1.tex
added heatmap pf cases for the last three epidemics
added some more to the results and discussion section.

@@
 Figure \ref{Fig:S0}, shows the model from (\ref{eq:model}) fitted to the data. 
+In it we can see that the susceptibles series in each epidemic starts at the 
+estimated level of $S_0$. The proportion of susceptibles may seem low, but we 
+must remember that these estimates are being affected by an unknown 
+underreporting factor, which experts suggest is somewhere between 5 and 10, 
+i.e. for every case observed there are 5 or 10 unobserved. Since this 
+underreporting ration affects both the numerator and denominator of 
+(\ref{eq:AR}), its effects should cancel out, giving us an unbiased attack ratio 
+estimate. One other possible source of bias which would lead to the 
+underestimation of $S_0$ could come from a significant part of the population 
+not being exposed to the disease. But, as we can see in figure xx\todo{add 
+figure with the four maps}, the disease spread pretty much homogeneously over 
+the entire city (in the last four epidemics, at least).
 
@@
 of infectious curves, represented by its Median and 95\% credible interval. 
-Scatter plot shows the observed cases.
+Scatter plot shows the observed cases. Prevalences are scaled as fractions 
+of the entire population.
 }

### 2014-11-19 draft_lm1.tex
Stochastic model working with cain via stochpy

@@
 To approximate the posterior in~(\ref{eq:S0post}) we use Markov chain Monte 
-Carlo techiniques implemented in the Bayesian inference with Python 
+Carlo techniques implemented in the Bayesian inference with Python 
 (BIP)~\cite{pone2011} available at 
@@
 posterior inference for $R(t)$. The authors are also grateful to Claudia T. 
-Codeco for helpful discussions about the manuscript.
+Codeço for helpful discussions about the manuscript.
 

### 2014-11-19 Modelo Estocastico/SDE_alt.py

@@ -1,0 +1,19 @@
+#!/usr/bin/env python
+
+"""
+Version of the dengue 4 serotype model implemented for Stochpy
+"""
+import stochpy
+import os
+
+
+smod = stochpy.SSA()
+smod.model_dir = os.getcwd()
+smod.output_dir = os.getcwd()
+smod.Model('Dengue_full.psc')
+
+#smod.DoStochSim(trajectories = 10,mode = 'time',end = 100, epsilon=0.03, method="TauLeaping", IsTrackPropensities=False)
+smod.DoCainStochSim(endtime=100,frames=10000,trajectories=10, solver="HomogeneousDirect2DSearch",IsTrackPropensities=False)
+smod.GetRegularGrid()
+smod.PlotAverageSpeciesTimeSeries()
+#smod.PlotAveragePropensitiesTimeSeries()

### 2014-11-19 Modelo Estocastico/dengue_full_SDE.py

@@ -20,7 +20,7 @@ tm = genfromtxt('tmat.csv', delimiter=',', skip_header=2, usecols=range(1, 56), 
 # Initial Conditions
 # Initial conditions for each of the four Serotype model
 
-N = 5000
+N = 500
 
 ini = (N, 50, 30, 0, 0, 0, 0, 0, 0, 0, 0, 5, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
 
@@ -121,7 +121,7 @@ assert len(propensity) == tm.shape[1]
 
 M = Model(vnames=vnames, rates=pars, inits=ini, tmat=tm, propensity=propensity)
 t0 = time.time()
-M.run(tmax=500, reps=1)
+M.run(tmax=1000, reps=1)
 print 'total time: {} seconds'.format(time.time() - t0)
 t, series, steps = M.getStats()
 
@@ -151,5 +151,8 @@ P.legend()
 # P.plot(t, ser[:, 1::3], 'g-^')  # I plots
 # P.plot(t, ser[:, 2::3], 'b-o')  # R plots
 # P.legend(M.vn[0::3] + M.vn[1::3] + M.vn[2::3], loc=0)
+P.figure()
+for prop in propensity:
+    P.plot(t, [prop(pars,i) for i in ser])
 P.show()
 

### 2014-11-19 Modelo Estocastico/propfun.py

@@ -1,8 +1,8 @@
 import numpy as np
 
 pars = (0.014285714285714285, 400, 1.0, 0.1, 1.0, 0.001)
-ini = (5000, 50, 30, 0, 0, 0, 0, 0, 0, 0, 0, 5, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
-N = 5000
+ini = (500, 50, 30, 0, 0, 0, 0, 0, 0, 0, 0, 5, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
+N = 500
 
 # infectives for each serotype
 

### 2014-11-19 Modelo Estocastico/SDE_alt.py
Stochastic model seems to be ok on stochpy

@@ -5,15 +5,37 @@ Version of the dengue 4 serotype model implemented for Stochpy
 """
 import stochpy
 import os
+from itertools import cycle
+import pylab as P
+import numpy as np
 
+def plot(t,s,l):
+    P.figure()
+    s = np.vstack(s).T
+    co = cycle(['b', 'g', 'r', 'c', 'm', 'y', 'k'])
+    sy = cycle(['o', '^', '>', '<', 's', '*', '+', '1'])
+    print s.shape, len(t)
+    for i in range(s.shape[1]):
+        P.plot(t, s[:,i], co.next()+sy.next()+'-')
+    P.legend(l, loc=0)
+    
+if __name__=="__main__":
+# Loading the model
+    smod = stochpy.SSA()
+    smod.model_dir = os.getcwd()
+    smod.output_dir = os.getcwd()
+    smod.Model('Dengue_full.psc')
 
-smod = stochpy.SSA()
-smod.model_dir = os.getcwd()
-smod.output_dir = os.getcwd()
-smod.Model('Dengue_full.psc')
-
-#smod.DoStochSim(trajectories = 10,mode = 'time',end = 100, epsilon=0.03, method="TauLeaping", IsTrackPropensities=False)
-smod.DoCainStochSim(endtime=100,frames=10000,trajectories=10, solver="HomogeneousDirect2DSearch",IsTrackPropensities=False)
-smod.GetRegularGrid()
-smod.PlotAverageSpeciesTimeSeries()
-#smod.PlotAveragePropensitiesTimeSeries()
+    #smod.DoStochSim(trajectories = 1,mode = 'time',end = 100, epsilon=0.03, method="TauLeaping", IsTrackPropensities=False)
+    smod.DoCainStochSim(endtime=100,frames=10000,trajectories=100, solver="HomogeneousDirect2DSearch",IsTrackPropensities=False)
+    smod.GetRegularGrid()
+    smod.PlotAverageSpeciesTimeSeries()
+    #smod.PlotAveragePropensitiesTimeSeries()
+    #smod.PlotAverageSpeciesAutocorrelations()
+    
+#    t = smod.data_stochsim_grid.time
+#    series = smod.data_stochsim_grid.species_means
+#    sds = smod.data_stochsim_grid.species_standard_deviations
+#    l = smod.data_stochsim.species_labels
+#    plot(t,series,l)
+#    P.show()

### 2014-11-19 Modelo Estocastico/SDE_alt.py
minor changes

@@ -27,7 +27,7 @@ if __name__=="__main__":
     smod.Model('Dengue_full.psc')
 
     #smod.DoStochSim(trajectories = 1,mode = 'time',end = 100, epsilon=0.03, method="TauLeaping", IsTrackPropensities=False)
-    smod.DoCainStochSim(endtime=100,frames=10000,trajectories=100, solver="HomogeneousDirect2DSearch",IsTrackPropensities=False)
+    smod.DoCainStochSim(endtime=100,frames=10000,trajectories=10, solver="HomogeneousDirect2DSearch",IsTrackPropensities=False)
     smod.GetRegularGrid()
     smod.PlotAverageSpeciesTimeSeries()
     #smod.PlotAveragePropensitiesTimeSeries()
@@ -37,5 +37,5 @@ if __name__=="__main__":
 #    series = smod.data_stochsim_grid.species_means
 #    sds = smod.data_stochsim_grid.species_standard_deviations
 #    l = smod.data_stochsim.species_labels
-#    plot(t,series,l)
+#    plot(t,seriegs,l)
 #    P.show()

### 2014-11-20 draft_lm1.tex
Minor fixes to table
it's != its

@@
 In this section we will start by describing how we estimated the effective 
-reproductive number, $R_t$, relative to the global dengue cases, and it's 
+reproductive number, $R_t$, relative to the global dengue cases, and its 
 posterior probability distribution. 
@@
 \hline
-Year & Median Attack Ratio \\
+Year & Median Attack Ratio (\%) \\
 \hline
-1996 & \textbf{38.64\%} (17.18\%-54.19\%)\\
-1997 & \textbf{87.31\%} (73.83\%-87.31\%)\\
-1998 & \textbf{49.86\%} (49.19\%-49.86\%)\\
-1999 & \textbf{10.76\%} (3.72\%-20.24\%)\\
-2000 & \textbf{25.49\%} (24.47\%-26.98\%)\\
-2001 & \textbf{48.12\%} (46.57\%-49.15\%)\\
-2005 & \textbf{14.74\%} (10.23\%-21.38\%)\\
-2006 & \textbf{11.22\%} (8.00\%-14.41\%)\\
-2007 & \textbf{15.08\%} (14.76\%-15.38\%)\\
-2008 & \textbf{13.73\%} (3.10\%-30.84\%)\\
-2010 & \textbf{18.30\%} (17.34\%-19.33\%)\\
-2011 & \textbf{8.59\%} (8.20\%-9.36\%)\\
-2012 & \textbf{13.92\%} (12.71\%-14.92\%)\\
+1996 & \textbf{38.64} (17.18-54.19)\\
+1997 & \textbf{87.31} (73.83-87.31)\\
+1998 & \textbf{49.86} (49.19-49.86)\\
+1999 & \textbf{10.76} (3.72-20.24)\\
+2000 & \textbf{25.49} (24.47-26.98)\\
+2001 & \textbf{48.12} (46.57-49.15)\\
+2005 & \textbf{14.74} (10.23-21.38)\\
+2006 & \textbf{11.22} (8.00-14.41)\\
+2007 & \textbf{15.08} (14.76-15.38)\\
+2008 & \textbf{13.73} (3.10-30.84)\\
+2010 & \textbf{18.30} (17.34-19.33)\\
+2011 & \textbf{8.59} (8.20-9.36)\\
+2012 & \textbf{13.92} (12.71-14.92)\\
 \hline
 \end{tabular}
-\begin{flushleft}Medians are presented with 95\% credible intervals.
+\begin{flushleft} We report medians and 95\% credibility intervals.
 \end{flushleft}

### 2014-11-20 draft_lm1.tex
organising the plots --  draft
removing redundant term

@@
 Many authors have derived $c_{\alpha}(\theta_t)$ following orthodox approaches 
-such as those of \cite{wilson} and \cite{clopper}, mainly for simplicity 
-reasons.
+such as those of \cite{wilson} and \cite{clopper}, mainly for simplicity.
 We choose instead to take a Bayesian approach and use the  $100\alpha \%$ 
@@
 \begin{center}
-\includegraphics[width=.9\textwidth]{./concat_SI.png}
+\includegraphics[width=.9\textwidth]{plots/concat_SI.png}
 \end{center}

### 2014-11-25 Modelo Estocastico/SDE_alt4.py
added full stochastic dengue model including up to four dengue infections

@@ -25,7 +25,7 @@ if __name__=="__main__":
     smod = stochpy.SSA()
     smod.model_dir = os.getcwd()
     smod.output_dir = os.getcwd()
-    smod.Model('Dengue_full.psc')
+    smod.Model('Dengue_full4.psc')
     
 #    InterfaceCain.getCainInputfile(smod.SSA, 100, 100, 1)
 #    with open(os.path.join(stochpy.temp_dir,'cain_in.txt')) as f:

### 2014-11-25 Modelo Estocastico/SDE_alt4.py
fixed the full stochastic model

@@ -22,23 +22,25 @@ def plot(t,s,l):
     
 if __name__=="__main__":
 # Loading the model
-    smod = stochpy.SSA()
+    smod = stochpy.SSA(IsInteractive=False)
     smod.model_dir = os.getcwd()
     smod.output_dir = os.getcwd()
     smod.Model('Dengue_full4.psc')
+
     
 #    InterfaceCain.getCainInputfile(smod.SSA, 100, 100, 1)
 #    with open(os.path.join(stochpy.temp_dir,'cain_in.txt')) as f:
 #        with open('export.txt','w') as g:
 #            g.write(f.read())
 
-    smod.DoStochSim(trajectories = 1,mode = 'time',end = 500,  method="direct", IsTrackPropensities=False)
+    smod.DoStochSim(trajectories = 1,mode = 'time',end = 200,  method="direct", IsTrackPropensities=False)
     #smod.DoCainStochSim(endtime=100,frames=10000,trajectories=1, solver="HomogeneousDirect2DSearch",IsTrackPropensities=False)
     smod.GetRegularGrid()
     smod.PlotSpeciesTimeSeries()
+    stochpy.plt.savefig(os.path.join(smod.output_dir,'full4.png'), dpi=300)
     #smod.PlotAveragePropensitiesTimeSeries()
     #smod.PlotAverageSpeciesAutocorrelations()
-    
+    stochpy.plt.show()
 #    t = smod.data_stochsim_grid.time
 #    series = smod.data_stochsim_grid.species_means
 #    sds = smod.data_stochsim_grid.species_standard_deviations

### 2014-11-26 S1_lm1.tex
started writing about the stochastic multi-strain model in S1

@@
 \newpage
+\section*{Multi-strain dynamics}
+
+In order to validate the estimates derived in the paper, we simulated 
+multi-strain Dengue dynamics by means of a stochastic 4-serotype SIR model with 
+cross-immunity as described below. The model permits up to 4 dengue infections 
+with reduced susceptibility after the first Dengue episode due to cross 
+immunity. Immunity to each serotype is considered complete and permanent. 
+Figure xx depicted all possible states and state-transitions included in the 
+model.
+
+Let $S$ be individuals susceptible to all 4 types of dengue, $I_i$ infectious 
+with Dengue type $i$ and $R_i$ individuals recovered from Dengue type $i$. 
+Infectious individuals already on their secondary and later Dengue infections 
+are represented by multiple indices. For example, $I_{[23]1}$ is an individual 
+which has had Dengues type 2 and 3 in the past -- and therefore is immune to 
+them -- and is currently transmitting Dengue 1. The index outside the bracket 
+denotes current infection. Recovered individuals indices denote their immunity, 
+so for instance $R_{123}$ is an individual which is immune to Dengue types 1, 2 
+and 3, but not to 4. Let $I_{*i} = \sum I_{[\ldots]i}$ with $[\ldots]$ 
+representing exposure history of the infected individual which can vary from 0 
+to 3 in length. All individuals are born to the $S$ state and birth and death 
+rates are equal.
+
+The possible state-transitions and their  propensities are listed in table 
+\ref{tab:trans}.
+\begin{table}
+\caption{
+\bf{State-transitions and propensities}}
+\label{tab:trans}
+\begin{center}
+\begin{tabular}[c]{l|l|l}
+\hline
+Transition & Propensity & Description\\
+\hline
+$S \rightarrow I_i$ & $\beta S I_{*i}$ & Primary infection \\
+$I_i \rightarrow R_i$ & $\sigma I_i$ & Primary recovery\\
+$R_i \rightarrow I_{[i]j}$ & $\beta \delta R_i I_{*j}$ & Secondary infection\\
+$I_{[i]j} \rightarrow R_{ij}$ & $\sigma I_{[i]j}$ & Secondary recovery\\
+$R_{ij} \rightarrow I_{[ij]k}$ & $\beta \delta R_{ij} I_{*k}$ & Tertiary 
+infection\\
+$I_{[ij]k} \rightarrow R_{ijk}$ & $\sigma I_{[ij]k}$ & Tertiary recovery\\
+$R_{ijk} \rightarrow I_{[ijk]l}$ & $\beta \delta R_{ijk} I_{*l}$ & Quaternary 
+infection\\
+$I_{[ijk]l} \rightarrow R_{ijkl}$ & $\sigma I_{[ijk]l}$ & Quaternary recovery\\
+$\rightarrow S$ & $\mu N$ & Birth\\
+$All \rightarrow$ & $\mu N$ & Death\\
+\hline
+\end{tabular}
+\end{center}
+
+\end{table}
+
 \bibliography{lm1}

### 2014-11-27 S1_lm1.tex
continuing the description of the stochastic model

@@
 multi-strain Dengue dynamics by means of a stochastic 4-serotype SIR model with 
-cross-immunity as described below. The model permits up to 4 dengue infections 
+cross-immunity as described below. 
+The model permits up to 4 dengue infections 
 with reduced susceptibility after the first Dengue episode due to cross 
-immunity. Immunity to each serotype is considered complete and permanent. 
-Figure xx depicted all possible states and state-transitions included in the 
+immunity. 
+Immunity to each serotype is considered complete and permanent. 
+Figure xx depicts all possible states and state-transitions included in the 
 model.
+
+          \begin{figure}
+ \centering
+ \includegraphics[width=16cm]{Dengue4.png}
+
+ \caption{Block diagram detailing the stochastic model}
+ \label{fig:sde_blocks}
+\end{figure}
 
@@
 which has had Dengues type 2 and 3 in the past -- and therefore is immune to 
-them -- and is currently transmitting Dengue 1. The index outside the bracket 
+them -- and is currently transmitting Dengue 1. 
+The index outside the bracket 
 denotes current infection. Recovered individuals indices denote their immunity, 
@@
 representing exposure history of the infected individual which can vary from 0 
-to 3 in length. All individuals are born to the $S$ state and birth and death 
+to 3 in length. 
+All individuals are born to the $S$ state and birth and death 
 rates are equal.
@@
 \caption{
-\bf{State-transitions and propensities}}
+\bf{State-transitions and propensities: $P(\Delta X(t)|X(t))$}. The 
+transitions are summarized below, Fully expanded the system comtemplates 64 
+possible state transitions, as can be verified in figure \ref{fig:sde_blocks}. 
+}
 \label{tab:trans}
 \begin{center}
-\begin{tabular}[c]{l|l|l}
+\begin{tabular}[c]{l|l|l|l}
 \hline
-Transition & Propensity & Description\\
+Transition & Propensity & State Change & Description\\
 \hline
-$S \rightarrow I_i$ & $\beta S I_{*i}$ & Primary infection \\
-$I_i \rightarrow R_i$ & $\sigma I_i$ & Primary recovery\\
-$R_i \rightarrow I_{[i]j}$ & $\beta \delta R_i I_{*j}$ & Secondary infection\\
-$I_{[i]j} \rightarrow R_{ij}$ & $\sigma I_{[i]j}$ & Secondary recovery\\
+$S \rightarrow I_i$ & $\beta S I_{*i}$ & $\Delta S(t)=-1,\, \Delta I_i(t) = 1$ 
+& Primary infection \\
+$I_i \rightarrow R_i$ & $\sigma I_i$ & $\Delta I_i(t)=-1,\, \Delta R_i(t) = 1$ 
+& Primary recovery\\
+$R_i \rightarrow I_{[i]j}$ & $\beta \delta R_i I_{*j}$ & $$\Delta R_i(t)=-1,\, 
+\Delta I_{[i]j}(t) = 1$$ & Secondary infection\\
+$I_{[i]j} \rightarrow R_{ij}$ & $\sigma I_{[i]j}$ &$$\Delta I_{[i]j}(t)=-1,\, 
+\Delta R_{ij}(t) = 1$$& Secondary recovery\\
 $R_{ij} \rightarrow I_{[ij]k}$ & $\beta \delta R_{ij} I_{*k}$ & Tertiary 
@@
 
+The model is implemented as a continuous time Markov jump process. Let
+$$\overrightarrow{X}(t) = [S(t), I_1(t), I_2(t), \ldots, R_{1234}(t)]$$ 
+be the 
+state of the system at the time $t$. 
+The system is written as a forward Kolmogorov differential equation, which in 
+matrix form looks like
+\begin{equation}
+dP_X(t) = Q P_X(t) 
+\end{equation}
+
+Where $P_X(t)$ is  the matrix of transition probabilities (given in table 
+\ref{tab:trans}) and Q is the generator matrix, whose non-zero values are also 
+given in table \ref{tab:trans} (in the state change column). The full formula 
+and matrices are ommited due to their large sizes.
+
 \bibliography{lm1}

### 2014-12-02 S1_lm1.tex
fixed many subtle bugs in the stochastic model

@@
 
- \caption{Block diagram detailing the stochastic model}
+ \caption{Block diagram detailing the stochastic model. Infected individuals 
+with different Dengue viruses are represented by different colors. Infections 
+are also represented by colored arrows matching the virus type.}
  \label{fig:sde_blocks}
@@
 \bf{State-transitions and propensities: $P(\Delta X(t)|X(t))$}. The 
-transitions are summarized below, Fully expanded the system comtemplates 64 
+transitions are summarized below. Fully expanded, the system contemplates 64 
 possible state transitions, as can be verified in figure \ref{fig:sde_blocks}. 
@@
 & Primary recovery\\
-$R_i \rightarrow I_{[i]j}$ & $\beta \delta R_i I_{*j}$ & $$\Delta R_i(t)=-1,\, 
-\Delta I_{[i]j}(t) = 1$$ & Secondary infection\\
-$I_{[i]j} \rightarrow R_{ij}$ & $\sigma I_{[i]j}$ &$$\Delta I_{[i]j}(t)=-1,\, 
-\Delta R_{ij}(t) = 1$$& Secondary recovery\\
-$R_{ij} \rightarrow I_{[ij]k}$ & $\beta \delta R_{ij} I_{*k}$ & Tertiary 
-infection\\
-$I_{[ij]k} \rightarrow R_{ijk}$ & $\sigma I_{[ij]k}$ & Tertiary recovery\\
-$R_{ijk} \rightarrow I_{[ijk]l}$ & $\beta \delta R_{ijk} I_{*l}$ & Quaternary 
-infection\\
-$I_{[ijk]l} \rightarrow R_{ijkl}$ & $\sigma I_{[ijk]l}$ & Quaternary recovery\\
-$\rightarrow S$ & $\mu N$ & Birth\\
-$All \rightarrow$ & $\mu N$ & Death\\
+$R_i \rightarrow I_{[i]j}$ & $\beta \delta R_i I_{*j}$ & $\Delta R_i(t)=-1,\, 
+\Delta I_{[i]j}(t) = 1$ & Secondary infection\\
+$I_{[i]j} \rightarrow R_{ij}$ & $\sigma I_{[i]j}$ &$\Delta I_{[i]j}(t)=-1,\, 
+\Delta R_{ij}(t) = 1$& Secondary recovery\\
+$R_{ij} \rightarrow I_{[ij]k}$ & $\beta \delta R_{ij} I_{*k}$ &$\Delta 
+R_{ij}(t)=-1,\, \Delta I_{[ij]k}(t) = 1$& Tertiary infection\\
+$I_{[ij]k} \rightarrow R_{ijk}$ & $\sigma I_{[ij]k}$ &$\Delta 
+I_{[ij]k}(t)=-1,\, \Delta R_{ijk}(t) = 1$& Tertiary recovery\\
+$R_{ijk} \rightarrow I_{[ijk]l}$ & $\beta \delta R_{ijk} I_{*l}$ &$\Delta 
+R_{ijk}(t)=-1,\, \Delta I_{[ijk]l}(t) = 1$& Quaternary infection\\
+$I_{[ijk]l} \rightarrow R_{ijkl}$ & $\sigma I_{[ijk]l}$ & $\Delta 
+I_{[ijk]l}(t)=-1,\, \Delta R_{ijkl}(t) = 1$ & Quaternary recovery\\
+$\rightarrow S$ & $\mu N$ & $\Delta S = 1$ & Birth\\
+$All \rightarrow$ & $\mu N$ &$\Delta All = -1$& Death\\
 \hline
@@
 \begin{equation}
-dP_X(t) = Q P_X(t) 
+\frac{dP_X(t)}{dt} = Q P_X(t) 
 \end{equation}

### 2014-12-02 Modelo Estocastico/SDE_alt4.py

@@ -1,5 +1,4 @@
 #!/usr/bin/env python
-
 """
 Version of the dengue 4 serotype model implemented for Stochpy
 """
@@ -10,40 +9,92 @@ import pylab as P
 import numpy as np
 from stochpy.modules import InterfaceCain
 
-def plot(t,s,l):
+
+def plot(t, s, l):
     P.figure()
     s = np.vstack(s).T
     co = cycle(['b', 'g', 'r', 'c', 'm', 'y', 'k'])
     sy = cycle(['o', '^', '>', '<', 's', '*', '+', '1'])
     print s.shape, len(t)
     for i in range(s.shape[1]):
-        P.plot(t, s[:,i], co.next()+sy.next()+'-')
+        P.plot(t, s[:, i], co.next() + sy.next() + '-')
     P.legend(l, loc=0)
-    
-if __name__=="__main__":
-# Loading the model
+
+
+def plot_4_types(t, s, l):
+    """
+    Plot all infectives for each serotype
+    """
+    print l
+    P.figure()
+    s = np.vstack(s).T
+    # type 1
+    i1 = [l.index(i) for i in l if i.endswith('1') and i.startswith('I')]
+    print i1
+    s1 = np.zeros(s.shape[0])
+    for i in i1:
+        s1 += s[:, i]
+    P.plot(t, s1, label='1')
+    # Type 2
+    i2 = [l.index(i) for i in l if i.endswith('2') and i.startswith('I')]
+    s2 = np.zeros(s.shape[0])
+    for i in i2:
+        s2 += s[:, i]
+    P.plot(t, s2, label='2')
+    # type 3
+    i3 = [l.index(i) for i in l if i.endswith('3') and i.startswith('I')]
+    print "I3: ", i3
+    s3 = np.zeros(s.shape[0])
+    for i in i3:
+        s3 += s[:, i]
+    P.plot(t, s3, label='3')
+    # Type 4
+    i4 = [l.index(i) for i in l if i.endswith('4') and i.startswith('I')]
+    s4 = np.zeros(s.shape[0])
+    print "I4: ", i4
+    for i in i4:
+        s4 += s[:, i]
+    P.plot(t, s4, label='4')
+    P.legend(loc=0)
+    P.xlabel('time')
+    P.ylabel('individuals')
+    P.savefig('4types.png', dpi=300)
+
+
+if __name__ == "__main__":
+    # Loading the model
[…]

### 2014-12-02 Modelo Estocastico/SDE_alt4.py
Yet another minor bug fix

@@ -20,45 +20,64 @@ def plot(t, s, l):
         P.plot(t, s[:, i], co.next() + sy.next() + '-')
     P.legend(l, loc=0)
 
-
-def plot_4_types(t, s, l):
+def agg_by_type(s, l):
     """
-    Plot all infectives for each serotype
+    return series aggregated by serotype
+    :param s: series
+    :param l: labels
     """
-    print l
-    P.figure()
     s = np.vstack(s).T
-    # type 1
     i1 = [l.index(i) for i in l if i.endswith('1') and i.startswith('I')]
-    print i1
     s1 = np.zeros(s.shape[0])
     for i in i1:
         s1 += s[:, i]
-    P.plot(t, s1, label='1')
-    # Type 2
+
     i2 = [l.index(i) for i in l if i.endswith('2') and i.startswith('I')]
     s2 = np.zeros(s.shape[0])
     for i in i2:
         s2 += s[:, i]
-    P.plot(t, s2, label='2')
-    # type 3
+
     i3 = [l.index(i) for i in l if i.endswith('3') and i.startswith('I')]
-    print "I3: ", i3
+    # print "I3: ", i3
     s3 = np.zeros(s.shape[0])
     for i in i3:
         s3 += s[:, i]
-    P.plot(t, s3, label='3')
-    # Type 4
+
     i4 = [l.index(i) for i in l if i.endswith('4') and i.startswith('I')]
     s4 = np.zeros(s.shape[0])
-    print "I4: ", i4
+    # print "I4: ", i4
     for i in i4:
         s4 += s[:, i]
+
+    return s1, s2, s3, s4
+
+
+def plot_4_types(t, s, l):
+    """
+    Plot all infectives for each serotype
+    :param t: Times
+    :param s: series
+    :param l: labels
+    """
+    # print l
+    P.figure()
+
+    s1, s2, s3, s4 = agg_by_type(s, l)
+
+    P.plot(t, s1, label='1')
+
+    P.plot(t, s2, label='2')
+
+    P.plot(t, s3, label='3')
+
     P.plot(t, s4, label='4')
     P.legend(loc=0)
     P.xlabel('time')
     P.ylabel('individuals')
     P.savefig('4types.png', dpi=300)
+
+def plot_xcorr(s1, s2):
+    P.figure("Cross Correlation")
 
 
 if __name__ == "__main__":
@@ -74,7 +93,7 @@ if __name__ == "__main__":
     #        with open('export.txt','w') as g:
[…]

### 2014-12-05 draft_lm1.tex
minor fixes in the bibtex

@@
 not being exposed to the disease. But, as we can see in figure xx\todo{add 
-figure with the four maps}, the disease spread pretty much homogeneously over 
+figure with the four maps}, the disease spreads pretty much homogeneously over 
 the entire city (in the last four epidemics, at least).

### 2014-12-05 fit_dengue_split_big_plot.py

@@ -158,6 +158,13 @@ def series(nam='Dengue_S0'):
         # ax2.xaxis.set_major_formatter(FormatStrFormatter('%g'))
         #print P.setp(ax2)
 
+def plot_rt_beta():
+    df = pd.read_csv('data_Rt_dengue_complete.csv', header=0, delimiter=',', skiprows=[1, 2, 3], parse_dates=True)
+    df.Rt.plot()
+    df.Rt2.plot()
+    df.lwr.plot()
+    df.upr.plot()
+
 def plot_concat_series(dbs):
     """
     Plot concatenated time series for susceptibles and infectious
@@ -216,8 +223,10 @@ if __name__ == "__main__":
     #~ series('Dengue_S0_big')
 
     dbs = glob.glob("DengueS*.sqlite")
-    plot_concat_series(dbs)
+    #plot_concat_series(dbs)
     # print create_tex_table(dbs)
+    plot_rt_beta()
+
 
 
     # sns.tsplot(series.S, time=series.time)

### 2014-12-06 draft_lm1.tex
fixing stuff; beginning methods

--- 2014-11-20 draft_lm1.tex
@@
 % Our defs 
-\def \rr {$R_{t}\ $}
+\def \rr {$R_{t}\:$}
 
@@
 vector reproduction, stands out as a strong predictor of 
-incidence\cite{honorio_temporal_2009,wu_higher_2009}.
+incidence~\cite{honorio_temporal_2009,wu_higher_2009}.
 In places with sufficient temperature variation dengue is predominantly a 
@@
 The Attack ratio of disease is a mesure of mobidity defined as the number of new
-cases divided byt the population at risk. For Dengue outbreaks and epidemics 
+cases divided byt the population at risk.
+For Dengue outbreaks and epidemics 
 this can be difficult to calculated due to the lack of information about the 
@@
 timeseries, which represent all the cases of Dengue reported by health 
-professionals with no discrimination of virus types. Only a very small sample 
+professionals with no discrimination of virus types.
+Only a very small sample 
 of the cases are ever examined to determined the virus type.
@@
 In order to calculate the attack ratio, we need to determine the number of 
-susceptibles to the circulating virus types right before the epidemic. Without 
+susceptibles to the circulating virus types right before the epidemic.
+Without 
 regular virological surveys, it is virtually impossible to determine the 
@@
 The attack ratio is also influenced by the reproductive number of the disease 
-\cite{bacaer_final_2009, katriel_attack_2012}. With Dengue, since transmission 
+~\cite{bacaer_final_2009, katriel_attack_2012}.
+With Dengue, since transmission 
 is affected by environmental conditions as discussed above, the reproductive 
@@
 
-Various challenges exist to assess the attack rate of a dengue epidemic. One of 
+Various challenges exist to assess the attack rate of a dengue epidemic. 
+One of 
 them is that there are a large number of asymptomatic cases in any epidemic, 
-which nevertheless acquire immunity. It is estimated that for 
+which nevertheless acquire immunity.
+It is estimated that for 
 every case reported, 10-20 are not seen by health authorities (personnal 
@@
 Thirdly, under-reporting is a serious issue in Brazil. 
-Duarte and Fran\c{c}a\cite{duarte_data_2006}, estimated 
+Duarte and Fran\c{c}a (2006)~\cite{duarte_data_2006}, estimated 
 the sensitivity of dengue reporting for hospitalized patients in 
@@
 In this paper, we estimate the number of susceptibles before every major 
-outbreak ($S_{0, year}$) in the last 18 years. Due to the alternation of 
+outbreak ($S_{0, year}$) in the last 18 years.
+Due to the alternation of 
 ciculating serotypes, we detected fluctuating levels of susceptibility the 
-circulating virus(es). To Estimate the susceptibles we have we have fixed the 
-tramissibility based on the $R_t$ series derived from data\cite{nishiura}. Then, 
+circulating virus(es).
+To estimate the susceptibles we have we have fixed the 
+tramissibility based on the $R_t$ series derived from data~\cite{nishiura}. 
+Then, 
 from the incidence series and the population at risk, we could then calculate 
@@
 To accomplish the estimation of the $S_0$, we use a simplified model of 
-dengue transmission based on a single-strain Susceptible-infecctious-Removed 
-(SIR) model. The reason for the choice of a single strain model, is the lack of 
+dengue transmission based on a single-strain Susceptible-Infecctious-Removed 
+(SIR) model.
+The reason for the choice of a single strain model, is the lack of 
 accurate information regarding the actual proportions of each virus in 
-circulation. However, some information is available about the predominant 
+circulation.
+However, some information is available about the predominant 
 serotypes for some epidemics in the period 
-of study\cite{macedo_virological_2013}.
+of study~\cite{macedo_virological_2013}.
 
 Methods for estimating the  number of susceptibles have been proposed 
-before, for other diseases\cite{bjornstad_dynamics_2002, 
-wallinga_reconstruction_2003}. These methods try to reconstruct the entire 
+before, for other diseases~\cite{bjornstad_dynamics_2002, 
+wallinga_reconstruction_2003}.
+These methods try to reconstruct the entire 
 series of infectious and susceptibles for measles 
-outbreaks from case data.\todo[inline]{anyone else?} In the case of Dengue the 
+outbreaks from case data.\todo[inline]{anyone else?}
+In the case of Dengue the 
 series of susceptibles to all possible serotypes, cannot be reconstructed based 
@@
 of susceptibles for the predominant strain of influenza at the beginning of  
-influenza seasons in europe\cite{pone2011}.
+influenza seasons in europe~\cite{pone2011}.
 
@@
 
-In this section we will start by describing how we estimated the effective 
-reproductive number, $R_t$, relative to the global dengue cases, and its 
-posterior probability distribution. 
-We then proceed to describe the SIR model 
-used to represent the global incidence and how the $R_t$ is integrated into the 
-model to allow for time varying transmission. 
-Then we describe the Bayesian 
-estimation methodology applied to find the posterior probability distributions 
-of the susceptibles to the main circulating Dengue viruses for each epidemic, 
-using Dengue incidence time series for Rio de Janeiro from 1996 to 2014.
-In the last subsection we describe how we estimated the Attack ratio of each 
-epidemic from the estimated susceptibility and observed incidence.
+In this section we will start by describing a method to estimate the 
+effective 
+reproductive number, $R_t$, and obtain its 
+posterior distribution. 
+We then proceed to describe Susceptible-infectious-recovered (SIR) model 
+used to represent the global disease incidence and how $R_t$ can be integrated 
+into the model to allow for time varying transmission. 
+Next, an approach to approximate the posterior distributions 
+of the numbers of susceptible to the main circulating Dengue viruses for each 
+epidemic is detailed.
+Finally, we discuss how to estimate the attack ratio of each 
+epidemic using the estimated susceptible fraction and observed incidence.
+We employ these methods to analyze Dengue incidence time series for the city of 
+Rio de Janeiro from 1996 to 2014.
 
-\subsection*{Estimating the effective reproductive number, $R_t$}
+\subsection*{Estimating the effective reproductive number ($R_t$)}
 
@@
 the average number of secondary cases per primary case at time $t$.
-Here we derive confidence/credibility bounds for the uncertainty about the 
-estimate of the effective reproductive number $R_t$ under the assumption that 
-reported disease counts are Poisson distributed.
-It is shown that one can construct a conservative confidence interval for \rr 
-using the induced probability distribution that results from the approach of 
-\cite{mantel}.
 
@@
 (0, T)$.
-\cite{nishiura} extend the theory developed by \cite{stallybrass} and propose 
-to estimate \rr as
+Nishiura el al. (2010)~\cite{nishiura} extend the theory developed by 
+Stallybrass et al. (1931)~\cite{stallybrass} and propose to estimate \rr as
 \begin{equation}
@@
 Here we are interested in the simpler case $n=1$.
-If \rr is to be used as a decision tool, one needs to be able to quantify the 
+If \rr is to be used as a decision tool, howeve, one needs to be able to 
+quantify the 
 uncertainty about estimate in equation~\ref{eq:Rtestimate}. 
-Here we detail how to obtain confidence intervals for \rr, exploring two 
-approaches under the assumption that the counts $Y_t$ are Poisson distributed 
-for all $t$.
+Here we detail how to obtain credibility intervals for \rr under the assumption 
+that the counts $Y_t$ are Poisson distributed for all $t$.
 
-First we explore the approach of \cite{mantel}, whose objective is to obtain 
+We explore the approach of Ederer and Mantel~\cite{mantel}, whose objective is 
+to obtain 
 confidence intervals for the ratio of two Poisson counts. 

### 2014-12-06 draft_lm1.tex
edited R_t estimation to improve clarity; hopefully closes #8

@@
 The authors have applied similar methods as used here, to estimate the number 
-of susceptibles for the predominant strain of influenza at the beginning of  
-influenza seasons in europe~\cite{pone2011}.
+of susceptibles for the predominant strain of Influenza at the beginning of  
+Influenza seasons in europe~\cite{pone2011}.
 
@@
 epidemic using the estimated susceptible fraction and observed incidence.
-We employ these methods to analyze Dengue incidence time series for the city of 
-Rio de Janeiro from 1996 to 2014.
+We employ these methods to analyze Dengue weekly incidence time series for the 
+city of Rio de Janeiro from 1996 to 2014.
 
@@
 Here we are interested in the simpler case $n=1$.
-If \rr is to be used as a decision tool, howeve, one needs to be able to 
+If \rr is to be used as a decision tool, however, one needs to be able to 
 quantify the 
@@
 and define $S = Y_{t} + Y_{t+1}$.
-The authors then proceed to note that by conditioning on the sum $S$
+The authors note that by conditioning on the sum $S$
 \begin{align}
@@
 $Pr(R_t^{(L)}<R_t<R_t^{(U)}) = \alpha$.
-\cite{mantel} show that one can construct a $100\alpha \%$ confidence interval 
-for \rr by noting that
+Ederer and Mantel (1974)~\cite{mantel} show that one can construct a $100\alpha 
+\%$ confidence interval for \rr by noting that
 \begin{align}
- R_t^{(L)} &= \frac{\theta_t^{(L)}}{(1-\theta_t^{(L)})}\\
+\label{eq:confRt}
+ R_t^{(L)} &= \frac{\theta_t^{(L)}}{(1-\theta_t^{(L)})}\nonumber \\
  R_t^{(U)} &= \frac{\theta_t^{(U)}}{(1-\theta_t^{(U)})}
 \end{align}
-
 Many authors have derived $c_{\alpha}(\theta_t)$ following orthodox approaches 
@@
 \end{equation}
-Equation~(\ref{eq:thetapost}) tells us that the induced distribution of $R_t$ is 
-a beta prime with parameters $ a_1 = Y_{t+1} + a_0$ and $b_1 =  Y_t + b_0$.
-The density of this induced distribution  $f_P(R_t)$ is then 
+Combining equations~(\ref{eq:confRt}) and~(\ref{eq:thetapost}) 
+tells us that the induced posterior distribution of $R_t$ is 
+a beta prime (or inverted beta) with parameters $ a_1 = Y_{t+1} + a_0$ and $b_1 
+=  Y_t + b_0$.
+The density of the induced distribution is then 
 \begin{equation}
@@
 \end{equation}
-Thus, the expectation of \rr under $f_P$ is $a_1/(b_1 - 1)$ and its variance is 
+Thus, the expectation of \rr is $a_1/(b_1 - 1)$ and its variance is 
 $a_1(a_1 + b_1 - 1)/\left((b_1 - 2)(b_1 - 1)^2 \right) $.
+Sampling from the posterior in~(\ref{eq:densityMantel}) can be made 
+straightforward by first sampling from~(\ref{eq:thetapost}) and then applying 
+the transform in~(\ref{eq:confRt}).
 
-Since when $R_t > 1$ indicates sustained transmission, one may be interested in 
-computing the probability of this event.
+Also, since $R_t > 1$ indicates sustained transmission, one may be 
+interested in computing the probability of this event.
+This can be easily achieved by integrating~(\ref{eq:densityMantel}) over the 
+appropriate interval.
 By noting that
@@
 \end{align}
-one can avoid dealing with the density in~(\ref{eq:densityMantel}) directly.
+one can compute the desired probability while avoiding dealing with the density 
+in~(\ref{eq:densityMantel}) directly.
 
-The data used to fit the model consists of weekly incidence from 1996 to 2014 
-in the city of Rio de Janeiro, Brazil.
+\subsection*{Mathematical modeling} % The model
 
-\subsection*{The model}
-
-A Susceptible-Infectious-Removed (SIR) model is proposed to model dengue dynamics.
+A Susceptible-Infectious-Removed (SIR) model is proposed to model dengue 
+dynamics.
 In the traditional formulation of the model, transmission is governed by a 
@@
 As the epidemic progresses, the effective transmission  rate changes and is 
+given by 
 \begin{equation} 

### 2014-12-06 draft_lm1.tex
'guaribating' ;0) the Methods

@@
 in~(\ref{eq:densityMantel}) directly.
-
+s
 \subsection*{Mathematical modeling} % The model
@@
 where $S + I + R = 1 \: \forall\: t$. % \in [T_0, T_1]$. 
-This is a rather simplified model, in which, for instance, the vector is left 
-out.
+Of course, this is a rather simplified model, in which, for instance, the 
+vector is omitted.
 The rationale for this simplification is based on the ability of the 
@@
 Also, although there are multiple circulating serotypes, our approach
-does not discriminate between them due to the lack of serotype-specific data.
+can not discriminate between them due to the lack of serotype-specific data.
+Nevertheless, this modeling strategy can still provide some insight into the 
+disease dynamics and allows us to estimate the initial fraction of susceptibles 
+$S_0$.
 
 \subsection*{Bayesian parameter estimation}
-
-For the fitting procedure the incidence time series was divided into $J=13$ 
-epidemic 
-windows that corresponded to significant raises in incidence and normalized to 
-lie on the $[0,1]$ interval.
-Also, as the numerical integration routine needs to treat $\beta(t)$ as a
-continuous function of time because of the variable step size, we used linear 
-interpolation to obtain values of $R_t$ for any time point.
-
-We take a Bayesian approach to the estimation of $S_0$ for each window.
+We take a Bayesian approach to the estimation of $S_0$.
+First the incidence time series was divided into $J=13$ 
+epidemic windows that corresponded to significant raises in incidence and 
+normalized to lie on the $[0,1]$ interval.
 For a given interval $j = \{ t_j^{\text{start}}, t_j^{\text{end}} \} $ we 
@@
 In this estimation procedure we kept $R_t$ fixed at fixed at the posterior mean 
-obtained as described in TextS1 and fixed $\tau = 1/7\: \text{days}^{-1}$.
-% \subsubsection*{Priors}
+obtained as described above and fixed $\tau = 1/7\: \text{days}^{-1}$.
 To complete the inference, we need to specify prior distributions for the 
@@
 We place a flat $\text{Beta}(1, 1)$ prior on $S_{0j}\:\forall j$
-% 
-% \begin{align}
-%  S_{0j} &\sim Beta(1, 1) \\
-%        m &\sim \\
-% \end{align}
-
-% \subsubsection*{Posterior approximation}
 
@@
 multiple adaptive chains running in parallel with delayed rejection.
+Also, as the numerical integration routine implemented within BIP needs 
+$\beta(t)$ to be available at 
+arbitrary values of $t$, i.e., as continuous function of time because of the 
+variable step size, we used linear interpolation to obtain values of $R_t$ for 
+any time point.
 In this study we used one chain per parameter, i.e, 3 chains for each run.
@@
 \label{eq:AR}
-A=\frac{cases}{Pop.\, at\, risk} 
+A=\frac{\text{\# cases}}{\text{Population at risk}} 
 \end{equation}
 
-Based on the what have been discussed so far, we can rewrite (\ref{eq:AR}) for 
+Based on the what has been discussed so far, we can rewrite (\ref{eq:AR}) for 
 each epidemic $j$ as
@@
 \end{equation}
-where $S_{0j}$ is the number of susceptibles before each epidemic $j$, 
+where $S_{0j}$ is the number of susceptibles before each epidemic $j$, which we
 estimated as described above.
 
-% Results and Discussion can be combined.
 \section*{Results and Discussion}
@@
 In it we can see that the susceptibles series in each epidemic starts at the 
-estimated level of $S_0$. The proportion of susceptibles may seem low, but we 
+estimated level of $S_0$.
+The proportion of susceptibles may seem low, but we 
 must remember that these estimates are being affected by an unknown 
 underreporting factor, which experts suggest is somewhere between 5 and 10, 
-i.e. for every case observed there are 5 or 10 unobserved. Since this 
+i.e. for every case observed there are 5 or 10 unobserved.
+Since this 
 underreporting ration affects both the numerator and denominator of 
 (\ref{eq:AR}), its effects should cancel out, giving us an unbiased attack ratio 
-estimate. One other possible source of bias which would lead to the 
+estimate.
+One other possible source of bias which would lead to the 
 underestimation of $S_0$ could come from a significant part of the population 
-not being exposed to the disease. But, as we can see in figure xx\todo{add 
+not being exposed to the disease.
+But, as we can see in figure xx\todo{add 
 figure with the four maps}, the disease spreads pretty much homogeneously over 
@@
 LMC is grateful to Dr. Leonardo Bastos for useful discussions on the 
-posterior inference for $R(t)$. The authors are also grateful to Claudia T. 
+posterior inference for $R(t)$.
+The authors are also grateful to Claudia T. 
 Codeço for helpful discussions about the manuscript.

### 2014-12-06 Modelo Estocastico/SDE_alt.py
Organising a bit, hope you don't mind

@@ -1,47 +1,0 @@
-#!/usr/bin/env python
-
-"""
-Version of the dengue 4 serotype model implemented for Stochpy
-"""
-import stochpy
-import os
-from itertools import cycle
-import pylab as P
-import numpy as np
-from stochpy.modules import InterfaceCain
-
-def plot(t,s,l):
-    P.figure()
-    s = np.vstack(s).T
-    co = cycle(['b', 'g', 'r', 'c', 'm', 'y', 'k'])
-    sy = cycle(['o', '^', '>', '<', 's', '*', '+', '1'])
-    print s.shape, len(t)
-    for i in range(s.shape[1]):
-        P.plot(t, s[:,i], co.next()+sy.next()+'-')
-    P.legend(l, loc=0)
-    
-if __name__=="__main__":
-# Loading the model
-    smod = stochpy.SSA()
-    smod.model_dir = os.getcwd()
-    smod.output_dir = os.getcwd()
-    smod.Model('Dengue_full.psc')
-    
-#    InterfaceCain.getCainInputfile(smod.SSA, 100, 100, 1)
-#    with open(os.path.join(stochpy.temp_dir,'cain_in.txt')) as f:
-#        with open('export.txt','w') as g:
-#            g.write(f.read())
-
-    smod.DoStochSim(trajectories = 1,mode = 'time',end = 500,  method="direct", IsTrackPropensities=False)
-    #smod.DoCainStochSim(endtime=100,frames=10000,trajectories=1, solver="HomogeneousDirect2DSearch",IsTrackPropensities=False)
-    smod.GetRegularGrid()
-    smod.PlotSpeciesTimeSeries()
-    #smod.PlotAveragePropensitiesTimeSeries()
-    #smod.PlotAverageSpeciesAutocorrelations()
-    
-#    t = smod.data_stochsim_grid.time
-#    series = smod.data_stochsim_grid.species_means
-#    sds = smod.data_stochsim_grid.species_standard_deviations
-#    l = smod.data_stochsim.species_labels
-#    plot(t,seriegs,l)
-#    P.show()

### 2014-12-06 Modelo Estocastico/SDE_alt4.py

@@ -1,119 +1,0 @@
-#!/usr/bin/env python
-"""
-Version of the dengue 4 serotype model implemented for Stochpy
-"""
-import stochpy
-import os
-from itertools import cycle
-import pylab as P
-import numpy as np
-from stochpy.modules import InterfaceCain
-
-
-def plot(t, s, l):
-    P.figure()
-    s = np.vstack(s).T
-    co = cycle(['b', 'g', 'r', 'c', 'm', 'y', 'k'])
-    sy = cycle(['o', '^', '>', '<', 's', '*', '+', '1'])
-    print s.shape, len(t)
-    for i in range(s.shape[1]):
-        P.plot(t, s[:, i], co.next() + sy.next() + '-')
-    P.legend(l, loc=0)
-
-def agg_by_type(s, l):
-    """
-    return series aggregated by serotype
-    :param s: series
-    :param l: labels
-    """
-    s = np.vstack(s).T
-    i1 = [l.index(i) for i in l if i.endswith('1') and i.startswith('I')]
-    s1 = np.zeros(s.shape[0])
-    for i in i1:
-        s1 += s[:, i]
-
-    i2 = [l.index(i) for i in l if i.endswith('2') and i.startswith('I')]
-    s2 = np.zeros(s.shape[0])
-    for i in i2:
-        s2 += s[:, i]
-
-    i3 = [l.index(i) for i in l if i.endswith('3') and i.startswith('I')]
-    # print "I3: ", i3
-    s3 = np.zeros(s.shape[0])
-    for i in i3:
-        s3 += s[:, i]
-
-    i4 = [l.index(i) for i in l if i.endswith('4') and i.startswith('I')]
-    s4 = np.zeros(s.shape[0])
-    # print "I4: ", i4
-    for i in i4:
-        s4 += s[:, i]
-
-    return s1, s2, s3, s4
-
-
-def plot_4_types(t, s, l):
-    """
-    Plot all infectives for each serotype
-    :param t: Times
-    :param s: series
-    :param l: labels
-    """
-    # print l
-    P.figure()
-
-    s1, s2, s3, s4 = agg_by_type(s, l)
-
-    P.plot(t, s1, label='1')
-
-    P.plot(t, s2, label='2')
-
-    P.plot(t, s3, label='3')
-
-    P.plot(t, s4, label='4')
-    P.legend(loc=0)
-    P.xlabel('time')
-    P.ylabel('individuals')
[…]

### 2014-12-06 aux/Rt_aux.R

@@ -26,7 +26,7 @@ Rt.beta <- function(d, gt = 3, alpha = .95, a0 = 2 , b0 = 3){
     ac <- ac + d[(gt-i):(N-i)] 
   } 
   K <- length(ac)
-  jk1 <- ac[2:K] # J[k+1] 
+  jk1 <- ac[2:K] # J[k+1] s
   jk  <- ac[1:(K-1)]# J[k]  
   Rt <- Rt2 <- NA
   Rt[(gt + 1):N] <- jk1/jk
@@ -39,6 +39,12 @@ Rt.beta <- function(d, gt = 3, alpha = .95, a0 = 2 , b0 = 3){
                                   n = jk1[k] + jk[k], a = a0, b = b0 ))
   } 
   return(data.frame(Rt, Rt2, CIs))
+}
+###################################################
+rRtpost <- function(n, a = , b = ){ # let's sample from the distribution of R_t
+  x <- rbeta(n, shape1 = a, shape2 = b)
+  y <- x/(1-x) # yup! simple as that!
+return(y)
 }
 ###################################################
 Rt.gamma <- function(d, gt = 3, alpha = .95){

### 2014-12-06 fit_dengue_my.py

@@ -1,260 +1,0 @@
-#
-# Multi-year fit of influenza epidemics
-#
-import pyximport;
-
-pyximport.install(pyimport=False)
-from BIP.Bayes.Melding import FitModel
-from scipy.integrate import odeint
-from scipy.interpolate import interp1d
-import scipy.stats as st
-import numpy as np
-import pylab as P
-import copy
-from collections import defaultdict
-import datetime
-import sys
-import pandas as pd
-
-
-beta = 1.0  # Transmission coefficient
-b1 = 19.9
-b0 = 1.5  # low beta during winters
-eta = .0  # infectivity of asymptomatic infections relative to clinical ones. FIXED
-epsilon = 2.8  # latency rate
-mu = 0  # nat/mortality rate
-
-tau = 1  # recovery rate. FIXED
-
-N = 1  # Population of Rio
-s0 = 0.9999*N  # fraction of susceptibles at the beginning of first epidemic
-s1 = 0.999*N;
-s2 = 0.999*N;
-
-Ss = {0: s0, 1: s1, 2: s2}  # Multiple Ss map
-
-
-# Initial conditions
-inits = np.array([s0, 100, 0.0])  # initial values for state variables.
-
-def model(theta):
-    # setting parameters
-    s0, s1, s2 = theta
-    Ss = {0: s0, 1: s1, 2: s2}  # Multiple Ss map
-
-    # ~ b1=19.9
-    def sir(y, t):
-        '''ODE model'''
-        S, I, R = y
-        
-        beta = 0 if t > 242 else  iRt(t) * tau/ N#(Ss[ycode[int(t)]])
-
-        #~ if (calc_R0s()<1.1).any():
-        #~ print "flat"
-        lamb = (beta * I * S)
-        return [-lamb,  #dS/dt
-                lamb - tau * I,  #dI/dt
-                tau * I,  #dR/dt
-        ]
-
-    Y = np.zeros((wl, 3))
-    # Initial conditions
-
-    for i in range(len(Ss)):
-        t0 = t0s[i]
-        tf = tfs[i]
-        #print t0,tf
-        if i>0:
-            #~ inits[1] = Y[t0-1,1]
-            #~ print inits
-            inits[1] = dt['I'][t0-1] if N-Ss[i] > dt['I'][t0-1] else N-Ss[i]
-            #~ print inits
-        inits[0] = Ss[i];  # Define S0
-        inits[-1] = N - sum(inits[:2])  # Define R(0)
[…]

### 2014-12-06 fit_dengue_my_big.py

@@ -1,244 +1,0 @@
-#
-# Multi-year fit of influenza epidemics
-#
-import pyximport;
-
-pyximport.install(pyimport=False)
-from BIP.Bayes.Melding import FitModel
-from scipy.integrate import odeint
-from scipy.interpolate import interp1d
-import scipy.stats as st
-import numpy as np
-import pylab as P
-import copy
-from collections import defaultdict
-import datetime
-import sys
-import pandas as pd
-import numba
-from numba import jit
-
-
-beta = 1.0  # Transmission coefficient
-b1 = 19.9
-b0 = 1.5  # low beta during winters
-eta = .0  # infectivity of asymptomatic infections relative to clinical ones. FIXED
-epsilon = 2.8  # latency rate
-mu = 0  # nat/mortality rate
-m = 1e-6 # influx of cases
-tau = 1  # recovery rate. FIXED
-
-N = 1  # Population of Rio
-
-
-#~ Ss = {0: s0, 1: s1, 2: s2}  # Multiple Ss map
-
-
-# Initial conditions
-inits = np.array([1, 0.001, 0.0])  # initial values for state variables.
-
-
-def model(theta):
-    # setting parameters
-    s0, s1, s2, s3, s4, s5, s6, s7, s8, s9 = theta
-    Ss = {0: s0, 1: s1, 2: s2, 3: s3, 4: s4, 5: s5, 6: s6, 7: s7, 8: s8, 9: s9}  # Multiple Ss map
-
-    @jit('f8[:](f8[:],f8)')
-    def sir(y, t):
-        '''ODE model'''
-        S, I, R = y
-        
-        beta = 0 if t > 728 else iRt(t) * tau#(Ss[ycode[int(t)]])
-
-        #~ if (calc_R0s()<1.1).any():
-        #~ print "flat"
-        lamb = (beta * (I+m) * S)
-        
-        return np.array([-lamb,  #dS/dt
-                lamb - tau * I,  #dI/dt
-                tau * I,  #dR/dt
-        ])
-    @jit('f8[:,:](f8[:],f8)')
-    def jac(y,t):
-        S, I, R = y
-        beta = 0 if t > 728 else iRt(t) * tau
-        return np.array([[-(I + m)*beta, -S*beta, 0],
-               [(I + m)*beta, S*beta - tau, 0],
-               [0, tau, 0]])
-               
-    Y = np.zeros((wl, 3))
-    # Initial conditions
-
[…]

### 2014-12-06 fit_dengue_my_big_plot.py

@@ -1,229 +1,0 @@
-""" 
-Script to generate the plots for the
-multi-year simulations
-"""
-import cPickle as CP
-from itertools import cycle
-from scipy import stats
-import glob
-from numpy import *
-from matplotlib import font_manager
-from matplotlib.ticker import FormatStrFormatter
-import matplotlib.dates as mdates
-import matplotlib.pyplot as P
-import Gnuplot
-from scipy.stats import gaussian_kde
-import datetime
-import matplotlib.collections as collections
-import seaborn as sns
-import pandas as pd
-from sqlalchemy import create_engine
-
-from BIP.Bayes.PlotMeld import violin_plot, pred_new_cases, peakdet, plot_par_violin
-
-
-def _read_results(nam):
-    """
-    read results from disk
-    """
-    pt, series, predseries = [], [], []
-    nf = len(glob.glob('%s*.pickle' % nam))
-    for w in range(nf):
-        fn = "%s_%s.pickle" % (nam, w)
-        print fn
-        with open(fn, 'r') as f:
-            a, b, obs, pred, samples = CP.load(f)
-        # f.close()
-        pt.append(a)
-        series.append(b)
-        predseries.append(pred)
-
-    return pt, series, predseries, obs, nf
-
-def read_data(dbname):
-    """
-    Returns dataframes from tables of the sqlite database
-    :param dbname: sqlite3 database name
-    :return: list of pandas dataframes
-    """
-    eng = create_engine('sqlite:///{}.sqlite'.format(dbname))
-    df_data = pd.read_sql_table('data', eng, index_col=['pk', 'time'], parse_dates={'time': '%Y-%m-%d %H:%M:%S'})
-    df_pt = pd.read_sql_table('post_theta', eng, index_col=['pk', 'time'], parse_dates={'time': '%Y-%m-%d %H:%M:%S'})
-    df_series = pd.read_sql_table('series', eng, index_col=['pk', 'time'], parse_dates={'time': '%Y-%m-%d %H:%M:%S'})
-    return df_data, df_pt, df_series
-
-
-def create_tex_table(nms):
-    pts = {}
-    for nm in nms:
-        pt, series, predseries, obs, weeks = _read_results(nm)
-        pts[nm.split("-")[0]] = pt
[…]

### 2014-12-06 fit_dengue_my_plot.py

@@ -1,208 +1,0 @@
-""" 
-Script to generate the plots for the
-multi-year simulations
-"""
-import cPickle as CP
-from itertools import cycle
-from scipy import stats
-import glob
-from numpy import *
-from matplotlib import font_manager
-from matplotlib.ticker import FormatStrFormatter
-import matplotlib.dates as mdates
-import matplotlib.pyplot as P
-import Gnuplot
-from scipy.stats import gaussian_kde
-import datetime
-import matplotlib.collections as collections
-import seaborn as sns
-import pandas as pd
-from sqlalchemy import create_engine
-
-from BIP.Bayes.PlotMeld import violin_plot, pred_new_cases, peakdet, plot_par_violin
-
-
-def _read_results(nam):
-    """
-    read results from disk
-    """
-    pt, series, predseries = [], [], []
-    nf = len(glob.glob('%s*.pickle' % nam))
-    for w in range(nf):
-        fn = "%s_%s.pickle" % (nam, w)
-        print fn
-        with open(fn, 'r') as f:
-            a, b, obs, pred, samples = CP.load(f)
-        # f.close()
-        pt.append(a)
-        series.append(b)
-        predseries.append(pred)
-
-    return pt, series, predseries, obs, nf
-
-def read_data(dbname):
-    """
-    Returns dataframes from tables of the sqlite database
-    :param dbname: sqlite3 database name
-    :return: list of pandas dataframes
-    """
-    eng = create_engine('sqlite:///{}.sqlite'.format(dbname))
-    df_data = pd.read_sql_table('data', eng, index_col=['pk', 'time'], parse_dates={'time': '%Y-%m-%d %H:%M:%S'})
-    df_pt = pd.read_sql_table('post_theta', eng, index_col=['pk', 'time'], parse_dates={'time': '%Y-%m-%d %H:%M:%S'})
-    df_series = pd.read_sql_table('series', eng, index_col=['pk', 'time'], parse_dates={'time': '%Y-%m-%d %H:%M:%S'})
-    return df_data, df_pt, df_series
-
-
-def create_tex_table(nms):
-    pts = {}
-    for nm in nms:
-        pt, series, predseries, obs, weeks = _read_results(nm)
-        pts[nm.split("-")[0]] = pt
-    head = r"""\begin{center}
[…]

### 2014-12-06 fit_dengue_split_big.py

@@ -1,275 +1,0 @@
-#
-# Multi-year fit of influenza epidemics
-#
-import pyximport;
-
-pyximport.install(pyimport=False)
-from BIP.Bayes.Melding import FitModel
-from scipy.integrate import odeint
-from scipy.interpolate import interp1d
-import scipy.stats as st
-import numpy as np
-import pylab as P
-import copy
-from collections import defaultdict
-import datetime
-import sys
-import pandas as pd
-import seaborn as sns
-# import numba
-#from numba import jit
-
-
-beta = 1.0  # Transmission coefficient
-b1 = 19.9
-b0 = 1.5  # low beta during winters
-eta = .0  # infectivity of asymptomatic infections relative to clinical ones. FIXED
-
-mu = 0  # nat/mortality rate
-m = 1e-6  # influx of cases
-tau = 1  # recovery rate. FIXED
-
-N = 1  # Population of Rio
-
-inicio = 0
-#~ Ss = {0: s0, 1: s1, 2: s2}  # Multiple Ss map
-
-
-# Initial conditions
-inits = np.array([0.999, 0.001, 0.0])  # initial values for state variables.
-
-
-#@jit
-def sir(y, t, *pars):
-    '''ODE model'''
-    S, I, R = y
-    s0, m, tau = pars
-    beta = 0 if t > 728 else iRt(t) * tau / s0
-    # print S, I, beta
-
-    lamb = beta * (I + m) * S
-
-    return np.array([-lamb,  #dS/dt
-                     lamb - tau * I,  #dI/dt
-                     tau * I,  #dR/dt
-    ])
-
-
-#@jit
-def jac(y, t, *pars):
-    S, I, R = y
-    s0, m, tau = pars
-    beta = 0 if t > 728 else iRt(t) * tau / s0
-    return np.array([[-(I + m) * beta, -S * beta, 0],
-                     [(I + m) * beta, S * beta - tau, 0],
-                     [0, tau, 0]])
-
-
-#@jit
-def model(theta):
-    # setting parameters
-    s0, m, tau = theta
-
-    t0 = 0
-    tf = fim
-    Y = np.zeros((tf - t0, 3))
-
-
-    # Initial conditions
-    #print inits, m
-    inits[0] = s0  # Define S0
-    inits[-1] = N - sum(inits[:2])  # Define R(0)
-    Y[t0:tf, :] = odeint(sir, inits, np.arange(t0, tf, 1), args=(s0, m, tau), Dfun=jac)  #,tcrit=tcrit)
[…]

### 2014-12-06 fit_dengue_split_big_plot.py

@@ -1,236 +1,0 @@
-""" 
-Script to generate the plots for the
-multi-year simulations
-"""
-import cPickle as CP
-from itertools import cycle
-from scipy import stats
-import glob
-from numpy import *
-from matplotlib import font_manager
-from matplotlib.ticker import FormatStrFormatter
-import matplotlib.dates as mdates
-import matplotlib.pyplot as P
-import Gnuplot
-from scipy.stats import gaussian_kde
-import datetime
-import matplotlib.collections as collections
-import seaborn as sns
-import pandas as pd
-from sqlalchemy import create_engine
-from collections import OrderedDict
-
-from BIP.Bayes.PlotMeld import violin_plot, pred_new_cases, peakdet, plot_par_violin
-
-
-
-def read_data(dbname):
-    """
-    Returns dataframes from tables of the sqlite database
-    :param dbname: sqlite3 database name
-    :return: list of pandas dataframes
-    """
-    eng = create_engine('sqlite:///{}'.format(dbname))
-    df_data = pd.read_sql_table('data', eng, index_col=['time'], parse_dates={'time': '%Y-%m-%d %H:%M:%S'})
-    df_pt = pd.read_sql_table('post_theta', eng, index_col=['time'], parse_dates={'time': '%Y-%m-%d %H:%M:%S'})
-    df_series = pd.read_sql_table('series', eng, index_col=['time'], parse_dates={'time': '%Y-%m-%d %H:%M:%S'})
-    return df_data, df_pt, df_series
-
-
-def create_tex_table(dbs):
-    """
-    Create Latex table with the Attack ratios for each epidemic
-    :return:
-    """
-    pts = OrderedDict()
-    series = OrderedDict()
-    obs = OrderedDict()
-    for db in dbs:
-        y = db.split('_')[0][-4:]
-        data, theta, srs = read_data(db)
-        #print data
-        series[y] = srs
-        obs[y] = data.I
-        pts[y] = theta
-
-    head = r"""\begin{center}
-\begin{tabular}{c|c}
-\hline
-"""
-    head += r"""Year & median Attack Ratio \\
-\hline
-"""
-    bot = r"""
-\hline
-\end{tabular}
-\end{center}
-        """
-    body = r""
-    st = []
[…]

### 2014-12-06 sirmodel.pyx

@@ -1,65 +1,0 @@
-from cython_gsl cimport *
-
-
-cdef int func (double t,  double y[], double f[], void *params) nogil:
-    cdef double beta, lamb, alpha, tau
-    beta = (<double *> params)[0]
-    alpha = (<double *> params)[1]
-    tau = (<double *> params)[2]
-    lamb = beta * (alpha * f[1]) * f[0]
-    f[0] = -lamb
-    f[1] = lamb - tau * f[1],  #dI/dt
-    f[2] = tau * f[1]
-    return GSL_SUCCESS
-
-cdef int jac (double t,  double y[], double *dfdy, double dfdt[], void *params) nogil:
-    cdef double mu
-    mu = (<double *>params)[0]
-    cdef gsl_matrix_view dfdy_mat
-    dfdy_mat  = gsl_matrix_view_array (dfdy, 2, 2)
-    cdef gsl_matrix * m
-    m = &dfdy_mat.matrix
-    gsl_matrix_set (m, 0, 0, 0.0)
-    gsl_matrix_set (m, 0, 1, 1.0)
-    gsl_matrix_set (m, 1, 0, -2.0*mu*y[0]*y[1] - 1.0)
-    gsl_matrix_set (m, 1, 1, -mu*(y[0]*y[0] - 1.0))
-    dfdt[0] = 0.0
-    dfdt[1] = 0.0
-    return GSL_SUCCESS
-
-def main ( ):
-    cdef double mu
-    mu = 10
-    cdef gsl_odeiv2_system sys
-    sys.function = func
-    sys.jacobian = jac
-    sys.dimension = 2
-    sys.params = &mu
-
-    cdef gsl_odeiv2_driver * d
-    d = gsl_odeiv2_driver_alloc_y_new(
-        &sys, gsl_odeiv2_step_rk8pd,
-        1e-6, 1e-6, 0.0)
-
-    cdef int i
-    cdef double t, t1, y[2]
-    t = 0.0
-    t1 = 100.0
-    y[0] = 1.0
-    y[1] = 0.0
-
-    cdef int status
-    cdef double ti
-    for i from 1 <= i <= 100:
-        ti = i * t1 / 100.0
-        status = gsl_odeiv2_driver_apply (d, &t, ti, y)
-
-        if (status != GSL_SUCCESS):
-            print("error, return value=%d\n" % status)
-            break
-
-        print("%.5e %.5e %.5e\n" %(t, y[0], y[1]))
-
-    gsl_odeiv2_driver_free(d)
-
-

### 2014-12-06 python/fit_dengue_my.py
Organising a bit, hope you don't mind II

@@ -1,0 +1,260 @@
+#
+# Multi-year fit of influenza epidemics
+#
+import pyximport;
+
+pyximport.install(pyimport=False)
+from BIP.Bayes.Melding import FitModel
+from scipy.integrate import odeint
+from scipy.interpolate import interp1d
+import scipy.stats as st
+import numpy as np
+import pylab as P
+import copy
+from collections import defaultdict
+import datetime
+import sys
+import pandas as pd
+
+
+beta = 1.0  # Transmission coefficient
+b1 = 19.9
+b0 = 1.5  # low beta during winters
+eta = .0  # infectivity of asymptomatic infections relative to clinical ones. FIXED
+epsilon = 2.8  # latency rate
+mu = 0  # nat/mortality rate
+
+tau = 1  # recovery rate. FIXED
+
+N = 1  # Population of Rio
+s0 = 0.9999*N  # fraction of susceptibles at the beginning of first epidemic
+s1 = 0.999*N;
+s2 = 0.999*N;
+
+Ss = {0: s0, 1: s1, 2: s2}  # Multiple Ss map
+
+
+# Initial conditions
+inits = np.array([s0, 100, 0.0])  # initial values for state variables.
+
+def model(theta):
+    # setting parameters
+    s0, s1, s2 = theta
+    Ss = {0: s0, 1: s1, 2: s2}  # Multiple Ss map
+
+    # ~ b1=19.9
+    def sir(y, t):
+        '''ODE model'''
+        S, I, R = y
+        
+        beta = 0 if t > 242 else  iRt(t) * tau/ N#(Ss[ycode[int(t)]])
+
+        #~ if (calc_R0s()<1.1).any():
+        #~ print "flat"
+        lamb = (beta * I * S)
+        return [-lamb,  #dS/dt
+                lamb - tau * I,  #dI/dt
+                tau * I,  #dR/dt
+        ]
+
+    Y = np.zeros((wl, 3))
+    # Initial conditions
+
+    for i in range(len(Ss)):
+        t0 = t0s[i]
+        tf = tfs[i]
+        #print t0,tf
+        if i>0:
+            #~ inits[1] = Y[t0-1,1]
+            #~ print inits
+            inits[1] = dt['I'][t0-1] if N-Ss[i] > dt['I'][t0-1] else N-Ss[i]
+            #~ print inits
+        inits[0] = Ss[i];  # Define S0
+        inits[-1] = N - sum(inits[:2])  # Define R(0)
[…]

### 2014-12-06 python/fit_dengue_my_big.py

@@ -1,0 +1,244 @@
+#
+# Multi-year fit of influenza epidemics
+#
+import pyximport;
+
+pyximport.install(pyimport=False)
+from BIP.Bayes.Melding import FitModel
+from scipy.integrate import odeint
+from scipy.interpolate import interp1d
+import scipy.stats as st
+import numpy as np
+import pylab as P
+import copy
+from collections import defaultdict
+import datetime
+import sys
+import pandas as pd
+import numba
+from numba import jit
+
+
+beta = 1.0  # Transmission coefficient
+b1 = 19.9
+b0 = 1.5  # low beta during winters
+eta = .0  # infectivity of asymptomatic infections relative to clinical ones. FIXED
+epsilon = 2.8  # latency rate
+mu = 0  # nat/mortality rate
+m = 1e-6 # influx of cases
+tau = 1  # recovery rate. FIXED
+
+N = 1  # Population of Rio
+
+
+#~ Ss = {0: s0, 1: s1, 2: s2}  # Multiple Ss map
+
+
+# Initial conditions
+inits = np.array([1, 0.001, 0.0])  # initial values for state variables.
+
+
+def model(theta):
+    # setting parameters
+    s0, s1, s2, s3, s4, s5, s6, s7, s8, s9 = theta
+    Ss = {0: s0, 1: s1, 2: s2, 3: s3, 4: s4, 5: s5, 6: s6, 7: s7, 8: s8, 9: s9}  # Multiple Ss map
+
+    @jit('f8[:](f8[:],f8)')
+    def sir(y, t):
+        '''ODE model'''
+        S, I, R = y
+        
+        beta = 0 if t > 728 else iRt(t) * tau#(Ss[ycode[int(t)]])
+
+        #~ if (calc_R0s()<1.1).any():
+        #~ print "flat"
+        lamb = (beta * (I+m) * S)
+        
+        return np.array([-lamb,  #dS/dt
+                lamb - tau * I,  #dI/dt
+                tau * I,  #dR/dt
+        ])
+    @jit('f8[:,:](f8[:],f8)')
+    def jac(y,t):
+        S, I, R = y
+        beta = 0 if t > 728 else iRt(t) * tau
+        return np.array([[-(I + m)*beta, -S*beta, 0],
+               [(I + m)*beta, S*beta - tau, 0],
+               [0, tau, 0]])
+               
+    Y = np.zeros((wl, 3))
+    # Initial conditions
+
[…]

### 2014-12-06 python/fit_dengue_my_plot.py

@@ -1,0 +1,208 @@
+""" 
+Script to generate the plots for the
+multi-year simulations
+"""
+import cPickle as CP
+from itertools import cycle
+from scipy import stats
+import glob
+from numpy import *
+from matplotlib import font_manager
+from matplotlib.ticker import FormatStrFormatter
+import matplotlib.dates as mdates
+import matplotlib.pyplot as P
+import Gnuplot
+from scipy.stats import gaussian_kde
+import datetime
+import matplotlib.collections as collections
+import seaborn as sns
+import pandas as pd
+from sqlalchemy import create_engine
+
+from BIP.Bayes.PlotMeld import violin_plot, pred_new_cases, peakdet, plot_par_violin
+
+
+def _read_results(nam):
+    """
+    read results from disk
+    """
+    pt, series, predseries = [], [], []
+    nf = len(glob.glob('%s*.pickle' % nam))
+    for w in range(nf):
+        fn = "%s_%s.pickle" % (nam, w)
+        print fn
+        with open(fn, 'r') as f:
+            a, b, obs, pred, samples = CP.load(f)
+        # f.close()
+        pt.append(a)
+        series.append(b)
+        predseries.append(pred)
+
+    return pt, series, predseries, obs, nf
+
+def read_data(dbname):
+    """
+    Returns dataframes from tables of the sqlite database
+    :param dbname: sqlite3 database name
+    :return: list of pandas dataframes
+    """
+    eng = create_engine('sqlite:///{}.sqlite'.format(dbname))
+    df_data = pd.read_sql_table('data', eng, index_col=['pk', 'time'], parse_dates={'time': '%Y-%m-%d %H:%M:%S'})
+    df_pt = pd.read_sql_table('post_theta', eng, index_col=['pk', 'time'], parse_dates={'time': '%Y-%m-%d %H:%M:%S'})
+    df_series = pd.read_sql_table('series', eng, index_col=['pk', 'time'], parse_dates={'time': '%Y-%m-%d %H:%M:%S'})
+    return df_data, df_pt, df_series
+
+
+def create_tex_table(nms):
+    pts = {}
+    for nm in nms:
+        pt, series, predseries, obs, weeks = _read_results(nm)
+        pts[nm.split("-")[0]] = pt
+    head = r"""\begin{center}
[…]

### 2014-12-06 python/fit_dengue_split_big.py

@@ -1,0 +1,275 @@
+#
+# Multi-year fit of influenza epidemics
+#
+import pyximport;
+
+pyximport.install(pyimport=False)
+from BIP.Bayes.Melding import FitModel
+from scipy.integrate import odeint
+from scipy.interpolate import interp1d
+import scipy.stats as st
+import numpy as np
+import pylab as P
+import copy
+from collections import defaultdict
+import datetime
+import sys
+import pandas as pd
+import seaborn as sns
+# import numba
+#from numba import jit
+
+
+beta = 1.0  # Transmission coefficient
+b1 = 19.9
+b0 = 1.5  # low beta during winters
+eta = .0  # infectivity of asymptomatic infections relative to clinical ones. FIXED
+
+mu = 0  # nat/mortality rate
+m = 1e-6  # influx of cases
+tau = 1  # recovery rate. FIXED
+
+N = 1  # Population of Rio
+
+inicio = 0
+#~ Ss = {0: s0, 1: s1, 2: s2}  # Multiple Ss map
+
+
+# Initial conditions
+inits = np.array([0.999, 0.001, 0.0])  # initial values for state variables.
+
+
+#@jit
+def sir(y, t, *pars):
+    '''ODE model'''
+    S, I, R = y
+    s0, m, tau = pars
+    beta = 0 if t > 728 else iRt(t) * tau / s0
+    # print S, I, beta
+
+    lamb = beta * (I + m) * S
+
+    return np.array([-lamb,  #dS/dt
+                     lamb - tau * I,  #dI/dt
+                     tau * I,  #dR/dt
+    ])
+
+
+#@jit
+def jac(y, t, *pars):
+    S, I, R = y
+    s0, m, tau = pars
+    beta = 0 if t > 728 else iRt(t) * tau / s0
+    return np.array([[-(I + m) * beta, -S * beta, 0],
+                     [(I + m) * beta, S * beta - tau, 0],
+                     [0, tau, 0]])
+
+
+#@jit
+def model(theta):
+    # setting parameters
+    s0, m, tau = theta
+
+    t0 = 0
+    tf = fim
+    Y = np.zeros((tf - t0, 3))
+
+
+    # Initial conditions
+    #print inits, m
+    inits[0] = s0  # Define S0
+    inits[-1] = N - sum(inits[:2])  # Define R(0)
+    Y[t0:tf, :] = odeint(sir, inits, np.arange(t0, tf, 1), args=(s0, m, tau), Dfun=jac)  #,tcrit=tcrit)
[…]

### 2014-12-06 python/fit_dengue_split_big_plot.py

@@ -1,0 +1,236 @@
+""" 
+Script to generate the plots for the
+multi-year simulations
+"""
+import cPickle as CP
+from itertools import cycle
+from scipy import stats
+import glob
+from numpy import *
+from matplotlib import font_manager
+from matplotlib.ticker import FormatStrFormatter
+import matplotlib.dates as mdates
+import matplotlib.pyplot as P
+import Gnuplot
+from scipy.stats import gaussian_kde
+import datetime
+import matplotlib.collections as collections
+import seaborn as sns
+import pandas as pd
+from sqlalchemy import create_engine
+from collections import OrderedDict
+
+from BIP.Bayes.PlotMeld import violin_plot, pred_new_cases, peakdet, plot_par_violin
+
+
+
+def read_data(dbname):
+    """
+    Returns dataframes from tables of the sqlite database
+    :param dbname: sqlite3 database name
+    :return: list of pandas dataframes
+    """
+    eng = create_engine('sqlite:///{}'.format(dbname))
+    df_data = pd.read_sql_table('data', eng, index_col=['time'], parse_dates={'time': '%Y-%m-%d %H:%M:%S'})
+    df_pt = pd.read_sql_table('post_theta', eng, index_col=['time'], parse_dates={'time': '%Y-%m-%d %H:%M:%S'})
+    df_series = pd.read_sql_table('series', eng, index_col=['time'], parse_dates={'time': '%Y-%m-%d %H:%M:%S'})
+    return df_data, df_pt, df_series
+
+
+def create_tex_table(dbs):
+    """
+    Create Latex table with the Attack ratios for each epidemic
+    :return:
+    """
+    pts = OrderedDict()
+    series = OrderedDict()
+    obs = OrderedDict()
+    for db in dbs:
+        y = db.split('_')[0][-4:]
+        data, theta, srs = read_data(db)
+        #print data
+        series[y] = srs
+        obs[y] = data.I
+        pts[y] = theta
+
+    head = r"""\begin{center}
+\begin{tabular}{c|c}
+\hline
+"""
+    head += r"""Year & median Attack Ratio \\
+\hline
+"""
+    bot = r"""
+\hline
+\end{tabular}
+\end{center}
+        """
+    body = r""
+    st = []
[…]

### 2014-12-06 stochastic_model/SDE_alt.py

@@ -1,0 +1,47 @@
+#!/usr/bin/env python
+
+"""
+Version of the dengue 4 serotype model implemented for Stochpy
+"""
+import stochpy
+import os
+from itertools import cycle
+import pylab as P
+import numpy as np
+from stochpy.modules import InterfaceCain
+
+def plot(t,s,l):
+    P.figure()
+    s = np.vstack(s).T
+    co = cycle(['b', 'g', 'r', 'c', 'm', 'y', 'k'])
+    sy = cycle(['o', '^', '>', '<', 's', '*', '+', '1'])
+    print s.shape, len(t)
+    for i in range(s.shape[1]):
+        P.plot(t, s[:,i], co.next()+sy.next()+'-')
+    P.legend(l, loc=0)
+    
+if __name__=="__main__":
+# Loading the model
+    smod = stochpy.SSA()
+    smod.model_dir = os.getcwd()
+    smod.output_dir = os.getcwd()
+    smod.Model('Dengue_full.psc')
+    
+#    InterfaceCain.getCainInputfile(smod.SSA, 100, 100, 1)
+#    with open(os.path.join(stochpy.temp_dir,'cain_in.txt')) as f:
+#        with open('export.txt','w') as g:
+#            g.write(f.read())
+
+    smod.DoStochSim(trajectories = 1,mode = 'time',end = 500,  method="direct", IsTrackPropensities=False)
+    #smod.DoCainStochSim(endtime=100,frames=10000,trajectories=1, solver="HomogeneousDirect2DSearch",IsTrackPropensities=False)
+    smod.GetRegularGrid()
+    smod.PlotSpeciesTimeSeries()
+    #smod.PlotAveragePropensitiesTimeSeries()
+    #smod.PlotAverageSpeciesAutocorrelations()
+    
+#    t = smod.data_stochsim_grid.time
+#    series = smod.data_stochsim_grid.species_means
+#    sds = smod.data_stochsim_grid.species_standard_deviations
+#    l = smod.data_stochsim.species_labels
+#    plot(t,seriegs,l)
+#    P.show()

### 2014-12-06 stochastic_model/SDE_alt4.py

@@ -1,0 +1,119 @@
+#!/usr/bin/env python
+"""
+Version of the dengue 4 serotype model implemented for Stochpy
+"""
+import stochpy
+import os
+from itertools import cycle
+import pylab as P
+import numpy as np
+from stochpy.modules import InterfaceCain
+
+
+def plot(t, s, l):
+    P.figure()
+    s = np.vstack(s).T
+    co = cycle(['b', 'g', 'r', 'c', 'm', 'y', 'k'])
+    sy = cycle(['o', '^', '>', '<', 's', '*', '+', '1'])
+    print s.shape, len(t)
+    for i in range(s.shape[1]):
+        P.plot(t, s[:, i], co.next() + sy.next() + '-')
+    P.legend(l, loc=0)
+
+def agg_by_type(s, l):
+    """
+    return series aggregated by serotype
+    :param s: series
+    :param l: labels
+    """
+    s = np.vstack(s).T
+    i1 = [l.index(i) for i in l if i.endswith('1') and i.startswith('I')]
+    s1 = np.zeros(s.shape[0])
+    for i in i1:
+        s1 += s[:, i]
+
+    i2 = [l.index(i) for i in l if i.endswith('2') and i.startswith('I')]
+    s2 = np.zeros(s.shape[0])
+    for i in i2:
+        s2 += s[:, i]
+
+    i3 = [l.index(i) for i in l if i.endswith('3') and i.startswith('I')]
+    # print "I3: ", i3
+    s3 = np.zeros(s.shape[0])
+    for i in i3:
+        s3 += s[:, i]
+
+    i4 = [l.index(i) for i in l if i.endswith('4') and i.startswith('I')]
+    s4 = np.zeros(s.shape[0])
+    # print "I4: ", i4
+    for i in i4:
+        s4 += s[:, i]
+
+    return s1, s2, s3, s4
+
+
+def plot_4_types(t, s, l):
+    """
+    Plot all infectives for each serotype
+    :param t: Times
+    :param s: series
+    :param l: labels
+    """
+    # print l
+    P.figure()
+
+    s1, s2, s3, s4 = agg_by_type(s, l)
+
+    P.plot(t, s1, label='1')
+
+    P.plot(t, s2, label='2')
+
+    P.plot(t, s3, label='3')
+
+    P.plot(t, s4, label='4')
+    P.legend(loc=0)
+    P.xlabel('time')
+    P.ylabel('individuals')
[…]

### 2014-12-06 stochastic_model/dengue_full_SDE.py

@@ -1,0 +1,158 @@
+# coding:utf8
+from __future__ import division
+from cgillespie import Model
+import time
+from numpy import array, genfromtxt
+import pylab as P
+import numpy as np
+
+from itertools import cycle
+
+vnames = ['S', 'I1', 'I2', 'I3', 'I4', 'R1', 'R2', 'R3', 'R4',
+          'I12', 'I13', 'I14', 'I21', 'I23', 'I24', 'I31', 'I32',
+            'I34', 'I41', 'I42', 'I43', 'R']
+
+# transitions
+# S->I1	S->I2	S->I3	S->I4	I1=>R1	I2=>R2	I3=>R3	I4=>R4	R1=>I12	R1=>I13	R1=>I14	R2=>I21	R2=>I23	R2=>I24	R3=>I31	R3=>I32	R3=>I34	R4=>I41	R4=>I42	R4=>I43	I12=R	I13=>R	I14=R	I21=>R	I23=>R	I24=R	I31=>R	I32=R	I34=R	I41=>R	I42=R	I43=>R
+
+tm = genfromtxt('tmat.csv', delimiter=',', skip_header=2, usecols=range(1, 56), dtype=int)
+
+# Initial Conditions
+# Initial conditions for each of the four Serotype model
+
+N = 500
+
+ini = (N, 50, 30, 0, 0, 0, 0, 0, 0, 0, 0, 5, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
+
+pars = (1 / 70.,  # mu
+        400,  # beta
+        1.,  # phi
+        .1,  # sigma
+        1., # gamma
+        0.001, #delta : cross immunity protection
+        )
+        
+        
+# infectives for each serotype
+inf_types = {1: [ini[1], ini[12], ini[15], ini[18]],
+             2: [ini[2], ini[9], ini[16], ini[19]],
+             3: [ini[3], ini[10], ini[13], ini[20]],
+             4: [ini[4], ini[11], ini[14], ini[17]]
+            }
+
+# Propensity functions
+def gen_prop_functs():
+    nat_code = "def fnat(r, ini): return r[0]*sum(ini)\n"
+    mu_code = "def fmu{}(r, ini): return r[0]*max(0,ini[{}])\n"
+    inf_code = """def f{}(r, ini): return ini[0]* lamb({})\n"""
[…]

### 2014-12-06 stochastic_model/propfun.py

@@ -1,0 +1,88 @@
+import numpy as np
+
+pars = (0.014285714285714285, 400, 1.0, 0.1, 1.0, 0.001)
+ini = (500, 50, 30, 0, 0, 0, 0, 0, 0, 0, 0, 5, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
+N = 500
+
+# infectives for each serotype
+
+inf_types = {1: [ini[1], ini[12], ini[15], ini[18]],
+             2: [ini[2], ini[9], ini[16], ini[19]],
+             3: [ini[3], ini[10], ini[13], ini[20]],
+             4: [ini[4], ini[11], ini[14], ini[17]]
+            }
+
+lamb = lambda i: 400*(ini[i+1] + sum(np.array(inf_types[i+1])*1.0))
+
+#Natality
+
+def fnat(r, ini): return r[0]*sum(ini)
+
+#Infection rates
+
+def fS_I1(r, ini): return ini[0]* lamb(0)
+def fS_I2(r, ini): return ini[0]* lamb(1)
+def fS_I3(r, ini): return ini[0]* lamb(2)
+def fS_I4(r, ini): return ini[0]* lamb(3)
+
+#Primary recovery rates
+
+def fI1_R1(r, ini): return r[3]*max(0,ini[1])
+def fI2_R2(r, ini): return r[3]*max(0,ini[2])
+def fI3_R3(r, ini): return r[3]*max(0,ini[3])
+def fI4_R4(r, ini): return r[3]*max(0,ini[4])
+
+#Secondary Infection rates
+
+def fR1_I12(r, ini): return r[4]*r[5]*ini[5]*lamb(1)
+def fR1_I13(r, ini): return r[4]*r[5]*ini[5]*lamb(2)
+def fR1_I14(r, ini): return r[4]*r[5]*ini[5]*lamb(3)
+def fR2_I21(r, ini): return r[4]*r[5]*ini[6]*lamb(0)
+def fR2_I23(r, ini): return r[4]*r[5]*ini[6]*lamb(2)
+def fR2_I24(r, ini): return r[4]*r[5]*ini[6]*lamb(3)
+def fR3_I31(r, ini): return r[4]*r[5]*ini[7]*lamb(0)
+def fR3_I32(r, ini): return r[4]*r[5]*ini[7]*lamb(1)
+def fR3_I34(r, ini): return r[4]*r[5]*ini[7]*lamb(3)
[…]

### 2014-12-06 aux/Rt_aux.R
added R_t plots and PSA for Rt

@@ -41,7 +41,7 @@ Rt.beta <- function(d, gt = 3, alpha = .95, a0 = 2 , b0 = 3){
   return(data.frame(Rt, Rt2, CIs))
 }
 ###################################################
-rRtpost <- function(n, a = , b = ){ # let's sample from the distribution of R_t
+rRt <- function(n, a, b){ # let's sample from the distribution of R_t
   x <- rbeta(n, shape1 = a, shape2 = b)
   y <- x/(1-x) # yup! simple as that!
 return(y)

### 2014-12-06 aux/sampling_Rt_posterior.R

@@ -1,0 +1,68 @@
+### Example code showing how to sample from posterior of R_t
+### A raw estimate of R_t is Y(t+1)/Y(t) but we can also take the  mean of the (inverse beta) posterior distribution
+### WARNING: please note that we know the posteriors exactly, so none of this random variate generation is actually needed.
+### We are just illustrating how one can sample from the prior/posterior if one so desires.
+source("Rt_aux.R")
+source("elicit_Rt_prior.R")
+set.seed(8632)
+## getting the data
+cases <- data.frame(read.table("../DATA/data_Rt_dengue.csv", sep = ",", header = TRUE))$cases
+
+# now lets choose some values of cases
+sorted <- sort(cases)
+low <- sample(head(sorted, 10), 1)
+medium <- sample(sorted[round(length(cases)/2):(round(length(cases)/2)+10)], 1)
+high <- sample(tail(sorted, 10), 1)
+values <- c(low, medium, high)
+post <- sapply(values, function(x) match(x, cases)) # Y_{t}
+post1 <- post + 1 # Y_{t+1}
+
+(hatRt <- cases[post1]/cases[post])
+
+## creating the priors
+
+(pars.cons <- elicit.Rt.hyperpars(m0 = .5, c = .5)) # conservative prior
+(pars.lax <- elicit.Rt.hyperpars(m0 = 1, c = 1)) # more lax prior
+(pars.crazy <- elicit.Rt.hyperpars(m0 = 2, c = .2)) # crazy prior
+
+# the actual priors
+N <- 1000
+cons.prior <- rRt(n = N, a = as.numeric(pars.cons["a_0"]), b = as.numeric(pars.cons["b_0"]))
+lax.prior <- rRt(n = N, a = as.numeric(pars.lax["a_0"]), b = as.numeric(pars.lax["b_0"]))
+crazy.prior <- rRt(n = N, a = as.numeric(pars.crazy["a_0"]), b = as.numeric(pars.crazy["b_0"]))
+
+# now lets get the posteriors, shall we?
+## with few cases
+cons.post.low <- rRt(n = N, a = cases[post1][1] + as.numeric(pars.cons["a_0"]), b = cases[post][1] + as.numeric(pars.cons["b_0"]))
[…]

### 2014-12-07 aux/Rt_plot_bands.R
Adding some of the R scripts

@@ -1,0 +1,31 @@
+### Code to calculate and plot R_t plus credibility intervals.
+### Rt is the raw estimate Y(t+1)/Y(t) and Rt2 is the mean of the inverse beta distribution
+### See the tech report for details: https://github.com/maxbiostat/CODE/blob/master/Alerta/confidence_Rt.pdf
+
+source("Rt_aux.R")
+source("elicit_Rt_prior.R")
+library(lubridate)
+#  
+dadosexp <- data.frame(read.table("../DATA/data_Rt_dengue_complete.csv", sep = ",", header = TRUE))
+dadoexp <- na.omit(dadosexp)
+library(ggplot2)
+library(gridExtra)
+dadoexp$start <- as.Date(dadoexp$start)
+svg("../plots/cases_and_Rt.svg")
+par(mar = c(0, 0, 0, 0))
+bottom.panel <- ggplot(dadoexp, aes(x = start, y = Rt2, ymin = lwr, ymax = upr)) +
+    geom_line() +
+  geom_ribbon(alpha = 0.25) + 
+  geom_line(data = dadoexp, size = 1.0, colour = "blue" ) +
+  labs(x = "Time (weeks)",
+       y = expression(R[t]),
+#        y = "Effective reproductive number",
+       title = "")
+top.panel <- ggplot(dadoexp, aes(x = start, y = cases)) +
+  geom_line() +
+  geom_line(data = dadoexp, size = 1.0, colour = "red" ) +
+  labs(x = "Time (weeks)",
+              y = "Dengue Cases",
+       title = "")
+grid.arrange(top.panel, bottom.panel, heights = c(1/2, 1/2)) 
+dev.off()

### 2014-12-07 aux/make_Rt_data_big.R

@@ -1,0 +1,22 @@
+### Code to export a CSV file with the disease cases and R_t plus credibility intervals.
+### Rt is the raw estimate Y(t+1)/Y(t) and Rt2 is the mean of the inverse beta distribution
+### See the tech report for details: https://github.com/maxbiostat/CODE/blob/master/Alerta/confidence_Rt.pdf
+
+source("Rt_aux.R")
+source("elicit_Rt_prior.R")
+library(lubridate)
+
+dat <- read.csv("../dengue_cases_Rio_perEW_1996-2009.csv", sep = "\t", header = TRUE)
+# please check these everytime you run the script with new data 
+# use http://www.cve.saude.sp.gov.br/htm/Cve_se03.htm to get 'start' and 'end' right
+start <- ymd("1996-01-08")
+end <- ymd("2009-12-20")
+starts <- format(seq(start, end, by = "week"), "%Y-%m-%d")
+ends <- format(seq(as.Date(start) + 6, as.Date(end) + 6, by = "week"), "%Y-%m-%d")
+seDates <- data.frame( start = starts, end = ends)
+#  
+cases <- dat$CASES
+pars <- elicit.Rt.hyperpars(m0 = 1/2, v0 = 2)  
+Rtb <- Rt.beta(cases, a0 = pars[1], b0 = pars[2])
+dadoexp <- data.frame(SE = rep(1:52, 14), seDates, cases, Rtb)
+write.csv(dadoexp, file = "data_Rt_dengue_big.csv", row.names = FALSE)

### 2014-12-07 S1_lm1.tex
Proving the tail behaviour claim

@@
 There are a number of approaches to deriving the distribution of \rr.
-One could for example  use the distribution in equation A6 of~\citet{nishiura} 
-or estimate \rr using the approach described in the main text~\citep{mantel}.
+Alternatively to the approach described in the main text~\citep{mantel}, one 
+could use the conditional distribution of \rr on 
+$Y_{t+1}$ and $Y_t$ as defined in equation A7 of \citet{nishiura}:
+\begin{equation}
+\label{seq:unorm}
+f_{R}(R_{t}) = (Y_tR_{t})^{Y_{t+1}} e^{-Y_tR_{t}}
+\end{equation}
+Noticing the kernel of (\ref{seq:unorm}) is that of a gamma distribution with 
+$a_2 = Y_{t+1}+1$ and $b_2 = Y_t$, we obtain a proper density from which to 
+construct $c_{\alpha}(R_t)$, simply by computing the appropriate quantiles of 
+said distribution.
+ This density is
+\begin{equation}
+\label{seq:densityNishiura}
+f_N(R_t| a_2, b_2) =  \frac{b_2^{a_2}}{\Gamma(a_2)} R_t^{a_2-1} e^{-b_2 R_t}
+\end{equation}
+
 In order to decide which approach to take, it may be of use analyzing the 
@@
 conservative confidence/credibility intervals.
+To see that one needs simply take the ratio of the Beta prime and Gamma 
+(unnormalized) densities and evaluate the limit as $R_t$ goes to infinity:
+\begin{equation}
+ \label{seq:densityratio}
+ \lim_{R_t\to\infty}\frac{f_P(R_t| a_1, b_1)}{f_N(R_t| a_2, b_2)} =  
+\lim_{R_t\to\infty}\frac{e^{Y_{t}R_t}}{(1 +R_t)^{Y_{t} + Y_{t +1}+2}} = \infty
+\end{equation}
+
 As a side note, the Bayesian approach presented in this 
@@
 \citet{wilson} and \citet{wilson} for $Y_{t+1}$ and $Y_t >> 1$.
-Under the flat  uniform prior for $\theta_t$, the Bayesian posterior credibility interval is nearly indistinguishable from the confidence interval proposed by \citet{clopper} for $Y_{t+1}, Y_t > 20$.
-Note that the $Beta(1, 1)$ uniform prior for $\theta_t$ constitutes a poor prior choice mainly because the induced distribution for \rr is only well-defined for $b_0 > 2$.
+Under the flat  uniform prior for $\theta_t$, the Bayesian posterior credibility 
+interval is nearly indistinguishable from the confidence interval proposed by 
+\citet{clopper} for $Y_{t+1}, Y_t > 20$.
+Note that the $Beta(1, 1)$ uniform prior for $\theta_t$ constitutes a poor prior 
+choice mainly because the induced distribution for \rr is only well-defined for 
+$b_0 > 2$.
 
@@
 and flexibility of the beta family of distributions.
-Prior elicitation can also be done for \rr and the hyperparameters directly plugged into the prior for $\theta_t$. 
-One can, for example, choose a priori mean and variance for \rr and find $a_0$ and $b_0$ that satisfy those conditions.
+Prior elicitation can also be done for \rr and the hyperparameters directly 
+plugged into the prior for $\theta_t$. 
+One can, for example, choose a priori mean and variance for \rr and find $a_0$ 
+and $b_0$ that satisfy those conditions.
 Let $m_0$ and $v_0$ be the prior expectation and variance for $R_t$. 
@@
 \begin{align}
-\label{eq:elicitation}
+\label{seq:elicitation}
 a_0 &= \frac{m_0v_0 + m_0^3 + m_0^2}{v_0} \\
@@
 \end{align}
-If one wants only to specify $m_0$ and a coefficient of variation $c$ \footnote{$c = \sqrt{v_0}/ m_0$.} for $R_t$ \textit{a priori}, some less boring algebra gives:
+If one wants only to specify $m_0$ and a coefficient of variation $c$ 
+\footnote{$c = \sqrt{v_0}/ m_0$.} for $R_t$ \textit{a priori}, some less boring 
+algebra gives:
 \begin{align}
-\label{eq:elicitationcv}
+\label{seq:elicitationcv}
 a_0 &= \frac{m_0^3c^2 + m_0^3 + m_0^2}{m_0^2c^2} \\
@@
 to the detection threshold.
+We provide an R script to perform the above elicitation at 
+\url{https://github.com/fccoelho/paperLM1/blob/master/aux/elicit_Rt_prior.R}.
 \newpage
@@
  \centering
- \includegraphics[width=16cm]{Dengue4.png}
+ \includegraphics[width=16cm]{plots/Dengue4.png}
 

### 2014-12-12 draft_lm1.tex
referencing to the elicitation; in the hopes that it FINALLY closes #8 :0)
little proof-reading

@@
 Many authors have derived $c_{\alpha}(\theta_t)$ following orthodox approaches  
-such as those of \cite{wilson} and \cite{clopper}, mainly for simplicity.
+(see for example~\cite{wilson} and~\cite{clopper}) mainly for simplicity.
 We choose instead to take a Bayesian approach and use the  $100\alpha \%$ 
@@
 the transform in~(\ref{eq:confRt}).
+Also, one can choose $a_0$ and $b_0$ so as to elicit meaningful prior distributions
+for $R_t$.
+We show how to elicit the prior for $R_t$ from specified prior mean and variance
+or coefficient of variation in Text S1.
 
@@
 in~(\ref{eq:densityMantel}) directly.
-s
+
 \subsection*{Mathematical modeling} % The model

### 2015-01-20 draft_lm1.tex
added figure with the maps

@@
 \documentclass[10pt]{article}
+\usepackage{todonotes}
+\usepackage{subfigure}
 % amsmath package, useful for mathematical formulas
@@
 \usepackage{microtype}
-\usepackage{todonotes}
 \DisableLigatures[f]{encoding = *, family = * }
@@
 {\Large
-\textbf{Estimating the Attack ratio of Dengue Epidemics without serotype 
-information}
+\textbf{Estimating the Attack ratio of Dengue Epidemics, under time-varying 
+force of infection, from aggregated case notification data}
 }
@@
 In places with sufficient temperature variation dengue is predominantly a 
-summer disease.
+summer disease. So it is fair to say that these environmental fluctuations play 
+a key role in determining beginning and end of epidemic periods through its 
+effects on the force of infection.
 
 The long term dynamics of dengue is also modulated by the alternation of virus 
-types in circulation Demographics also plays a role in replenishing the 
+types in circulation. Demographics also plays a role in replenishing the 
 population of susceptibles.
 
-The Attack ratio of disease is a mesure of mobidity defined as the number of new
-cases divided byt the population at risk.
-For Dengue outbreaks and epidemics 
-this can be difficult to calculated due to the lack of information about the 
-population at risk. The population at risk in the case of dengue, is the number 
-of susceptibles to the circulating virus type(s) befor a given epidemic.
+The Attack ratio of disease is a measure of morbidity defined as the number of 
+new cases divided by the population at risk.
+For Dengue outbreaks and epidemics this can be difficult to calculated due to 
+the lack of information about the population at risk. The population at risk in 
+the case of dengue, is the number of susceptibles to the circulating virus 
+type(s) before a given epidemic.
 
 The main source of information about Dengue dynamics are the incidence 
-timeseries, which represent all the cases of Dengue reported by health 
+time series, which represent all the cases of Dengue reported by health 
 professionals with no discrimination of virus types.
@@
 Belo-Horizonte, Brazil to be of 63\%, 
-meaning that aproximately 37\% of the suspected dengue cases go unreported.  
+meaning that approximately 37\% of the suspected dengue cases go unreported.  
 Lastly, demography and migrations affect the number of susceptible in ways which 
@@
 Due to the alternation of 
-ciculating serotypes, we detected fluctuating levels of susceptibility the 
+circulating serotypes, we detected fluctuating levels of susceptibility the 
 circulating virus(es).
 To estimate the susceptibles we have we have fixed the 
-tramissibility based on the $R_t$ series derived from data~\cite{nishiura}. 
+transmissibility based on the $R_t$ series derived from data~\cite{nishiura}. 
 Then, 
@@
 
+\begin{figure}
+ \subfigure[2010]{\includegraphics[width=8cm, 
+height=4.82cm]{./plots/heatmap2010.jpg}
+% heatmap2010.jpg: 1075x589 pixel, 96dpi, 28.44x15.58 cm, bb=0 0 806 442
+}
+  \subfigure[2011]{\includegraphics[width=8cm]{./plots/heatmap2011.jpg}}
+  \subfigure[2012]{\includegraphics[width=8cm]{./plots/heatmap2012.jpg}
+% heatmap2010.jpg: 1075x589 pixel, 96dpi, 28.44x15.58 cm, bb=0 0 806 442
+}
+  \subfigure[2013]{\includegraphics[width=8cm]{./plots/heatmap2013.jpg}}
+  \caption{{\bf Maps showing the incidence of dengue in the city of Rio de 
+Janeiro from 2010 to 2013.} Circles indicate individual notified cases. A 
+heatmap is overlayed on the maps showing absolute density of cases.}
+  \label{fig:mapas}
+\end{figure}
+
 Table \ref{tab:AR} contains the attack ratios calculated for each 

### 2015-01-20 draft_lm1.tex
added a subsection describing the data. Closes issue #4.

@@
 
-In this section we will start by describing a method to estimate the 
-effective 
-reproductive number, $R_t$, and obtain its 
-posterior distribution. 
+In this section we will start by describing the data and then the method used 
+to estimate the effective reproductive number, $R_t$, from the data and obtain 
+its posterior distribution. 
 We then proceed to describe Susceptible-infectious-recovered (SIR) model 
@@
 epidemic using the estimated susceptible fraction and observed incidence.
-We employ these methods to analyze Dengue weekly incidence time series for the 
-city of Rio de Janeiro from 1996 to 2014.
+
+\subsection{The dataset} 
+The data used in this paper consists of  Dengue weekly notified cases time 
+series for the city of Rio de Janeiro from 1996 to 2014. The cases are notified 
+based only on clinical symptoms. Laboratory confirmation and serotype 
+information are available only for a very small sample and only on recent 
+years (2010-2013).
 

### 2015-01-27 draft_lm1.tex
improved code to generate the table; added script to generate rt plot improved manuscript

@@
 considered permanent after the exposure of humans to each of the viruses and 
-cross immunity is considered very limited[ref].
+cross immunity is considered very limited\cite{halstead_dengue_2007}.
 Thus, the proportions of viral types co-circulating at any point in time is 
@@
 a key role in determining beginning and end of epidemic periods through its 
-effects on the force of infection.
+effects on the force of infection, which cannot be taken as 
+constant\cite{reiner_time-varying_2014}
 
@@
 outbreak ($S_{0, year}$) in the last 18 years.
-Due to the alternation of 
-circulating serotypes, we detected fluctuating levels of susceptibility the 
-circulating virus(es).
+Due to the alternation of circulating serotypes, we detected fluctuating levels 
+of susceptibility the circulating virus(es).
 To estimate the susceptibles we have we have fixed the 
@@
 series of infectious and susceptibles for measles 
-outbreaks from case data.\todo[inline]{anyone else?}
-In the case of Dengue the 
+outbreaks from case data.
+In the case of Dengue, the full (multi-year/multi-epidemic)
 series of susceptibles to all possible serotypes, cannot be reconstructed based 
@@
 information are available only for a very small sample and only on recent 
-years (2010-2013).
+years (2010-2013). The incidence is normalized for the parameter estimation 
+procedures. The normalizatio was done by dividing the number os cases reported 
+by the total city population at each year as given by the census.
 
@@
 In the traditional formulation of the model, transmission is governed by a 
-transmission rate $\beta$ and recovery happens at a rate $\tau$.
+constant transmission rate $\beta$ and recovery happens at a rate $\tau$.
 
-As the epidemic progresses, the effective transmission  rate changes and is 
+For our analysis we chose to let the force of infection vary with time, just 
+as it does in the actual epidemics, as seen in the data. So as the epidemic 
+progresses, the effective transmission  rate changes and is 
 given by 
@@
 \end{equation}
-where $R_t$ is the effective reproductive number as defined 
+where $R_t$ is the effective reproductive number, estimated as 
 in~\ref{eq:Rtestimate}.
-We parameterize our model with the varying transmission using the system of 
-ordinary differential equations
+The complete model with the time-verying force of infection is given by
+the system of ordinary differential equations below.
 \begin{align}
@@
 
-The time series used for all the analyses consists of officially notified 
-dengue cases aggregated weekly. 
-Figure XX  shows the $R_t$ series, according to 
+Figure \ref{fig:rtseries}  shows the $R_t$ series, according to 
 (\ref{eq:Rtestimate}). 
@@
 indistinguishable from 1.
+
+\begin{figure}
+ \centering
+ \includegraphics[width=16cm]{./plots/rt_series.png}
+
+ \caption{{\bf Estimated time-series for $R_t$, along with 95\% credible 
+intervals.} Top panel show reported cases from which $R_t$ is estimated.}
+ \label{fig:rtseries}
+\end{figure}
 
@@
 not being exposed to the disease.
-But, as we can see in figure xx\todo{add 
-figure with the four maps}, the disease spreads pretty much homogeneously over 
+However, as we can see in figure \ref{fig:mapas} , the disease spreads pretty 
+much homogeneously over 
 the entire city (in the last four epidemics, at least).
@@
 Table \ref{tab:AR} contains the attack ratios calculated for each 
-epidemic/outbreak.
+epidemic/outbreak. It is interesting to notice that the larger epidemics, in 
+terms of total number of cases are not the one with the greater attack ratios. 
 % We only support three levels of headings, please do not create a heading level below \subsubsection.
@@
 \begin{center}
-\includegraphics[width=.9\textwidth]{plots/concat_SI.png}
+\includegraphics[width=\textwidth]{plots/concat_SI.png}
 \end{center}
@@
 \caption{
-\bf{Median Attack ration calculated according to (\ref{eq:AR2})}}
+\bf{Median Attack ration calculated according to (\ref{eq:AR2})}. $^\dag$: Year 
+corresponds to the start of the epidemic. Sometimes the peak of cases occur in 
+the following year.}
 \begin{center}
@@
 \hline
-Year & Median Attack Ratio (\%) \\
+Year$^\dag$ & Median Attack Ratio (\%) \\
 \hline

### 2015-01-27 python/fit_dengue_split_big_plot.py

@@ -42,6 +42,7 @@ def create_tex_table(dbs):
     Create Latex table with the Attack ratios for each epidemic
     :return:
     """
+    p = {}; s = {}; o = {}
     pts = OrderedDict()
     series = OrderedDict()
     obs = OrderedDict()
@@ -49,15 +50,19 @@ def create_tex_table(dbs):
         y = db.split('_')[0][-4:]
         data, theta, srs = read_data(db)
         #print data
-        series[y] = srs
-        obs[y] = data.I
-        pts[y] = theta
-
+        s[y] = srs
+        o[y] = data.I
+        p[y] = theta
+    for y in sorted(s.keys()):
+        series[y] = s[y]
+        obs[y] = o[y]
+        pts[y] = p[y]
+    del p, s, o
     head = r"""\begin{center}
-\begin{tabular}{c|c}
+\begin{tabular}{c|c|c}
 \hline
 """
-    head += r"""Year & median Attack Ratio \\
+    head += r"""Year & median Attack Ratio $ $S_0$ \\
 \hline
 """
     bot = r"""
@@ -68,16 +73,19 @@ def create_tex_table(dbs):
     body = r""
     st = []
     # years = sorted(list(series.keys()))
-    for i, Y in enumerate(series.keys()):
+    print series.keys()
+    for i, (Y, V) in enumerate(series.items()):
         cases = obs[Y].sum()
-        first_week = series.values()[i].index[0]
+        first_week = V.index[0]
+        s0 = array(series[Y].S.ix[first_week])
         try:
-            ratio = 1.0*cases/array(series[Y].S.ix[first_week])
-            body += Y + r" & {:.2%} ({:.2%}-{:.2%})\\".format(nanmedian(ratio), stats.scoreatpercentile(ratio, 2.5), stats.scoreatpercentile(ratio, 97.5))
+            ratio = 1.0*cases/s0
+            body += Y + r" & {:.2%} ({:.2%}-{:.2%}) & {:.3%}\\".format(nanmedian(ratio), stats.scoreatpercentile(ratio, 2.5), stats.scoreatpercentile(ratio, 97.5), nanmedian(s0))
             body += "\n"
         except KeyError as e:
             print Y, first_week, e
-        
+        except ValueError as e:
+            print s0, e
 
     return head + body + bot
 
@@ -222,10 +230,10 @@ if __name__ == "__main__":
[…]

### 2015-01-27 python/plotRtbands.py

@@ -1,0 +1,36 @@
+# -*- coding:utf-8 -*-
+u"""
+Created on 22/01/15
+by fccoelho
+license: GPL V3 or Later
+"""
+
+
+import pandas as pd
+import seaborn as sns
+import matplotlib.pyplot as P
+import datetime
+import numpy as np
+
+
+
+
+DF = pd.read_csv("../DATA/data_Rt_dengue_complete.csv", header=0, delimiter=',', skiprows=[1, 2, 3], parse_dates=True)
+# Creating a date-time index  to facilitate date axes below.
+DF.index = pd.to_datetime(DF.start)# [datetime.datetime.strptime(d, "%Y-%m-%d") for d in DF.start]
+# DF.set_index(['start'])
+print DF.info()
+
+# create two subplots with  shared x axes
+fig, (ax1, ax2) = P.subplots(2, 1, sharex=True)
+ax1.plot(DF.index, DF.cases,'-')
+ax1.set_ylabel('Reported cases')
+# print DF.start
+ax2.fill_between(DF.index, DF.lwr, DF.upr, facecolor='blue', alpha=.2)
+ax2.plot(DF.index, DF.Rt2, 'k-', alpha=.8, lw=.5)
+ax2.set_ylabel('$R_t$')
+ax2.set_ylim(0, 5)
+fig.tight_layout()
+fig.autofmt_xdate()
+P.savefig("../plots/rt_series.png", dpi=400)
+P.show()

### 2015-01-27 draft_lm1.tex
updated table 1 an improved results and discussion

@@
 i.e. for every case observed there are 5 or 10 unobserved.
-Since this 
-underreporting ration affects both the numerator and denominator of 
+Since this underreporting affects both the numerator and denominator of 
 (\ref{eq:AR}), its effects should cancel out, giving us an unbiased attack ratio 
@@
 
-Table \ref{tab:AR} contains the attack ratios calculated for each 
-epidemic/outbreak. It is interesting to notice that the larger epidemics, in 
-terms of total number of cases are not the one with the greater attack ratios. 
+Table \ref{tab:AR} contains the attack ratios and medians of the $S_0$ 
+estimated for each epidemic/outbreak. It is interesting to notice that the 
+larger epidemics, in terms of total number of cases are not the one with the 
+greater attack ratios. This stresses the importance of knowing the imunological 
+structure of the population in order to truly assess the impact of eny 
+epidemic.
+
+The authors hope that this evidence will motivate public health 
+outhorities to invest in annual serological surveys, to determine the profile 
+of susceptibility to each dengue virus as well as to estimate the 
+underreporting factor of the notification system.
 % We only support three levels of headings, please do not create a heading level below \subsubsection.
 %\subsection*{Subsection 1}
-\section*{Discussion}
+%\section*{Discussion}
 % Do NOT remove this, even if you are not including acknowledgments.
@@
 The authors are also grateful to Claudia T. 
-Codeço for helpful discussions about the manuscript.
+Code\c{c}o for helpful discussions about the manuscript.
 
@@
 \caption{
-\bf{Median Attack ration calculated according to (\ref{eq:AR2})}. $^\dag$: Year 
-corresponds to the start of the epidemic. Sometimes the peak of cases occur in 
-the following year.}
+\bf{Median Attack ratio and 95\% credibility intervals calculated according to 
+(\ref{eq:AR2})}. $^\dag$: Year corresponds to the start of the epidemic. 
+Sometimes the peak of cases occur in the following year.}
 \begin{center}
-\begin{tabular}{l|r}
+\begin{tabular}{c|c|c}
 \hline
-Year$^\dag$ & Median Attack Ratio (\%) \\
+Year$^\dag$ & median Attack Ratio & $S_0$ \\
 \hline
-1996 & \textbf{38.64} (17.18-54.19)\\
-1997 & \textbf{87.31} (73.83-87.31)\\
-1998 & \textbf{49.86} (49.19-49.86)\\
-1999 & \textbf{10.76} (3.72-20.24)\\
-2000 & \textbf{25.49} (24.47-26.98)\\
-2001 & \textbf{48.12} (46.57-49.15)\\
-2005 & \textbf{14.74} (10.23-21.38)\\
-2006 & \textbf{11.22} (8.00-14.41)\\
-2007 & \textbf{15.08} (14.76-15.38)\\
-2008 & \textbf{13.73} (3.10-30.84)\\
-2010 & \textbf{18.30} (17.34-19.33)\\
-2011 & \textbf{8.59} (8.20-9.36)\\
-2012 & \textbf{13.92} (12.71-14.92)\\
+1996 & 38.64\% (17.18\%-54.19\%) & 0.171\%\\
+1997 & 87.31\% (73.83\%-87.31\%) & 0.273\%\\
+1998 & 49.86\% (49.19\%-49.86\%) & 0.142\%\\
+1999 & 10.76\% (3.72\%-20.24\%) & 0.345\%\\
+2000 & 25.49\% (24.47\%-26.98\%) & 1.547\%\\
+2001 & 48.12\% (46.57\%-49.15\%) & 4.951\%\\
+2005 & 14.74\% (10.23\%-21.38\%) & 1.472\%\\
+2006 & 11.22\% (8.00\%-14.41\%) & 2.810\%\\
+2007 & 15.08\% (14.76\%-15.38\%) & 13.451\%\\
+2008 & 13.73\% (3.10\%-30.84\%) & 0.672\%\\
+2010 & 18.30\% (17.34\%-19.33\%) & 4.540\%\\
+2011 & 8.59\% (8.20\%-9.36\%) & 21.482\%\\
+2012 & 13.92\% (12.71\%-14.92\%) & 6.208\%\\
+
 \hline
 \end{tabular}
-\begin{flushleft} We report medians and 95\% credibility intervals.
-\end{flushleft}
 

### 2015-01-28 draft_lm1.tex
moved the supplementary material to an appendix to the main article.

@@
 
+\section*{Appendix}
+\subsection*{A remark on prior distributions and tail behavior of the 
+distribution of $R_t$}
+\label{sec:tails}
+There are a number of approaches to deriving the distribution of \rr.
+Alternatively to the approach described in the main text\cite{mantel}, one 
+could use the conditional distribution of \rr on 
+$Y_{t+1}$ and $Y_t$ as defined in equation A7 of Nishiura et al.\cite{nishiura}:
+\begin{equation}
+\label{seq:unorm}
+f_{R}(R_{t}) = (Y_tR_{t})^{Y_{t+1}} e^{-Y_tR_{t}}
+\end{equation}
+Noticing the kernel of (\ref{seq:unorm}) is that of a gamma distribution with 
+$a_2 = Y_{t+1}+1$ and $b_2 = Y_t$, we obtain a proper density from which to 
+construct $c_{\alpha}(R_t)$, simply by computing the appropriate quantiles of 
+said distribution.
+ This density is
+\begin{equation}
+\label{seq:densityNishiura}
+f_N(R_t| a_2, b_2) =  \frac{b_2^{a_2}}{\Gamma(a_2)} R_t^{a_2-1} e^{-b_2 R_t}
+\end{equation}
+
+In order to decide which approach to take, it may be of use analyzing the 
+variance and tail behavior of the derived distributions for \rr. 
+Consider the case of using a flat $Uniform(0, 1)$ prior for $\theta_t$.
+With $a_0 = b_0 = 1$, $a_1 = a_2$ and $b_1 = b_2 + 1$.
+The beta prime (inverse beta distribution) will have heavier tails compared to 
+the conditional distribution proposed by~\cite{nishiura}, thus providing more 
+conservative confidence/credibility intervals.
+To see that one needs simply take the ratio of the Beta prime and Gamma 
+(unnormalized) densities and evaluate the limit as $R_t$ goes to infinity:
+\begin{equation}
+ \label{seq:densityratio}
+ \lim_{R_t\to\infty}\frac{f_P(R_t| a_1, b_1)}{f_N(R_t| a_2, b_2)} =  
+\lim_{R_t\to\infty}\frac{e^{Y_{t}R_t}}{(1 +R_t)^{Y_{t} + Y_{t +1}+2}} = \infty
+\end{equation}
+
+As a side note, the Bayesian approach presented in this 
+paper will give similar results to those of 
+\cite{wilson} and \cite{wilson} for $Y_{t+1}$ and $Y_t >> 1$.
+Under the flat  uniform prior for $\theta_t$, the Bayesian posterior 
+credibility 
+interval is nearly indistinguishable from the confidence interval proposed by 
+\citet{clopper} for $Y_{t+1}, Y_t > 20$.
+Note that the $Beta(1, 1)$ uniform prior for $\theta_t$ constitutes a poor 
+prior 
+choice mainly because the induced distribution for \rr is only well-defined for 
+$b_0 > 2$.
+
+An advantage of the Bayesian approach is that one can devise prior 
+distributions for $\theta_t$ taking advantage of the intuitive parameterization 
+and flexibility of the beta family of distributions.
+Prior elicitation can also be done for \rr and the hyperparameters directly 
+plugged into the prior for $\theta_t$. 
+One can, for example, choose a priori mean and variance for \rr and find $a_0$ 
+and $b_0$ that satisfy those conditions.
+Let $m_0$ and $v_0$ be the prior expectation and variance for $R_t$. 
+After some tedious algebra one finds
+\begin{align}
+\label{seq:elicitation}
+a_0 &= \frac{m_0v_0 + m_0^3 + m_0^2}{v_0} \\
+b_0 &= \frac{2v_0 + m_0^2 + m_0}{v_0}
+\end{align}
+If one wants only to specify $m_0$ and a coefficient of variation $c$ 
+\footnote{$c = \sqrt{v_0}/ m_0$.} for $R_t$ \textit{a priori}, some less boring 
+algebra gives:
+\begin{align}
+\label{seq:elicitationcv}
+a_0 &= \frac{m_0^3c^2 + m_0^3 + m_0^2}{m_0^2c^2} \\
+b_0 &= \frac{2m_0^2c^2 + m^2 + m}{m_0^2c^2}
+\end{align}
+
+This approach thus makes possible to incorporate epidemiological knowledge 
+about disease biology (e.g. the magnitude of $R_0$) into the computation of \rr.
+This may prove particularly important when disease counts are low and/or close 
+to the detection threshold.
+We provide an R script to perform the above elicitation at 
+\url{https://github.com/fccoelho/paperLM1/blob/master/aux/elicit_Rt_prior.R}.
+
 \bibliography{lm1}

### 2015-01-28 python/plotRtbands.py
update figure 3 and code to generate it

@@ -17,7 +17,7 @@ import numpy as np
 
 DF = pd.read_csv("../DATA/data_Rt_dengue_complete.csv", header=0, delimiter=',', skiprows=[1, 2, 3], parse_dates=True)
 # Creating a date-time index  to facilitate date axes below.
-DF.index = pd.to_datetime(DF.start)# [datetime.datetime.strptime(d, "%Y-%m-%d") for d in DF.start]
+DF.index = pd.to_datetime(DF.start)
 # DF.set_index(['start'])
 print DF.info()
 

### 2015-01-28 draft_lm1.tex
update figure 3 and code to generate it
Sorry for that; fixes #11
updated discussion

@@
 
-Based on the what has been discussed so far, we can rewrite (\ref{eq:AR}) for 
+Based on what has been discussed so far, we can rewrite (\ref{eq:AR}) for 
 each epidemic $j$ as
@@
 where $S_{0j}$ is the number of susceptibles before each epidemic $j$, which we
-estimated as described above.
+estimated before.
 
@@
 It can be seen that the inter-epidemic periods are characterized by $R_t$ being 
-indistinguishable from 1.
+indistinguishable from 1. Due to the instrinsic variability of the \rr series, 
+the exmination of its credible intervals is essential to identify periods 
+of sustained transmission. The wider credible intervals for the interepidemic 
+periods are due to the scarsity of cases. Our methodology to derive the 
+probability distribution for \rr is an original contribution applicable to the 
+analysis of any incidence time-series of  any transmissible disease.
 
@@
 The authors hope that this evidence will motivate public health 
-outhorities to invest in annual serological surveys, to determine the profile 
-of susceptibility to each dengue virus as well as to estimate the 
+outhorities to invest in annual serological surveys, to determine the 
+susceptibility profile to each dengue virus as well as to estimate the 
 underreporting factor of the notification system.
@@
 interval is nearly indistinguishable from the confidence interval proposed by 
-\citet{clopper} for $Y_{t+1}, Y_t > 20$.
+\cite{clopper} for $Y_{t+1}, Y_t > 20$.
 Note that the $Beta(1, 1)$ uniform prior for $\theta_t$ constitutes a poor 
@@
 each epidemic/outbreak. In the lower panel see the posterior distribution 
-of infectious curves, represented by its Median and 95\% credible interval. 
+of infectious curves, represented by its Median and 95\% credible interval. The 
+credible intervals are very narrow, and can be hard to distiguish from the 
+median line. 
 Scatter plot shows the observed cases. Prevalences are scaled as fractions 

### 2015-01-28 draft_lm1.tex
fixed table formatting
Some small tweaks; #9 still pending; will do later today

@@
 
-\subsection{The dataset} 
+\subsection*{The data set} 
 The data used in this paper consists of  Dengue weekly notified cases time 
@@
 years (2010-2013). The incidence is normalized for the parameter estimation 
-procedures. The normalizatio was done by dividing the number os cases reported 
+procedures. The normalization was done by dividing the number os cases reported 
 by the total city population at each year as given by the census.
@@
 \%$ confidence interval for \rr by noting that
-\begin{align}
+\begin{equation}
 \label{eq:confRt}
- R_t^{(L)} &= \frac{\theta_t^{(L)}}{(1-\theta_t^{(L)})}\nonumber \\
- R_t^{(U)} &= \frac{\theta_t^{(U)}}{(1-\theta_t^{(U)})}
-\end{align}
+ R_t^{(L)} = \frac{\theta_t^{(L)}}{(1-\theta_t^{(L)})} \quad \text{and} \quad R_t^{(U)} = \frac{\theta_t^{(U)}}{(1-\theta_t^{(U)})}\nonumber \\
+\end{equation}
+
 Many authors have derived $c_{\alpha}(\theta_t)$ following orthodox approaches  
@@
 paper will give similar results to those of 
-\cite{wilson} and \cite{wilson} for $Y_{t+1}$ and $Y_t >> 1$.
+\cite{wilson} and \cite{clopper} for $Y_{t+1}$ and $Y_t >> 1$.
 Under the flat  uniform prior for $\theta_t$, the Bayesian posterior 
@@
 \caption{
-\bf{Median Attack ratio and 95\% credibility intervals calculated according to 
-(\ref{eq:AR2})}. $^\dag$: Year corresponds to the start of the epidemic. 
-Sometimes the peak of cases occur in the following year.}
+{\bf Median Attack ratio and 95\% credibility intervals calculated according to 
+(\ref{eq:AR2})}. Values are presented as percentage of total population. 
+$^\dag$: Year corresponds to the start of the epidemic. 
+Sometimes the peak of cases occur in the following year.
+$^\ddag$: Susceptible fraction.}
 \begin{center}
@@
 \hline
-Year$^\dag$ & median Attack Ratio & $S_0$ \\
+Year$^\dag$ & Median attack ratio (\%) & $S_0^\ddag$ (\%)\\
 \hline
-1996 & 38.64\% (17.18\%-54.19\%) & 0.171\%\\
-1997 & 87.31\% (73.83\%-87.31\%) & 0.273\%\\
-1998 & 49.86\% (49.19\%-49.86\%) & 0.142\%\\
-1999 & 10.76\% (3.72\%-20.24\%) & 0.345\%\\
-2000 & 25.49\% (24.47\%-26.98\%) & 1.547\%\\
-2001 & 48.12\% (46.57\%-49.15\%) & 4.951\%\\
-2005 & 14.74\% (10.23\%-21.38\%) & 1.472\%\\
-2006 & 11.22\% (8.00\%-14.41\%) & 2.810\%\\
-2007 & 15.08\% (14.76\%-15.38\%) & 13.451\%\\
-2008 & 13.73\% (3.10\%-30.84\%) & 0.672\%\\
-2010 & 18.30\% (17.34\%-19.33\%) & 4.540\%\\
-2011 & 8.59\% (8.20\%-9.36\%) & 21.482\%\\
-2012 & 13.92\% (12.71\%-14.92\%) & 6.208\%\\
+1996 & 38.64 (17.18-54.19) & 0.171\\
+1997 & 87.31 (73.83-87.31) & 0.273\\
+1998 & 49.86 (49.19-49.86) & 0.142\\
+1999 & 10.76 (3.72-20.24) & 0.345\\
+2000 & 25.49 (24.47-26.98) & 1.547\\
+2001 & 48.12 (46.57-49.15) & 4.951\\
+2005 & 14.74 (10.23-21.38) & 1.472\\
+2006 & 11.22 (8.00-14.41) & 2.810\\
+2007 & 15.08 (14.76-15.38) & 13.451\\
+2008 & 13.73 (3.10-30.84) & 0.672\\
+2010 & 18.30 (17.34-19.33) & 4.540\\
+2011 & 8.59 (8.20-9.36) & 21.482\\
+2012 & 13.92 (12.71-14.92) & 6.208\\
 
@@
 
-\section*{Supporting Information Legends}
+%\section*{Supporting Information Legends}
 %

### 2015-01-28 draft_lm1.tex
Confidence interval is not credibility interval; cleared this confusion in the manuscript

@@
 \end{equation}
+Because the transform from $\theta$ to \rr is monotonically 
+increasing, the result holds for confidence and credibility intervals alike.
 
-Many authors have derived $c_{\alpha}(\theta_t)$ following orthodox approaches  
-(see for example~\cite{wilson} and~\cite{clopper}) mainly for simplicity.
+Many authors have chosen to quantify the uncertainty about $\theta$ 
+following orthodox approaches  (see for example~\cite{wilson} 
+and~\cite{clopper}) mainly for simplicity.
 We choose instead to take a Bayesian approach and use the  $100\alpha \%$ 
@@
 We show how to elicit the prior for $R_t$ from specified prior mean and variance
-or coefficient of variation in Text S1.
+or coefficient of variation in the Appendix.
 
@@
 The complete model with the time-verying force of infection is given by
-the system of ordinary differential equations below.
+the system of ordinary differential equations:
 \begin{align}
@@
 disease dynamics and allows us to estimate the initial fraction of susceptibles 
-$S_0$.
+$S_0$, a key epidemiological parameter.
 
@@
 The authors hope that this evidence will motivate public health 
-outhorities to invest in annual serological surveys, to determine the 
+authorities to invest in annual serological surveys, to determine the 
 susceptibility profile to each dengue virus as well as to estimate the 
@@
 There are a number of approaches to deriving the distribution of \rr.
-Alternatively to the approach described in the main text\cite{mantel}, one 
+Alternatively to the approach described in the main text~\cite{mantel}, one 
 could use the conditional distribution of \rr on 
@@
 In order to decide which approach to take, it may be of use analyzing the 
-variance and tail behavior of the derived distributions for \rr. 
+tail behavior of the derived distributions for \rr. 
 Consider the case of using a flat $Uniform(0, 1)$ prior for $\theta_t$.
@@
 \end{equation}
+Finally, note that we deliberately construct $c_{\alpha}(R_{t})$ as a 
+equal-tailed $100\alpha\%$ credible set, rather than a less conservative 
+highest posterior density (HPD) interval.
 
 As a side note, the Bayesian approach presented in this 
-paper will give similar results to those of 
-\cite{wilson} and \cite{clopper} for $Y_{t+1}$ and $Y_t >> 1$.
+paper will give similar results to orthodox confidence intervals~\cite{wilson} 
+and~\cite{clopper} for $Y_{t+1}$ and $Y_t >> 1$.
 Under the flat  uniform prior for $\theta_t$, the Bayesian posterior 
@@
 interval is nearly indistinguishable from the confidence interval proposed by 
-\cite{clopper} for $Y_{t+1}, Y_t > 20$.
-Note that the $Beta(1, 1)$ uniform prior for $\theta_t$ constitutes a poor 
+Clopper \& Pearson (1931)~\cite{clopper} for $Y_{t+1}, Y_t > 20$.
+Note also that the uniform prior ($Beta(1, 1)$) for $\theta_t$ constitutes a 
+poor 
 prior 
@@
 plugged into the prior for $\theta_t$. 
-One can, for example, choose a priori mean and variance for \rr and find $a_0$ 
+One can, for example, choose prior mean and variance for \rr and find $a_0$ 
 and $b_0$ that satisfy those conditions.

### 2015-01-30 draft_lm1.tex
Small fix in eqs 4/5
another review of the text

@@
 
-Dengue transmission is modulated by environmental conditions, of which 
+Dengue transmission is also modulated by environmental conditions, among which, 
 temperature, due to its effects on the 
@@
 
-The Attack ratio of disease is a measure of morbidity defined as the number of 
-new cases divided by the population at risk.
-For Dengue outbreaks and epidemics this can be difficult to calculated due to 
-the lack of information about the population at risk. The population at risk in 
-the case of dengue, is the number of susceptibles to the circulating virus 
+The Attack ratio (AR) of a disease is a measure of morbidity defined as the 
+number of new cases divided by the population at risk.
+For Dengue outbreaks and epidemics, it can be difficult to calculate the AR due 
+to the lack of knowledge of the population at risk. The population at risk 
+in the case of dengue, is the number of susceptibles to the circulating virus 
 type(s) before a given epidemic.
-
-The main source of information about Dengue dynamics are the incidence 
-time series, which represent all the cases of Dengue reported by health 
-professionals with no discrimination of virus types.
-Only a very small sample 
-of the cases are ever examined to determined the virus type.
 
@@
 susceptibles to the circulating virus types right before the epidemic.
-Without 
-regular virological surveys, it is virtually impossible to determine the 
-population at risk.
+Without regular virological surveys, it is virtually impossible to determine 
+the population at risk.
 
 The attack ratio is also influenced by the reproductive number of the disease 
-~\cite{bacaer_final_2009, katriel_attack_2012}.
-With Dengue, since transmission 
-is affected by environmental conditions as discussed above, the reproductive 
-number is also expected to vary with time.
+~\cite{bacaer_final_2009, katriel_attack_2012}, which is closely associated 
+with the force of infection. Thus the incorporation of the effective 
+reproduction number, \rr, as a function of time, is crucial to an accurate 
+estimation of the AR.
 
-Various challenges exist to assess the attack rate of a dengue epidemic. 
-One of 
-them is that there are a large number of asymptomatic cases in any epidemic, 
-which nevertheless acquire immunity.
-It is estimated that for 
-every case reported, 10-20 are not seen by health authorities (personnal 
-communication).
-Thirdly, under-reporting is a serious issue in Brazil. 
+There are other sources of uncertainty impinging on the estimation of the AR
+of a dengue epidemic. One of them is that there are a large number of 
+asymptomatic cases in any epidemic, which nevertheless acquire immunity.
+It is estimated that for every case reported, 10-20 are not seen by health 
+authorities (personal communication).
+Another is under-reporting, which is a serious issue in Brazil. 
 Duarte and Fran\c{c}a (2006)~\cite{duarte_data_2006}, estimated 
@@
 Lastly, demography and migrations affect the number of susceptible in ways which 
-are not easy to determine.
+are not easy to fully determine.
 
@@
 Due to the alternation of circulating serotypes, we detected fluctuating levels 
-of susceptibility the circulating virus(es).
-To estimate the susceptibles we have we have fixed the 
-transmissibility based on the $R_t$ series derived from data~\cite{nishiura}. 
+of susceptibility to the circulating virus(es).
+To estimate the susceptibles we have set the 
+transmissibility from the $R_t$ series derived from data~\cite{nishiura} 
+(\ref{eq:effbeta}). 
 Then, 
@@
 To accomplish the estimation of the $S_0$, we use a simplified model of 
-dengue transmission based on a single-strain Susceptible-Infecctious-Removed 
+dengue transmission based on a single-strain Susceptible-Infectious-Removed 
 (SIR) model.
 The reason for the choice of a single strain model, is the lack of 
-accurate information regarding the actual proportions of each virus in 
+ information regarding the actual proportions of each virus in 
 circulation.
@@
 its posterior distribution. 
-We then proceed to describe Susceptible-infectious-recovered (SIR) model 
-used to represent the global disease incidence and how $R_t$ can be integrated 
-into the model to allow for time varying transmission. 
+We then proceed to describe the Susceptible-Infectious-Recovered (SIR) model 
+used to represent the aggregated disease incidence and how $R_t$ can be 
+integrated into the model to allow for time varying force of infection. 
 Next, an approach to approximate the posterior distributions 
@@
 Finally, we discuss how to estimate the attack ratio of each 
-epidemic using the estimated susceptible fraction and observed incidence.
+epidemic using the estimated susceptible fraction and the observed incidence.
 
 \subsection*{The data set} 
-The data used in this paper consists of  Dengue weekly notified cases time 
-series for the city of Rio de Janeiro from 1996 to 2014. The cases are notified 
-based only on clinical symptoms. Laboratory confirmation and serotype 
+The data used in this paper consists of time 
+series of weekly notified cases of Dengue for the 
+city of Rio de Janeiro from 1996 to 2014. The cases are notified 
+based only on clinical criteria. Laboratory confirmation and serotype 
 information are available only for a very small sample and only on recent 
 years (2010-2013). The incidence is normalized for the parameter estimation 
-procedures. The normalization was done by dividing the number os cases reported 
+procedures. The normalization was done by dividing the number of cases reported 
 by the total city population at each year as given by the census.
@@
 \label{eq:confRt}
- R_t^{(L)} = \frac{\theta_t^{(L)}}{(1-\theta_t^{(L)})} \quad \text{and} \quad R_t^{(U)} = \frac{\theta_t^{(U)}}{(1-\theta_t^{(U)})}\nonumber \\
+ R_t^{(L)} = \frac{\theta_t^{(L)}}{(1-\theta_t^{(L)})} \quad \text{and} \quad 
+R_t^{(U)} = \frac{\theta_t^{(U)}}{(1-\theta_t^{(U)})}\\
 \end{equation}
@@
 larger epidemics, in terms of total number of cases are not the one with the 
-greater attack ratios. This stresses the importance of knowing the imunological 
-structure of the population in order to truly assess the impact of eny 
-epidemic.
+greater attack ratios. This stresses the importance of knowing the 
+immunological structure of the population in order to truly assess the impact 
+of any epidemic.
 

### 2015-01-30 draft_lm1.tex
further edits

@@
 
-The attack ratio of an epidemic is define by the number of infections divided 
+The attack ratio of an epidemic is defined by the number of infections divided 
 by the size of the population at risk.
@@
 It can be seen that the inter-epidemic periods are characterized by $R_t$ being 
-indistinguishable from 1. Due to the instrinsic variability of the \rr series, 
-the exmination of its credible intervals is essential to identify periods 
-of sustained transmission. The wider credible intervals for the interepidemic 
-periods are due to the scarsity of cases. Our methodology to derive the 
-probability distribution for \rr is an original contribution applicable to the 
+indistinguishable from 1. Due to the intrinsic variability of the \rr series, 
+the examination of its credible intervals is essential to identify periods 
+of sustained transmission. The wider intervals between epidemics are due to the 
+scarcity of cases during these periods. Our methodology to derive the 
+probability distribution of \rr is an original contribution applicable to the 
 analysis of any incidence time-series of  any transmissible disease.
@@
 Since this underreporting affects both the numerator and denominator of 
-(\ref{eq:AR}), its effects should cancel out, giving us an unbiased attack ratio 
+(\ref{eq:AR2}), its effects should cancel out, giving us an unbiased attack 
+ratio 
 estimate.
@@
 estimated for each epidemic/outbreak. It is interesting to notice that the 
-larger epidemics, in terms of total number of cases are not the one with the 
+larger epidemics, in terms of peak size are not the one with the 
 greater attack ratios. This stresses the importance of knowing the 
@@
 and flexibility of the beta family of distributions.
-Prior elicitation can also be done for \rr and the hyperparameters directly 
+Prior elicitation can also be done for \rr and the hyper-parameters directly 
 plugged into the prior for $\theta_t$. 
@@
 of infectious curves, represented by its Median and 95\% credible interval. The 
-credible intervals are very narrow, and can be hard to distiguish from the 
+credible intervals are very narrow, and can be hard to distinguish from the 
 median line. 

### 2015-01-30 draft_lm1.tex
initial version of abstract

@@
 \section*{Abstract}
-Our paper is just awesome.
-% Please keep the Author Summary between 150 and 200 words
-% Use first person. PLOS ONE authors please skip this step. 
-% Author Summary not valid for PLOS ONE submissions.   
+A method for estimation of the Attack ratio of an epidemic and the initial 
+fraction of susceptibles, $S_0$, is proposed. The estimation methods obtain the 
+Bayesian posterior distributions of the parameters by fitting a 
+single-strain SIR transmission model with time-varying force of infection to 
+the aggregated incidence data.  The data consists of 18 years of dengue 
+incidence in the city of Rio de Janeiro, Brazil. An analytical method for 
+deriving the probability distribution of the effective reproductive number, \rr.
 \section*{Author Summary}

### 2015-01-31 draft_lm1.tex
cleaning up + improving abstract; much more to come

--- 2015-01-30 draft_lm1.tex
@@
 %\doublespacing
-% Text layout
 \topmargin 0.0cm
@@
 }
-% Insert Author names, affiliations and corresponding author email.
 \\
@@
 \\
-\bf{2} Programa de Computa\c{c}\~ao Cient\'ifica (PROCC), Funda\c{c}\~ao Oswaldo Cruz, Rio de Janeiro -- RJ, Brazil.
+\bf{2} Programa de Computa\c{c}\~ao Cient\'ifica (PROCC), Funda\c{c}\~ao Oswaldo 
+Cruz, Rio de Janeiro -- RJ, Brazil.
 \\
@@
 
-% Please keep the abstract between 250 and 300 words
 \section*{Abstract}
-Our paper is just awesome.
-% Please keep the Author Summary between 150 and 200 words
-% Use first person. PLOS ONE authors please skip this step. 
-% Author Summary not valid for PLOS ONE submissions.   
+
+Quantifying the attack ratio of disease is key to epidemiological inference and 
+Public Health planning.
+For multi-serotype pathogens, however, different levels of 
+serotype-specific immunity make it difficult to assess the population at risk. 
+In this paper we propose a Bayesian method for estimation of the attack ratio 
+of an epidemic and the initial fraction of susceptibles using aggregated 
+incidence data. 
+We derive the probability distribution of the effective reproductive number, 
+\rr, and use MCMC to obtain posterior 
+distributions of the parameters of a single-strain SIR transmission model with 
+time-varying force of infection.
+Our method is showcased in a data set consisting of 18 years of dengue 
+incidence in the city of Rio de Janeiro, Brazil.
+We demonstrate that it is possible to learn about the initial fraction of 
+susceptibles and the attack ratio even in the absence of serotype specific data.
+
 \section*{Author Summary}
@@
 the transform in~(\ref{eq:confRt}).
-Also, one can choose $a_0$ and $b_0$ so as to elicit meaningful prior distributions
-for $R_t$.
+Also, one can choose $a_0$ and $b_0$ so as to elicit meaningful prior 
+distributions for $R_t$.
 We show how to elicit the prior for $R_t$ from specified prior mean and variance
@@
 \subsection*{Bayesian parameter estimation}
+
 We take a Bayesian approach to the estimation of $S_0$.
@@
 underreporting factor of the notification system.
-% We only support three levels of headings, please do not create a heading level below \subsubsection.
-%\subsection*{Subsection 1}
-%\section*{Discussion}
-% Do NOT remove this, even if you are not including acknowledgments.
+
 \section*{Acknowledgments}
@@
 \section*{Figure Legends}
-% This section is for figure legends only, do not include
-% graphics in your manuscript file.
-%
-%\begin{figure}
-%\caption{
-%{\bf Bold the first sentence.}  Rest of figure caption.  
-%}
-%\label{Figure_label}
-%\end{figure}
 \begin{figure}
@@
 \section*{Tables}
-% 
-% See introductory notes if you wish to include sideways tables.
-%
-% NOTE: Please look over our table guidelines at http://www.plosone.org/static/figureGuidelines#tables to make sure that your tables meet our requirements.
-Certain types of spacing, cell merging, and other formatting tricks may have unintended results and will be returned for revision.
-%
-%\begin{table}[!ht]
-%\caption{
-%\bf{Table title}}
-%\begin{tabular}{|c|c|c|}
-%table information
-%\end{tabular}
-%\begin{flushleft}Table caption
-%\end{flushleft}
-%\label{tab:label}
-% \end{table}
 
@@
 
-%\section*{Supporting Information Legends}
-%
-% Please enter your Supporting Information captions below in the following format:
-%\item{\bf Figure SX. Enter mandatory title here.} Enter optional descriptive information here.
-% 
-%\begin{description}
-%\item {\bf}
-%\item {\bf}
-%\end{description}
-
 \end{document}

### 2015-01-31 draft_lm1.tex
credit where it is due: Dubey (1970) characterised the inverted beta

@@
 a beta prime (or inverted beta) with parameters $ a_1 = Y_{t+1} + a_0$ and $b_1 
-=  Y_t + b_0$.
+=  Y_t + b_0$~\cite{dubey1970}.
 The density of the induced distribution is then 

### 2015-01-31 draft_lm1.tex
An important (caveat) note
small tweaks

@@
 {\Large
-\textbf{Estimating the Attack ratio of Dengue Epidemics, under time-varying 
+\textbf{Estimating the attack ratio of Dengue Epidemics, under time-varying 
 force of infection, from aggregated case notification data}
@@
 
-\section*{Author Summary}
-
 \section*{Introduction}
@@
 
-The Attack ratio (AR) of a disease is a measure of morbidity defined as the 
+The attack ratio (AR) of a disease is a measure of morbidity defined as the 
 number of new cases divided by the population at risk.
@@
 $a_1(a_1 + b_1 - 1)/\left((b_1 - 2)(b_1 - 1)^2 \right) $.
+Note that this result holds only for $n = 1$.
 Sampling from the posterior in~(\ref{eq:densityMantel}) can be made 
@@
 \caption{
-{\bf Median Attack ratio and 95\% credibility intervals calculated according to 
+{\bf Median attack ratio and 95\% credibility intervals calculated according to 
 (\ref{eq:AR2})}. Values are presented as percentage of total population. 

### 2015-01-31 draft_lm1.tex
completely re-written introduction

@@
 %
-% -- IMPORTANT NOTE
-%
-% Be advised that this is merely a template 
-% designed to facilitate accurate translation of manuscript content 
-% into our production files. 
-%
-% This template contains extensive comments intended 
-% to minimize problems and delays during our production 
-% process. Please follow the template 
-% whenever possible.
-%
-% % % % % % % % % % % % % % % % % % % % % % % 
-%
-% Once your paper is accepted for publication and enters production, 
-% PLEASE REMOVE ALL TRACKED CHANGES in this file and leave only
-% the final text of your manuscript.
-%
-% DO NOT ADD EXTRA PACKAGES TO THIS TEMPLATE unless absolutely necessary.
-% Packages included in this template are intentionally
-% limited and basic in order to reduce the possibility
-% of issues during our production process.
-%
-% % % % % % % % % % % % % % % % % % % % % % %
-%
-% -- FIGURES AND TABLES
-%
-% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
-% - Figures should be uploaded separately from your manuscript file. 
-% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
-% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
-% See http://www.plosone.org/static/figureGuidelines for PLOS figure guidelines.
-%
-% Tables should be cell-based and may not contain:
-% - tabs/spacing/line breaks within cells to alter layout
-% - vertically-merged cells (no tabular environments within tabular environments, do not use \multirow)
-% - colors, shading, or graphic objects
-% See http://www.plosone.org/static/figureGuidelines#tables for table guidelines.
-%
-% For sideways tables, use the {rotating} package and use \begin{sidewaystable} instead of \begin{table} in the appropriate section. PLOS guidelines do not accomodate sideways figures.
-%
-% % % % % % % % % % % % % % % % % % % % % % % %
-%
-% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
-%
-% IMPORTANT
-% Below are a few tips to help format your equations and other special characters according to our specifications.
-For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://www.plosone.org/static/latexGuidelines
-%
-% Please be sure to include all portions of an equation in the math environment, and for any superscripts or subscripts also include the base number/text.
-For example, use $mathrm{mm}^2$ instead of mm$^2$ (do not use \textsuperscript command).
-%
-% DO NOT USE the \rm command to render mathmode characters in roman font, instead use $\mathrm{}$
-% For bolding characters in mathmode, please use $\mathbf{}$ 
-%
-% Please add line breaks to long equations when possible in order to fit our 2-column layout. 
-%
-% For inline equations, please do not include punctuation within the math environment unless this is part of the equation.
-%
-% For spaces within the math environment please use the \; or \: commands, even within \text{} (do not use smaller spacing as this does not convert well).
-%
-%
-% % % % % % % % % % % % % % % % % % % % % % % %
-
 \documentclass[10pt]{article}
@@
 \usepackage{microtype}
+\usepackage{indentfirst}
 \DisableLigatures[f]{encoding = *, family = * }
@@
 {\Large
-\textbf{Estimating the attack ratio of Dengue Epidemics, under time-varying 
-force of infection, from aggregated case notification data}
+\textbf{Estimating the Attack Ratio of Dengue Epidemics under Time-varying 
+Force of Infection using Aggregated Notification Data}
 }
@@
 
-Dengue is a disease which is caused by 4 different types of viruses 
-which co-circulate between the vector and human populations. Immunity is 
-considered permanent after the exposure of humans to each of the viruses and 
-cross immunity is considered very limited\cite{halstead_dengue_2007}.
-Thus, the proportions of viral types co-circulating at any point in time is 
-strongly dependent on previous incidence of the disease, which determines the 
-number of susceptibles for each viral type at any point in time.
+Dengue is an arthropod-borne febrile disease caused by a flavivirus with four 
+serotypes which causes an estimated  $50$ million infections each 
+year~\cite{Guzman2010}.
+In humans, immunity against  a particular serotype is  considered permanent 
+after the exposure and cross immunity to there serotypes is considered short 
+lived~\cite{halstead_dengue_2007}.
+As a consequence, the proportions of viral serotypes co-circulating at any 
+point in time are strongly dependent on previous incidence patterns of the 
+disease, which determine the number of individuals susceptible to each serotype 
+at any point in time.
+% LM: Maybe a citation here?
 
@@
 incidence~\cite{honorio_temporal_2009,wu_higher_2009}.
-In places with sufficient temperature variation dengue is predominantly a 
-summer disease. So it is fair to say that these environmental fluctuations play 
-a key role in determining beginning and end of epidemic periods through its 
-effects on the force of infection, which cannot be taken as 
-constant\cite{reiner_time-varying_2014}
-
+In places with sufficient seasonal temperature variation, dengue is 
+predominantly a  summer disease.
+So it is fair to say that these environmental 
+fluctuations play a key role in determining beginning and end of epidemic 
+periods.
+This climatic influence is exerted mainly through its effects on the force of 
+infection, which cannot be taken as constant~\cite{reiner_time-varying_2014} 
+but rather as a seasonal (oscillating) function of time.
 The long term dynamics of dengue is also modulated by the alternation of virus 
@@
 number of new cases divided by the population at risk.
-For Dengue outbreaks and epidemics, it can be difficult to calculate the AR due 
-to the lack of knowledge of the population at risk. The population at risk 
-in the case of dengue, is the number of susceptibles to the circulating virus 
-type(s) before a given epidemic.
-
-In order to calculate the attack ratio, we need to determine the number of 
-susceptibles to the circulating virus types right before the epidemic.
-Without regular virological surveys, it is virtually impossible to determine 
-the population at risk.
+% LM: Maybe a citation here? It's ridiculous I know, but referees can be a pain.
+For dengue epidemics, it can be difficult to calculate the AR due 
+to the lack of knowledge of the population at risk.
+The population at risk in this case is the number of susceptibles 
+to the circulating virus type(s) before a given epidemic. 
+Thus, in order to calculate the attack ratio, we need to determine the number 
+of susceptibles to the circulating virus types right before the epidemic, which 
+is virtually impossible without regular virological surveys.
 
@@
 ~\cite{bacaer_final_2009, katriel_attack_2012}, which is closely associated 
-with the force of infection. Thus the incorporation of the effective 
-reproduction number, \rr, as a function of time, is crucial to an accurate 
-estimation of the AR.
+with the force of infection.
+Thus the incorporation of the effective reproduction number, \rr, as a function 
+of time, is crucial to an accurate estimation of the AR of seasonal diseases 
+like dengue and Influenza.
 
-There are other sources of uncertainty impinging on the estimation of the AR
-of a dengue epidemic. One of them is that there are a large number of 
-asymptomatic cases in any epidemic, which nevertheless acquire immunity.
-It is estimated that for every case reported, 10-20 are not seen by health 
-authorities (personal communication).
-Another is under-reporting, which is a serious issue in Brazil. 
+It is also important to consider other sources of uncertainty impinging on the 
+estimation of the AR of a dengue epidemic. 
+For instance, in any given epidemic there is a large number of 
+mild and asymptomatic cases, which nevertheless acquire immunity.
+Depending on surveillance structure it is estimated that for every case 
+reported, up to 10-20 are not seen by health authorities (personal 
+communication).
+%LM: Best to name who gave that. Oswaldo perhaps?
+Another source of uncertainty is under-reporting, which is a serious issue in
+the health care systems of many underdeveloped countries such as Brazil. 
 Duarte and Fran\c{c}a (2006)~\cite{duarte_data_2006}, estimated 
 the sensitivity of dengue reporting for hospitalized patients in 
-Belo-Horizonte, Brazil to be of 63\%, 
-meaning that approximately 37\% of the suspected dengue cases go unreported.  
-Lastly, demography and migrations affect the number of susceptible in ways which 
+Belo-Horizonte, Brazil to be of 63\%, meaning that approximately 37\% of the 
+suspected dengue cases go unreported.  
+Lastly, demography and migrations affect the number of susceptible in ways 
+which 
 are not easy to fully determine.
+%LM: Since this paragraph pretty much deals with possible limitations we may 
+% want to move it to the discussion. see #14
 
-In this paper, we estimate the number of susceptibles before every major 
-outbreak ($S_{0, year}$) in the last 18 years.
-Due to the alternation of circulating serotypes, we detected fluctuating levels 
-of susceptibility to the circulating virus(es).
-To estimate the susceptibles we have set the 
-transmissibility from the $R_t$ series derived from data~\cite{nishiura} 
-(\ref{eq:effbeta}). 
-Then, 
-from the incidence series and the population at risk, we could then calculate 
-the attack ratio for each epidemic.
+Other methods for estimating the  number of susceptibles while 
+accommodating time-varying force of infection have been proposed 
+before, for measles~\cite{bjornstad_dynamics_2002, 
+wallinga_reconstruction_2003}, a disease that shows remarkable seasonality.
+These methods try to reconstruct the entire 
+series of infectious and susceptibles from case data using deterministic models 
+and generally work well for measles because there is a one-to-one 
+relationship between exposure and immunity, since measles is caused by a 
+single-strain pathogen.
+In the case of dengue, the full (multi-year/multi-epidemic)
+series of susceptibles to all possible serotypes, cannot be reconstructed based 
+solely on a deterministic transmission model, since the arrival/re-emergence of 
+new serotypes, an intrinsically stochastic events, can drastically change the 
+pool of susceptibles, throwing off any sequential estimation based on the 
+incidence dynamics.
 
-To accomplish the estimation of the $S_0$, we use a simplified model of 
-dengue transmission based on a single-strain Susceptible-Infectious-Removed 
-(SIR) model.
-The reason for the choice of a single strain model, is the lack of 
- information regarding the actual proportions of each virus in 
-circulation.
-However, some information is available about the predominant 
-serotypes for some epidemics in the period 
-of study~\cite{macedo_virological_2013}.
-
-Methods for estimating the  number of susceptibles have been proposed 
-before, for other diseases~\cite{bjornstad_dynamics_2002, 
-wallinga_reconstruction_2003}.
-These methods try to reconstruct the entire 
-series of infectious and susceptibles for measles 
-outbreaks from case data.
-In the case of Dengue, the full (multi-year/multi-epidemic)
-series of susceptibles to all possible serotypes, cannot be reconstructed based 
-solely on a deterministic transmission model, since the arrival/re-emergenge of 
-new serotypes (which are a stochastic events) can change drastically the pool 
-of susceptibles throwing off any sequential estimation based on the incidence
-dynamics.
-
-The authors have applied similar methods as used here, to estimate the number 
-of susceptibles for the predominant strain of Influenza at the beginning of  
-Influenza seasons in europe~\cite{pone2011}.
+In this paper, we propose a new approach to estimate the number (fraction) of 
+susceptibles using a simplified model of dengue transmission based on a 
+single-strain Susceptible-Infectious-Removed (SIR) model with time-varying 
+infection rate.
+In order to bypass the limitations of not knowing the serotype-specific 
+seroprevalence and the exact behavior of the force of infection through time, we
+propose to inform the time-varying transmissibility using the \rr series 
+derived from the notification data~\cite{nishiura}.
+We extend a Bayesian framework previously used to estimate the number 
+of susceptibles in Influenza epidemics in Europe~\cite{pone2011} to include 
+time-varying force of infection and derive a probability distribution  for 
+\rr to accommodate uncertainty in the estimates. 
+Then, from the incidence series and the population at risk, we  calculate the 
+attack ratio for each epidemic.
+We apply our method to estimate $S_0$ before every major dengue epidemic in the 
+city of Rio de Janeiro, Brazil in the last 18 years.
 
@@
 Next, an approach to approximate the posterior distributions 
-of the numbers of susceptible to the main circulating Dengue viruses for each 
+of the numbers of susceptible to the main circulating dengue viruses for each 
 epidemic is detailed.
@@
 
-\subsection*{The data set} 
+\subsection*{Data} 
 The data used in this paper consists of time 
-series of weekly notified cases of Dengue for the 
+series of weekly notified cases of dengue for the 
 city of Rio de Janeiro from 1996 to 2014. The cases are notified 
@@
 \label{eq:densityMantel}
-f_P(R_t| a_1, b_1) = \frac{\Gamma(a_1 + b_1)}{\Gamma(a_1)\Gamma(b_1)} R_t^{a_1 - 
+f_P(R_t| a_1, b_1) = \frac{\Gamma(a_1 + b_1)}{\Gamma(a_1)\Gamma(b_1)} R_t^{a_1 
+- 
 1} (1 + R_t)^{-(a_1 + b_1)}
@@
 
+Methods for estimating the  number of susceptibles have been proposed 
+before, for other diseases~\cite{bjornstad_dynamics_2002, 
+wallinga_reconstruction_2003}.
+These methods try to reconstruct the entire 
+series of infectious and susceptibles for measles 
+outbreaks from case data.
+In the case of dengue, the full (multi-year/multi-epidemic)
+series of susceptibles to all possible serotypes, cannot be reconstructed based 
+solely on a deterministic transmission model, since the arrival/re-emergenge of 
+new serotypes (which are a stochastic events) can change drastically the pool 
+of susceptibles throwing off any sequential estimation based on the incidence
+dynamics.
+
+The reason for the choice of a single strain model, is the lack of 
+information regarding the actual proportions of each virus in 
+circulation.
+However, some information is available about the predominant 
+serotypes for some epidemics in the period 
+of study~\cite{macedo_virological_2013}.
+
 \begin{figure}
@@
 LMC is grateful to Dr. Leonardo Bastos for useful discussions on the 
-posterior inference for $R(t)$.
+posterior inference for \rr.
 The authors are also grateful to Claudia T. 
@@
 \end{align}
-If one wants only to specify $m_0$ and a coefficient of variation $c$ 
-\footnote{$c = \sqrt{v_0}/ m_0$.} for $R_t$ \textit{a priori}, some less boring 
+If one wants only to specify $m_0$ and the coefficient of variation $c = 
+\sqrt{v_0}/ m_0$ for $R_t$ \textit{a priori}, some less boring 
 algebra gives:
@@
 
-This approach thus makes possible to incorporate epidemiological knowledge 
-about disease biology (e.g. the magnitude of $R_0$) into the computation of \rr.
+This approach thus makes it possible to incorporate epidemiological knowledge 
+about disease Biology (e.g. the magnitude of $R_0$) into the computation of \rr.
 This may prove particularly important when disease counts are low and/or close 
@@
 We provide an R script to perform the above elicitation at 
-\url{https://github.com/fccoelho/paperLM1/blob/master/aux/elicit_Rt_prior.R}.
+\url{https://github.com/fccoelho/paperLM1/blob/master/R/elicit_Rt_prior.R}.
 

### 2015-01-31 draft_lm1.tex
important reference

@@
 single-strain pathogen.
-In the case of dengue, the full (multi-year/multi-epidemic)
-series of susceptibles to all possible serotypes, cannot be reconstructed based 
-solely on a deterministic transmission model, since the arrival/re-emergence of 
-new serotypes, an intrinsically stochastic events, can drastically change the 
-pool of susceptibles, throwing off any sequential estimation based on the 
-incidence dynamics.
+Recently, methods in the same fashion were developed for dengue when 
+serotype-specific data is available~\cite{Reiner2014}.
+When such data is not available, the series of susceptibles to all possible
+serotypes, cannot be reconstructed based solely on a deterministic transmission 
+model, since the arrival/re-emergence of  new serotypes, an intrinsically 
+stochastic events, can drastically change the pool of susceptibles, throwing off 
+any sequential estimation based on the  incidence dynamics.
 
@@
 \subsection*{Data}
+
 The data used in this paper consists of time 
@@
 city of Rio de Janeiro from 1996 to 2014. The cases are notified 
-based only on clinical criteria. Laboratory confirmation and serotype 
+based only on clinical criteria.
+Laboratory confirmation and serotype 
 information are available only for a very small sample and only on recent 
-years (2010-2013). The incidence is normalized for the parameter estimation 
-procedures. The normalization was done by dividing the number of cases reported 
-by the total city population at each year as given by the census.
+years (2010-2013).
+For the parameter estimation  procedures incidence was normalized by dividing 
+the number of cases reported by the total city population at each year as given 
+by the census (Census Bureau, Brazilian Institute of Geography and Statistics, 
+\url{http://www.ibge.gov.br/english/}).
 
@@
 Convergence of the parallel chains was verified at every 100 iterations by the 
-calculation of the Gelman-Rubin's R convergence diagnostic~\cite{brooks1998}.
+calculation of the Gelman-Rubin's R (potential scale 
+reduction), that approaches $1$ at convergence~\cite{brooks1998}.
 

### 2015-02-01 python/fit_dengue_split_big_plot.py
changed to code to solve issue #15.

@@ -80,7 +80,13 @@ def create_tex_table(dbs):
         s0 = array(series[Y].S.ix[first_week])
         try:
             ratio = 1.0*cases/s0
-            body += Y + r" & {:.2%} ({:.2%}-{:.2%}) & {:.3%}\\".format(nanmedian(ratio), stats.scoreatpercentile(ratio, 2.5), stats.scoreatpercentile(ratio, 97.5), nanmedian(s0))
+            body += Y + r" & {:.2} ({:.2}-{:.2}) & {:.3}({:.2}-{:.2})\\".format(nanmedian(ratio),
+                                                                                      stats.scoreatpercentile(ratio, 2.5),
+                                                                                      stats.scoreatpercentile(ratio, 97.5),
+                                                                                      nanmedian(s0),
+                                                                                      stats.scoreatpercentile(s0, 2.5),
+                                                                                      stats.scoreatpercentile(s0, 97.2)
+                                                                                      )
             body += "\n"
         except KeyError as e:
             print Y, first_week, e
@@ -232,8 +238,8 @@ if __name__ == "__main__":
     #~ series('Dengue_S0_big')
 
     dbs = glob.glob("../DengueS*.sqlite")
-    plot_concat_series(dbs)
-    # print create_tex_table(dbs)
+    #plot_concat_series(dbs)
+    print create_tex_table(dbs)
     #plot_rt_beta()
 
 

### 2015-02-01 stochastic_model/SDE_alt4.py

@@ -7,7 +7,7 @@ import os
 from itertools import cycle
 import pylab as P
 import numpy as np
-from stochpy.modules import InterfaceCain
+# from stochpy.modules import InterfaceCain
 
 
 def plot(t, s, l):

### 2015-02-01 draft_lm1.tex
moved the paragraph discussing uncertaities to discusion. Closes #14

@@
 like dengue and Influenza.
-
-It is also important to consider other sources of uncertainty impinging on the 
-estimation of the AR of a dengue epidemic. 
-For instance, in any given epidemic there is a large number of 
-mild and asymptomatic cases, which nevertheless acquire immunity.
-Depending on surveillance structure it is estimated that for every case 
-reported, up to 10-20 are not seen by health authorities (personal 
-communication).
-%LM: Best to name who gave that. Oswaldo perhaps?
-Another source of uncertainty is under-reporting, which is a serious issue in
-the health care systems of many underdeveloped countries such as Brazil. 
-Duarte and Fran\c{c}a (2006)~\cite{duarte_data_2006}, estimated 
-the sensitivity of dengue reporting for hospitalized patients in 
-Belo-Horizonte, Brazil to be of 63\%, meaning that approximately 37\% of the 
-suspected dengue cases go unreported.  
-Lastly, demography and migrations affect the number of susceptible in ways 
-which 
-are not easy to fully determine.
-%LM: Since this paragraph pretty much deals with possible limitations we may 
-% want to move it to the discussion. see #14
 
@@
 
+We sought to deal with all important sources of uncertainty impinging on the 
+estimation of the AR of a dengue epidemic, but not all could be satisfactorily 
+addressed in this analysis.
+For instance, in any given epidemic there is a large number of 
+mild and asymptomatic cases, which nevertheless acquire immunity.
+It is estimated that for every case reported, up to 10-20 are not seen by health 
+authorities (Oswaldo G. Cruz, personal communication).
+
+Another source of uncertainty is under-reporting, which is a serious issue in
+the health care systems of many developing countries such as Brazil. 
+Duarte and Fran\c{c}a (2006)~\cite{duarte_data_2006}, estimated 
+the sensitivity of Dengue reporting for hospitalized patients in 
+Belo-Horizonte, Brazil to be of 63\%, meaning that approximately 37\% of the 
+suspected Dengue cases go unreported.  
+Lastly, demography and migrations affect the number of susceptible in ways 
+which are not easy to fully determine.
+
 \begin{figure}

### 2015-02-01 draft_lm1.tex
fixed code for labels of fig3. Closes #13

@@
 reflect the prevalence of fraction of susceptibles to circulating strain(s) for 
-each epidemic/outbreak. In the lower panel see the posterior distribution 
+each epidemic/outbreak. In the lower panel, we see the posterior distribution 
 of infectious curves, represented by its Median and 95\% credible interval. The 
@@
 median line. 
-Scatter plot shows the observed cases. Prevalences are scaled as fractions 
-of the entire population.
+Scatter plot shows the observed cases. Prevalences (y-axes) are scaled as 
+fractions of the entire population.
 }

### 2015-02-01 python/fit_dengue_split_big_plot.py

@@ -194,6 +194,7 @@ def plot_concat_series(dbs):
     lwr = lambda x: stats.scoreatpercentile(x, 2.5)
     c = cycle(['b', 'g', 'r', 'c', 'm', 'y', 'k'])
     ax1 = P.subplot(211)
+
     for srs in series:
         co = c.next()
         s_median = srs.S.groupby(level='time').median()
@@ -204,6 +205,7 @@ def plot_concat_series(dbs):
         P.fill_between(s_median.index, s_lwr, s_upr, facecolor=co, alpha=.2)
 
     ax2 = P.subplot(212, sharex=ax1)
+    ax2.set_xlabel("Time (weeks)")
     for srs, da in zip(series, obs):
         co = c.next()
         i_median = srs.I.groupby(level='time').median()

### 2015-02-01 stochastic_model/SDE_alt4.py
removed suplemment, and stochastic model code. Closes #12

@@ -1,119 +1,0 @@
-#!/usr/bin/env python
-"""
-Version of the dengue 4 serotype model implemented for Stochpy
-"""
-import stochpy
-import os
-from itertools import cycle
-import pylab as P
-import numpy as np
-# from stochpy.modules import InterfaceCain
-
-
-def plot(t, s, l):
-    P.figure()
-    s = np.vstack(s).T
-    co = cycle(['b', 'g', 'r', 'c', 'm', 'y', 'k'])
-    sy = cycle(['o', '^', '>', '<', 's', '*', '+', '1'])
-    print s.shape, len(t)
-    for i in range(s.shape[1]):
-        P.plot(t, s[:, i], co.next() + sy.next() + '-')
-    P.legend(l, loc=0)
-
-def agg_by_type(s, l):
-    """
-    return series aggregated by serotype
-    :param s: series
-    :param l: labels
-    """
-    s = np.vstack(s).T
-    i1 = [l.index(i) for i in l if i.endswith('1') and i.startswith('I')]
-    s1 = np.zeros(s.shape[0])
-    for i in i1:
-        s1 += s[:, i]
-
-    i2 = [l.index(i) for i in l if i.endswith('2') and i.startswith('I')]
-    s2 = np.zeros(s.shape[0])
-    for i in i2:
-        s2 += s[:, i]
-
-    i3 = [l.index(i) for i in l if i.endswith('3') and i.startswith('I')]
-    # print "I3: ", i3
-    s3 = np.zeros(s.shape[0])
-    for i in i3:
-        s3 += s[:, i]
-
-    i4 = [l.index(i) for i in l if i.endswith('4') and i.startswith('I')]
-    s4 = np.zeros(s.shape[0])
-    # print "I4: ", i4
-    for i in i4:
-        s4 += s[:, i]
-
-    return s1, s2, s3, s4
-
-
-def plot_4_types(t, s, l):
-    """
-    Plot all infectives for each serotype
-    :param t: Times
-    :param s: series
-    :param l: labels
-    """
-    # print l
-    P.figure()
-
-    s1, s2, s3, s4 = agg_by_type(s, l)
-
-    P.plot(t, s1, label='1')
-
-    P.plot(t, s2, label='2')
-
-    P.plot(t, s3, label='3')
-
-    P.plot(t, s4, label='4')
-    P.legend(loc=0)
-    P.xlabel('time')
-    P.ylabel('individuals')
[…]

### 2015-02-01 stochastic_model/dengue_full_SDE.py

@@ -1,158 +1,0 @@
-# coding:utf8
-from __future__ import division
-from cgillespie import Model
-import time
-from numpy import array, genfromtxt
-import pylab as P
-import numpy as np
-
-from itertools import cycle
-
-vnames = ['S', 'I1', 'I2', 'I3', 'I4', 'R1', 'R2', 'R3', 'R4',
-          'I12', 'I13', 'I14', 'I21', 'I23', 'I24', 'I31', 'I32',
-            'I34', 'I41', 'I42', 'I43', 'R']
-
-# transitions
-# S->I1	S->I2	S->I3	S->I4	I1=>R1	I2=>R2	I3=>R3	I4=>R4	R1=>I12	R1=>I13	R1=>I14	R2=>I21	R2=>I23	R2=>I24	R3=>I31	R3=>I32	R3=>I34	R4=>I41	R4=>I42	R4=>I43	I12=R	I13=>R	I14=R	I21=>R	I23=>R	I24=R	I31=>R	I32=R	I34=R	I41=>R	I42=R	I43=>R
-
-tm = genfromtxt('tmat.csv', delimiter=',', skip_header=2, usecols=range(1, 56), dtype=int)
-
-# Initial Conditions
-# Initial conditions for each of the four Serotype model
-
-N = 500
-
-ini = (N, 50, 30, 0, 0, 0, 0, 0, 0, 0, 0, 5, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
-
-pars = (1 / 70.,  # mu
-        400,  # beta
-        1.,  # phi
-        .1,  # sigma
-        1., # gamma
-        0.001, #delta : cross immunity protection
-        )
-        
-        
-# infectives for each serotype
-inf_types = {1: [ini[1], ini[12], ini[15], ini[18]],
-             2: [ini[2], ini[9], ini[16], ini[19]],
-             3: [ini[3], ini[10], ini[13], ini[20]],
-             4: [ini[4], ini[11], ini[14], ini[17]]
-            }
-
-# Propensity functions
-def gen_prop_functs():
-    nat_code = "def fnat(r, ini): return r[0]*sum(ini)\n"
-    mu_code = "def fmu{}(r, ini): return r[0]*max(0,ini[{}])\n"
-    inf_code = """def f{}(r, ini): return ini[0]* lamb({})\n"""
[…]

### 2015-02-01 stochastic_model/propfun.py

@@ -1,88 +1,0 @@
-import numpy as np
-
-pars = (0.014285714285714285, 400, 1.0, 0.1, 1.0, 0.001)
-ini = (500, 50, 30, 0, 0, 0, 0, 0, 0, 0, 0, 5, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
-N = 500
-
-# infectives for each serotype
-
-inf_types = {1: [ini[1], ini[12], ini[15], ini[18]],
-             2: [ini[2], ini[9], ini[16], ini[19]],
-             3: [ini[3], ini[10], ini[13], ini[20]],
-             4: [ini[4], ini[11], ini[14], ini[17]]
-            }
-
-lamb = lambda i: 400*(ini[i+1] + sum(np.array(inf_types[i+1])*1.0))
-
-#Natality
-
-def fnat(r, ini): return r[0]*sum(ini)
-
-#Infection rates
-
-def fS_I1(r, ini): return ini[0]* lamb(0)
-def fS_I2(r, ini): return ini[0]* lamb(1)
-def fS_I3(r, ini): return ini[0]* lamb(2)
-def fS_I4(r, ini): return ini[0]* lamb(3)
-
-#Primary recovery rates
-
-def fI1_R1(r, ini): return r[3]*max(0,ini[1])
-def fI2_R2(r, ini): return r[3]*max(0,ini[2])
-def fI3_R3(r, ini): return r[3]*max(0,ini[3])
-def fI4_R4(r, ini): return r[3]*max(0,ini[4])
-
-#Secondary Infection rates
-
-def fR1_I12(r, ini): return r[4]*r[5]*ini[5]*lamb(1)
-def fR1_I13(r, ini): return r[4]*r[5]*ini[5]*lamb(2)
-def fR1_I14(r, ini): return r[4]*r[5]*ini[5]*lamb(3)
-def fR2_I21(r, ini): return r[4]*r[5]*ini[6]*lamb(0)
-def fR2_I23(r, ini): return r[4]*r[5]*ini[6]*lamb(2)
-def fR2_I24(r, ini): return r[4]*r[5]*ini[6]*lamb(3)
-def fR3_I31(r, ini): return r[4]*r[5]*ini[7]*lamb(0)
-def fR3_I32(r, ini): return r[4]*r[5]*ini[7]*lamb(1)
-def fR3_I34(r, ini): return r[4]*r[5]*ini[7]*lamb(3)
[…]

### 2015-02-02 python/fit_dengue_split_big_plot.py
Update table an figure 3 ( closes #15).
also improved plotting code.

@@ -96,6 +96,42 @@ def create_tex_table(dbs):
     return head + body + bot
 
 
+def plot_AR_S0(dbs):
+    """
+    Scatter plot of Attack ratio by S0 by year.
+    :param ar:
+    :param s0:
+    :return:
+    """
+    p = {}; s = {}; o = {}
+    pts = OrderedDict()
+    series = OrderedDict()
+    obs = OrderedDict()
+    for db in dbs:
+        y = db.split('_')[0][-4:]
+        data, theta, srs = read_data(db)
+        #print data
+        s[y] = srs
+        o[y] = data.I
+        p[y] = theta
+    for y in sorted(s.keys()):
+        series[y] = s[y]
+        obs[y] = o[y]
+        pts[y] = p[y]
+    del p, s, o
+    f, axes = P.subplots(7, 2, sharex='col', sharey='row')
+    for i, (Y, V) in enumerate(series.items()):
+        ax = axes.ravel()[i]
+        cases = obs[Y].sum()
+        first_week = V.index[0]
+        s0 = array(series[Y].S.ix[first_week])
+        ratio = 1.0*cases/s0
+        ax.set_title("${}$".format(Y))
+        ax.scatter(s0, ratio)
+    axes.ravel()[-2].set_xlabel("$S_0$")
+    axes.ravel()[-1].set_xlabel("$S_0$")
+
+
 
 
 def multiplot():
@@ -205,7 +241,7 @@ def plot_concat_series(dbs):
         P.fill_between(s_median.index, s_lwr, s_upr, facecolor=co, alpha=.2)
 
     ax2 = P.subplot(212, sharex=ax1)
-    ax2.set_xlabel("Time (weeks)")
+
     for srs, da in zip(series, obs):
         co = c.next()
         i_median = srs.I.groupby(level='time').median()
@@ -218,6 +254,7 @@ def plot_concat_series(dbs):
         # P.plot_date(pd.to_datetime(da.index), da.I, 'b.', alpha=0.8, label='Observations')
     # P.legend(['Median', 'obs'])
     P.tight_layout()
+    ax2.set_xlabel("Time (weeks)")
     P.savefig('../plots/concat_SI.svg')
     P.savefig('../plots/concat_SI.png', dpi=400)
 
@@ -240,14 +277,10 @@ if __name__ == "__main__":
     #~ series('Dengue_S0_big')
 
     dbs = glob.glob("../DengueS*.sqlite")
[…]

### 2015-02-02 draft_lm1.tex
Update table an figure 3 ( closes #15).
also improved plotting code.
applied UK spell checker

@@
 In order to bypass the limitations of not knowing the serotype-specific 
-seroprevalence and the exact behavior of the force of infection through time, we
+seroprevalence and the exact behaviour of the force of infection through time, 
+we
 propose to inform the time-varying transmissibility using the \rr series 
@@
 
-\subsection*{Mathematical modeling} % The model
+\subsection*{Mathematical modelling} % The model
 
@@
 in~\ref{eq:Rtestimate}.
-The complete model with the time-verying force of infection is given by
+The complete model with the time-varying force of infection is given by
 the system of ordinary differential equations:
@@
 can not discriminate between them due to the lack of serotype-specific data.
-Nevertheless, this modeling strategy can still provide some insight into the 
+Nevertheless, this modelling strategy can still provide some insight into the 
 disease dynamics and allows us to estimate the initial fraction of susceptibles 
@@
 series of susceptibles to all possible serotypes, cannot be reconstructed based 
-solely on a deterministic transmission model, since the arrival/re-emergenge of 
+solely on a deterministic transmission model, since the arrival/re-emergence of 
 new serotypes (which are a stochastic events) can change drastically the pool 
@@
 must remember that these estimates are being affected by an unknown 
-underreporting factor, which experts suggest is somewhere between 5 and 10, 
+under-reporting factor, which experts suggest is somewhere between 5 and 10, 
 i.e. for every case observed there are 5 or 10 unobserved.
-Since this underreporting affects both the numerator and denominator of 
+Since this under-reporting affects both the numerator and denominator of 
 (\ref{eq:AR2}), its effects should cancel out, giving us an unbiased attack 
@@
 susceptibility profile to each dengue virus as well as to estimate the 
-underreporting factor of the notification system.
+under-reporting factor of the notification system.
 
-\section*{Acknowledgments}
+\section*{Acknowledgements}
 LMC is grateful to Dr. Leonardo Bastos for useful discussions on the 
@@
 \section*{Appendix}
-\subsection*{A remark on prior distributions and tail behavior of the 
+\subsection*{A remark on prior distributions and tail behaviour of the 
 distribution of $R_t$}
@@
 
-In order to decide which approach to take, it may be of use analyzing the 
-tail behavior of the derived distributions for \rr. 
+In order to decide which approach to take, it may be of use analysing the 
+tail behaviour of the derived distributions for \rr. 
 Consider the case of using a flat $Uniform(0, 1)$ prior for $\theta_t$.
@@
 An advantage of the Bayesian approach is that one can devise prior 
-distributions for $\theta_t$ taking advantage of the intuitive parameterization 
+distributions for $\theta_t$ taking advantage of the intuitive parametrization 
 and flexibility of the beta family of distributions.
@@
 \hline
-Year$^\dag$ & Median attack ratio (\%) & $S_0^\ddag$ (\%)\\
+Year$^\dag$ & median Attack Ratio & $S_0^\ddag$ \\
 \hline
-1996 & 38.64 (17.18-54.19) & 0.171\\
-1997 & 87.31 (73.83-87.31) & 0.273\\
-1998 & 49.86 (49.19-49.86) & 0.142\\
-1999 & 10.76 (3.72-20.24) & 0.345\\
-2000 & 25.49 (24.47-26.98) & 1.547\\
-2001 & 48.12 (46.57-49.15) & 4.951\\
-2005 & 14.74 (10.23-21.38) & 1.472\\
-2006 & 11.22 (8.00-14.41) & 2.810\\
-2007 & 15.08 (14.76-15.38) & 13.451\\
-2008 & 13.73 (3.10-30.84) & 0.672\\
-2010 & 18.30 (17.34-19.33) & 4.540\\
-2011 & 8.59 (8.20-9.36) & 21.482\\
-2012 & 13.92 (12.71-14.92) & 6.208\\
+1996 & 0.39 (0.17-0.54) & 0.00171(0.0012-0.0038)\\
+1997 & 0.87 (0.74-0.87) & 0.00273(0.0027-0.0032)\\
+1998 & 0.5 (0.49-0.5) & 0.00142(0.0014-0.0014)\\
+1999 & 0.11 (0.037-0.2) & 0.00345(0.0018-0.01)\\
+2000 & 0.25 (0.24-0.27) & 0.0155(0.015-0.016)\\
+2001 & 0.48 (0.47-0.49) & 0.0495(0.048-0.051)\\
+2005 & 0.15 (0.1-0.21) & 0.0147(0.01-0.021)\\
+2006 & 0.11 (0.08-0.14) & 0.0281(0.022-0.037)\\
+2007 & 0.15 (0.15-0.15) & 0.135(0.13-0.14)\\
+2008 & 0.14 (0.031-0.31) & 0.00672(0.003-0.024)\\
+2010 & 0.18 (0.17-0.19) & 0.0454(0.043-0.048)\\
+2011 & 0.086 (0.082-0.094) & 0.215(0.2-0.23)\\
+2012 & 0.14 (0.13-0.15) & 0.0621(0.058-0.068)\\
 

### 2015-02-02 python/fit_dengue_split_big_plot.py
added axes labels to figure 3. closes #13

@@ -83,9 +83,9 @@ def create_tex_table(dbs):
             body += Y + r" & {:.2} ({:.2}-{:.2}) & {:.3}({:.2}-{:.2})\\".format(nanmedian(ratio),
                                                                                       stats.scoreatpercentile(ratio, 2.5),
                                                                                       stats.scoreatpercentile(ratio, 97.5),
-                                                                                      nanmedian(s0),
-                                                                                      stats.scoreatpercentile(s0, 2.5),
-                                                                                      stats.scoreatpercentile(s0, 97.2)
+                                                                                      nanmedian(s0)*100,
+                                                                                      stats.scoreatpercentile(s0, 2.5)*100,
+                                                                                      stats.scoreatpercentile(s0, 97.2)*100
                                                                                       )
             body += "\n"
         except KeyError as e:
@@ -239,7 +239,7 @@ def plot_concat_series(dbs):
         s_lwr = srs.S.groupby(level='time').aggregate(lwr)
         # P.legend(loc=0)
         P.fill_between(s_median.index, s_lwr, s_upr, facecolor=co, alpha=.2)
-
+    ax1.set_ylabel("Fraction of population")
     ax2 = P.subplot(212, sharex=ax1)
 
     for srs, da in zip(series, obs):
@@ -254,6 +254,7 @@ def plot_concat_series(dbs):
         # P.plot_date(pd.to_datetime(da.index), da.I, 'b.', alpha=0.8, label='Observations')
     # P.legend(['Median', 'obs'])
     P.tight_layout()
+    ax2.set_ylabel("Fraction of population")
     ax2.set_xlabel("Time (weeks)")
     P.savefig('../plots/concat_SI.svg')
     P.savefig('../plots/concat_SI.png', dpi=400)
@@ -278,7 +279,7 @@ if __name__ == "__main__":
 
     dbs = glob.glob("../DengueS*.sqlite")
     plot_concat_series(dbs)
-    # print create_tex_table(dbs)
+    print create_tex_table(dbs)
     # plot_AR_S0(dbs)
 
 

### 2015-02-03 draft_lm1.tex
expanded the discussion; added information to figure/table legends; see recent issues

@@
 susceptibles and the attack ratio even in the absence of serotype specific data.
+On the other hand, the information provided by this approach is limited, 
+stressing the need for detailed serological surveys to characterise the 
+distribution of serotype-specific immunity in the population.
 
@@
 calculation of the Gelman-Rubin's R (potential scale 
-reduction), that approaches $1$ at convergence~\cite{brooks1998}.
+reduction factor), which approaches $1$ at convergence~\cite{brooks1998}.
 
@@
 
+Python and R code to perform all the analyses described above is publicly 
+available at~\url{https://github.com/fccoelho/paperLM1}.
+
 \section*{Results and Discussion}
 
-Figure \ref{fig:rtseries}  shows the $R_t$ series, according to 
-(\ref{eq:Rtestimate}). 
+In this paper we propose a method to bypass the lack of serotype-specific case 
+data by informing the time-varying force of infection with the instantaneous 
+reproductive number, \rr which we calculate from aggregated data.
+Our contribution is two-fold: (i) we show a method to quantify uncertainty 
+about \rr that is readily applicable to other diseases and; (ii) we propose to 
+use \rr to inform a dynamic epidemic model with time-varying force of infection 
+in order to gain insight into the attack ratio of each epidemic.
+
+Figure~\ref{fig:rtseries}  shows the $R_t$ series, according to 
+(\ref{eq:Rtestimate})~\cite{nishiura} along with the confidence bands derived 
+in this paper. 
 It can be seen that the inter-epidemic periods are characterized by $R_t$ being 
-indistinguishable from 1. Due to the intrinsic variability of the \rr series, 
-the examination of its credible intervals is essential to identify periods 
-of sustained transmission. The wider intervals between epidemics are due to the 
-scarcity of cases during these periods. Our methodology to derive the 
-probability distribution of \rr is an original contribution applicable to the 
-analysis of any incidence time-series of  any transmissible disease.
+indistinguishable from 1.
+Due to the intrinsic variability of the \rr series, the examination of its 
+credible intervals is essential to identify periods of sustained transmission.
+The wider intervals between epidemics are due to the scarcity of cases during 
+these periods.
+The method to quantify uncertainty proposed here provides more conservative 
+credibility intervals, and therefore offers protection against false alarms.
 
+\begin{center}
+\begin{figure}[!h]
+ \centering
+ \includegraphics[width=16cm]{./plots/rt_series.png}
+ \caption{{\bf Estimated time-series for $R_t$, along with 95\% credible 
+intervals.} Top panel show reported cases from which $R_t$ is estimated.
+As expected, uncertainty about \rr is greater when the case counts are low, for 
+instance in the period $2003$--$2006$, which represented a big hiatus between 
+major epidemics.
+The intrinsic variability of \rr can be used to inform the time-varying force 
+of infection, since it reflects variation in the vector population and other 
+environmental factors such as temperature and seasonal variation.}
+\label{fig:rtseries}
+\end{figure}
+\end{center}
+
+A key epidemiological quantity is the attack ratio (AR) of an epidemic, a 
+measure of morbidity and speed of spread which can be used to predict epidemic 
+size and help efficient Public Health planning.
+The AR depends fundamentally on the population at risk, which in the case of 
+dengue is every naive (to a particular serotype) individual in the population.
+Estimating the initial susceptible fraction $S_0$ for each epidemic is thus 
+central to the estimation of the AR.
 Methods for estimating the  number of susceptibles have been proposed 
@@
 
-The reason for the choice of a single strain model, is the lack of 
-information regarding the actual proportions of each virus in 
-circulation.
-However, some information is available about the predominant 
-serotypes for some epidemics in the period
-of study~\cite{macedo_virological_2013}.
-
+Since there is very limited information regarding the actual proportions of 
+each virus in circulation and most information available is about the 
+predominant serotypes for some epidemics in the period of 
+study only~\cite{macedo_virological_2013}, we propose the use of a 
+simplified a single strain model.
+The main argument we put forward is that by conditioning on the \rr series, we 
+implicitly take into account the variability introduced by the 
+co-circulation of multiple serotypes and heterogeneous levels of immunity in 
+the general population.
 We sought to deal with all important sources of uncertainty impinging on the 
@@
 authorities (Oswaldo G. Cruz, personal communication).
-
-Another source of uncertainty is under-reporting, which is a serious issue in
-the health care systems of many developing countries such as Brazil. 
+Another source of uncertainty is under-reporting of diagnosed cases, which is a 
+serious issue in the health care systems of many developing countries such as 
+Brazil. 
 Duarte and Fran\c{c}a (2006)~\cite{duarte_data_2006}, estimated 
@@
 
-\begin{figure}
- \centering
- \includegraphics[width=16cm]{./plots/rt_series.png}
-
- \caption{{\bf Estimated time-series for $R_t$, along with 95\% credible 
-intervals.} Top panel show reported cases from which $R_t$ is estimated.}
- \label{fig:rtseries}
-\end{figure}
-
-Figure \ref{Fig:S0}, shows the model from (\ref{eq:model}) fitted to the data. 
+Figure~\ref{Fig:S0}, shows the model from (\ref{eq:model}) fitted to the data.
+Despite its limitations, our simplified model fits the data well.
 In it we can see that the susceptibles series in each epidemic starts at the 
@@
 (\ref{eq:AR2}), its effects should cancel out, giving us an unbiased attack 
-ratio 
-estimate.
-One other possible source of bias which would lead to the 
-underestimation of $S_0$ could come from a significant part of the population 
-not being exposed to the disease.
-However, as we can see in figure \ref{fig:mapas} , the disease spreads pretty 
-much homogeneously over 
-the entire city (in the last four epidemics, at least).
+ratio estimate.
+One other possible source of bias which would lead to the underestimation of 
+$S_0$ could come from a significant part of the population not being exposed to 
+the disease.
+However, as we can see in Figure~\ref{fig:mapas}, despite the differences in 
+intensity (incidence), the entire city seems to be at risk, with no 
+particularly ``protected'' areas, at least in the last four epidemics.
 
+\begin{center}
 \begin{figure}
@@
 Janeiro from 2010 to 2013.} Circles indicate individual notified cases. A 
-heatmap is overlayed on the maps showing absolute density of cases.}
+heatmap is overlayed on the maps showing absolute density of cases.
+It can be seen that several areas of the city were affected and no region seems 
+to be free of transmission risk.
+This suggests that although transmission risk varies spatially, there is 
+significant exposure over the entire city.}
   \label{fig:mapas}
 \end{figure}
+\end{center}
 
-Table \ref{tab:AR} contains the attack ratios and medians of the $S_0$ 
-estimated for each epidemic/outbreak. It is interesting to notice that the 
-larger epidemics, in terms of peak size are not the one with the 
-greater attack ratios. This stresses the importance of knowing the 
-immunological structure of the population in order to truly assess the impact 
-of any epidemic.
+Table~\ref{tab:AR} contains the attack ratios and medians of the $S_0$ 
+estimated for each epidemic/outbreak.
+It is interesting to notice that the larger epidemics, in terms of peak size are 
+not the one with the greater attack ratios.
+This stresses the importance of knowing the immunological structure of the 
+population in order to truly assess the impact of any epidemic. [WHY?]
 
-The authors hope that this evidence will motivate public health 
-authorities to invest in annual serological surveys, to determine the 
+We hope that the results presented in paper will motivate public 
+health authorities to invest in annual serological surveys, to determine the 
 susceptibility profile to each dengue virus as well as to estimate the 
@@
 
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[width=\textwidth]{plots/concat_SI.png}
+\end{center}
+\caption{
+{\bf Susceptibles and Infectious posterior curves.}  The curves were estimated 
+only for the periods where $R_t> 1$.  The susceptible curves in the top panel 
+reflect the prevalence of fraction of susceptibles to circulating strain(s) for 
+each epidemic/outbreak.
+In the lower panel, we see the posterior distribution 
+of infectious curves, represented by its median and 95\% credible interval.
+Credible intervals are very narrow, and can be hard to distinguish from the 
+median line. 
+Dots show the observed cases, scaled as fractions of the entire 
+population.
+}
+\label{Fig:S0}
+\end{figure}
+
+\begin{table}[!ht]
+\caption{
+{\bf Median attack ratio and 95\% credibility intervals calculated according to 
+(\ref{eq:AR2})}. 
+Values are presented as percentage of total population. 
+$^\dag$: Year corresponds to the start of the epidemic. 
+Sometimes the peak of cases occur in the following year.
+$^\ddag$: Susceptible fraction.}
+\begin{center}
+\begin{tabular}{c|c|c}
+\hline
+Year$^\dag$ & median Attack Ratio & $S_0^\ddag$ \\
+\hline
+1996 & 0.39 (0.17-0.54) & 0.00171(0.0012-0.0038)\\
+1997 & 0.87 (0.74-0.87) & 0.00273(0.0027-0.0032)\\
+1998 & 0.5 (0.49-0.5) & 0.00142(0.0014-0.0014)\\
+1999 & 0.11 (0.037-0.2) & 0.00345(0.0018-0.01)\\
+2000 & 0.25 (0.24-0.27) & 0.0155(0.015-0.016)\\
+2001 & 0.48 (0.47-0.49) & 0.0495(0.048-0.051)\\
+2005 & 0.15 (0.1-0.21) & 0.0147(0.01-0.021)\\
+2006 & 0.11 (0.08-0.14) & 0.0281(0.022-0.037)\\
+2007 & 0.15 (0.15-0.15) & 0.135(0.13-0.14)\\
+2008 & 0.14 (0.031-0.31) & 0.00672(0.003-0.024)\\
+2010 & 0.18 (0.17-0.19) & 0.0454(0.043-0.048)\\
+2011 & 0.086 (0.082-0.094) & 0.215(0.2-0.23)\\
+2012 & 0.14 (0.13-0.15) & 0.0621(0.058-0.068)\\
+\hline
+\end{tabular}
+\end{center}
+\label{tab:AR}
+\end{table}
+
+\newpage
 \section*{Acknowledgements}
@@
 Code\c{c}o for helpful discussions about the manuscript.
+
+\newpage
+\bibliography{lm1}
 
@@
 could use the conditional distribution of \rr on 
-$Y_{t+1}$ and $Y_t$ as defined in equation A7 of Nishiura et al.\cite{nishiura}:
+$Y_{t+1}$ and $Y_t$ as defined in equation A7 of Nishiura et 
+al.~\cite{nishiura}:
 \begin{equation}
@@
 
-\bibliography{lm1}
-\section*{Figure Legends}
-\begin{figure}
-\begin{center}
-\includegraphics[width=\textwidth]{plots/concat_SI.png}
-\end{center}
-\caption{
-{\bf Susceptibles and Infectious posterior curves.}  The curves were estimated 
-only for the periods where $R_t> 1$.  The susceptible curves in the top panel 
-reflect the prevalence of fraction of susceptibles to circulating strain(s) for 
-each epidemic/outbreak. In the lower panel, we see the posterior distribution 
-of infectious curves, represented by its Median and 95\% credible interval. The 
-credible intervals are very narrow, and can be hard to distinguish from the 
-median line. 
-Scatter plot shows the observed cases. Prevalences (y-axes) are scaled as 
-fractions of the entire population.
-}
-\label{Fig:S0}
-\end{figure}
-
-\section*{Tables}
-
-\begin{table}[!ht]
-\caption{
-{\bf Median attack ratio and 95\% credibility intervals calculated according to 
-(\ref{eq:AR2})}. Values are presented as percentage of total population. 
-$^\dag$: Year corresponds to the start of the epidemic. 
-Sometimes the peak of cases occur in the following year.
-$^\ddag$: Susceptible fraction.}
-\begin{center}
-\begin{tabular}{c|c|c}
-\hline
-Year$^\dag$ & median Attack Ratio & $S_0^\ddag$ \\
-\hline
-1996 & 0.39 (0.17-0.54) & 0.00171(0.0012-0.0038)\\
-1997 & 0.87 (0.74-0.87) & 0.00273(0.0027-0.0032)\\
-1998 & 0.5 (0.49-0.5) & 0.00142(0.0014-0.0014)\\
-1999 & 0.11 (0.037-0.2) & 0.00345(0.0018-0.01)\\
-2000 & 0.25 (0.24-0.27) & 0.0155(0.015-0.016)\\
-2001 & 0.48 (0.47-0.49) & 0.0495(0.048-0.051)\\
-2005 & 0.15 (0.1-0.21) & 0.0147(0.01-0.021)\\
-2006 & 0.11 (0.08-0.14) & 0.0281(0.022-0.037)\\
-2007 & 0.15 (0.15-0.15) & 0.135(0.13-0.14)\\
-2008 & 0.14 (0.031-0.31) & 0.00672(0.003-0.024)\\
-2010 & 0.18 (0.17-0.19) & 0.0454(0.043-0.048)\\
-2011 & 0.086 (0.082-0.094) & 0.215(0.2-0.23)\\
-2012 & 0.14 (0.13-0.15) & 0.0621(0.058-0.068)\\
-
-\hline
-\end{tabular}
-
-\end{center}
-
-\label{tab:AR}
-\end{table}
-
 \end{document}

### 2015-02-03 python/fit_dengue_split_big_plot.py
added a function to calculate the MSEs of the fit: 1/n*sum((obs- median(I))**2)

@@ -37,19 +37,22 @@ def read_data(dbname):
     return df_data, df_pt, df_series
 
 
-def create_tex_table(dbs):
+def get_ordered_series(dbs):
     """
-    Create Latex table with the Attack ratios for each epidemic
+    Return three ordered dicts with the data ordered by year
+    :param dbs: list of filenames of the sqlite databases
     :return:
     """
-    p = {}; s = {}; o = {}
+    p = {}
+    s = {}
+    o = {}
     pts = OrderedDict()
     series = OrderedDict()
     obs = OrderedDict()
     for db in dbs:
         y = db.split('_')[0][-4:]
         data, theta, srs = read_data(db)
-        #print data
+        # print data
         s[y] = srs
         o[y] = data.I
         p[y] = theta
@@ -57,7 +60,16 @@ def create_tex_table(dbs):
         series[y] = s[y]
         obs[y] = o[y]
         pts[y] = p[y]
-    del p, s, o
+    return obs, series, pts
+
+
+def create_tex_table(dbs):
+    """
+    Create Latex table with the Attack ratios for each epidemic
+    :return:
+    """
+    obs, series, pts = get_ordered_series(dbs)
+
     head = r"""\begin{center}
 \begin{tabular}{c|c|c}
 \hline
@@ -103,22 +115,8 @@ def plot_AR_S0(dbs):
     :param s0:
     :return:
     """
-    p = {}; s = {}; o = {}
-    pts = OrderedDict()
-    series = OrderedDict()
-    obs = OrderedDict()
-    for db in dbs:
-        y = db.split('_')[0][-4:]
-        data, theta, srs = read_data(db)
-        #print data
-        s[y] = srs
-        o[y] = data.I
-        p[y] = theta
-    for y in sorted(s.keys()):
-        series[y] = s[y]
-        obs[y] = o[y]
-        pts[y] = p[y]
-    del p, s, o
+    obs, series, pts = get_ordered_series(dbs)
+
     f, axes = P.subplots(7, 2, sharex='col', sharey='row')
     for i, (Y, V) in enumerate(series.items()):
         ax = axes.ravel()[i]
@@ -131,47 +129,6 @@ def plot_AR_S0(dbs):
     axes.ravel()[-2].set_xlabel("$S_0$")
     axes.ravel()[-1].set_xlabel("$S_0$")
 
-
[…]

### 2015-02-03 python/fit_dengue_split_big_plot.py
extended function to calculate errors and added cover letter.

@@ -232,11 +232,15 @@ def plot_pt_corr(df):
 def calc_mse(dbs):
     obs, series, pts = get_ordered_series(dbs)
     mse = OrderedDict()
+    mae = OrderedDict()
+    mape = OrderedDict()
     for Y, srs in series.iteritems():
         n = len(obs[Y])
         i_median = srs.I.groupby(level='time').median()
-        mse[Y] = (1./n) * sum((i_median-obs[Y])**2)
-    return mse
+        mse[Y] = (1./n) * sum(((i_median-obs[Y])**2)/obs[Y]**2)
+        mae[Y] = (1./n) * sum(abs(i_median-obs[Y]))
+        mape[Y] = (1./n) * sum(abs((i_median-obs[Y])/obs[Y]))
+    return mse, mae, mape
 
 if __name__ == "__main__":
     sns.set(style="darkgrid", palette="Set2")

### 2015-02-04 draft_lm1.tex
improved letter and a bit of the discussion

@@
 reproductive number, \rr which we calculate from aggregated data.
-Our contribution is two-fold: (i) we show a method to quantify uncertainty 
+The main contribution of this paper can be summarized in the following items: 
+(i) we show a method to quantify uncertainty 
 about \rr that is readily applicable to other diseases and; (ii) we propose to 
 use \rr to inform a dynamic epidemic model with time-varying force of infection 
-in order to gain insight into the attack ratio of each epidemic.
+in order to gain insight into the attack ratio of each epidemic; (iii) We 
+propose an estimation procedure for circulating serotype's $S_0$ from 
+aggregate case data, which is robust to epidemic sizes; We estimate the AR 
+for 18 years of Dengue epidemics.
 
@@
 It is estimated that for every case reported, up to 10-20 are not seen by health 
-authorities (Oswaldo G. Cruz, personal communication).
+authorities~\cite{luz_disability_2009}.
 Another source of uncertainty is under-reporting of diagnosed cases, which is a 
@@
 This stresses the importance of knowing the immunological structure of the 
-population in order to truly assess the impact of any epidemic. [WHY?]
+population. Knowing the $S_0$ for the circulating viruses we can order to more 
+accurately assess the potential impact of a coming epidemic, since particularly 
+virulent types, can be rendered less of a threat by a low $S_0$.
 

### 2015-02-04 draft_lm1.tex
killing some white space
small edit to Table legend

@@
 credibility intervals, and therefore offers protection against false alarms.
-
 \begin{center}
@@
 Values are presented as percentage of total population. 
-$^\dag$: Year corresponds to the start of the epidemic. 
-Sometimes the peak of cases occur in the following year.
-$^\ddag$: Susceptible fraction.}
+$^\dag$: Year corresponds to the start of the epidemic, however the peak of cases
+may occur in the following year.
+$^\ddag$: Susceptible fraction.
+These results show considerable variation in AR between epidemics, consistent with
+the accquiring and loss of serotype-specific immunity.}
 \begin{center}

### 2015-02-12 Presentations/Dinamica de Vetores-SP-2015/dengueARS0.tex
Added presentation.

\documentclass[10pt,compress,notheorems]{beamer}

%\usepackage{ucs}
\usepackage[utf8]{inputenc}
\usepackage[dark]{beamerthemesidebar}

%\usepackage{babel}
\usepackage{fontenc}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{url}

\def \rr {$R_{t}\:$}

\author{Flávio Codeço Coelho and Luiz Max Carvalho}
\title{Estimating the Attack Ratio of Dengue Epidemics under Time-varying 
Force of Infection using Aggregated Notification Data}
\date{11-02-2015}
\institute [EMAp, FGV]{Escola de Matemática Aplicada,   Fundação Getúlio Vargas}

%\logo{\includegraphics[height=2cm]{./fb1.png}}
\begin{document}
 \begin{frame}
\titlepage
\end{frame}

\begin{frame}[fragile]
\frametitle{Summary}
\tableofcontents
\end{frame}

\section{Motivation}
\begin{frame}
\frametitle{Dengue Dynamics}
\begin{itemize}[<+->]
  \item Dengue is a Multi-Strain  vector-borne disease
  \item 4 major viral strains in circulation in Brazil
  \item Case-notification data is aggregated, i.e., does not discriminate 
serotype except for a handful of cases.
  \item It's a Seasonal disease, but recurrence pattern is hard to predict
  \item Vector population dynamics plays a major role in the modulation of 
incidence
  \item Imunological structure of the population is also a key factor, but is 
mostly unknown.
\end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{4 epidemics}
\begin{flushleft}
 \begin{figure}
\subfigure[2010]{\includegraphics[width=4.5cm, 
height=2.75cm]{./heatmap2010.jpg}}
  \subfigure[2011]{\includegraphics[width=4.5cm]{./heatmap2011.jpg}}
  \subfigure[2012]{\includegraphics[width=4.5cm]{./heatmap2012.jpg}}
  \subfigure[2013]{\includegraphics[width=4.5cm]{./heatmap2013.jpg}}
  \end{figure}
\end{flushleft}

\end{frame}

\section{Building blocks}

\subsection{Variable Force of Infection}
\begin{frame}
\frametitle{Effective Reproductive number(\rr)}
The effective reproductive number can be easily estimated from the incidence 
time-series, $Y_t$:
\begin{equation}
\label{eq:Rtestimate}
R_t = \left( \frac{Y_{t+1}}{Y_t}\right)^{1/n}
\end{equation}
Where $n$ is the ration between the length of reporting interval and 
the mean generation time of the disease.
\begin{flushright}
 Nishiura et. al. (2010)
\end{flushright}
\end{frame}

\begin{frame}
\frametitle{\rr's uncertainty}
But what about the uncertainty about \rr?

If we assume that the counts $Y_t$ are Poisson distributed for all $t$, we can 
derive the probability distribution of \rr\footnote{Coelho, FC and Carvalho, LM 
(Submitted)}
\begin{center}
\begin{figure}
  \includegraphics[width=9cm]{./rt_series.png}
\end{figure}
\end{center}

\end{frame}

\subsection{Vector dynamics}
\begin{frame}
\frametitle{Environmental determinants}
\begin{itemize}[<+->]
 \item A. Aegypti population dynamics display marked seasonality
 \item Temperature, Humidity and rainfall are important factors
 \item Environmental stock of eggs
 \item Effects on mosquito reproduction are non-linear
 \item Delayed influence
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{\rr \emph{vs.} Temperature}
\begin{columns}
 \column{5cm}
\includegraphics[width=5cm]{./rt_min_t.png}

\includegraphics[width=5cm]{./rt_t24.png}
 \column{6cm}
\includegraphics[width=5cm]{./transmissao_temp.png}
\end{columns}
\end{frame}

\section{Modeling Dengue}
\subsection{Single-strain model}
\begin{frame}
\frametitle{Single Strain SIR}
Why not multi-strain? No Multi-strain data!!
\begin{align}
   \label{eq:model}
 \frac{dS}{dt} &= -\beta(t)SI \\     \nonumber
 \frac{dI}{dt} &= \beta(t)SI - \tau I&\\      \nonumber
 \frac{dR}{dt} &= \tau I&
\end{align}  
where $S(t) + I(t) + R(t) = 1 \: \forall\: t$. % \in [T_0, T_1]$.

\end{frame}
\subsection{Variable Force of Infection}
\begin{frame}
\frametitle{Variable Force of Infection}
From \rr, we can define a force of infection which varies with time:
\begin{equation} 
 \label{eq:effbeta}
 \beta(t) = \frac{R_t\cdot\tau}{S}
\end{equation}
But how do we get the value of $S$? we need to estimate $S_0$.

\end{frame}

\section{Parameter estimation}
\subsection{Estimating $S_0$}
\begin{frame}
\frametitle{Estimating $S_0$}
Bayesian framework:
\begin{itemize}[<+->]
 \item Define priors for $S_0$ in the range (0,1)
 \item Samples from prior, calculate $\beta(t)$ and run the model
 \item calculate Likelihood of data given current parameterization
 \item Determine posterior probability of parameterization
\end{itemize}
\begin{equation}
 \label{eq:S0post}
 p(S_{0j}|\mathbf{Y_{j}}) \propto L(\mathbf{Y_{j}}|S_{0j}, R_t, m, \tau 
)\pi(S_{0j}) 
\end{equation}
\end{frame}

\begin{frame}[fragile]
\frametitle{Models \textit{vs} Data}
fiting the model to data (Rio de janeiro) to estimate $S_0$\footnote{Coelho 
FC et al., 2011}.

 \includegraphics[width=\textwidth]{concat_SI.png}

 \begin{small}Posterior distribution for Susceptible (S) and infectious (I) 
individuous. Blue dots are data. \end{small}

\end{frame}

\subsection{Attack Ratio}
\begin{frame}
\frametitle{Attack Ratio}
Once we have $S_0$, we can caculate the attack ratio:
\begin{equation}
\label{eq:AR2}
 A_{j}=\frac{\sum Y_j}{S_{0j}}
\end{equation}
\end{frame}

\begin{frame}
\frametitle{Attack ratio}
\begin{table}[!ht]
\begin{tiny}\caption{
{\bf Median attack ratio and 95\% credibility intervals calculated according to 
(\ref{eq:AR2})}. 
Values are presented as percentage of total population. 
$^\dag$: Year corresponds to the start of the epidemic, however the peak of 
cases
may occur in the following year.
$^\ddag$: Susceptible fraction.
These results show considerable variation in AR between epidemics, consistent 
with
the accquiring and loss of serotype-specific immunity.}                          
                             \end{tiny}
\begin{center}
\begin{footnotesize}\begin{tabular}{c|c|c}
\hline
Year$^\dag$ & median Attack Ratio & $S_0^\ddag$ \\
\hline
1996 & 0.39 (0.17-0.54) & 0.00171(0.0012-0.0038)\\
1997 & 0.87 (0.74-0.87) & 0.00273(0.0027-0.0032)\\
1998 & 0.5 (0.49-0.5) & 0.00142(0.0014-0.0014)\\
1999 & 0.11 (0.037-0.2) & 0.00345(0.0018-0.01)\\
2000 & 0.25 (0.24-0.27) & 0.0155(0.015-0.016)\\
2001 & 0.48 (0.47-0.49) & 0.0495(0.048-0.051)\\
2005 & 0.15 (0.1-0.21) & 0.0147(0.01-0.021)\\
2006 & 0.11 (0.08-0.14) & 0.0281(0.022-0.037)\\
2007 & 0.15 (0.15-0.15) & 0.135(0.13-0.14)\\
2008 & 0.14 (0.031-0.31) & 0.00672(0.003-0.024)\\
2010 & 0.18 (0.17-0.19) & 0.0454(0.043-0.048)\\
2011 & 0.086 (0.082-0.094) & 0.215(0.2-0.23)\\
2012 & 0.14 (0.13-0.15) & 0.0621(0.058-0.068)\\
\hline
\end{tabular}\end{footnotesize}
\end{center}
\label{tab:AR}
\end{table}
\end{frame}
\begin{frame}[fragile]
\frametitle{References}
\begin{thebibliography}{10}
 \bibitem{nishiura}
Nishiura H, Chowell G, Heesterbeek H, Wallinga J (2010) {{T}he ideal reporting
  interval for an epidemic to objectively interpret the epidemiological time
  course}.
\newblock J R Soc Interface 7: 297--307.
%\bibAnnoteFile{nishiura}

\bibitem{pone2011}
Coelho FC, Code\c{c}o CT, Gomes MG (2011) {A} {B}ayesian framework for
  parameter estimation in dynamical models.
\newblock PLoS ONE 6: e19616.

\bibitem{pone2011}
Coelho FC, Code\c{c}o CT, Carvalho, LM {E}stimating the Attack Ratio of Dengue 
Epidemics under Time-varying Force of Infection using Aggregated Notification 
Data
\newblock Arxiv. \url{http://arxiv.org/abs/1502.01236}
%\bibAnnoteFile{pone2011}
\end{thebibliography}

\end{frame}
\end{document}

### 2015-02-12 draft_lm1.tex
Added presentation.

@@
 propose an estimation procedure for circulating serotype's $S_0$ from 
-aggregate case data, which is robust to epidemic sizes; We estimate the AR 
+aggregate case data, which is robust to epidemic sizes; (iv) We estimate the AR 
 for 18 years of Dengue epidemics.

### 2015-02-12 Presentations/Dinamica de Vetores-SP-2015/dengueARS0.tex
fixed typo in the references

@@
 \bibitem{pone2011}
-Coelho FC, Code\c{c}o CT, Carvalho, LM {E}stimating the Attack Ratio of Dengue 
+Coelho FC, Carvalho, LM {E}stimating the Attack Ratio of Dengue 
 Epidemics under Time-varying Force of Infection using Aggregated Notification 

### 2015-02-19 draft_lm1.tex
joined the maps into a single figure

@@
 % cite package, to clean up citations in the main text. Do not remove.
-\usepackage{cite}
+\usepackage[numbers]{natbib}
 \usepackage{hyperref}
@@
 % Use the PLoS provided BiBTeX style
-\bibliographystyle{plos2009}
+\bibliographystyle{naturemag}
 % Remove brackets from numbering in List of References
@@
 serotypes which causes an estimated  $50$ million infections each 
-year~\cite{Guzman2010}.
+year~\citep{Guzman2010}.
 In humans, immunity against  a particular serotype is  considered permanent 
 after the exposure and cross immunity to there serotypes is considered short 
-lived~\cite{halstead_dengue_2007}.
+lived~\citep{halstead_dengue_2007}.
 As a consequence, the proportions of viral serotypes co-circulating at any 
@@
 vector reproduction, stands out as a strong predictor of 
-incidence~\cite{honorio_temporal_2009,wu_higher_2009}.
+incidence~\citep{honorio_temporal_2009,wu_higher_2009}.
 In places with sufficient seasonal temperature variation, dengue is 
@@
 This climatic influence is exerted mainly through its effects on the force of 
-infection, which cannot be taken as constant~\cite{reiner_time-varying_2014} 
+infection, which cannot be taken as constant~\citep{reiner_time-varying_2014} 
 but rather as a seasonal (oscillating) function of time.
@@
 The attack ratio is also influenced by the reproductive number of the disease 
-~\cite{bacaer_final_2009, katriel_attack_2012}, which is closely associated 
+~\citep{bacaer_final_2009, katriel_attack_2012}, which is closely associated 
 with the force of infection.
@@
 accommodating time-varying force of infection have been proposed 
-before, for measles~\cite{bjornstad_dynamics_2002, 
+before, for measles~\citep{bjornstad_dynamics_2002, 
 wallinga_reconstruction_2003}, a disease that shows remarkable seasonality.
@@
 Recently, methods in the same fashion were developed for dengue when 
-serotype-specific data is available~\cite{Reiner2014}.
+serotype-specific data is available~\citep{Reiner2014}.
 When such data is not available, the series of susceptibles to all possible
@@
 propose to inform the time-varying transmissibility using the \rr series 
-derived from the notification data~\cite{nishiura}.
+derived from the notification data~\citep{nishiura}.
 We extend a Bayesian framework previously used to estimate the number 
-of susceptibles in Influenza epidemics in Europe~\cite{pone2011} to include 
+of susceptibles in Influenza epidemics in Europe~\citep{pone2011} to include 
 time-varying force of infection and derive a probability distribution  for 
@@
 (0, T)$.
-Nishiura el al. (2010)~\cite{nishiura} extend the theory developed by 
-Stallybrass et al. (1931)~\cite{stallybrass} and propose to estimate \rr as
+Nishiura el al. (2010)~\citep{nishiura} extend the theory developed by 
+Stallybrass et al. (1931)~\citep{stallybrass} and propose to estimate \rr as
 \begin{equation}
@@
 
-We explore the approach of Ederer and Mantel~\cite{mantel}, whose objective is 
+We explore the approach of Ederer and Mantel~\citep{mantel}, whose objective is 
 to obtain 
@@
 $Pr(R_t^{(L)}<R_t<R_t^{(U)}) = \alpha$.
-Ederer and Mantel (1974)~\cite{mantel} show that one can construct a $100\alpha 
+Ederer and Mantel (1974)~\citep{mantel} show that one can construct a 
+$100\alpha 
 \%$ confidence interval for \rr by noting that
@@
 Many authors have chosen to quantify the uncertainty about $\theta$ 
-following orthodox approaches  (see for example~\cite{wilson} 
-and~\cite{clopper}) mainly for simplicity.
+following orthodox approaches  (see for example~\citep{wilson} 
+and~\citep{clopper}) mainly for simplicity.
 We choose instead to take a Bayesian approach and use the  $100\alpha \%$ 
@@
 a beta prime (or inverted beta) with parameters $ a_1 = Y_{t+1} + a_0$ and $b_1 
-=  Y_t + b_0$~\cite{dubey1970}.
+=  Y_t + b_0$~\citep{dubey1970}.
 The density of the induced distribution is then 
@@
 Carlo techniques implemented in the Bayesian inference with Python 
-(BIP)~\cite{pone2011} available at  
+(BIP)~\citep{pone2011} available at  
 \url{http://code.google.com/p/bayesian-inference/}.
-BIP uses a Differential Evolution Adaptive Metropolis (DREAM)~\cite{vrugt2008} 
+BIP uses a Differential Evolution Adaptive Metropolis (DREAM)~\citep{vrugt2008} 
 scheme that efficiently samples from high-dimensional joint distributions using 
@@
 calculation of the Gelman-Rubin's R (potential scale 
-reduction factor), which approaches $1$ at convergence~\cite{brooks1998}.
+reduction factor), which approaches $1$ at convergence~\citep{brooks1998}.
 
@@
 Figure~\ref{fig:rtseries}  shows the $R_t$ series, according to 
-(\ref{eq:Rtestimate})~\cite{nishiura} along with the confidence bands derived 
+(\ref{eq:Rtestimate})~\citep{nishiura} along with the confidence bands derived 
 in this paper. 
@@
 Methods for estimating the  number of susceptibles have been proposed 
-before, for other diseases~\cite{bjornstad_dynamics_2002, 
+before, for other diseases~\citep{bjornstad_dynamics_2002, 
 wallinga_reconstruction_2003}.
@@
 predominant serotypes for some epidemics in the period of 
-study only~\cite{macedo_virological_2013}, we propose the use of a 
+study only~\citep{macedo_virological_2013}, we propose the use of a 
 simplified a single strain model.
@@
 It is estimated that for every case reported, up to 10-20 are not seen by health 
-authorities~\cite{luz_disability_2009}.
+authorities~\citep{luz_disability_2009}.
 Another source of uncertainty is under-reporting of diagnosed cases, which is a 
@@
 Brazil. 
-Duarte and Fran\c{c}a (2006)~\cite{duarte_data_2006}, estimated 
+Duarte and Fran\c{c}a (2006)~\citep{duarte_data_2006}, estimated 
 the sensitivity of Dengue reporting for hospitalized patients in 
@@
 There are a number of approaches to deriving the distribution of \rr.
-Alternatively to the approach described in the main text~\cite{mantel}, one 
+Alternatively to the approach described in the main text~\citep{mantel}, one 
 could use the conditional distribution of \rr on 
 $Y_{t+1}$ and $Y_t$ as defined in equation A7 of Nishiura et 
-al.~\cite{nishiura}:
+al.~\citep{nishiura}:
 \begin{equation}
@@
 The beta prime (inverse beta distribution) will have heavier tails compared to 
-the conditional distribution proposed by~\cite{nishiura}, thus providing more 
+the conditional distribution proposed by~\citep{nishiura}, thus providing more 
 conservative confidence/credibility intervals.
@@
 As a side note, the Bayesian approach presented in this 
-paper will give similar results to orthodox confidence intervals~\cite{wilson} 
-and~\cite{clopper} for $Y_{t+1}$ and $Y_t >> 1$.
+paper will give similar results to orthodox confidence intervals~\citep{wilson} 
+and~\citep{clopper} for $Y_{t+1}$ and $Y_t >> 1$.
 Under the flat  uniform prior for $\theta_t$, the Bayesian posterior 
@@
 interval is nearly indistinguishable from the confidence interval proposed by 
-Clopper \& Pearson (1931)~\cite{clopper} for $Y_{t+1}, Y_t > 20$.
+Clopper \& Pearson (1931)~\citep{clopper} for $Y_{t+1}, Y_t > 20$.
 Note also that the uniform prior ($Beta(1, 1)$) for $\theta_t$ constitutes a 

### 2015-02-19 suppl.tex
recreated the supplementary material

--- 2015-02-04 draft_lm1.tex
@@
-% Template for PLoS
-% Version 2.0 July 2014
-%
-% To compile to pdf, run:
-% latex plos.template
-% bibtex plos.template
-% latex plos.template
-% latex plos.template
-% dvipdf plos.template
-%
-% % % % % % % % % % % % % % % % % % % % % %
-%
-\documentclass[10pt]{article}
-\usepackage{todonotes}
-\usepackage{subfigure}
+\documentclass[a4paper,10pt]{article}
+\usepackage[utf8]{inputenc}
+\usepackage[numbers]{natbib}
+\bibliographystyle{naturemag}
 % amsmath package, useful for mathematical formulas
@@
 \usepackage{amssymb}
-% cite package, to clean up citations in the main text. Do not remove.
-\usepackage{cite}
-\usepackage{hyperref}
-% line numbers
-\usepackage{lineno}
-% ligatures disabled
-\usepackage{microtype}
-\usepackage{indentfirst}
-\DisableLigatures[f]{encoding = *, family = * }
-% rotating package for sideways tables
-%\usepackage{rotating}
-% If you wish to include algorithms, please use one of the packages below.
-Also, please see the algorithm section of our LaTeX guidelines (http://www.plosone.org/static/latexGuidelines) for important information about required formatting.
-%\usepackage{algorithmic}
-%\usepackage{algorithmicx}
-% Use doublespacing - comment out for single spacing
-%\usepackage{setspace} 
-%\doublespacing
-\topmargin 0.0cm
-\oddsidemargin 0.5cm
-\evensidemargin 0.5cm
-\textwidth 16cm 
-\textheight 21cm
-% Bold the 'Figure #' in the caption and separate it with a period
-% Captions will be left justified
-\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}
-% Use the PLoS provided BiBTeX style
-\bibliographystyle{plos2009}
-% Remove brackets from numbering in List of References
-\makeatletter
-\renewcommand{\@biblabel}[1]{\quad#1.}
-\makeatother
-
+\usepackage{url}
 % Our defs 
 \def \rr {$R_{t}\:$}
-
-% Leave date blank
-\date{}
-
-\pagestyle{myheadings}
-
-%% Include all macros below. Please limit the use of macros.
-
-%% END MACROS SECTION
+%opening
+\title{Supplementary material to ``Estimating the Attack Ratio of Dengue 
+Epidemics under Time-varying 
+Force of Infection using Aggregated Notification Data''}
+\author{Flavio Code\c{c}o Coelho and Luiz Max de Carvalho}
 
@@
 
-% Title must be 150 characters or less
-\begin{flushleft}
-{\Large
-\textbf{Estimating the Attack Ratio of Dengue Epidemics under Time-varying 
-Force of Infection using Aggregated Notification Data}
-}
-\\
-Flavio Code\c{c}o Coelho$^{1\ast}$, 
-Luiz Max de Carvalho$^{2}$, 
-\\
-\bf{1} Escola de Matem\'atica Aplicada , Funda\c{c}\~ao Getulio Vargas (FGV), 
-Rio de Janeiro -- RJ, Brazil.
-\\
-\bf{2} Programa de Computa\c{c}\~ao Cient\'ifica (PROCC), Funda\c{c}\~ao Oswaldo 
-Cruz, Rio de Janeiro -- RJ, Brazil.
-\\
-$\ast$ E-mail: fccoelho@fgv.br
-\end{flushleft}
+\maketitle
 
-\section*{Abstract}
-
-Quantifying the attack ratio of disease is key to epidemiological inference and 
-Public Health planning.
-For multi-serotype pathogens, however, different levels of 
-serotype-specific immunity make it difficult to assess the population at risk. 
-In this paper we propose a Bayesian method for estimation of the attack ratio 
-of an epidemic and the initial fraction of susceptibles using aggregated 
-incidence data. 
-We derive the probability distribution of the effective reproductive number, 
-\rr, and use MCMC to obtain posterior 
-distributions of the parameters of a single-strain SIR transmission model with 
-time-varying force of infection.
-Our method is showcased in a data set consisting of 18 years of dengue 
-incidence in the city of Rio de Janeiro, Brazil.
-We demonstrate that it is possible to learn about the initial fraction of 
-susceptibles and the attack ratio even in the absence of serotype specific data.
-On the other hand, the information provided by this approach is limited, 
-stressing the need for detailed serological surveys to characterise the 
-distribution of serotype-specific immunity in the population.
-
-\section*{Introduction}
-
-Dengue is an arthropod-borne febrile disease caused by a flavivirus with four 
-serotypes which causes an estimated  $50$ million infections each 
-year~\cite{Guzman2010}.
-In humans, immunity against  a particular serotype is  considered permanent 
-after the exposure and cross immunity to there serotypes is considered short 
-lived~\cite{halstead_dengue_2007}.
-As a consequence, the proportions of viral serotypes co-circulating at any 
-point in time are strongly dependent on previous incidence patterns of the 
-disease, which determine the number of individuals susceptible to each serotype 
-at any point in time.
-% LM: Maybe a citation here?
-
-Dengue transmission is also modulated by environmental conditions, among which, 
-temperature, due to its effects on the 
-vector reproduction, stands out as a strong predictor of 
-incidence~\cite{honorio_temporal_2009,wu_higher_2009}.
-In places with sufficient seasonal temperature variation, dengue is 
-predominantly a  summer disease.
-So it is fair to say that these environmental 
-fluctuations play a key role in determining beginning and end of epidemic 
-periods.
-This climatic influence is exerted mainly through its effects on the force of 
-infection, which cannot be taken as constant~\cite{reiner_time-varying_2014} 
-but rather as a seasonal (oscillating) function of time.
-The long term dynamics of dengue is also modulated by the alternation of virus 
-types in circulation. Demographics also plays a role in replenishing the 
-population of susceptibles.
-
-The attack ratio (AR) of a disease is a measure of morbidity defined as the 
-number of new cases divided by the population at risk.
-% LM: Maybe a citation here? It's ridiculous I know, but referees can be a pain.
-For dengue epidemics, it can be difficult to calculate the AR due 
-to the lack of knowledge of the population at risk.
-The population at risk in this case is the number of susceptibles 
-to the circulating virus type(s) before a given epidemic. 
-Thus, in order to calculate the attack ratio, we need to determine the number 
-of susceptibles to the circulating virus types right before the epidemic, which 
-is virtually impossible without regular virological surveys.
-
-The attack ratio is also influenced by the reproductive number of the disease 
-~\cite{bacaer_final_2009, katriel_attack_2012}, which is closely associated 
-with the force of infection.
-Thus the incorporation of the effective reproduction number, \rr, as a function 
-of time, is crucial to an accurate estimation of the AR of seasonal diseases 
-like dengue and Influenza.
-
-Other methods for estimating the  number of susceptibles while 
-accommodating time-varying force of infection have been proposed 
-before, for measles~\cite{bjornstad_dynamics_2002, 
-wallinga_reconstruction_2003}, a disease that shows remarkable seasonality.
-These methods try to reconstruct the entire 
-series of infectious and susceptibles from case data using deterministic models 
-and generally work well for measles because there is a one-to-one 
-relationship between exposure and immunity, since measles is caused by a 
-single-strain pathogen.
-Recently, methods in the same fashion were developed for dengue when 
-serotype-specific data is available~\cite{Reiner2014}.
-When such data is not available, the series of susceptibles to all possible
-serotypes, cannot be reconstructed based solely on a deterministic transmission 
-model, since the arrival/re-emergence of  new serotypes, an intrinsically 
-stochastic events, can drastically change the pool of susceptibles, throwing off 
-any sequential estimation based on the  incidence dynamics.
-
-In this paper, we propose a new approach to estimate the number (fraction) of 
-susceptibles using a simplified model of dengue transmission based on a 
-single-strain Susceptible-Infectious-Removed (SIR) model with time-varying 
-infection rate.
-In order to bypass the limitations of not knowing the serotype-specific 
-seroprevalence and the exact behaviour of the force of infection through time, 
-we
-propose to inform the time-varying transmissibility using the \rr series 
-derived from the notification data~\cite{nishiura}.
-We extend a Bayesian framework previously used to estimate the number 
-of susceptibles in Influenza epidemics in Europe~\cite{pone2011} to include 
-time-varying force of infection and derive a probability distribution  for 
-\rr to accommodate uncertainty in the estimates. 
-Then, from the incidence series and the population at risk, we  calculate the 
-attack ratio for each epidemic.
-We apply our method to estimate $S_0$ before every major dengue epidemic in the 
-city of Rio de Janeiro, Brazil in the last 18 years.
-
-\section*{Methods}
-
-In this section we will start by describing the data and then the method used 
-to estimate the effective reproductive number, $R_t$, from the data and obtain 
-its posterior distribution. 
-We then proceed to describe the Susceptible-Infectious-Recovered (SIR) model 
-used to represent the aggregated disease incidence and how $R_t$ can be 
-integrated into the model to allow for time varying force of infection. 
-Next, an approach to approximate the posterior distributions 
-of the numbers of susceptible to the main circulating dengue viruses for each 
-epidemic is detailed.
-Finally, we discuss how to estimate the attack ratio of each 
-epidemic using the estimated susceptible fraction and the observed incidence.
-
-\subsection*{Data}
-
-The data used in this paper consists of time 
-series of weekly notified cases of dengue for the 
-city of Rio de Janeiro from 1996 to 2014. The cases are notified 
-based only on clinical criteria.
-Laboratory confirmation and serotype 
-information are available only for a very small sample and only on recent 
-years (2010-2013).
-For the parameter estimation  procedures incidence was normalized by dividing 
-the number of cases reported by the total city population at each year as given 
-by the census (Census Bureau, Brazilian Institute of Geography and Statistics, 
-\url{http://www.ibge.gov.br/english/}).
-
-\subsection*{Estimating the effective reproductive number ($R_t$)}
-
-In monitoring of infectious diseases, it is important to assess whether the 
-incidence of a  particular disease is increasing significantly, in order to 
-decide to take preventive measures.
-The effective reproductive number at time $t$, \rr, can be understood as a 
-real-time estimate of the basic reproductive number ($R_{0}$) and is defined as 
-the average number of secondary cases per primary case at time $t$.
-
-Let $Y_t$ be the number of reported disease cases for a particular time $t \in 
-(0, T)$.
-Nishiura el al. (2010)~\cite{nishiura} extend the theory developed by 
-Stallybrass et al. (1931)~\cite{stallybrass} and propose to estimate \rr as
-\begin{equation}
-\label{eq:Rtestimate}
-R_t = \left( \frac{Y_{t+1}}{Y_t}\right)^{1/n}
-\end{equation}
-where $n$ is taken to be the ratio between the length of reporting interval and 
-the mean generation time of the disease.
-Here we are interested in the simpler case $n=1$.
-If \rr is to be used as a decision tool, however, one needs to be able to 
-quantify the 
-uncertainty about estimate in equation~\ref{eq:Rtestimate}. 
-Here we detail how to obtain credibility intervals for \rr under the assumption 
-that the counts $Y_t$ are Poisson distributed for all $t$.
-
-We explore the approach of Ederer and Mantel~\cite{mantel}, whose objective is 
-to obtain 
-confidence intervals for the ratio of two Poisson counts. 
-Let $Y_{t} \sim Poisson(\lambda_t)$ and $Y_{t+1} \sim Poisson(\lambda_{t+1})$ 
-and define $S = Y_{t} + Y_{t+1}$.
-The authors note that by conditioning on the sum $S$
-\begin{align}
-\label{eq:binlike}
-Y_{t+1} | S &\sim Binomial(S, \theta_t) \\
-\theta_t &= \frac{\lambda_{t+1}}{\lambda_{t} + \lambda_{t+1}}
-\end{align}
-Let $c_{\alpha}(\theta_t) = \{\theta_t^{(L)} , \theta_t^{(U)} \}$ be such that 
-$Pr(\theta_t^{(L)}<\theta_t <\theta_t^{(U)}) = \alpha$.
-Analogously, define $c_{\alpha}(R_t) = \{R_t^{(L)} , R_t^{(U)} \}$ such that 
-$Pr(R_t^{(L)}<R_t<R_t^{(U)}) = \alpha$.
-Ederer and Mantel (1974)~\cite{mantel} show that one can construct a $100\alpha 
-\%$ confidence interval for \rr by noting that
-\begin{equation}
-\label{eq:confRt}
- R_t^{(L)} = \frac{\theta_t^{(L)}}{(1-\theta_t^{(L)})} \quad \text{and} \quad 
-R_t^{(U)} = \frac{\theta_t^{(U)}}{(1-\theta_t^{(U)})}\\
-\end{equation}
-Because the transform from $\theta$ to \rr is monotonically 
-increasing, the result holds for confidence and credibility intervals alike.
-
-Many authors have chosen to quantify the uncertainty about $\theta$ 
-following orthodox approaches  (see for example~\cite{wilson} 
-and~\cite{clopper}) mainly for simplicity.
-We choose instead to take a Bayesian approach and use the  $100\alpha \%$ 
-posterior credibility interval for $\theta_t$ as $c_{\alpha}(\theta_t)$.
-If we choose the conjugate beta prior with parameters $a_0$ and $b_0$ for the 
-binomial likelihood in (\ref{eq:binlike}), the posterior distribution for 
-$\theta_t$ is
-\begin{equation}
-\label{eq:thetapost}
-p(\theta_t| Y_{t+1}, S) \sim Beta(Y_{t+1} + a_0, Y_t + b_0)
-\end{equation}
-Combining equations~(\ref{eq:confRt}) and~(\ref{eq:thetapost}) 
-tells us that the induced posterior distribution of $R_t$ is 
-a beta prime (or inverted beta) with parameters $ a_1 = Y_{t+1} + a_0$ and $b_1 
-=  Y_t + b_0$~\cite{dubey1970}.
-The density of the induced distribution is then 
-\begin{equation}
-\label{eq:densityMantel}
-f_P(R_t| a_1, b_1) = \frac{\Gamma(a_1 + b_1)}{\Gamma(a_1)\Gamma(b_1)} R_t^{a_1 
-- 
-1} (1 + R_t)^{-(a_1 + b_1)}
-\end{equation}
-Thus, the expectation of \rr is $a_1/(b_1 - 1)$ and its variance is 
-$a_1(a_1 + b_1 - 1)/\left((b_1 - 2)(b_1 - 1)^2 \right) $.
-Note that this result holds only for $n = 1$.
-Sampling from the posterior in~(\ref{eq:densityMantel}) can be made 
-straightforward by first sampling from~(\ref{eq:thetapost}) and then applying 
-the transform in~(\ref{eq:confRt}).
-Also, one can choose $a_0$ and $b_0$ so as to elicit meaningful prior 
-distributions for $R_t$.
-We show how to elicit the prior for $R_t$ from specified prior mean and variance
-or coefficient of variation in the Appendix.
-
-Also, since $R_t > 1$ indicates sustained transmission, one may be 
-interested in computing the probability of this event.
-This can be easily achieved by integrating~(\ref{eq:densityMantel}) over the 
-appropriate interval.
-By noting that
-\begin{align}
-\label{cumprobMantel}
-Pr(R_t > 1) &= 1 - \int_0^1 f_P(r)dr \\
-            &= 1- Pr(\theta_t < \frac{1}{2})
-\end{align}
-one can compute the desired probability while avoiding dealing with the density 
-in~(\ref{eq:densityMantel}) directly.
-
-\subsection*{Mathematical modelling} % The model
-
-A Susceptible-Infectious-Removed (SIR) model is proposed to model dengue 
-dynamics.
-In the traditional formulation of the model, transmission is governed by a 
-constant transmission rate $\beta$ and recovery happens at a rate $\tau$.
-
-For our analysis we chose to let the force of infection vary with time, just 
-as it does in the actual epidemics, as seen in the data. So as the epidemic 
-progresses, the effective transmission  rate changes and is 
-given by 
-\begin{equation} 
- \label{eq:effbeta}
- \beta(t) = \frac{R_t\cdot\tau}{S_0}
-\end{equation}
-where $R_t$ is the effective reproductive number, estimated as 
-in~\ref{eq:Rtestimate}.
-The complete model with the time-varying force of infection is given by
-the system of ordinary differential equations:
-\begin{align}
-   \label{eq:model}
- \frac{dS}{dt} &= -\beta(t)SI \\     \nonumber
- \frac{dI}{dt} &= \beta(t)SI - \tau I&\\      \nonumber
- \frac{dR}{dt} &= \tau I&
-\end{align}  
-where $S + I + R = 1 \: \forall\: t$. % \in [T_0, T_1]$. 
-Of course, this is a rather simplified model, in which, for instance, the 
-vector is omitted.
-The rationale for this simplification is based on the ability of the 
-empirically derived $R_t$  to incorporate the effects of the fluctuating vector 
-populations.
-Also, although there are multiple circulating serotypes, our approach
-can not discriminate between them due to the lack of serotype-specific data.
-Nevertheless, this modelling strategy can still provide some insight into the 
-disease dynamics and allows us to estimate the initial fraction of susceptibles 
-$S_0$, a key epidemiological parameter.
-
-\subsection*{Bayesian parameter estimation}
-
-We take a Bayesian approach to the estimation of $S_0$.
-First the incidence time series was divided into $J=13$ 
-epidemic windows that corresponded to significant raises in incidence and 
-normalized to lie on the $[0,1]$ interval.
-For a given interval $j = \{ t_j^{\text{start}}, t_j^{\text{end}} \} $ we 
-observe an incidence time series $\mathbf{Y_{j}}$.
-We are thus interested in the posterior distribution
-\begin{equation}
- \label{eq:S0post}
- p(S_{0j}|\mathbf{Y_{j}}) \propto l(\mathbf{Y_{j}}|S_{0j}, R_t, m, \tau 
-)\pi(S_{0j}) 
-\end{equation}
-The likelihood $l(\mathbf{Y_{j}}|\cdot)$ is assumed to be a Normal distribution 
-with fixed variance $\sigma^2$.
-In this estimation procedure we kept $R_t$ fixed at fixed at the posterior mean 
-obtained as described above and fixed $\tau = 1/7\: \text{days}^{-1}$.
-To complete the inference, we need to specify prior distributions for the 
-parameters of interest.
-We place a flat $\text{Beta}(1, 1)$ prior on $S_{0j}\:\forall j$
-
-To approximate the posterior in~(\ref{eq:S0post}) we use Markov chain Monte 
-Carlo techniques implemented in the Bayesian inference with Python 
-(BIP)~\cite{pone2011} available at  
-\url{http://code.google.com/p/bayesian-inference/}.
-BIP uses a Differential Evolution Adaptive Metropolis (DREAM)~\cite{vrugt2008} 
-scheme that efficiently samples from high-dimensional joint distributions using 
-multiple adaptive chains running in parallel with delayed rejection.
-Also, as the numerical integration routine implemented within BIP needs 
-$\beta(t)$ to be available at 
-arbitrary values of $t$, i.e., as continuous function of time because of the 
-variable step size, we used linear interpolation to obtain values of $R_t$ for 
-any time point.
-In this study we used one chain per parameter, i.e, 3 chains for each run.
-The chains were run until $5000$ samples were obtained after discarding $500$ 
-burn-in samples.
-Convergence of the parallel chains was verified at every 100 iterations by the 
-calculation of the Gelman-Rubin's R (potential scale 
-reduction factor), which approaches $1$ at convergence~\cite{brooks1998}.
-
-\subsection*{Calculating the attack ratio}
-
-The attack ratio of an epidemic is defined by the number of infections divided 
-by the size of the population at risk.
-
-\begin{equation}
-\label{eq:AR}
-A=\frac{\text{\# cases}}{\text{Population at risk}} 
-\end{equation}
-
-Based on what has been discussed so far, we can rewrite (\ref{eq:AR}) for 
-each epidemic $j$ as
-\begin{equation}
-\label{eq:AR2}
- A_{j}=\frac{\sum Y_j}{S_{0j}}
-\end{equation}
-where $S_{0j}$ is the number of susceptibles before each epidemic $j$, which we
-estimated before.
-
-Python and R code to perform all the analyses described above is publicly 
-available at~\url{https://github.com/fccoelho/paperLM1}.
-
-\section*{Results and Discussion}
-
-In this paper we propose a method to bypass the lack of serotype-specific case 
-data by informing the time-varying force of infection with the instantaneous 
-reproductive number, \rr which we calculate from aggregated data.
-The main contribution of this paper can be summarized in the following items: 
-(i) we show a method to quantify uncertainty 
-about \rr that is readily applicable to other diseases and; (ii) we propose to 
-use \rr to inform a dynamic epidemic model with time-varying force of infection 
-in order to gain insight into the attack ratio of each epidemic; (iii) We 
-propose an estimation procedure for circulating serotype's $S_0$ from 
-aggregate case data, which is robust to epidemic sizes; We estimate the AR 
-for 18 years of Dengue epidemics.
-
-Figure~\ref{fig:rtseries}  shows the $R_t$ series, according to 
-(\ref{eq:Rtestimate})~\cite{nishiura} along with the confidence bands derived 
-in this paper. 
-It can be seen that the inter-epidemic periods are characterized by $R_t$ being 
-indistinguishable from 1.
-Due to the intrinsic variability of the \rr series, the examination of its 
-credible intervals is essential to identify periods of sustained transmission.
-The wider intervals between epidemics are due to the scarcity of cases during 
-these periods.
-The method to quantify uncertainty proposed here provides more conservative 
-credibility intervals, and therefore offers protection against false alarms.
-\begin{center}
-\begin{figure}[!h]
- \centering
- \includegraphics[width=16cm]{./plots/rt_series.png}
- \caption{{\bf Estimated time-series for $R_t$, along with 95\% credible 
-intervals.} Top panel show reported cases from which $R_t$ is estimated.
-As expected, uncertainty about \rr is greater when the case counts are low, for 
-instance in the period $2003$--$2006$, which represented a big hiatus between 
-major epidemics.
-The intrinsic variability of \rr can be used to inform the time-varying force 
-of infection, since it reflects variation in the vector population and other 
-environmental factors such as temperature and seasonal variation.}
-\label{fig:rtseries}
-\end{figure}
-\end{center}
-
-A key epidemiological quantity is the attack ratio (AR) of an epidemic, a 
-measure of morbidity and speed of spread which can be used to predict epidemic 
-size and help efficient Public Health planning.
-The AR depends fundamentally on the population at risk, which in the case of 
-dengue is every naive (to a particular serotype) individual in the population.
-Estimating the initial susceptible fraction $S_0$ for each epidemic is thus 
-central to the estimation of the AR.
-Methods for estimating the  number of susceptibles have been proposed 
-before, for other diseases~\cite{bjornstad_dynamics_2002, 
-wallinga_reconstruction_2003}.
-These methods try to reconstruct the entire 
-series of infectious and susceptibles for measles 
-outbreaks from case data.
-In the case of dengue, the full (multi-year/multi-epidemic)
-series of susceptibles to all possible serotypes, cannot be reconstructed based 
-solely on a deterministic transmission model, since the arrival/re-emergence of 
-new serotypes (which are a stochastic events) can change drastically the pool 
-of susceptibles throwing off any sequential estimation based on the incidence
-dynamics.
-
-Since there is very limited information regarding the actual proportions of 
-each virus in circulation and most information available is about the 
-predominant serotypes for some epidemics in the period of 
-study only~\cite{macedo_virological_2013}, we propose the use of a 
-simplified a single strain model.
-The main argument we put forward is that by conditioning on the \rr series, we 
-implicitly take into account the variability introduced by the 
-co-circulation of multiple serotypes and heterogeneous levels of immunity in 
-the general population.
-We sought to deal with all important sources of uncertainty impinging on the 
-estimation of the AR of a dengue epidemic, but not all could be satisfactorily 
-addressed in this analysis.
-For instance, in any given epidemic there is a large number of 
-mild and asymptomatic cases, which nevertheless acquire immunity.
-It is estimated that for every case reported, up to 10-20 are not seen by health 
-authorities~\cite{luz_disability_2009}.
-Another source of uncertainty is under-reporting of diagnosed cases, which is a 
-serious issue in the health care systems of many developing countries such as 
-Brazil. 
-Duarte and Fran\c{c}a (2006)~\cite{duarte_data_2006}, estimated 
-the sensitivity of Dengue reporting for hospitalized patients in 
-Belo-Horizonte, Brazil to be of 63\%, meaning that approximately 37\% of the 
-suspected Dengue cases go unreported.  
-Lastly, demography and migrations affect the number of susceptible in ways 
-which are not easy to fully determine.
-
-Figure~\ref{Fig:S0}, shows the model from (\ref{eq:model}) fitted to the data.
-Despite its limitations, our simplified model fits the data well.
-In it we can see that the susceptibles series in each epidemic starts at the 
-estimated level of $S_0$.
-The proportion of susceptibles may seem low, but we 
-must remember that these estimates are being affected by an unknown 
-under-reporting factor, which experts suggest is somewhere between 5 and 10, 
-i.e. for every case observed there are 5 or 10 unobserved.
-Since this under-reporting affects both the numerator and denominator of 
-(\ref{eq:AR2}), its effects should cancel out, giving us an unbiased attack 
-ratio estimate.
-One other possible source of bias which would lead to the underestimation of 
-$S_0$ could come from a significant part of the population not being exposed to 
-the disease.
-However, as we can see in Figure~\ref{fig:mapas}, despite the differences in 
-intensity (incidence), the entire city seems to be at risk, with no 
-particularly ``protected'' areas, at least in the last four epidemics.
-
-\begin{center}
-\begin{figure}
- \subfigure[2010]{\includegraphics[width=8cm, 
-height=4.82cm]{./plots/heatmap2010.jpg}
-% heatmap2010.jpg: 1075x589 pixel, 96dpi, 28.44x15.58 cm, bb=0 0 806 442
-}
-  \subfigure[2011]{\includegraphics[width=8cm]{./plots/heatmap2011.jpg}}
-  \subfigure[2012]{\includegraphics[width=8cm]{./plots/heatmap2012.jpg}
-% heatmap2010.jpg: 1075x589 pixel, 96dpi, 28.44x15.58 cm, bb=0 0 806 442
-}
-  \subfigure[2013]{\includegraphics[width=8cm]{./plots/heatmap2013.jpg}}
-  \caption{{\bf Maps showing the incidence of dengue in the city of Rio de 
-Janeiro from 2010 to 2013.} Circles indicate individual notified cases. A 
-heatmap is overlayed on the maps showing absolute density of cases.
-It can be seen that several areas of the city were affected and no region seems 
-to be free of transmission risk.
-This suggests that although transmission risk varies spatially, there is 
-significant exposure over the entire city.}
-  \label{fig:mapas}
-\end{figure}
-\end{center}
-
-Table~\ref{tab:AR} contains the attack ratios and medians of the $S_0$ 
-estimated for each epidemic/outbreak.
-It is interesting to notice that the larger epidemics, in terms of peak size are 
-not the one with the greater attack ratios.
-This stresses the importance of knowing the immunological structure of the 
-population. Knowing the $S_0$ for the circulating viruses we can order to more 
-accurately assess the potential impact of a coming epidemic, since particularly 
-virulent types, can be rendered less of a threat by a low $S_0$.
-
-We hope that the results presented in paper will motivate public 
-health authorities to invest in annual serological surveys, to determine the 
-susceptibility profile to each dengue virus as well as to estimate the 
-under-reporting factor of the notification system.
-
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics[width=\textwidth]{plots/concat_SI.png}
-\end{center}
-\caption{
-{\bf Susceptibles and Infectious posterior curves.}  The curves were estimated 
-only for the periods where $R_t> 1$.  The susceptible curves in the top panel 
-reflect the prevalence of fraction of susceptibles to circulating strain(s) for 
-each epidemic/outbreak.
-In the lower panel, we see the posterior distribution 
-of infectious curves, represented by its median and 95\% credible interval.
-Credible intervals are very narrow, and can be hard to distinguish from the 
-median line. 
-Dots show the observed cases, scaled as fractions of the entire 
-population.
-}
-\label{Fig:S0}
-\end{figure}
-
-\begin{table}[!ht]
-\caption{
-{\bf Median attack ratio and 95\% credibility intervals calculated according to 
-(\ref{eq:AR2})}. 
-Values are presented as percentage of total population. 
-$^\dag$: Year corresponds to the start of the epidemic, however the peak of cases
-may occur in the following year.
-$^\ddag$: Susceptible fraction.
-These results show considerable variation in AR between epidemics, consistent with
-the accquiring and loss of serotype-specific immunity.}
-\begin{center}
-\begin{tabular}{c|c|c}
-\hline
-Year$^\dag$ & median Attack Ratio & $S_0^\ddag$ \\
-\hline
-1996 & 0.39 (0.17-0.54) & 0.00171(0.0012-0.0038)\\
-1997 & 0.87 (0.74-0.87) & 0.00273(0.0027-0.0032)\\
-1998 & 0.5 (0.49-0.5) & 0.00142(0.0014-0.0014)\\
-1999 & 0.11 (0.037-0.2) & 0.00345(0.0018-0.01)\\
-2000 & 0.25 (0.24-0.27) & 0.0155(0.015-0.016)\\
-2001 & 0.48 (0.47-0.49) & 0.0495(0.048-0.051)\\
-2005 & 0.15 (0.1-0.21) & 0.0147(0.01-0.021)\\
-2006 & 0.11 (0.08-0.14) & 0.0281(0.022-0.037)\\
-2007 & 0.15 (0.15-0.15) & 0.135(0.13-0.14)\\
-2008 & 0.14 (0.031-0.31) & 0.00672(0.003-0.024)\\
-2010 & 0.18 (0.17-0.19) & 0.0454(0.043-0.048)\\
-2011 & 0.086 (0.082-0.094) & 0.215(0.2-0.23)\\
-2012 & 0.14 (0.13-0.15) & 0.0621(0.058-0.068)\\
-\hline
-\end{tabular}
-\end{center}
-\label{tab:AR}
-\end{table}
-
-\newpage
-\section*{Acknowledgements}
-LMC is grateful to Dr. Leonardo Bastos for useful discussions on the 
-posterior inference for \rr.
-The authors are also grateful to Claudia T. 
-Code\c{c}o for helpful discussions about the manuscript.
-
-\newpage
-\bibliography{lm1}
-
-\section*{Appendix}
-\subsection*{A remark on prior distributions and tail behaviour of the 
+\section*{A remark on prior distributions and tail behaviour of the 
 distribution of $R_t$}
@@
 
+\bibliography{lm1}
+
 \end{document}

### 2015-02-19 supplSR.tex
recreated the supplementary material

--- 2015-02-19 suppl.tex
@@
 
-\bibliography{lm1}
+\begin{thebibliography}{1}
+\expandafter\ifx\csname url\endcsname\relax
+  \def\url#1{\texttt{#1}}\fi
+\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
+\providecommand{\bibinfo}[2]{#2}
+\providecommand{\eprint}[2][]{\url{#2}}
+
+\bibitem{mantel}
+\bibinfo{author}{Ederer, F.} \& \bibinfo{author}{Mantel, N.}
+\newblock \bibinfo{title}{Confidence limits on the ratio of two poisson
+  variables}.
+\newblock \emph{\bibinfo{journal}{American Journal of Epidemiology}}
+  \textbf{\bibinfo{volume}{100}}, \bibinfo{pages}{165--167}
+  (\bibinfo{year}{1974}).
+
+\bibitem{nishiura}
+\bibinfo{author}{Nishiura, H.}, \bibinfo{author}{Chowell, G.},
+  \bibinfo{author}{Heesterbeek, H.} \& \bibinfo{author}{Wallinga, J.}
+\newblock \bibinfo{title}{{{T}he ideal reporting interval for an epidemic to
+  objectively interpret the epidemiological time course}}.
+\newblock \emph{\bibinfo{journal}{J R Soc Interface}}
+  \textbf{\bibinfo{volume}{7}}, \bibinfo{pages}{297--307}
+  (\bibinfo{year}{2010}).
+
+\bibitem{wilson}
+\bibinfo{author}{Wilson, E.~B.}
+\newblock \bibinfo{title}{Probable inference, the law of succession, and
+  statistical inference}.
+\newblock \emph{\bibinfo{journal}{Journal of the American Statistical
+  Association}} \textbf{\bibinfo{volume}{22}}, \bibinfo{pages}{209--212}
+  (\bibinfo{year}{1927}).
+
+\bibitem{clopper}
+\bibinfo{author}{Clopper, C.} \& \bibinfo{author}{Pearson, E.~S.}
+\newblock \bibinfo{title}{The use of confidence or fiducial limits illustrated
+  in the case of the binomial}.
+\newblock \emph{\bibinfo{journal}{Biometrika}} \bibinfo{pages}{404--413}
+  (\bibinfo{year}{1934}).
+
+\end{thebibliography}
 

### 2015-05-29 Presentations/Dinamica de Vetores-SP-2015/dengueARS0.tex
Updated presentation

@@
 \end{equation}
-Where $n$ is the ration between the length of reporting interval and 
+Where $n$ is the ratio between the length of reporting interval and 
 the mean generation time of the disease.
@@
 \frametitle{\rr's uncertainty}
-But what about the uncertainty about \rr?
+But what about the uncertainty about \rr\footnote{Coelho, FC and Carvalho, LM 
+(Submitted)}?
 
-If we assume that the counts $Y_t$ are Poisson distributed for all $t$, we can 
-derive the probability distribution of \rr\footnote{Coelho, FC and Carvalho, LM 
-(Submitted)}
+We explore the approach of Ederer and Mantel\cite{mantel}, whose objective is 
+to obtain 
+confidence intervals for the ratio of two Poisson counts. 
+Let $Y_{t} \sim Poisson(\lambda_t)$ and $Y_{t+1} \sim Poisson(\lambda_{t+1})$ 
+and define $S = Y_{t} + Y_{t+1}$.
+The authors note that by conditioning on the sum $S$
+\begin{align}
+\label{eq:binlike}
+Y_{t+1} | S &\sim Binomial(S, \theta_t) \\
+\theta_t &= \frac{\lambda_{t+1}}{\lambda_{t} + \lambda_{t+1}}
+\end{align}
+\end{frame}
+
+\begin{frame}
+\frametitle{\rr's Uncertainty}
+Let $c_{\alpha}(\theta_t) = \{\theta_t^{(L)} , \theta_t^{(U)} \}$ be such that 
+$Pr(\theta_t^{(L)}<\theta_t <\theta_t^{(U)}) = \alpha$.
+Analogously, define $c_{\alpha}(R_t) = \{R_t^{(L)} , R_t^{(U)} \}$ such that 
+$Pr(R_t^{(L)}<R_t<R_t^{(U)}) = \alpha$.
+
+Ederer and Mantel (1974)~\cite{mantel} show that one can construct a $100\alpha 
+\%$ confidence interval for \rr by noting that
+\begin{equation}
+\label{eq:confRt}
+ R_t^{(L)} = \frac{\theta_t^{(L)}}{(1-\theta_t^{(L)})} \quad \text{and} \quad 
+R_t^{(U)} = \frac{\theta_t^{(U)}}{(1-\theta_t^{(U)})}\\
+\end{equation}
+\end{frame}
+
+\begin{frame}
+\frametitle{\rr's Uncertainty}
+Taking a Bayesian conjugate distribution approach, If we choose a
+Beta conjugate prior with parameters $a_0$ and $b_0$ for the 
+Binomial likelihood in (\ref{eq:binlike}), the posterior distribution for 
+$\theta_t$ is
+\begin{equation}
+\label{eq:thetapost}
+p(\theta_t| Y_{t+1}, S) \sim Beta(Y_{t+1} + a_0, Y_t + b_0)
+\end{equation}
+Combining equations~(\ref{eq:confRt}) and~(\ref{eq:thetapost}) 
+tells us that the induced posterior distribution of $R_t$ is 
+a Beta prime (or inverted Beta) with parameters $ a_1 = Y_{t+1} + a_0$ and $b_1 
+=  Y_t + b_0$~\cite{dubey1970}.
+The density of the induced distribution is then 
+\begin{equation}
+\label{eq:densityMantel}
+f_P(R_t| a_1, b_1) = \frac{\Gamma(a_1 + b_1)}{\Gamma(a_1)\Gamma(b_1)} R_t^{a_1 
+- 
+1} (1 + R_t)^{-(a_1 + b_1)}
+\end{equation}
+Thus, the expectation of \rr is $a_1/(b_1 - 1)$ and its variance is 
+$a_1(a_1 + b_1 - 1)/\left((b_1 - 2)(b_1 - 1)^2 \right) $.
+\end{frame}
+
+\begin{frame}
+\frametitle{\rr's Uncertainty}
 \begin{center}
@@
 \end{center}
-
 \end{frame}
-
 \subsection{Vector dynamics}
@@
 %\bibAnnoteFile{pone2011}
+
+\bibitem{mantel}
+Ederer, F. and Mantel, N. (1974)
+\newblock Confidence limits on the ratio of two poisson
+  variables.
+\newblock \emph{American Journal of Epidemiology}
+  \textbf{volume 100:}, pages 165--167
+
 \end{thebibliography}

### 2015-05-29 scirep_version.tex
Updated presentation

--- 2015-02-12 draft_lm1.tex
@@
-% Template for PLoS
-% Version 2.0 July 2014
-%
-% To compile to pdf, run:
-% latex plos.template
-% bibtex plos.template
-% latex plos.template
-% latex plos.template
-% dvipdf plos.template
-%
-% % % % % % % % % % % % % % % % % % % % % %
-%
 \documentclass[10pt]{article}
@@
 \begin{figure}
- \subfigure[2010]{\includegraphics[width=8cm, 
-height=4.82cm]{./plots/heatmap2010.jpg}
-% heatmap2010.jpg: 1075x589 pixel, 96dpi, 28.44x15.58 cm, bb=0 0 806 442
-}
-  \subfigure[2011]{\includegraphics[width=8cm]{./plots/heatmap2011.jpg}}
-  \subfigure[2012]{\includegraphics[width=8cm]{./plots/heatmap2012.jpg}
-% heatmap2010.jpg: 1075x589 pixel, 96dpi, 28.44x15.58 cm, bb=0 0 806 442
-}
-  \subfigure[2013]{\includegraphics[width=8cm]{./plots/heatmap2013.jpg}}
+\includegraphics[width=\textwidth]{./plots/figure-2-coelho.jpg}
+
   \caption{{\bf Maps showing the incidence of dengue in the city of Rio de 
-Janeiro from 2010 to 2013.} Circles indicate individual notified cases. A 
+Janeiro from 2010 to 2013(left-to-right, rowise).} Circles indicate individual 
+notified cases. A 
 heatmap is overlayed on the maps showing absolute density of cases.
@@
 This suggests that although transmission risk varies spatially, there is 
-significant exposure over the entire city.}
+significant exposure over the entire city. Maps were made with QGis GIS 
+software(\url{qgis.org})}
   \label{fig:mapas}
@@
 \newpage
-\bibliography{lm1}
+\begin{thebibliography}{10}
 
-\section*{Appendix}
-\subsection*{A remark on prior distributions and tail behaviour of the 
-distribution of $R_t$}
-\label{sec:tails}
-There are a number of approaches to deriving the distribution of \rr.
-Alternatively to the approach described in the main text~\cite{mantel}, one 
-could use the conditional distribution of \rr on 
-$Y_{t+1}$ and $Y_t$ as defined in equation A7 of Nishiura et 
-al.~\cite{nishiura}:
-\begin{equation}
-\label{seq:unorm}
-f_{R}(R_{t}) = (Y_tR_{t})^{Y_{t+1}} e^{-Y_tR_{t}}
-\end{equation}
-Noticing the kernel of (\ref{seq:unorm}) is that of a gamma distribution with 
-$a_2 = Y_{t+1}+1$ and $b_2 = Y_t$, we obtain a proper density from which to 
-construct $c_{\alpha}(R_t)$, simply by computing the appropriate quantiles of 
-said distribution.
- This density is
-\begin{equation}
-\label{seq:densityNishiura}
-f_N(R_t| a_2, b_2) =  \frac{b_2^{a_2}}{\Gamma(a_2)} R_t^{a_2-1} e^{-b_2 R_t}
-\end{equation}
+\expandafter\ifx\csname url\endcsname\relax
+  \def\url#1{\texttt{#1}}\fi
+\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
+\providecommand{\bibinfo}[2]{#2}
+\providecommand{\eprint}[2][]{\url{#2}}
 
-In order to decide which approach to take, it may be of use analysing the 
-tail behaviour of the derived distributions for \rr. 
-Consider the case of using a flat $Uniform(0, 1)$ prior for $\theta_t$.
-With $a_0 = b_0 = 1$, $a_1 = a_2$ and $b_1 = b_2 + 1$.
-The beta prime (inverse beta distribution) will have heavier tails compared to 
-the conditional distribution proposed by~\cite{nishiura}, thus providing more 
-conservative confidence/credibility intervals.
-To see that one needs simply take the ratio of the Beta prime and Gamma 
-(unnormalized) densities and evaluate the limit as $R_t$ goes to infinity:
-\begin{equation}
- \label{seq:densityratio}
- \lim_{R_t\to\infty}\frac{f_P(R_t| a_1, b_1)}{f_N(R_t| a_2, b_2)} =  
-\lim_{R_t\to\infty}\frac{e^{Y_{t}R_t}}{(1 +R_t)^{Y_{t} + Y_{t +1}+2}} = \infty
-\end{equation}
-Finally, note that we deliberately construct $c_{\alpha}(R_{t})$ as a 
-equal-tailed $100\alpha\%$ credible set, rather than a less conservative 
-highest posterior density (HPD) interval.
+\bibitem{Guzman2010}
+\bibinfo{author}{Guzman, M.~G.} \emph{et~al.}
+\newblock \bibinfo{title}{{{D}engue: a continuing global threat}}.
+\newblock \emph{\bibinfo{journal}{Nat. Rev. Microbiol.}}
+  \textbf{\bibinfo{volume}{8}}, \bibinfo{pages}{7--16} (\bibinfo{year}{2010}).
 
-As a side note, the Bayesian approach presented in this 
-paper will give similar results to orthodox confidence intervals~\cite{wilson} 
-and~\cite{clopper} for $Y_{t+1}$ and $Y_t >> 1$.
-Under the flat  uniform prior for $\theta_t$, the Bayesian posterior 
-credibility 
-interval is nearly indistinguishable from the confidence interval proposed by 
-Clopper \& Pearson (1931)~\cite{clopper} for $Y_{t+1}, Y_t > 20$.
-Note also that the uniform prior ($Beta(1, 1)$) for $\theta_t$ constitutes a 
-poor 
-prior 
-choice mainly because the induced distribution for \rr is only well-defined for 
-$b_0 > 2$.
+\bibitem{halstead_dengue_2007}
+\bibinfo{author}{Halstead, S.~B.}
+\newblock \bibinfo{title}{Dengue}.
+\newblock \emph{\bibinfo{journal}{The Lancet}} \textbf{\bibinfo{volume}{370}},
+  \bibinfo{pages}{1644--1652} (\bibinfo{year}{2007}).
+\newblock
 
-An advantage of the Bayesian approach is that one can devise prior 
-distributions for $\theta_t$ taking advantage of the intuitive parametrization 
-and flexibility of the beta family of distributions.
-Prior elicitation can also be done for \rr and the hyper-parameters directly 
-plugged into the prior for $\theta_t$. 
-One can, for example, choose prior mean and variance for \rr and find $a_0$ 
-and $b_0$ that satisfy those conditions.
-Let $m_0$ and $v_0$ be the prior expectation and variance for $R_t$. 
-After some tedious algebra one finds
-\begin{align}
-\label{seq:elicitation}
-a_0 &= \frac{m_0v_0 + m_0^3 + m_0^2}{v_0} \\
-b_0 &= \frac{2v_0 + m_0^2 + m_0}{v_0}
-\end{align}
-If one wants only to specify $m_0$ and the coefficient of variation $c = 
-\sqrt{v_0}/ m_0$ for $R_t$ \textit{a priori}, some less boring 
-algebra gives:
-\begin{align}
-\label{seq:elicitationcv}
-a_0 &= \frac{m_0^3c^2 + m_0^3 + m_0^2}{m_0^2c^2} \\
-b_0 &= \frac{2m_0^2c^2 + m^2 + m}{m_0^2c^2}
-\end{align}
+\urlprefix\url{
+http://www.sciencedirect.com/science/article/pii/S0140673607616870}.
 
-This approach thus makes it possible to incorporate epidemiological knowledge 
-about disease Biology (e.g. the magnitude of $R_0$) into the computation of \rr.
-This may prove particularly important when disease counts are low and/or close 
-to the detection threshold.
-We provide an R script to perform the above elicitation at 
-\url{https://github.com/fccoelho/paperLM1/blob/master/R/elicit_Rt_prior.R}.
+\bibitem{honorio_temporal_2009}
+\bibinfo{author}{Hon\'orio, N.~A.}, \bibinfo{author}{Code\c{c}o, C.~T.},
+  \bibinfo{author}{Alves, F.~C.}, \bibinfo{author}{Magalh\~aes, M.} \&
+  \bibinfo{author}{Louren\c{c}o-De-Oliveira, R.}
+\newblock \bibinfo{title}{Temporal distribution of aedes aegypti in different
+  districts of rio de janeiro, brazil, measured by two types of traps}.
+\newblock \emph{\bibinfo{journal}{Journal of Medical Entomology}}
+  \textbf{\bibinfo{volume}{46}}, \bibinfo{pages}{1001--1014}
+  (\bibinfo{year}{2009}).
+\newblock \urlprefix\url{http://www.bioone.org/doi/abs/10.1603/033.046.0505}.
+
+\bibitem{wu_higher_2009}
+\bibinfo{author}{Wu, P.-C.} \emph{et~al.}
+\newblock \bibinfo{title}{Higher temperature and urbanization affect the
+  spatial patterns of dengue fever transmission in subtropical taiwan}.
+\newblock \emph{\bibinfo{journal}{Science of The Total Environment}}
+  \textbf{\bibinfo{volume}{407}}, \bibinfo{pages}{2224--2233}
+  (\bibinfo{year}{2009}).
+\newblock
+
+\urlprefix\url{
+http://www.sciencedirect.com/science/article/pii/S0048969708011509}.
+
+\bibitem{reiner_time-varying_2014}
+\bibinfo{author}{Reiner, R.~C.} \emph{et~al.}
+\newblock \bibinfo{title}{Time-varying, serotype-specific force of infection of
+  dengue virus}.
+\newblock \emph{\bibinfo{journal}{Proceedings of the National Academy of
+  Sciences of the United States of America}} \textbf{\bibinfo{volume}{111}},
+  \bibinfo{pages}{E2694--E2702} (\bibinfo{year}{2014}).
+\newblock \urlprefix\url{http://www.ncbi.nlm.nih.gov/pmc/articles/PMC4084484/}.
+
+\bibitem{bacaer_final_2009}
+\bibinfo{author}{Baca\"er, N.} \& \bibinfo{author}{Gomes, M. G.~M.}
+\newblock \bibinfo{title}{On the final size of epidemics with seasonality}.
+\newblock \emph{\bibinfo{journal}{Bulletin of Mathematical Biology}}
+  \textbf{\bibinfo{volume}{71}}, \bibinfo{pages}{1954--1966}
+  (\bibinfo{year}{2009}).
+\newblock
+  \urlprefix\url{http://link.springer.com/article/10.1007/s11538-009-9433-7}.
+
+\bibitem{katriel_attack_2012}
+\bibinfo{author}{Katriel, G.} \& \bibinfo{author}{Stone, L.}
+\newblock \bibinfo{title}{Attack rates of seasonal epidemics}.
+\newblock \emph{\bibinfo{journal}{Mathematical Biosciences}}
+  \textbf{\bibinfo{volume}{235}}, \bibinfo{pages}{56--65}
+  (\bibinfo{year}{2012}).
+\newblock
+
+\urlprefix\url{
+http://www.sciencedirect.com/science/article/pii/S0025556411001532}.
+
+\bibitem{bjornstad_dynamics_2002}
+\bibinfo{author}{Bjørnstad, O.}, \bibinfo{author}{Finkenstädt, B.} \&
+  \bibinfo{author}{Grenfell, B.}
+\newblock \bibinfo{title}{Dynamics of measles epidemics: Estimating scaling of
+  transmission rates using a time series {SIR} model}.
+\newblock \emph{\bibinfo{journal}{Ecological Monographs}}
+  \textbf{\bibinfo{volume}{72}}, \bibinfo{pages}{169--184}
+  (\bibinfo{year}{2002}).
+
+\bibitem{wallinga_reconstruction_2003}
+\bibinfo{author}{Wallinga, J.}, \bibinfo{author}{Teunis, P.} \&
+  \bibinfo{author}{Kretzschmar, M.}
+\newblock \bibinfo{title}{Reconstruction of measles dynamics in a vaccinated
+  population}.
+\newblock \emph{\bibinfo{journal}{Vaccine}} \textbf{\bibinfo{volume}{21}},
+  \bibinfo{pages}{2643--2650} (\bibinfo{year}{2003}).
+
+\bibitem{Reiner2014}
+\bibinfo{author}{Reiner, R.~C.} \emph{et~al.}
+\newblock \bibinfo{title}{{{T}ime-varying, serotype-specific force of infection
+  of dengue virus}}.
+\newblock \emph{\bibinfo{journal}{Proc. Natl. Acad. Sci. U.S.A.}}
+  \textbf{\bibinfo{volume}{111}}, \bibinfo{pages}{E2694--2702}
+  (\bibinfo{year}{2014}).
+
+\bibitem{nishiura}
+\bibinfo{author}{Nishiura, H.}, \bibinfo{author}{Chowell, G.},
+  \bibinfo{author}{Heesterbeek, H.} \& \bibinfo{author}{Wallinga, J.}
+\newblock \bibinfo{title}{{{T}he ideal reporting interval for an epidemic to
+  objectively interpret the epidemiological time course}}.
+\newblock \emph{\bibinfo{journal}{J R Soc Interface}}
+  \textbf{\bibinfo{volume}{7}}, \bibinfo{pages}{297--307}
+  (\bibinfo{year}{2010}).
+
+\bibitem{pone2011}
+\bibinfo{author}{Coelho, F.~C.}, \bibinfo{author}{Code\c{c}o, C.~T.} \&
+  \bibinfo{author}{Gomes, M.~G.}
+\newblock \bibinfo{title}{{A} {B}ayesian framework for parameter estimation in
+  dynamical models}.
+\newblock \emph{\bibinfo{journal}{PLoS ONE}} \textbf{\bibinfo{volume}{6}},
+  \bibinfo{pages}{e19616} (\bibinfo{year}{2011}).
+
+\bibitem{stallybrass}
+\bibinfo{author}{Stallybrass, C.~O.} \emph{et~al.}
+\newblock \bibinfo{title}{The principles of epidemiology and the process of
+  infection.}
+\newblock \emph{\bibinfo{journal}{The Principles of Epidemiology and the
+  Process of Infection.}}  (\bibinfo{year}{1931}).
+
+\bibitem{mantel}
+\bibinfo{author}{Ederer, F.} \& \bibinfo{author}{Mantel, N.}
+\newblock \bibinfo{title}{Confidence limits on the ratio of two poisson
+  variables}.
+\newblock \emph{\bibinfo{journal}{American Journal of Epidemiology}}
+  \textbf{\bibinfo{volume}{100}}, \bibinfo{pages}{165--167}
+  (\bibinfo{year}{1974}).
+
+\bibitem{wilson}
+\bibinfo{author}{Wilson, E.~B.}
+\newblock \bibinfo{title}{Probable inference, the law of succession, and
+  statistical inference}.
+\newblock \emph{\bibinfo{journal}{Journal of the American Statistical
+  Association}} \textbf{\bibinfo{volume}{22}}, \bibinfo{pages}{209--212}
+  (\bibinfo{year}{1927}).
+
+\bibitem{clopper}
+\bibinfo{author}{Clopper, C.} \& \bibinfo{author}{Pearson, E.~S.}
+\newblock \bibinfo{title}{The use of confidence or fiducial limits illustrated
+  in the case of the binomial}.
+\newblock \emph{\bibinfo{journal}{Biometrika}} \bibinfo{pages}{404--413}
+  (\bibinfo{year}{1934}).
+
+\bibitem{dubey1970}
+\bibinfo{author}{Dubey, S.~D.}
+\newblock \bibinfo{title}{Compound gamma, beta and f distributions}.
+\newblock \emph{\bibinfo{journal}{Metrika}} \textbf{\bibinfo{volume}{16}},
+  \bibinfo{pages}{27--31} (\bibinfo{year}{1970}).
+
+\bibitem{vrugt2008}
+\bibinfo{author}{Vrugt, J.~A.} \emph{et~al.}
+\newblock \bibinfo{title}{Accelerating markov chain monte carlo simulation by
+  differential evolution with self-adaptive randomized subspace sampling}.
+\newblock \bibinfo{type}{Tech. Rep.}, \bibinfo{institution}{Citeseer}
+  (\bibinfo{year}{2008}).
+
+\bibitem{brooks1998}
+\bibinfo{author}{Brooks, S.~P.} \& \bibinfo{author}{Gelman, A.}
+\newblock \bibinfo{title}{General methods for monitoring convergence of
+  iterative simulations}.
+\newblock \emph{\bibinfo{journal}{Journal of computational and graphical
+  statistics}} \textbf{\bibinfo{volume}{7}}, \bibinfo{pages}{434--455}
+  (\bibinfo{year}{1998}).
+
+\bibitem{macedo_virological_2013}
+\bibinfo{author}{Macedo, G.~A.} \emph{et~al.}
+\newblock \bibinfo{title}{Virological surveillance for early warning of dengue
+  epidemics in the state of rio de janeiro, brazil}.
+\newblock \emph{\bibinfo{journal}{Transactions of the Royal Society of Tropical
+  Medicine and Hygiene}} \textbf{\bibinfo{volume}{107}},
+  \bibinfo{pages}{141--146} (\bibinfo{year}{2013}).
+
+\bibitem{luz_disability_2009}
+\bibinfo{author}{Luz, P.~M.}, \bibinfo{author}{Grinsztejn, B.} \&
+  \bibinfo{author}{Galvani, A.~P.}
+\newblock \bibinfo{title}{Disability adjusted life years lost to dengue in
+  brazil}.
+\newblock \emph{\bibinfo{journal}{Tropical Medicine \& International Health}}
+  \textbf{\bibinfo{volume}{14}}, \bibinfo{pages}{237--246}
+  (\bibinfo{year}{2009}).
+\newblock
+
+\urlprefix\url{
+http://onlinelibrary.wiley.com/doi/10.1111/j.1365-3156.2008.02203.x/abstract}.
+
+\bibitem{duarte_data_2006}
+\bibinfo{author}{Duarte, H. H.~P.} \& \bibinfo{author}{Fran\c{c}a, E.~B.}
+\newblock \bibinfo{title}{Data quality of dengue epidemiological surveillance
+  in belo horizonte, southeastern brazil}.
+\newblock \emph{\bibinfo{journal}{Revista de Sa\'ude P\'ublica}}
+  \textbf{\bibinfo{volume}{40}}, \bibinfo{pages}{134--142}
+  (\bibinfo{year}{2006}).
+\newblock
+
+\urlprefix\url{
+http://www.scielosp.org/scielo.php?script=sci_abstract&pid=S0034-891020060001000
+  21&lng=en&nrm=iso&tlng=en}.
+
+\end{thebibliography}
+
+\section*{Authors Contributions}
+FCC and LMC wrote the main manuscript. LMC and FCC did the analyses. FCC 
+prepared the figures. All authors reviewed the manuscript.
+
+\section*{Additional Information}
+\subsection*{Competing financial interests}
+The author(s) declare no competing financial interests.
 

### 2015-06-01 Presentations/Dinamica de Vetores-SP-2015/dengueARS0.tex
updated presentation for Talk at UPMC

@@
 Force of Infection using Aggregated Notification Data}
-\date{11-02-2015}
-\institute [EMAp, FGV]{Escola de Matemática Aplicada,   Fundação Getúlio Vargas}
+\date{May 1st, 2015}
+\institute [EMAp, FGV]{Applied Mathematics School,   Fundação Getúlio Vargas}
 
-%\logo{\includegraphics[height=2cm]{./fb1.png}}
+\logo{\includegraphics[width=\beamer@sidebarwidth]{./fb1.png}}
 \begin{document}
@@
  R_t^{(L)} = \frac{\theta_t^{(L)}}{(1-\theta_t^{(L)})} \quad \text{and} \quad 
-R_t^{(U)} = \frac{\theta_t^{(U)}}{(1-\theta_t^{(U)})}\\
+R_t^{(U)} = \frac{\theta_t^{(U)}}{(1-\theta_t^{(U)})}
 \end{equation}
@@
 \end{frame}
+\begin{frame}
+ \frametitle{Collaboration possibilities}
+ \begin{itemize}
+  \item Relaxing other simplifying assumptions in classical epidemiological 
+models.
+\item Looking into modeling spatio-temporal dynamics taking into account 
+Environmental stock of eggs.
+\item Separating noise from relevant varibility in observational data, and 
+finding new ways of integrating it into models.
+ \end{itemize}
+
+\end{frame}
+\begin{frame}
+
+ \begin{center}
+\begin{Huge}Thank you!           \end{Huge} \end{center}
+\end{frame}
+
 \end{document}

### 2015-06-14 python/fit_dengue_split_big_plot.py
minor changes

@@ -191,7 +191,7 @@ def plot_concat_series(dbs):
     for srs in series:
         co = c.next()
         s_median = srs.S.groupby(level='time').median()
-        s_median.plot(style='k-', label='Median')
+        s_median.plot(style='k-', label='Median S')
         s_upr = srs.S.groupby(level='time').aggregate(upr)
         s_lwr = srs.S.groupby(level='time').aggregate(lwr)
         # P.legend(loc=0)
@@ -202,7 +202,7 @@ def plot_concat_series(dbs):
     for srs, da in zip(series, obs):
         co = c.next()
         i_median = srs.I.groupby(level='time').median()
-        i_median.plot(style='k-')
+        i_median.plot(style='k-', label="Median I")
         # da.I.plot(style='ro', alpha=.5, label='obs')
         i_upr = srs.I.groupby(level='time').aggregate(upr)
         i_lwr = srs.I.groupby(level='time').aggregate(lwr)

### 2015-06-14 python/fit_simulated_data.py

@@ -16,6 +16,7 @@ from collections import defaultdict
 import datetime
 import sys
 import pandas as pd
+from sqlalchemy import create_engine
 import seaborn as sns
 # import numba
 #from numba import jit
@@ -60,10 +61,10 @@ def sir(y, t, *pars):
 def jac(y, t, *pars):
     """
     Jacobian of the model
-    :param y:
-    :param t:
-    :param pars:
-    :return:
+    :param y: State of the system
+    :param t: current time
+    :param pars: tuple of parameters
+    :return: the jacobian matrix at time t
     """
     S, I, R = y
     s0, m, tau = pars
@@ -84,12 +85,12 @@ def model(theta):
     s0, m, tau = theta
 
     t0 = 0
-    tf = fim
+    tf = fim-inicio
     Y = np.zeros((tf - t0, 3))
 
 
     # Initial conditions
-    #print inits, m
+
     inits[0] = s0  # Define S0
     inits[-1] = N - sum(inits[:2])  # Define R(0)
     Y[t0:tf, :] = odeint(sir, inits, np.arange(t0, tf, 1), args=(s0, m, tau), Dfun=jac)  #,tcrit=tcrit)
@@ -180,6 +181,20 @@ def fix_rt(rt):
     else:
         return rt
 
+
+def read_data(dbname):
+    """
+    Returns dataframes from tables of the sqlite database with the results of the inference
+    :param dbname: sqlite3 database name
+    :return: list of pandas dataframes
+    """
+    eng = create_engine('sqlite:///{}'.format(dbname))
+    df_data = pd.read_sql_table('data', eng, index_col=['time'], parse_dates={'time': '%Y-%m-%d %H:%M:%S'})
+    df_pt = pd.read_sql_table('post_theta', eng, index_col=['time'], parse_dates={'time': '%Y-%m-%d %H:%M:%S'})
+    df_series = pd.read_sql_table('series', eng, index_col=['time'], parse_dates={'time': '%Y-%m-%d %H:%M:%S'})
+    return df_data, df_pt, df_series
+
+
 # # running the analysys
 if __name__ == "__main__":
     # dt = prepdata('aux/data_Rt_dengue_big.csv', 0, 728, 1)
@@ -218,18 +233,9 @@ if __name__ == "__main__":
[…]

### 2015-06-15 draft_lm1.tex
fixes #36
Yes it is, now in boldface to make clear. fixes #38
Hopefully closes #35

@@
 We show how to elicit the prior for $R_t$ from specified prior mean and variance
-or coefficient of variation in the Appendix.
+or coefficient of variation in the \textbf{Appendix}.
 
@@
 scheme that efficiently samples from high-dimensional joint distributions using 
-multiple adaptive chains running in parallel with delayed rejection.
+multiple adaptive chains running in parallel with delayed rejection (see 
+\textbf{Appendix} for details).
 Also, as the numerical integration routine implemented within BIP needs 
@@
 
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[width=\textwidth]{plots/concat_SI.png}
+\end{center}
+\caption{
+{\bf Susceptibles and Infectious posterior curves.}  The curves were estimated 
+only for the periods where $R_t> 1$.  The susceptible curves in the top panel 
+reflect the prevalence of fraction of susceptibles to circulating strain(s) for 
+each epidemic/outbreak.
+In the lower panel, we see the posterior distribution 
+of infectious curves, represented by its median and 95\% credible interval.
+Credible intervals are very narrow, and can be hard to distinguish from the 
+median line. 
+Dots show the observed cases, scaled as fractions of the entire 
+population.
+}
+\label{Fig:S0}
+\end{figure}
+
+Table~\ref{tab:AR} contains the attack ratios and medians of the $S_0$ 
+estimated for each epidemic/outbreak.
+It is interesting to notice that the larger epidemics, in terms of peak size are 
+not the one with the greater attack ratios.
+This stresses the importance of knowing the immunological structure of the 
+population. Knowing the $S_0$ for the circulating viruses we can order to more 
+accurately assess the potential impact of a coming epidemic, since particularly 
+virulent types, can be rendered less of a threat by a low $S_0$.
+
+We hope that the results presented in paper will motivate public 
+health authorities to invest in annual serological surveys, to determine the 
+susceptibility profile to each dengue virus as well as to estimate the 
+under-reporting factor of the notification system.
+
 \begin{center}
@@
 \end{center}
-
-Table~\ref{tab:AR} contains the attack ratios and medians of the $S_0$ 
-estimated for each epidemic/outbreak.
-It is interesting to notice that the larger epidemics, in terms of peak size are 
-not the one with the greater attack ratios.
-This stresses the importance of knowing the immunological structure of the 
-population. Knowing the $S_0$ for the circulating viruses we can order to more 
-accurately assess the potential impact of a coming epidemic, since particularly 
-virulent types, can be rendered less of a threat by a low $S_0$.
-
-We hope that the results presented in paper will motivate public 
-health authorities to invest in annual serological surveys, to determine the 
-susceptibility profile to each dengue virus as well as to estimate the 
-under-reporting factor of the notification system.
-
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics[width=\textwidth]{plots/concat_SI.png}
-\end{center}
-\caption{
-{\bf Susceptibles and Infectious posterior curves.}  The curves were estimated 
-only for the periods where $R_t> 1$.  The susceptible curves in the top panel 
-reflect the prevalence of fraction of susceptibles to circulating strain(s) for 
-each epidemic/outbreak.
-In the lower panel, we see the posterior distribution 
-of infectious curves, represented by its median and 95\% credible interval.
-Credible intervals are very narrow, and can be hard to distinguish from the 
-median line. 
-Dots show the observed cases, scaled as fractions of the entire 
-population.
-}
-\label{Fig:S0}
-\end{figure}
 
@@
 to the detection threshold.
-We provide an R script to perform the above elicitation at 
+We provide an R script to perform the elicitation described above at 
 \url{https://github.com/fccoelho/paperLM1/blob/master/R/elicit_Rt_prior.R}.
 
+\subsection*{DiffeRential Evolution Adaptive Metropolis (DREAM)}
+
+Since the posterior distributions desired in this paper are not available in 
+closed-form, we need to resort to numerical methods to obtain 
+approximations.
+We employ an adaptive Markov chain Monte Carlo (MCMC) algorithm, proposed
+by~\cite{vrugt2008}, called  DiffeRential Evolution Adaptive Metropolis (DREAM).
+
+DREAM draws on the basic structure of Differential Evolution Markov Chain 
+(DE-MC) which runs $N$ independent chains in parallel and accepts moves 
+proportional to the difference of two randomly sampled members (chains), thus 
+differentially evolving conditional on the proposal variances.
+An additional aspect of DREAM is the so-called delayed rejection.
+This procedure adapts the original Metropolis-Hastings algorithm by attempting 
+new moves whenever a proposal is rejected (instead of setting the current state 
+to the previously accepted one), thus delaying the rejection of proposals. 
+This potentially decreases autocorrelation and improves mixing 
+(see~\cite{mira2001} for details).
+
+A desirable property of DREAM is that one can scale the algorithm with model 
+complexity, meaning one can initialise as many parallel chains as there are 
+parameters in the model, thus exploiting the multi-core architecture of modern 
+computers to speed up convergence of the MCMC.
+
+A Python implementation of DREAM for dynamic models is available in the package 
+Bayesian inference with Python (BIP)~\citep{pone2011} available at 
+\url{http://code.google.com/p/bayesian-inference/}.
+The code to produce the results presented in this paper is available 
+at~\url{https://github.com/fccoelho/paperLM1/tree/master/python}.
+
 \end{document}

### 2015-06-15 draft_lm1.tex
defined S_0 in equation \label{eq:effbeta}; closes #33
resolving confusing notation; fixes #31

@@
 Let $Y_{t} \sim Poisson(\lambda_t)$ and $Y_{t+1} \sim Poisson(\lambda_{t+1})$ 
-and define $S = Y_{t} + Y_{t+1}$.
-The authors note that by conditioning on the sum $S$
+and define $W = Y_{t} + Y_{t+1}$.
+The authors note that by conditioning on the sum $W$
 \begin{align}
 \label{eq:binlike}
-Y_{t+1} | S &\sim Binomial(S, \theta_t) \\
+Y_{t+1} | S &\sim Binomial(W, \theta_t) \\
 \theta_t &= \frac{\lambda_{t+1}}{\lambda_{t} + \lambda_{t+1}}
@@
 \label{eq:thetapost}
-p(\theta_t| Y_{t+1}, S) \sim Beta(Y_{t+1} + a_0, Y_t + b_0)
+p(\theta_t| Y_{t+1}, W) \sim Beta(Y_{t+1} + a_0, Y_t + b_0)
 \end{equation}
@@
 tells us that the induced posterior distribution of $R_t$ is 
-a beta prime (or inverted beta) with parameters $ a_1 = Y_{t+1} + a_0$ and $b_1 
+a beta prime (or inverted beta) with parameters $a_1 = Y_{t+1} + a_0$ and $b_1 
 =  Y_t + b_0$~\citep{dubey1970}.
@@
 where $R_t$ is the effective reproductive number, estimated as 
-in~\ref{eq:Rtestimate}.
+in~\ref{eq:Rtestimate} and $S_0$ is the initial fraction of susceptible 
+individuals.
 The complete model with the time-varying force of infection is given by

### 2015-06-15 draft_lm1.tex
sort of solves #30

@@
 In humans, immunity against a particular serotype is considered permanent 
-after the exposure and cross immunity to there serotypes is considered short 
+after the exposure and cross immunity to other serotypes is considered short 
 lived~\citep{halstead_dengue_2007}.

### 2015-06-15 draft_lm1.tex
fixes #26

@@
 Figure~\ref{fig:rtseries}  shows the $R_t$ series, according to 
-(\ref{eq:Rtestimate})~\citep{nishiura} along with the confidence bands derived 
+Equation~\ref{eq:Rtestimate}~\citep{nishiura} along with the confidence bands 
+derived 
 in this paper. 

### 2015-06-16 python/fit_simulated_data.py
minor tweaks to the script

@@ -19,7 +19,7 @@ import pandas as pd
 from sqlalchemy import create_engine
 import seaborn as sns
 # import numba
-#from numba import jit
+# from numba import jit
 
 
 beta = 1.0  # Transmission coefficient
@@ -34,14 +34,14 @@ tau = 1  # recovery rate. FIXED
 N = 1  # Population of Rio
 
 inicio = 0
-#~ Ss = {0: s0, 1: s1, 2: s2}  # Multiple Ss map
+# ~ Ss = {0: s0, 1: s1, 2: s2}  # Multiple Ss map
 
 
 # Initial conditions
 inits = np.array([0.999, 0.001, 0.0])  # initial values for state variables.
 
 
-#@jit
+# @jit
 def sir(y, t, *pars):
     '''ODE model'''
     S, I, R = y
@@ -51,13 +51,13 @@ def sir(y, t, *pars):
 
     lamb = beta * (I + m) * S
 
-    return np.array([-lamb,  #dS/dt
-                     lamb - tau * I,  #dI/dt
-                     tau * I,  #dR/dt
-    ])
+    return np.array([-lamb,  # dS/dt
+                     lamb - tau * I,  # dI/dt
+                     tau * I,  # dR/dt
+                     ])
 
 
-#@jit
+# @jit
 def jac(y, t, *pars):
     """
     Jacobian of the model
@@ -74,7 +74,7 @@ def jac(y, t, *pars):
                      [0, tau, 0]])
 
 
-#@jit
+# @jit
 def model(theta):
     """
     Wrap-up function which is called at every step of the MCMC by BIP's FitModel.
@@ -85,7 +85,7 @@ def model(theta):
     s0, m, tau = theta
 
     t0 = 0
-    tf = fim-inicio
+    tf = fim - inicio
     Y = np.zeros((tf - t0, 3))
 
 
@@ -93,8 +93,8 @@ def model(theta):
 
     inits[0] = s0  # Define S0
     inits[-1] = N - sum(inits[:2])  # Define R(0)
-    Y[t0:tf, :] = odeint(sir, inits, np.arange(t0, tf, 1), args=(s0, m, tau), Dfun=jac)  #,tcrit=tcrit)
-    #inits = Y[-1, :]
+    Y[t0:tf, :] = odeint(sir, inits, np.arange(t0, tf, 1), args=(s0, m, tau), Dfun=jac)  # ,tcrit=tcrit)
+    # inits = Y[-1, :]
[…]

### 2015-06-16 draft_lm1.tex
added citations suggested by the reviewer. closes #24

@@
 after the exposure and cross immunity to other serotypes is considered short 
-lived~\citep{halstead_dengue_2007}.
+lived~\citep{halstead_dengue_2007} although some 
+studies argue for a longer duration of cross-immunity \citep{Reich2013, 
+salje2012}.
 As a consequence, the proportions of viral serotypes co-circulating at any 

### 2015-06-16 scirep_version.tex
added citations suggested by the reviewer. closes #24

@@
 In humans, immunity against  a particular serotype is  considered permanent 
-after the exposure and cross immunity to there serotypes is considered short 
-lived~\cite{halstead_dengue_2007}.
+after the exposure and cross-immunity to the other serotypes is 
+generally considered short lived~\cite{halstead_dengue_2007} although some 
+studies argue for a longer duration of cross-immunity \cite{Reich2013, 
+salje2012}.
 As a consequence, the proportions of viral serotypes co-circulating at any 
@@
 \begin{thebibliography}{10}
-
 \expandafter\ifx\csname url\endcsname\relax
@@
 http://www.sciencedirect.com/science/article/pii/S0140673607616870}.
+
+\bibitem{Reich2013}
+\bibinfo{author}{Reich, N.~G.} \emph{et~al.}
+\newblock \bibinfo{title}{{{I}nteractions between serotypes of dengue highlight
+  epidemiological impact of cross-immunity}}.
+\newblock \emph{\bibinfo{journal}{J R Soc Interface}}
+  \textbf{\bibinfo{volume}{10}}, \bibinfo{pages}{20130414}
+  (\bibinfo{year}{2013}).
+
+\bibitem{salje2012}
+\bibinfo{author}{Salje, H.} \emph{et~al.}
+\newblock \bibinfo{title}{Revealing the microscale spatial signature of dengue
+  transmission and immunity in an urban population}.
+\newblock \emph{\bibinfo{journal}{Proceedings of the National Academy of
+  Sciences}} \textbf{\bibinfo{volume}{109}}, \bibinfo{pages}{9535--9538}
+  (\bibinfo{year}{2012}).
 
@@
   differential evolution with self-adaptive randomized subspace sampling}.
-\newblock \bibinfo{type}{Tech. Rep.}, \bibinfo{institution}{Citeseer}
-  (\bibinfo{year}{2008}).
+\newblock \emph{\bibinfo{journal}{International Journal of Nonlinear Sciences
+  and Numerical Simulation}} \textbf{\bibinfo{volume}{10}},
+  \bibinfo{pages}{271--288} (\bibinfo{year}{2008}).
 
@@
 
+\bibitem{mira2001}
+\bibinfo{author}{Mira, A.}
+\newblock \bibinfo{title}{On metropolis-hastings algorithms with delayed
+  rejection}.
+\newblock \emph{\bibinfo{journal}{Metron}} \textbf{\bibinfo{volume}{59}},
+  \bibinfo{pages}{231--241} (\bibinfo{year}{2001}).
+
 \end{thebibliography}

### 2015-06-22 python/fit_dengue_split_big.py
Finished the script for the simulated data.

@@ -120,12 +120,12 @@ def prepdata(fname, sday=0, eday=None, mao=7):
     data = pd.read_csv(fname, header=0, delimiter=',', skiprows=[1, 2, 3], parse_dates=True)
     # slicing to the desired period
     data = data[sday:eday]
-    pop = pd.read_csv("pop_rio_1980-2012.csv", header=0, delimiter=',', index_col=0)
+    pop = pd.read_csv("../DATA/pop_rio_1980-2012.csv", header=0, delimiter=',', index_col=0)
 
     dates = [datetime.datetime.strptime(d, "%Y-%m-%d") for d in data.start]
     pop_d = np.array([pop.loc[d.year] for d in dates])  # population on each date
 
-    eday = len(df) if eday is None else eday
+    eday = len(data) if eday is None else eday
     # print data.dtype.names
     incidence = data.cases  # daily incidence
     # Converting incidence to Prevalence
@@ -166,7 +166,7 @@ def fix_rt(rt):
 # # running the analysys
 if __name__ == "__main__":
     # dt = prepdata('aux/data_Rt_dengue_big.csv', 0, 728, 1)
-    dt = prepdata('data_Rt_dengue_complete.csv', 0, 971, 1)
+    dt = prepdata('../DATA/data_Rt_dengue_complete.csv', 0, 971, 1)
 
 
     # Defining start and end of the simulations
@@ -206,7 +206,7 @@ if __name__ == "__main__":
     iRt = interp1d(np.arange(dt['Rt'].size), np.array(dt['Rt']), kind='linear', bounds_error=False, fill_value=0)
 
     P.plot(dt['Rt'], '*', label='$R_t$')
-    print dt['I']/dt['I'].max()
+    print (dt['I']/dt['I'].max())
     P.plot(dt['I']/dt['I'].max(), 'k-+', label='log(Cases)')
     P.plot(np.arange(0, dt['Rt'].size, .2), [iRt(t) for t in np.arange(0, dt['Rt'].size, .2)])
     P.legend()
@@ -238,8 +238,8 @@ if __name__ == "__main__":
 
 
 
-    for inicio, fim in zip(t0s, tfs)[3:4]:  # Slice to force start from a different point
-        dt = prepdata('data_Rt_dengue_complete.csv', inicio, fim, 1)
+    for inicio, fim in zip(t0s, tfs):#[3:4]:  # Slice to force start from a different point
[…]

### 2015-06-22 python/fit_dengue_split_big_plot.py

@@ -11,7 +11,6 @@ from matplotlib import font_manager
 from matplotlib.ticker import FormatStrFormatter
 import matplotlib.dates as mdates
 import matplotlib.pyplot as P
-import Gnuplot
 from scipy.stats import gaussian_kde
 import datetime
 import matplotlib.collections as collections
@@ -85,7 +84,7 @@ def create_tex_table(dbs):
     body = r""
     st = []
     # years = sorted(list(series.keys()))
-    print series.keys()
+    print (series.keys())
     for i, (Y, V) in enumerate(series.items()):
         cases = obs[Y].sum()
         first_week = V.index[0]
@@ -101,9 +100,9 @@ def create_tex_table(dbs):
                                                                                       )
             body += "\n"
         except KeyError as e:
-            print Y, first_week, e
+            print (Y, first_week, e)
         except ValueError as e:
-            print s0, e
+            print (s0, e)
 
     return head + body + bot
 
@@ -251,7 +250,7 @@ if __name__ == "__main__":
     #plot_concat_series(dbs)
     #print create_tex_table(dbs)
     # plot_AR_S0(dbs)
-    print calc_mse(dbs)
+    print (calc_mse(dbs))
 
 
     P.show()

### 2015-06-22 python/fit_simulated_data.py

@@ -29,7 +29,7 @@ eta = .0  # infectivity of asymptomatic infections relative to clinical ones. FI
 
 mu = 0  # nat/mortality rate
 m = 1e-6  # influx of cases
-tau = 1  # recovery rate. FIXED
+tau = 1.  # recovery rate. FIXED
 
 N = 1  # Population of Rio
 
@@ -45,9 +45,8 @@ inits = np.array([0.999, 0.001, 0.0])  # initial values for state variables.
 def sir(y, t, *pars):
     '''ODE model'''
     S, I, R = y
-    s0, m, tau = pars
-    beta = 0 if t > 728 else iRt(t) * tau / s0
-    # print S, I, beta
+    s0, m = pars
+    beta = 0 if t > 728 else iRt(t)*tau / 0.0621
 
     lamb = beta * (I + m) * S
 
@@ -67,8 +66,8 @@ def jac(y, t, *pars):
     :return: the jacobian matrix at time t
     """
     S, I, R = y
-    s0, m, tau = pars
-    beta = 0 if t > 728 else iRt(t) * tau / s0
+    s0, m = pars
+    beta = 0 if t > 728 else iRt(t)*tau / 0.0621
     return np.array([[-(I + m) * beta, -S * beta, 0],
                      [(I + m) * beta, S * beta - tau, 0],
                      [0, tau, 0]])
@@ -82,18 +81,18 @@ def model(theta):
     :return: the output of the model
     """
     # setting parameters
-    s0, m, tau = theta
+    s0, m = theta
 
     t0 = 0
     tf = fim - inicio
     Y = np.zeros((tf - t0, 3))
 
-
+    # print(iRt(np.arange(t0, tf, 1)))
     # Initial conditions
 
     inits[0] = s0  # Define S0
     inits[-1] = N - sum(inits[:2])  # Define R(0)
-    Y[t0:tf, :] = odeint(sir, inits, np.arange(t0, tf, 1), args=(s0, m, tau), Dfun=jac)  # ,tcrit=tcrit)
+    Y[t0:tf, :] = odeint(sir, inits, np.arange(t0, tf, 1), args=(s0, m), Dfun=jac)  # ,tcrit=tcrit)
     # inits = Y[-1, :]
     Y[t0:tf, 1] = Y[t0:tf, 1] / 1  # Adjusting the output to just the reported I to compare with data
     return Y
@@ -121,7 +120,7 @@ def prepdata(fname, sday=0, eday=None, mao=7):
 
     incidence = data.cases  # daily incidence
[…]

### 2015-06-22 Presentations/Dinamica de Vetores-SP-2015/dengueARS0.tex

@@
 \end{flushleft}
+
+\end{frame}
+
+\subsection{Vector dynamics}
+
+\begin{frame}
+\frametitle{Environmental determinants}
+\begin{block}{Vector dynamics}
+\begin{itemize}[<+->]
+ \item A. Aegypti population dynamics display marked seasonality
+ \item Temperature, Humidity and rainfall are important factors
+ \item Environmental stock of eggs
+ \item Effects on mosquito reproduction are non-linear
+ \item Delayed influence
+\end{itemize}
+\end{block}
 
@@
 \end{center}
-\end{frame}
-\subsection{Vector dynamics}
-\begin{frame}
-\frametitle{Environmental determinants}
-\begin{itemize}[<+->]
- \item A. Aegypti population dynamics display marked seasonality
- \item Temperature, Humidity and rainfall are important factors
- \item Environmental stock of eggs
- \item Effects on mosquito reproduction are non-linear
- \item Delayed influence
-\end{itemize}
-
 \end{frame}
@@
 \end{frame}
-\begin{frame}
- \frametitle{Collaboration possibilities}
- \begin{itemize}
-  \item Relaxing other simplifying assumptions in classical epidemiological 
-models.
-\item Looking into modeling spatio-temporal dynamics taking into account 
-Environmental stock of eggs.
-\item Separating noise from relevant varibility in observational data, and 
-finding new ways of integrating it into models.
- \end{itemize}
-
-\end{frame}
+% \begin{frame}
+%  \frametitle{Collaboration possibilities}
+%  \begin{itemize}
+%   \item Relaxing other simplifying assumptions in classical epidemiological 
+% models.
+% \item Looking into modeling spatio-temporal dynamics taking into account 
+% Environmental stock of eggs.
+% \item Separating noise from relevant varibility in observational data, and 
+% finding new ways of integrating it into models.
+%  \end{itemize}
+% 
+% 
+% \end{frame}
 \begin{frame}

### 2015-06-22 suppl.tex
Added section to supplementary material fixing #21. closes #21

@@
 \usepackage{url}
+\usepackage{graphicx}
 % Our defs 
@@
 
+\section*{Estimating $S_0$ from simulated data}
+In order to make absolutely clear that the inference methodology proposed can 
+recover the true parameter values of the underlying simulation model, we have 
+devised a simple simulation experiment. Using the SIR model presented in the 
+main paper we have simulated the incidence curve of the  2013 epidemic (2012 in 
+table 1 of the article), using $S_0 = 0.0621$ and \rr, as estimated from the 
+actual incidence data, and $\tau=1$. Figure \ref{fig:sim_data} shows the 
+posterior $S$ and $I$ alongside with the simulated incidence data. It is clear 
+the inference can recover the correct value for $S_0$ used to   simulate 
+incidence. The script to generate the simulated data and run the inference is in 
+the paper's Github repository(\verb|fit_simulated_data.py|).
+
+\begin{figure}
+ \centering
+ \includegraphics[width=14cm]{./plots/Sim_DengueS2012_11_series.png}
+ % Sim_DengueS2012_11_series.png: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=
+ \caption{$S$ and $I$ series fitted to simulated incidence data(blue dots) 
+with $S_0=0.0621$, and \rr estimated from real incidence data.}
+ \label{fig:sim_data}
+\end{figure}
+
 \bibliography{lm1}

### 2015-06-22 supplSR.tex
Added section to supplementary material fixing #21. closes #21

@@
 
+\section*{Estimating $S_0$ from simulated data}
+In order to make absolutely clear that the inference methodology proposed can 
+recover the true parameter values of the underlying simulation model, we have 
+devised a simple simulation experiment. Using the SIR model presented in the 
+main paper we have simulated the incidence curve of the  2013 epidemic (2012 in 
+table 1 of the article), using $S_0 = 0.0621$ and \rr, as estimated from the 
+actual incidence data, and $\tau=1$. Figure \ref{fig:sim_data} shows the 
+posterior $S$ and $I$ alongside with the simulated incidence data. It is clear 
+the inference can recover the correct value for $S_0$ used to   simulate 
+incidence. The script to generate the simulated data and run the inference is in 
+the paper's Github repository(\verb|fit_simulated_data.py|).
+
+\begin{figure}
+ \centering
+ \includegraphics[width=14cm]{./plots/Sim_DengueS2012_11_series.png}
+ % Sim_DengueS2012_11_series.png: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=
+ \caption{$S$ and $I$ series fitted to simulated incidence data(blue dots) 
+with $S_0=0.0621$, and \rr estimated from real incidence data.}
+ \label{fig:sim_data}
+\end{figure}
+
 \begin{thebibliography}{1}

### 2015-06-23 Correspondence/revisions.tex
added response letter; added some responses to the manuscript. changes the code which generates the table in the paper.

\documentclass[a4paper,10pt]{letter}
\usepackage[utf8]{inputenc}

\begin{document}
% If you want headings on subsequent pages,
% remove the ``%'' on the next line:
% \pagestyle{headings}

\begin{letter}{Laura White\\Editorial Board Member\\Scientific Reports }
\address{Escola de Matem\'atica Aplicada\\Funda\c{c}\~ao Getulio Vargas 
(FGV)\\Rio de Janeiro -- RJ\\Brazil.}

\opening{Dear editor,}

In response to the comments of the reviewers of our manuscript, we have made 
the following modifications to the manuscript:

\begin{description}
 \item[Reviewer 1:] responses are listed below.

 \begin{enumerate}
  \item A simulation experiment was done, as requested and included as a new 
section in the supplementary material. The simulations shows that the inference 
methodology can recover the exact parameters user to simulate  the data.
\item The text of the methods section was improved to answer the four issues 
raised in this comment.
 \end{enumerate}

 \item [Reviewer 2:]
\end{description}

\signature{Flávio Codeço Coelho\\Professor}

\closing{Sincerely}

%enclosure listing
%\encl{}

\end{letter}
\end{document}

### 2015-06-23 draft_lm1.tex
Removed m from equation; closes #39
added response letter; added some responses to the manuscript. changes the code which generates the table in the paper.

@@
 empirically derived $R_t$  to incorporate the effects of the fluctuating vector 
-populations.
+populations. Although demography is not included in the model -- population is 
+assumed to remain constant throughout an epidemic -- the population variation 
+from year to year is taken into consideration as the prevalences are calculated 
+for each year by dividing the number of cases by the official population 
+estimated by the census. In Brazil, dengue affects all age groups which still 
+haven't been exposed to all 4 serotypes. This is in contrast with countries in 
+which the four serotypes have been endemic for a very long period of time, 
+where dengue mostly affect the youth.
+
 Also, although there are multiple circulating serotypes, our approach
@@
  \label{eq:S0post}
- p(S_{0j}|\mathbf{Y_{j}}) \propto l(\mathbf{Y_{j}}|S_{0j}, R_t, m, \tau 
+ p(S_{0j}|\mathbf{Y_{j}}) \propto l(\mathbf{Y_{j}}|S_{0j}, R_t, \tau 
 )\pi(S_{0j}) 
@@
 Values are presented as percentage of total population. 
-$^\dag$: Year corresponds to the start of the epidemic, however the peak of cases
+$^\dag$: Year corresponds to the start of the epidemic, but the peak of cases
 may occur in the following year.
-$^\ddag$: Susceptible fraction.
+$^\ddag$: percentage of population.
 These results show considerable variation in AR between epidemics, consistent with
@@
 \begin{center}
-\begin{tabular}{c|c|c}
+\begin{tabular}{l|c|c|c}
 \hline
-Year$^\dag$ & median Attack Ratio & $S_0^\ddag$ \\
+$^\dag$Year & $^\ddag$Cases & median Attack Ratio & $^\ddag$$S_0$ \\
 \hline
-1996 & 0.39 (0.17-0.54) & 0.00171(0.0012-0.0038)\\
-1997 & 0.87 (0.74-0.87) & 0.00273(0.0027-0.0032)\\
-1998 & 0.5 (0.49-0.5) & 0.00142(0.0014-0.0014)\\
-1999 & 0.11 (0.037-0.2) & 0.00345(0.0018-0.01)\\
-2000 & 0.25 (0.24-0.27) & 0.0155(0.015-0.016)\\
-2001 & 0.48 (0.47-0.49) & 0.0495(0.048-0.051)\\
-2005 & 0.15 (0.1-0.21) & 0.0147(0.01-0.021)\\
-2006 & 0.11 (0.08-0.14) & 0.0281(0.022-0.037)\\
-2007 & 0.15 (0.15-0.15) & 0.135(0.13-0.14)\\
-2008 & 0.14 (0.031-0.31) & 0.00672(0.003-0.024)\\
-2010 & 0.18 (0.17-0.19) & 0.0454(0.043-0.048)\\
-2011 & 0.086 (0.082-0.094) & 0.215(0.2-0.23)\\
-2012 & 0.14 (0.13-0.15) & 0.0621(0.058-0.068)\\
+1996 & 0.066 & 0.39 (0.17-0.54) & 0.171(0.12-0.38)\\
+1997 & 0.238 & 0.87 (0.74-0.87) & 0.273(0.27-0.32)\\
+1998 & 0.0708 & 0.5 (0.49-0.5) & 0.142(0.14-0.14)\\
+1999 & 0.0371 & 0.11 (0.037-0.2) & 0.345(0.18-1.0)\\
+2000 & 0.394 & 0.25 (0.24-0.27) & 1.55(1.5-1.6)\\
+2001 & 2.38 & 0.48 (0.47-0.49) & 4.95(4.8-5.1)\\
+2005 & 0.217 & 0.15 (0.1-0.21) & 1.47(1.0-2.1)\\
+2006 & 0.315 & 0.11 (0.08-0.14) & 2.81(2.2-3.7)\\
+2007 & 2.03 & 0.15 (0.15-0.15) & 13.5(1.3e+01-1.4e+01)\\
+2008 & 0.0923 & 0.14 (0.031-0.31) & 0.672(0.3-2.4)\\
+2010 & 0.831 & 0.18 (0.17-0.19) & 4.54(4.3-4.8)\\
+2011 & 1.85 & 0.086 (0.082-0.094) & 21.5(2e+01-2.3e+01)\\
+2012 & 0.864 & 0.14 (0.13-0.15) & 6.21(5.8-6.8)\\
+
 \hline

### 2015-06-23 python/fit_dengue_split_big_plot.py

@@ -2,7 +2,7 @@
 Script to generate the plots for the
 multi-year simulations
 """
-import cPickle as CP
+import pickle as CP
 from itertools import cycle
 from scipy import stats
 import glob
@@ -70,10 +70,10 @@ def create_tex_table(dbs):
     obs, series, pts = get_ordered_series(dbs)
 
     head = r"""\begin{center}
-\begin{tabular}{c|c|c}
+\begin{tabular}{l|c|c|c}
 \hline
 """
-    head += r"""Year & median Attack Ratio $ $S_0$ \\
+    head += r"""Year & Cases & median Attack Ratio $ $S_0$ \\
 \hline
 """
     bot = r"""
@@ -91,7 +91,7 @@ def create_tex_table(dbs):
         s0 = array(series[Y].S.ix[first_week])
         try:
             ratio = 1.0*cases/s0
-            body += Y + r" & {:.2} ({:.2}-{:.2}) & {:.3}({:.2}-{:.2})\\".format(nanmedian(ratio),
+            body += Y + r" & {:.3} & {:.2} ({:.2}-{:.2}) & {:.3}({:.2}-{:.2})\\".format(cases*100, nanmedian(ratio),
                                                                                       stats.scoreatpercentile(ratio, 2.5),
                                                                                       stats.scoreatpercentile(ratio, 97.5),
                                                                                       nanmedian(s0)*100,
@@ -233,7 +233,7 @@ def calc_mse(dbs):
     mse = OrderedDict()
     mae = OrderedDict()
     mape = OrderedDict()
-    for Y, srs in series.iteritems():
+    for Y, srs in series.items():
         n = len(obs[Y])
         i_median = srs.I.groupby(level='time').median()
         mse[Y] = (1./n) * sum(((i_median-obs[Y])**2)/obs[Y]**2)
@@ -248,7 +248,7 @@ if __name__ == "__main__":
 
     dbs = glob.glob("../DengueS*.sqlite")
     #plot_concat_series(dbs)
-    #print create_tex_table(dbs)
+    print(create_tex_table(dbs))
     # plot_AR_S0(dbs)
     print (calc_mse(dbs))
 

### 2015-06-23 python/fit_dengue_split_big.py
updated letter

@@ -11,14 +11,8 @@ import scipy.stats as st
 import numpy as np
 import pylab as P
 import copy
-from collections import defaultdict
 import datetime
-import sys
 import pandas as pd
-import seaborn as sns
-# import numba
-#from numba import jit
-
 
 beta = 1.0  # Transmission coefficient
 b1 = 19.9
@@ -32,14 +26,13 @@ tau = 1  # recovery rate. FIXED
 N = 1  # Population of Rio
 
 inicio = 0
-#~ Ss = {0: s0, 1: s1, 2: s2}  # Multiple Ss map
+# ~ Ss = {0: s0, 1: s1, 2: s2}  # Multiple Ss map
 
 
 # Initial conditions
 inits = np.array([0.999, 0.001, 0.0])  # initial values for state variables.
 
 
-#@jit
 def sir(y, t, *pars):
     '''ODE model'''
     S, I, R = y
@@ -49,13 +42,12 @@ def sir(y, t, *pars):
 
     lamb = beta * (I + m) * S
 
-    return np.array([-lamb,  #dS/dt
-                     lamb - tau * I,  #dI/dt
-                     tau * I,  #dR/dt
-    ])
+    return np.array([-lamb,  # dS/dt
+                     lamb - tau * I,  # dI/dt
+                     tau * I,  # dR/dt
+                     ])
 
 
-#@jit
 def jac(y, t, *pars):
     S, I, R = y
     s0, m, tau = pars
@@ -65,7 +57,6 @@ def jac(y, t, *pars):
                      [0, tau, 0]])
 
 
-#@jit
 def model(theta):
     # setting parameters
     s0, m, tau = theta
@@ -76,11 +67,11 @@ def model(theta):
 
 
     # Initial conditions
-    #print inits, m
+    # print inits, m
     inits[0] = s0  # Define S0
     inits[-1] = N - sum(inits[:2])  # Define R(0)
-    Y[t0:tf, :] = odeint(sir, inits, np.arange(t0, tf, 1), args=(s0, m, tau), Dfun=jac)  #,tcrit=tcrit)
-    #inits = Y[-1, :]
+    Y[t0:tf, :] = odeint(sir, inits, np.arange(t0, tf, 1), args=(s0, m, tau), Dfun=jac)  # ,tcrit=tcrit)
+    # inits = Y[-1, :]
     Y[t0:tf, 1] = Y[t0:tf, 1] / 1  # Adjusting the output to just the reported I to compare with data
     return Y
 
[…]

### 2015-06-23 draft_lm1.tex
minor comment 6 closes #34

@@
 Quantifying the attack ratio of disease is key to epidemiological inference and 
-Public Health planning.
+public health planning.
 For multi-serotype pathogens, however, different levels of 
@@
 \end{equation}
-The likelihood $l(\mathbf{Y_{j}}|\cdot)$ is assumed to be a Normal distribution 
-with fixed variance $\sigma^2$.
+The likelihood $l(\mathbf{Y_{j}}|\cdot)$ is a approximated by a Normal 
+distribution with fixed variance $\sigma^2$. This approximation is a numerical 
+convenience since it confers better stability to the MCMC sampling.
 In this estimation procedure we kept $R_t$ fixed at fixed at the posterior mean 

### 2015-07-06 Correspondence/revisions.tex
updated letter
fixed typo
minor comment 6 closes #34
Done responding to comments 2 and 3 of reviewer 1 and described the changes in the letter. Closes #22 Closes #23

@@
 \begin{description}
- \item[Reviewer 1:] responses are listed below.
+ \item[Reviewer 1:] Responses to the main comments are listed below. The minor 
+comments where also accepted, namely the addition of readme do the code 
+repository as well as the citation of two  additional papers.
 
@@
 raised in this comment.
+\item The introduction and discussion sections were expanded to included the 
+additional information required by this comment.
  \end{enumerate}
 
- \item [Reviewer 2:]
+ \item [Reviewer 2:] \textbf{Minor comments:} Fixed typo in the abstract; 
+ \begin{enumerate}
+  \item 
+ \end{enumerate}
+
 \end{description}

### 2015-07-06 draft_lm1.tex
Done responding to comments 2 and 3 of reviewer 1 and described the changes in the letter. Closes #22 Closes #23

@@
 Dengue is an arthropod-borne febrile disease caused by a flavivirus with four 
-serotypes which causes an estimated  $50$ million infections each 
+serotypes (DEN-1, 2, 3 and 4) which causes an estimated  $50$ million 
+infections each 
 year~\citep{Guzman2010}.
@@
 We apply our method to estimate $S_0$ before every major dengue epidemic in the 
-city of Rio de Janeiro, Brazil in the last 18 years.
+city of Rio de Janeiro, Brazil in the last 18 years. Dengue has been 
+reintroduced in Rio de Janeiro in 1986 after being absent for 68 
+years\citep{nogueira1999dengue}. During the last decades of the 20th century 
+only DEN-1 and DEN-2 serotypes
 
@@
 estimated for each epidemic/outbreak.
-It is interesting to notice that the larger epidemics, in terms of peak size are 
-not the one with the greater attack ratios.
+ Underreporting of cases, which is known to exist but of which exact figures 
+cannot be determined, will lead to underestimation $S_0$. However, The attack 
+ratio shall remain unbiased as the underreporting affects both the numerator and 
+the denominator of equation \ref{eq:AR2}.
+It is interesting to notice that the larger epidemics, in terms of peak size 
+are not the one with the greater attack ratios.
 This stresses the importance of knowing the immunological structure of the 
@@
 accurately assess the potential impact of a coming epidemic, since particularly 
-virulent types, can be rendered less of a threat by a low $S_0$.
+virulent types, can be rendered less of a threat by a low $S_0$. Hon\'orio et 
+al.\citep{honorio2009spatial} conducted a serological survey in three separate 
+localities within the city, right before the 2008 epidemic, the authors report 
+seroprevalences varying from 56-77.4\% which is compatible with our prediction 
+of 87.5\% ($1-S_{0,2007}$) for the entire city, considering that we 
+underestimate $S_0$ due to the underreporting of cases.
 
@@
 \begin{center}
-\begin{tabular}{l|c|c|c}
+\begin{tabular}{l|c|c|c|c}
 \hline
-$^\dag$Year & $^\ddag$Cases & median Attack Ratio & $^\ddag$$S_0$ \\
+$^\dag$Year & $^\ddag$Cases & median Attack Ratio & $^\ddag$$S_0$ & Circulating 
+ Serotypes \\
 \hline
-1996 & 0.066 & 0.39 (0.17-0.54) & 0.171(0.12-0.38)\\
-1997 & 0.238 & 0.87 (0.74-0.87) & 0.273(0.27-0.32)\\
-1998 & 0.0708 & 0.5 (0.49-0.5) & 0.142(0.14-0.14)\\
-1999 & 0.0371 & 0.11 (0.037-0.2) & 0.345(0.18-1.0)\\
-2000 & 0.394 & 0.25 (0.24-0.27) & 1.55(1.5-1.6)\\
-2001 & 2.38 & 0.48 (0.47-0.49) & 4.95(4.8-5.1)\\
-2005 & 0.217 & 0.15 (0.1-0.21) & 1.47(1.0-2.1)\\
-2006 & 0.315 & 0.11 (0.08-0.14) & 2.81(2.2-3.7)\\
-2007 & 2.03 & 0.15 (0.15-0.15) & 13.5(1.3e+01-1.4e+01)\\
-2008 & 0.0923 & 0.14 (0.031-0.31) & 0.672(0.3-2.4)\\
-2010 & 0.831 & 0.18 (0.17-0.19) & 4.54(4.3-4.8)\\
-2011 & 1.85 & 0.086 (0.082-0.094) & 21.5(2e+01-2.3e+01)\\
-2012 & 0.864 & 0.14 (0.13-0.15) & 6.21(5.8-6.8)\\
+1996 & 0.066 & 0.39 (0.17-0.54) & 0.171(0.12-0.38)& DEN-1 and 
+2\citep{nogueira1999dengue}\\
+1997 & 0.238 & 0.87 (0.74-0.87) & 0.273(0.27-0.32)& DEN-1 and 
+2\citep{nogueira1999dengue}\\
+1998 & 0.0708 & 0.5 (0.49-0.5) & 0.142(0.14-0.14)& DEN-1 and 
+2\citep{nogueira1999dengue}\\
+1999 & 0.0371 & 0.11 (0.037-0.2) & 0.345(0.18-1.0)&DEN-1, 2 and 
+3\citep{de2004dengue}\\
+2000 & 0.394 & 0.25 (0.24-0.27) & 1.55(1.5-1.6)&DEN-1, 2 and 
+3\citep{de2004dengue}\\
+2001 & 2.38 & 0.48 (0.47-0.49) & 4.95(4.8-5.1)&DEN-1, 2 and 
+3\citep{de2004dengue}\\
+2005 & 0.217 & 0.15 (0.1-0.21) & 1.47(1.0-2.1)&\\
+2006 & 0.315 & 0.11 (0.08-0.14) & 
+2.81(2.2-3.7)&DEN-2\citep{nogueira2011dengue}\\
+2007 & 2.03 & 0.15 (0.15-0.15) & 
+13.5(13.0-14.01)&DEN-2\citep{nogueira2011dengue}\\
+2008 & 0.0923 & 0.14 (0.031-0.31) & 
+0.672(0.3-2.4)&DEN-2\citep{nogueira2011dengue}\\
+2010 & 0.831 & 0.18 (0.17-0.19) & 
+4.54(4.3-4.8)&DEN-2\citep{nogueira2011dengue}\\
+2011 & 1.85 & 0.086 (0.082-0.094) & 
+21.5(20.0-23.0)&DEN-1, 2 and 4\citep{nogueira2011dengue}\\
+2012 & 0.864 & 0.14 (0.13-0.15) & 6.21(5.8-6.8)&\\
 

### 2015-07-06 Correspondence/revisions.tex
finished responding to reviewer 2's comments. Closes #27, closes #28, closes #37

@@
 
- \item [Reviewer 2:] \textbf{Minor comments:} Fixed typo in the abstract; 
+ \item [Reviewer 2:] \textbf{Minor comments:} Fixed typo in the abstract; Added 
+text to the discussion stating that we expect the effects of underreporting to 
+cancel out in the calculation of the Attack ratio. 
  \begin{enumerate}
-  \item 
+  \item Reference 5, despite having similar concepts as those discussed in our 
+paper, has quite different goals. we have however, cited another publication 
+(ref 26), which fro the same group and based on the same field study, and use 
+their findings about seroprevalence in the 'results and discussion' section.
+\item We have not found articles combining the use of aggregated data to study 
+variable force of infection, however we have reviewed other works modeling 
+variable FoI (refs 7, 8, 9, 10, 12).
  \end{enumerate}

### 2015-07-06 Correspondence/revisions.tex
improved letter

@@
 paper, has quite different goals. we have however, cited another publication 
-(ref 26), which fro the same group and based on the same field study, and use 
-their findings about seroprevalence in the 'results and discussion' section.
+(ref 26), which is from the same group and based on the same field study, and 
+use their findings about seroprevalence in the 'results and discussion' section.
 \item We have not found articles combining the use of aggregated data to study 
@@
 \end{description}
+
+Having fulfilled all the demands of the reviewers we re-submit the modified 
+manuscript and await the the final decision. As always, please do not hesitate 
+to contact us if further questions arise.
 

### 2015-07-25 R/Rt_aux.R
code for simulation study

@@ -1,16 +1,16 @@
 #############################
-## obtain 100\lapha confidence/credibility intervals for the success probability \theta
+## obtain 100\% alpha confidence/credibility intervals for the success probability \theta
 betaconf <- function(alpha = .95, x, n, a = 1, b = 1, CP = "FALSE"){
   if(CP=="TRUE"){  
     lower <- 1 - qbeta((1-alpha)/2, n + x - 1, x)
     upper <- 1 - qbeta((1+alpha)/2, n - x, x + 1)
   }else{
-    lower <- qbeta( (1-alpha)/2, a + x, b + n - x)
-    upper <- qbeta(( 1+alpha)/2, a + x, b + n - x)  
+    lower <- qbeta((1-alpha)/2, a + x, b + n - x)
+    upper <- qbeta((1+alpha)/2, a + x, b + n - x)  
   } 
   return(c(lower, upper))
-  #CP stands for Clopper-Pearson
-  #Default is 'Bayesian' with an uniform prior over p (a=b=1)
+  # CP stands for Clopper-Pearson
+  # Default is 'Bayesian' with an uniform prior over p (a=b=1)
 }
 #############################
 # R = theta/(1-theta)
@@ -34,9 +34,9 @@ Rt.beta <- function(d, gt = 3, alpha = .95, a0 = 2 , b0 = 3){
   CIs <- data.frame (p1 = rep(NA, N), lwr = rep(NA, N), upr = rep(NA, N) )
   ## p1 = Pr(R>1) 
   for( k in 1: (N-gt)){
-    CIs[k+gt, 1] <- 1 - pbeta(.5, shape1 = jk1[k], shape2 = jk[k])
+    CIs[k+gt, 1] <- 1 - pbeta(.5, shape1 = jk1[k] + a0 , shape2 = jk[k] + b0)
     CIs[k+gt, 2:3] <- ll(betaconf(alpha = alpha, x = jk1[k], 
-                                  n = jk1[k] + jk[k], a = a0, b = b0 ))
+                                  n = jk1[k] + jk[k], a = a0, b = b0))
   } 
   return(data.frame(Rt, Rt2, CIs))
 }
@@ -73,4 +73,31 @@ Rt.gamma <- function(d, gt = 3, alpha = .95){
                              shape = jk1[k] + 1, rate = jk[k] )
   } 
   return(data.frame(Rt, CIs))
+}
+###
[…]

### 2015-07-25 R/credibility_comparison.R

@@ -1,0 +1,51 @@
+## This script will compare the credibility intervals obtained using the methods of
+## Nishiura et al. (2010) [slightly modified] and Coelho & Carvalho (2015)
+# generate the data
+set.seed(666)
+lambdas <- expand.grid(l1 = seq(2,  1000, length.out = 10),
+                       l2 = seq(2, 1000, length.out = 10))
+counts <- apply(lambdas, 1, function(x){
+  data.frame(Y1 = rpois(10000, lambda = x[1]),
+             Y2 = rpois(10000, lambda = x[2]))
+})
+
+# calculate empirical "Rt"
+getempRt <- function(dt){
+  dtt <- subset(subset(dt, Y1 > 0), Y2 > 0)
+  return(dtt$Y2/dtt$Y1)
+}
+empRts <- lapply(counts, getempRt)
+# lapply(empRts, hist)
+
+# calculating posterior means and credible intervals
+source("Rt_aux.R")
+nzcounts <- lapply(counts, function(d) subset(subset(d, Y1 > 0), Y2 > 0))
+
+betas <- lapply(nzcounts,
+                function(d){
+                  apply(d, 1, 
+                        function(x) estBeta(y1 = x[1], y2 = x[2]))})
+
+gammas <- lapply(nzcounts,
+                function(d){apply(d, 1,
+                                  function(x) estGamma(y1 = x[1], y2 = x[2]))})
+
+CIcomparisons <- vector(length(betas), mode = "list")
+for (i in 1:length(betas)){
+  indexes <- 1:length(betas[[i]])
+  tempcompz <- sapply(indexes, function(j) compareCI(c1 = betas[[i]][[j]]$CI,
+                                                           c2 = gammas[[i]][[j]]$CI))
+  CIcomparisons[[i]] <- table(tempcompz)/sum(table(tempcompz))
+}
+CIcomparisons
+## integrating over the grid
+betas.total <- unlist(betas, recursive = FALSE)
+gammas.total <- unlist(gammas, recursive = FALSE)
+inds <- 1:length(betas.total)
+compz.total <- sapply(inds, function(j) compareCI(c1 = betas.total[[j]]$CI,
+                                            c2 = gammas.total[[j]]$CI))
+prob.compz.total <- sapply(inds,
+                           function(j) betas.total[[j]]$pmt1 < gammas.total[[j]]$pmt1)
+
+table(prob.compz.total)/sum(table(prob.compz.total))
+table(compz.total)/sum(table(compz.total))

### 2015-07-25 draft_lm1.tex
first commit of LM reviews; description of the simulation study

@@
 % % % % % % % % % % % % % % % % % % % % % %
-%
 \documentclass[10pt]{article}
@@
 \def \rr {$R_{t}\:$}
-
 % Leave date blank
 \date{}
-
 \pagestyle{myheadings}
@@
 We show how to elicit the prior for $R_t$ from specified prior mean and variance
-or coefficient of variation in the \textbf{Appendix}.
+or coefficient of variation in the~\textbf{Appendix}.
 
@@
 indistinguishable from 1.
-Due to the intrinsic variability of the \rr series, the examination of its 
+Due to the intrinsic variability of the~\rr series, the examination of its 
 credible intervals is essential to identify periods of sustained transmission.
@@
 these periods. 
-The method to quantify uncertainty proposed here provides more conservative 
-credibility intervals, and therefore offers protection against false alarms.
+credibility intervals, and therefore offers protection against false alarms (see 
+the section on tail behaviour in the \textbf{Appendix} for a detailed 
+explanation).
 \begin{center}
@@
 ratio shall remain unbiased as the underreporting affects both the numerator and 
-the denominator of equation \ref{eq:AR2}.
+the denominator of equation~\ref{eq:AR2}.
 It is interesting to notice that the larger epidemics, in terms of peak size 
@@
 \label{sec:tails}
-There are a number of approaches to deriving the distribution of \rr.
+There are a number of approaches to deriving the distribution of~\rr.
 Alternatively to the approach described in the main text~\citep{mantel}, one 
@@
 tail behaviour of the derived distributions for \rr. 
-Consider the case of using a flat $Uniform(0, 1)$ prior for $\theta_t$.
+Consider the case of using a flat $\text{Uniform}(0, 1)$ prior for $\theta_t$.
 With $a_0 = b_0 = 1$, $a_1 = a_2$ and $b_1 = b_2 + 1$.
@@
 \end{equation}
-Finally, note that we deliberately construct $c_{\alpha}(R_{t})$ as a 
+Note also that we deliberately construct $c_{\alpha}(R_{t})$ as a 
 equal-tailed $100\alpha\%$ credible set, rather than a less conservative 
 highest posterior density (HPD) interval.
+Finally, we performed a simple simulation study to assess the result in equation
+~\ref{seq:densityratio}.
+The simulations were carried out as follows:
+\begin{enumerate}
+ \item Create a two-dimensional grid of values for $\lambda_1$ and $\lambda_2$, 
+with rate values from $2$ to $1000$;
+\item For each point $(\lambda_1^{(j)}, \lambda_2^{(j)})$ in the grid,
+\begin{itemize}
+ \item Generate $1000$ realisations of the random variables $Y_{1}^{(ij)} \sim 
+\text{Poisson}(\lambda_1^{(j)})$ and $Y_{2}^{(ij)} \sim 
+\text{Poisson}(\lambda_2^{(j)})$, with $i= 1, 2, \ldots, 1000$  and compute 
+$R_{t}^{(ij)} = Y_{2}^{(ij)}/
+Y_{1}^{(ij)}$;
+\item Compute the $\alpha\%$ credibility/confidence intervals using the method 
+proposed here $(\beta)$ and using the distribution proposed by Nishiura 
+(2010)~\cite{nishiura} ($\Gamma$), denoted by $c_{\beta}(R_{t}^{(ij)}; 
+\alpha)$ and 
+$c_{\Gamma}(R_{t}^{(ij)}; \alpha)$ respectively;
+\item Determine whether
+\begin{itemize}
+ \item [(i)] $ c_{\Gamma}(R_{t}^{(ij)}; \alpha) \subseteq 
+c_{\beta}(R_{t}^{(ij)}; \alpha)$;
+\item [(ii)] $c_{\beta}(R_{t}^{(ij)}; \alpha) \subseteq 
+c_{\Gamma}(R_{t}^{(ij)}; \alpha)$ or;
+\item [(iii)] there is overlap
+\end{itemize}
+\item Calculate $Pr(R_{t}^{(ij)} > 1)$ using both distributions
+\end{itemize}
+\end{enumerate}
+The results show that the credibility intervals obtained using the distribution 
+proposed here were larger than (i.e. contained) those computed using the 
+distribution in~\ref{seq:densityNishiura} in $90.69\%$ of the simulations 
+(integrating over the grid), while the converse was observed in $1.14\%$ of the 
+cases.
+Moreover, our method yielded lower $Pr(R_{t} > 1)$  in $79.48\%$ of the 
+simulations, indicating it is indeed more robust to false alarm.
+An R script to perform the simulation study described above can be found 
+at~\url{
+https://github.com/fccoelho/paperLM1/blob/master/R/credibility_comparison.R}.
 
@@
 Under the flat  uniform prior for $\theta_t$, the Bayesian posterior 
-credibility 
-interval is nearly indistinguishable from the confidence interval proposed by 
-Clopper \& Pearson (1931)~\citep{clopper} for $Y_{t+1}, Y_t > 20$.
+credibility interval is nearly indistinguishable from the confidence interval 
+proposed by Clopper \& Pearson (1931)~\citep{clopper} for $Y_{t+1}, Y_t > 20$.
 Note also that the uniform prior ($Beta(1, 1)$) for $\theta_t$ constitutes a 
-poor 
-prior 
-choice mainly because the induced distribution for \rr is only well-defined for 
-$b_0 > 2$.
+poor prior choice mainly because the induced distribution for \rr is only 
+well-defined for $b_0 > 2$.
 
@@
 and flexibility of the beta family of distributions.
-Prior elicitation can also be done for \rr and the hyper-parameters directly 
+Prior elicitation can also be done for~\rr and the hyper-parameters directly 
 plugged into the prior for $\theta_t$. 
-One can, for example, choose prior mean and variance for \rr and find $a_0$ 
+One can, for example, choose prior mean and variance for~\rr and find $a_0$ 
 and $b_0$ that satisfy those conditions.

### 2015-07-25 draft_lm1.tex
LM reviews to the manuscript

@@
 model, since the arrival/re-emergence of  new serotypes, an intrinsically 
-stochastic events, can drastically change the pool of susceptibles, throwing off 
+stochastic event, can drastically change the pool of susceptibles, throwing off 
 any sequential estimation based on the  incidence dynamics.
+
+Dengue has been reintroduced in Rio de Janeiro in 1986 after being absent for 
+68 years~\citep{nogueira1999dengue}.
+During the last decades of the 20th century only DEN-1 and DEN-2 serotypes
 
@@
 seroprevalence and the exact behaviour of the force of infection through time, 
-we
-propose to inform the time-varying transmissibility using the \rr series 
+we propose to inform the time-varying transmissibility using the \rr series 
 derived from the notification data~\citep{nishiura}.
@@
 We apply our method to estimate $S_0$ before every major dengue epidemic in the 
-city of Rio de Janeiro, Brazil in the last 18 years. Dengue has been 
-reintroduced in Rio de Janeiro in 1986 after being absent for 68 
-years\citep{nogueira1999dengue}. During the last decades of the 20th century 
-only DEN-1 and DEN-2 serotypes
+city of Rio de Janeiro, Brazil in the last 18 years.
 
@@
 
-We explore the approach of Ederer and Mantel~\citep{mantel}, whose objective is 
-to obtain 
-confidence intervals for the ratio of two Poisson counts. 
+We explore the approach of Ederer and Mantel~\citep{mantel}, whose objective 
+was  to obtain confidence intervals for the ratio of two Poisson counts. 
 Let $Y_{t} \sim Poisson(\lambda_t)$ and $Y_{t+1} \sim Poisson(\lambda_{t+1})$ 
@@
 The main contribution of this paper can be summarized in the following items: 
-(i) we show a method to quantify uncertainty 
-about \rr that is readily applicable to other diseases and; (ii) we propose to 
-use \rr to inform a dynamic epidemic model with time-varying force of infection 
-in order to gain insight into the attack ratio of each epidemic; (iii) We 
+(i) we develop a method to quantify uncertainty 
+about \rr that is readily applicable to other diseases and; (ii) \rr is used to 
+inform a dynamic epidemic model with time-varying force of infection 
+in order to gain insight into the attack ratio of each epidemic; (iii) we 
 propose an estimation procedure for circulating serotype's $S_0$ from 
-aggregate case data, which is robust to epidemic sizes; (iv) We estimate the AR 
-for 18 years of Dengue epidemics.
+aggregate case data, which is robust to epidemic sizes; (iv) AR estimates are 
+provided for 18 years of Dengue epidemics in Rio de Janeiro, Brazil.
 
@@
 wallinga_reconstruction_2003}.
-These methods try to reconstruct the entire 
+These methods attempt to reconstruct the entire 
 series of infectious and susceptibles for measles 
@@
  Underreporting of cases, which is known to exist but of which exact figures 
-cannot be determined, will lead to underestimation $S_0$. However, The attack 
-ratio shall remain unbiased as the underreporting affects both the numerator and 
-the denominator of equation~\ref{eq:AR2}.
+cannot be determined, will lead to underestimation $S_0$.
+However, the attack ratio shall remain unbiased as the underreporting affects 
+both the numerator and the denominator of equation~\ref{eq:AR2}.
 It is interesting to notice that the larger epidemics, in terms of peak size 
@@
 This stresses the importance of knowing the immunological structure of the 
-population. Knowing the $S_0$ for the circulating viruses we can order to more 
+population.
+Knowing the $S_0$ for the circulating viruses we can order to more 
 accurately assess the potential impact of a coming epidemic, since particularly 
-virulent types, can be rendered less of a threat by a low $S_0$. Hon\'orio et 
-al.\citep{honorio2009spatial} conducted a serological survey in three separate 
-localities within the city, right before the 2008 epidemic, the authors report 
-seroprevalences varying from 56-77.4\% which is compatible with our prediction 
-of 87.5\% ($1-S_{0,2007}$) for the entire city, considering that we 
-underestimate $S_0$ due to the underreporting of cases.
+virulent types, can be rendered less of a threat by a low $S_0$.
+Hon\'orio et al.\citep{honorio2009spatial} conducted a serological survey in 
+three separate localities within the city, right before the 2008 epidemic, the 
+authors report seroprevalences varying from 56-77.4\% which is compatible with 
+our prediction of 87.5\% ($1-S_{0,2007}$) for the entire city, considering that 
+we underestimate $S_0$ due to the underreporting of cases.
 
@@
 
+\section*{Estimating $S_0$ from simulated data}
+
+In order to determine whether the inference methodology proposed can 
+recover the true parameter values of the underlying simulation model, we have 
+devised a simple simulation experiment.
+Using the SIR model presented in the main paper we have simulated the incidence 
+curve of the  2013 epidemic (2012 in table~\ref{eq:AR2}), using 
+$S_0 = 0.0621$ and \rr, as estimated from actual incidence data, and $\tau=1$. 
+Figure~\ref{fig:sim_data} shows the posterior $S$ and $I$ alongside with 
+simulated incidence data.
+It is clear that the method can recover the correct value for $S_0$ 
+used to   simulate incidence.
+The script to generate the simulated data and run the inference is available at 
+the paper's Github repository 
+(\url{
+https://github.com/fccoelho/paperLM1/blob/master/python/fit_simulated_data.py}).
+
+\begin{figure}
+ \centering
+ \includegraphics[width=14cm]{./plots/Sim_DengueS2012_11_series.png}
+ % Sim_DengueS2012_11_series.png: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=
+ \caption{$S$ and $I$ series fitted to simulated incidence data (blue dots) 
+with $S_0=0.0621$, and \rr estimated from real incidence data.}
+ \label{fig:sim_data}
+\end{figure}
+
 \subsection*{DiffeRential Evolution Adaptive Metropolis (DREAM)}
@@
 Since the posterior distributions desired in this paper are not available in 
-closed-form, we need to resort to numerical methods to obtain 
-approximations.
+closed-form, we need to resort to numerical methods to obtain approximations.
 We employ an adaptive Markov chain Monte Carlo (MCMC) algorithm, proposed

### 2015-07-25 Correspondence/revisions.tex
LM edits to the response letter

@@
 
-\opening{Dear editor,}
+\opening{Dear Editor,}
 
-In response to the comments of the reviewers of our manuscript, we have made 
+In response to the comments of the reviewers on our manuscript, we have made 
 the following modifications to the manuscript:
@@
  \item[Reviewer 1:] Responses to the main comments are listed below. The minor 
-comments where also accepted, namely the addition of readme do the code 
-repository as well as the citation of two  additional papers.
+comments where also integrally accepted, namely the  improvement of the README 
+in the code repository as well as the citation of two  additional papers.
 
  \begin{enumerate}
-  \item A simulation experiment was done, as requested and included as a new 
-section in the supplementary material. The simulations shows that the inference 
-methodology can recover the exact parameters user to simulate  the data.
+  \item The response to this comment is two part: first, we performed 
+a simple study to assess whether our method of estimation of $R_{t}$ provides 
+more conservative credibility intervals and thus offers protection against 
+false alarm.
+In agreement with the analytical result shown in equation (16), 
+our simulation study -- described at the end of the section on 
+the  tail behaviour of the distribution of $R_t$ section -- shows that our 
+method provided wider CIs in $\approx 90\%$ of the time.
+Then, in order to assess the ability to recover the parameters, simulation 
+experiment was done, as requested and included as a new section in the  
+supplementary material. The simulations shows that the inference 
+methodology can recover the exact parameters used to simulate  the data.
 \item The text of the methods section was improved to answer the four issues 
 raised in this comment.
-\item The introduction and discussion sections were expanded to included the 
+\item The Introduction and discussion sections were expanded to included the 
 additional information required by this comment.
@@
 text to the discussion stating that we expect the effects of underreporting to 
-cancel out in the calculation of the Attack ratio. 
+cancel out in the calculation of the attack ratio.
+Moreover, a section explaining the details of DREAM (delayed rejection, 
+adaptative MCMC, etc.) was added to the Appendix.
  \begin{enumerate}

### 2015-08-03 draft_lm1.tex
finished the incomplete paragraph. closes #45

@@
 68 years~\citep{nogueira1999dengue}.
-During the last decades of the 20th century only DEN-1 and DEN-2 serotypes
+During the last decades of the 20th century only DEN-1 and DEN-2 serotypes were 
+in circulation. The remaining serotypes DEN-3 and DEN-4, arrived respectively 
+in 2000 and 2011 \citep{de2004dengue, nogueira2011dengue}. Due to this patterns 
+of recent re-introduction of the disease, its incidence dynamics is still 
+dominated by the introduction events and environmental determinants of 
+transmission.
 

### 2015-08-03 scirep_version.tex
updated supplSR content

@@
 Dengue is an arthropod-borne febrile disease caused by a flavivirus with four 
-serotypes which causes an estimated  $50$ million infections each 
-year~\cite{Guzman2010}.
+serotypes (DEN-1, 2, 3 and 4) which causes an estimated  $50$ million 
+infections each 
+year~\citep{Guzman2010}.
 In humans, immunity against a particular serotype is considered permanent 
-after the exposure and cross-immunity to the other serotypes is 
-generally considered short lived~\cite{halstead_dengue_2007} although some 
-studies argue for a longer duration of cross-immunity \cite{Reich2013, 
+after the exposure and cross immunity to other serotypes is considered short 
+lived~\citep{halstead_dengue_2007} although some 
+studies argue for a longer duration of cross-immunity \citep{Reich2013, 
 salje2012}.
@@
 model, since the arrival/re-emergence of  new serotypes, an intrinsically 
-stochastic events, can drastically change the pool of susceptibles, throwing off 
+stochastic event, can drastically change the pool of susceptibles, throwing off 
 any sequential estimation based on the  incidence dynamics.
+
+Dengue has been reintroduced in Rio de Janeiro in 1986 after being absent for 
+68 years~\citep{nogueira1999dengue}.
+During the last decades of the 20th century only DEN-1 and DEN-2 serotypes were 
+in circulation. The remaining serotypes DEN-3 and DEN-4, arrived respectively 
+in 2000 and 2011 \citep{de2004dengue, nogueira2011dengue}. Due to this patterns 
+of recent re-introduction of the disease, its incidence dynamics is still 
+dominated by the introduction events and environmental determinants of 
+transmission.
 
@@
 seroprevalence and the exact behaviour of the force of infection through time, 
-we
-propose to inform the time-varying transmissibility using the \rr series 
-derived from the notification data~\cite{nishiura}.
+we propose to inform the time-varying transmissibility using the \rr series 
+derived from the notification data~\citep{nishiura}.
 We extend a Bayesian framework previously used to estimate the number 
-of susceptibles in Influenza epidemics in Europe~\cite{pone2011} to include 
+of susceptibles in Influenza epidemics in Europe~\citep{pone2011} to include 
 time-varying force of infection and derive a probability distribution  for 
@@
 
-We explore the approach of Ederer and Mantel~\cite{mantel}, whose objective is 
-to obtain 
-confidence intervals for the ratio of two Poisson counts. 
+We explore the approach of Ederer and Mantel~\citep{mantel}, whose objective 
+was  to obtain confidence intervals for the ratio of two Poisson counts. 
 Let $Y_{t} \sim Poisson(\lambda_t)$ and $Y_{t+1} \sim Poisson(\lambda_{t+1})$ 
-and define $S = Y_{t} + Y_{t+1}$.
-The authors note that by conditioning on the sum $S$
+and define $W = Y_{t} + Y_{t+1}$.
+The authors note that by conditioning on the sum $W$
 \begin{align}
 \label{eq:binlike}
-Y_{t+1} | S &\sim Binomial(S, \theta_t) \\
+Y_{t+1} | S &\sim Binomial(W, \theta_t) \\
 \theta_t &= \frac{\lambda_{t+1}}{\lambda_{t} + \lambda_{t+1}}
@@
 tells us that the induced posterior distribution of $R_t$ is 
-a beta prime (or inverted beta) with parameters $ a_1 = Y_{t+1} + a_0$ and $b_1 
-=  Y_t + b_0$~\cite{dubey1970}.
+a beta prime (or inverted beta) with parameters $a_1 = Y_{t+1} + a_0$ and $b_1 
+=  Y_t + b_0$~\citep{dubey1970}.
 The density of the induced distribution is then 
@@
 We show how to elicit the prior for $R_t$ from specified prior mean and variance
-or coefficient of variation in the Appendix.
+or coefficient of variation in the~\textbf{Appendix}.
 
@@
  \label{eq:S0post}
- p(S_{0j}|\mathbf{Y_{j}}) \propto l(\mathbf{Y_{j}}|S_{0j}, R_t, m, \tau 
+ p(S_{0j}|\mathbf{Y_{j}}) \propto l(\mathbf{Y_{j}}|S_{0j}, R_t, \tau 
 )\pi(S_{0j}) 
 \end{equation}
-The likelihood $l(\mathbf{Y_{j}}|\cdot)$ is assumed to be a Normal distribution 
-with fixed variance $\sigma^2$.
+The likelihood $l(\mathbf{Y_{j}}|\cdot)$ is a approximated by a Normal 
+distribution with fixed variance $\sigma^2$. This approximation is a numerical 
+convenience since it confers better stability to the MCMC sampling.
 In this estimation procedure we kept $R_t$ fixed at fixed at the posterior mean 
@@
 Carlo techniques implemented in the Bayesian inference with Python 
-(BIP)~\cite{pone2011} available at  
+(BIP)~\citep{pone2011} available at  
 \url{http://code.google.com/p/bayesian-inference/}.
-BIP uses a Differential Evolution Adaptive Metropolis (DREAM)~\cite{vrugt2008} 
+BIP uses a Differential Evolution Adaptive Metropolis (DREAM)~\citep{vrugt2008} 
 scheme that efficiently samples from high-dimensional joint distributions using 
-multiple adaptive chains running in parallel with delayed rejection.
+multiple adaptive chains running in parallel with delayed rejection (see 
+\textbf{Appendix} for details).
 Also, as the numerical integration routine implemented within BIP needs 
@@
 The main contribution of this paper can be summarized in the following items: 
-(i) we show a method to quantify uncertainty 
-about \rr that is readily applicable to other diseases and; (ii) we propose to 
-use \rr to inform a dynamic epidemic model with time-varying force of infection 
-in order to gain insight into the attack ratio of each epidemic; (iii) We 
+(i) we develop a method to quantify uncertainty 
+about \rr that is readily applicable to other diseases and; (ii) \rr is used to 
+inform a dynamic epidemic model with time-varying force of infection 
+in order to gain insight into the attack ratio of each epidemic; (iii) we 
 propose an estimation procedure for circulating serotype's $S_0$ from 
-aggregate case data, which is robust to epidemic sizes; (iv) We estimate the AR 
-for 18 years of Dengue epidemics.
+aggregate case data, which is robust to epidemic sizes; (iv) AR estimates are 
+provided for 18 years of Dengue epidemics in Rio de Janeiro, Brazil.
 
 Figure~\ref{fig:rtseries} shows the $R_t$ series, according to 
-(\ref{eq:Rtestimate})~\cite{nishiura} along with the confidence bands derived 
+Equation~\ref{eq:Rtestimate}~\citep{nishiura} along with the confidence bands 
+derived 
 in this paper. 
@@
 these periods. 
-The method to quantify uncertainty proposed here provides more conservative 
-credibility intervals, and therefore offers protection against false alarms.
+credibility intervals, and therefore offers protection against false alarms 
+(see 
+the section on tail behaviour in the \textbf{Appendix} for a detailed 
+explanation).
 \begin{center}
@@
 Methods for estimating the  number of susceptibles have been proposed 
-before, for other diseases~\cite{bjornstad_dynamics_2002, 
+before, for other diseases~\citep{bjornstad_dynamics_2002, 
 wallinga_reconstruction_2003}.
-These methods try to reconstruct the entire 
+These methods attempt to reconstruct the entire 
 series of infectious and susceptibles for measles 
@@
 mild and asymptomatic cases, which nevertheless acquire immunity.
-It is estimated that for every case reported, up to 10-20 are not seen by health 
-authorities~\cite{luz_disability_2009}.
+It is estimated that for every case reported, up to 10-20 are not seen by 
+health 
+authorities~\citep{luz_disability_2009}.
 Another source of uncertainty is under-reporting of diagnosed cases, which is a 
@@
 
-\begin{center}
-\begin{figure}
-\includegraphics[width=\textwidth]{./plots/figure-2-coelho.jpg}
-
-  \caption{{\bf Maps showing the incidence of dengue in the city of Rio de 
-Janeiro from 2010 to 2013(left-to-right, rowise).} Circles indicate individual 
-notified cases. A 
-heatmap is overlayed on the maps showing absolute density of cases.
-It can be seen that several areas of the city were affected and no region seems 
-to be free of transmission risk.
-This suggests that although transmission risk varies spatially, there is 
-significant exposure over the entire city. Maps were made with QGis GIS 
-software(\url{qgis.org})}
-  \label{fig:mapas}
-\end{figure}
-\end{center}
-
-Table~\ref{tab:AR} contains the attack ratios and medians of the $S_0$ 
-estimated for each epidemic/outbreak.
-It is interesting to notice that the larger epidemics, in terms of peak size are 
-not the one with the greater attack ratios.
-This stresses the importance of knowing the immunological structure of the 
-population. Knowing the $S_0$ for the circulating viruses we can order to more 
-accurately assess the potential impact of a coming epidemic, since particularly 
-virulent types, can be rendered less of a threat by a low $S_0$.
-
-We hope that the results presented in paper will motivate public 
-health authorities to invest in annual serological surveys, to determine the 
-susceptibility profile to each dengue virus as well as to estimate the 
-under-reporting factor of the notification system.
-
 \begin{figure}[!ht]
@@
 
+Table~\ref{tab:AR} contains the attack ratios and medians of the $S_0$ 
+estimated for each epidemic/outbreak.
+ Underreporting of cases, which is known to exist but of which exact figures 
+cannot be determined, will lead to underestimation $S_0$.
+However, the attack ratio shall remain unbiased as the underreporting affects 
+both the numerator and the denominator of equation~\ref{eq:AR2}.
+It is interesting to notice that the larger epidemics, in terms of peak size 
+are not the one with the greater attack ratios.
+This stresses the importance of knowing the immunological structure of the 
+population.
+Knowing the $S_0$ for the circulating viruses we can order to more 
+accurately assess the potential impact of a coming epidemic, since particularly 
+virulent types, can be rendered less of a threat by a low $S_0$.
+Hon\'orio et al.\citep{honorio2009spatial} conducted a serological survey in 
+three separate localities within the city, right before the 2008 epidemic, the 
+authors report seroprevalences varying from 56-77.4\% which is compatible with 
+our prediction of 87.5\% ($1-S_{0,2007}$) for the entire city, considering that 
+we underestimate $S_0$ due to the underreporting of cases.
+
+We hope that the results presented in paper will motivate public 
+health authorities to invest in annual serological surveys, to determine the 
+susceptibility profile to each dengue virus as well as to estimate the 
+under-reporting factor of the notification system.
+
+\begin{center}
+\begin{figure}
+ \subfigure[2010]{\includegraphics[width=8cm, 
+height=4.82cm]{./plots/heatmap2010.jpg}
+% heatmap2010.jpg: 1075x589 pixel, 96dpi, 28.44x15.58 cm, bb=0 0 806 442
+}
+  \subfigure[2011]{\includegraphics[width=8cm]{./plots/heatmap2011.jpg}}
+  \subfigure[2012]{\includegraphics[width=8cm]{./plots/heatmap2012.jpg}
+% heatmap2010.jpg: 1075x589 pixel, 96dpi, 28.44x15.58 cm, bb=0 0 806 442
+}
+  \subfigure[2013]{\includegraphics[width=8cm]{./plots/heatmap2013.jpg}}
+  \caption{{\bf Maps showing the incidence of dengue in the city of Rio de 
+Janeiro from 2010 to 2013.} Circles indicate individual notified cases. A 
+heatmap is overlayed on the maps showing absolute density of cases.
+It can be seen that several areas of the city were affected and no region seems 
+to be free of transmission risk.
+This suggests that although transmission risk varies spatially, there is 
+significant exposure over the entire city.}
+  \label{fig:mapas}
+\end{figure}
+\end{center}
+
 \begin{table}[!ht]
@@
 Values are presented as percentage of total population. 
-$^\dag$: Year corresponds to the start of the epidemic, however the peak of cases
+$^\dag$: Year corresponds to the start of the epidemic, but the peak of cases
 may occur in the following year.
-$^\ddag$: Susceptible fraction.
-These results show considerable variation in AR between epidemics, consistent with
+$^\ddag$: percentage of population.
+These results show considerable variation in AR between epidemics, consistent 
+with
 the accquiring and loss of serotype-specific immunity.}
 \begin{center}
-\begin{tabular}{c|c|c}
+\begin{tabular}{l|c|c|c|c}
 \hline
-Year$^\dag$ & median Attack Ratio & $S_0^\ddag$ \\
+$^\dag$Year & $^\ddag$Cases & median Attack Ratio & $^\ddag$$S_0$ & Circulating 
+ Serotypes \\
 \hline
-1996 & 0.39 (0.17-0.54) & 0.00171(0.0012-0.0038)\\
-1997 & 0.87 (0.74-0.87) & 0.00273(0.0027-0.0032)\\
-1998 & 0.5 (0.49-0.5) & 0.00142(0.0014-0.0014)\\
-1999 & 0.11 (0.037-0.2) & 0.00345(0.0018-0.01)\\
-2000 & 0.25 (0.24-0.27) & 0.0155(0.015-0.016)\\
-2001 & 0.48 (0.47-0.49) & 0.0495(0.048-0.051)\\
-2005 & 0.15 (0.1-0.21) & 0.0147(0.01-0.021)\\
-2006 & 0.11 (0.08-0.14) & 0.0281(0.022-0.037)\\
-2007 & 0.15 (0.15-0.15) & 0.135(0.13-0.14)\\
-2008 & 0.14 (0.031-0.31) & 0.00672(0.003-0.024)\\
-2010 & 0.18 (0.17-0.19) & 0.0454(0.043-0.048)\\
-2011 & 0.086 (0.082-0.094) & 0.215(0.2-0.23)\\
-2012 & 0.14 (0.13-0.15) & 0.0621(0.058-0.068)\\
+1996 & 0.066 & 0.39 (0.17-0.54) & 0.171(0.12-0.38)& DEN-1 and 
+2\citep{nogueira1999dengue}\\
+1997 & 0.238 & 0.87 (0.74-0.87) & 0.273(0.27-0.32)& DEN-1 and 
+2\citep{nogueira1999dengue}\\
+1998 & 0.0708 & 0.5 (0.49-0.5) & 0.142(0.14-0.14)& DEN-1 and 
+2\citep{nogueira1999dengue}\\
+1999 & 0.0371 & 0.11 (0.037-0.2) & 0.345(0.18-1.0)&DEN-1, 2 and 
+3\citep{de2004dengue}\\
+2000 & 0.394 & 0.25 (0.24-0.27) & 1.55(1.5-1.6)&DEN-1, 2 and 
+3\citep{de2004dengue}\\
+2001 & 2.38 & 0.48 (0.47-0.49) & 4.95(4.8-5.1)&DEN-1, 2 and 
+3\citep{de2004dengue}\\
+2005 & 0.217 & 0.15 (0.1-0.21) & 1.47(1.0-2.1)&\\
+2006 & 0.315 & 0.11 (0.08-0.14) & 
+2.81(2.2-3.7)&DEN-2\citep{nogueira2011dengue}\\
+2007 & 2.03 & 0.15 (0.15-0.15) & 
+13.5(13.0-14.01)&DEN-2\citep{nogueira2011dengue}\\
+2008 & 0.0923 & 0.14 (0.031-0.31) & 
+0.672(0.3-2.4)&DEN-2\citep{nogueira2011dengue}\\
+2010 & 0.831 & 0.18 (0.17-0.19) & 
+4.54(4.3-4.8)&DEN-2\citep{nogueira2011dengue}\\
+2011 & 1.85 & 0.086 (0.082-0.094) & 
+21.5(20.0-23.0)&DEN-1, 2 and 4\citep{nogueira2011dengue}\\
+2012 & 0.864 & 0.14 (0.13-0.15) & 6.21(5.8-6.8)&\\
+
 \hline

### 2015-08-03 supplSR.tex
updated supplSR content

--- 2015-02-19 supplSR.tex
@@
 \label{sec:tails}
-There are a number of approaches to deriving the distribution of \rr.
-Alternatively to the approach described in the main text~\cite{mantel}, one 
+There are a number of approaches to deriving the distribution of~\rr.
+Alternatively to the approach described in the main text~\citep{mantel}, one 
 could use the conditional distribution of \rr on 
 $Y_{t+1}$ and $Y_t$ as defined in equation A7 of Nishiura et 
-al.~\cite{nishiura}:
+al.~\citep{nishiura}:
 \begin{equation}
@@
 tail behaviour of the derived distributions for \rr. 
-Consider the case of using a flat $Uniform(0, 1)$ prior for $\theta_t$.
+Consider the case of using a flat $\text{Uniform}(0, 1)$ prior for $\theta_t$.
 With $a_0 = b_0 = 1$, $a_1 = a_2$ and $b_1 = b_2 + 1$.
 The beta prime (inverse beta distribution) will have heavier tails compared to 
-the conditional distribution proposed by~\cite{nishiura}, thus providing more 
+the conditional distribution proposed by~\citep{nishiura}, thus providing more 
 conservative confidence/credibility intervals.
@@
 \end{equation}
-Finally, note that we deliberately construct $c_{\alpha}(R_{t})$ as a 
+Note also that we deliberately construct $c_{\alpha}(R_{t})$ as a 
 equal-tailed $100\alpha\%$ credible set, rather than a less conservative 
 highest posterior density (HPD) interval.
+Finally, we performed a simple simulation study to assess the result in equation
+~\ref{seq:densityratio}.
+The simulations were carried out as follows:
+\begin{enumerate}
+ \item Create a two-dimensional grid of values for $\lambda_1$ and $\lambda_2$, 
+with rate values from $2$ to $1000$;
+\item For each point $(\lambda_1^{(j)}, \lambda_2^{(j)})$ in the grid,
+\begin{itemize}
+ \item Generate $1000$ realisations of the random variables $Y_{1}^{(ij)} \sim 
+\text{Poisson}(\lambda_1^{(j)})$ and $Y_{2}^{(ij)} \sim 
+\text{Poisson}(\lambda_2^{(j)})$, with $i= 1, 2, \ldots, 1000$  and compute 
+$R_{t}^{(ij)} = Y_{2}^{(ij)}/
+Y_{1}^{(ij)}$;
+\item Compute the $\alpha\%$ credibility/confidence intervals using the method 
+proposed here $(\beta)$ and using the distribution proposed by Nishiura 
+(2010)~\cite{nishiura} ($\Gamma$), denoted by $c_{\beta}(R_{t}^{(ij)}; 
+\alpha)$ and 
+$c_{\Gamma}(R_{t}^{(ij)}; \alpha)$ respectively;
+\item Determine whether
+\begin{itemize}
+ \item [(i)] $ c_{\Gamma}(R_{t}^{(ij)}; \alpha) \subseteq 
+c_{\beta}(R_{t}^{(ij)}; \alpha)$;
+\item [(ii)] $c_{\beta}(R_{t}^{(ij)}; \alpha) \subseteq 
+c_{\Gamma}(R_{t}^{(ij)}; \alpha)$ or;
+\item [(iii)] there is overlap
+\end{itemize}
+\item Calculate $Pr(R_{t}^{(ij)} > 1)$ using both distributions
+\end{itemize}
+\end{enumerate}
+The results show that the credibility intervals obtained using the distribution 
+proposed here were larger than (i.e. contained) those computed using the 
+distribution in~\ref{seq:densityNishiura} in $90.69\%$ of the simulations 
+(integrating over the grid), while the converse was observed in $1.14\%$ of the 
+cases.
+Moreover, our method yielded lower $Pr(R_{t} > 1)$  in $79.48\%$ of the 
+simulations, indicating it is indeed more robust to false alarm.
+An R script to perform the simulation study described above can be found 
+at~\url{
+https://github.com/fccoelho/paperLM1/blob/master/R/credibility_comparison.R}.
 
 As a side note, the Bayesian approach presented in this 
-paper will give similar results to orthodox confidence intervals~\cite{wilson} 
-and~\cite{clopper} for $Y_{t+1}$ and $Y_t >> 1$.
+paper will give similar results to orthodox confidence intervals~\citep{wilson} 
+and~\citep{clopper} for $Y_{t+1}$ and $Y_t >> 1$.
 Under the flat  uniform prior for $\theta_t$, the Bayesian posterior 
-credibility 
-interval is nearly indistinguishable from the confidence interval proposed by 
-Clopper \& Pearson (1931)~\cite{clopper} for $Y_{t+1}, Y_t > 20$.
+credibility interval is nearly indistinguishable from the confidence interval 
+proposed by Clopper \& Pearson (1931)~\citep{clopper} for $Y_{t+1}, Y_t > 20$.
 Note also that the uniform prior ($Beta(1, 1)$) for $\theta_t$ constitutes a 
-poor 
-prior 
-choice mainly because the induced distribution for \rr is only well-defined for 
-$b_0 > 2$.
+poor prior choice mainly because the induced distribution for \rr is only 
+well-defined for $b_0 > 2$.
 
@@
 to the detection threshold.
-We provide an R script to perform the above elicitation at 
+We provide an R script to perform the elicitation described above at 
 \url{https://github.com/fccoelho/paperLM1/blob/master/R/elicit_Rt_prior.R}.
+
+\section*{Estimating $S_0$ from simulated data}
+
+In order to determine whether the inference methodology proposed can 
+recover the true parameter values of the underlying simulation model, we have 
+devised a simple simulation experiment.
+Using the SIR model presented in the main paper we have simulated the incidence 
+curve of the  2012-13 epidemic, using 
+$S_0 = 0.0621$ and \rr, as estimated from actual incidence data, and $\tau=1$. 
+Figure~\ref{fig:sim_data} shows the posterior $S$ and $I$ alongside with 
+simulated incidence data.
+It is clear that the method can recover the correct value for $S_0$ 
+used to   simulate incidence.
+The script to generate the simulated data and run the inference is available at 
+the paper's Github repository 
+(\url{
+https://github.com/fccoelho/paperLM1/blob/master/python/fit_simulated_data.py}).
+
+\begin{figure}
+ \centering
+ \includegraphics[width=14cm]{./plots/Sim_DengueS2012_11_series.png}
+ % Sim_DengueS2012_11_series.png: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=
+ \caption{$S$ and $I$ series fitted to simulated incidence data (blue dots) 
+with $S_0=0.0621$, and \rr estimated from real incidence data.}
+ \label{fig:sim_data}
+\end{figure}
+
+\subsection*{DiffeRential Evolution Adaptive Metropolis (DREAM)}
+
+Since the posterior distributions desired in this paper are not available in 
+closed-form, we need to resort to numerical methods to obtain approximations.
+We employ an adaptive Markov chain Monte Carlo (MCMC) algorithm, proposed
+by~\cite{vrugt2008}, called  DiffeRential Evolution Adaptive Metropolis (DREAM).
+
+DREAM draws on the basic structure of Differential Evolution Markov Chain 
+(DE-MC) which runs $N$ independent chains in parallel and accepts moves 
+proportional to the difference of two randomly sampled members (chains), thus 
+differentially evolving conditional on the proposal variances.
+An additional aspect of DREAM is the so-called delayed rejection.
+This procedure adapts the original Metropolis-Hastings algorithm by attempting 
+new moves whenever a proposal is rejected (instead of setting the current state 
+to the previously accepted one), thus delaying the rejection of proposals. 
+This potentially decreases autocorrelation and improves mixing 
+(see~\cite{mira2001} for details).
+
+A desirable property of DREAM is that one can scale the algorithm with model 
+complexity, meaning one can initialise as many parallel chains as there are 
+parameters in the model, thus exploiting the multi-core architecture of modern 
+computers to speed up convergence of the MCMC.
+
+A Python implementation of DREAM for dynamic models is available in the package 
+Bayesian inference with Python (BIP)~\cite{pone2011} available at 
+\url{http://code.google.com/p/bayesian-inference/}.
+The code to produce the results presented in this paper is available 
+at~\url{https://github.com/fccoelho/paperLM1/tree/master/python}.
 
@@
 
+\bibitem{vrugt2008}
+\bibinfo{author}{Vrugt, J.~A.} \emph{et~al.}
+\newblock \bibinfo{title}{Accelerating markov chain monte carlo simulation by
+  differential evolution with self-adaptive randomized subspace sampling}.
+\newblock \emph{\bibinfo{journal}{International Journal of Nonlinear Sciences
+  and Numerical Simulation}} \textbf{\bibinfo{volume}{10}},
+  \bibinfo{pages}{271--288} (\bibinfo{year}{2008}).
+
+\bibitem{mira2001}
+\bibinfo{author}{Mira, A.}
+\newblock \bibinfo{title}{On metropolis-hastings algorithms with delayed
+  rejection}.
+\newblock \emph{\bibinfo{journal}{Metron}} \textbf{\bibinfo{volume}{59}},
+  \bibinfo{pages}{231--241} (\bibinfo{year}{2001}).
+
+\bibitem{pone2011}
+\bibinfo{author}{Coelho, F.~C.}, \bibinfo{author}{Code\c{c}o, C.~T.} \&
+  \bibinfo{author}{Gomes, M.~G.}
+\newblock \bibinfo{title}{{A} {B}ayesian framework for parameter estimation in
+  dynamical models}.
+\newblock \emph{\bibinfo{journal}{PLoS ONE}} \textbf{\bibinfo{volume}{6}},
+  \bibinfo{pages}{e19616} (\bibinfo{year}{2011}).
+
 \end{thebibliography}

### 2015-08-03 draft_lm1.tex
deleted appendix from draft_lm1 and updated scirep_version. closes #42

@@
 
-\section*{Appendix}
-\subsection*{A remark on prior distributions and tail behaviour of the 
-distribution of $R_t$}
-\label{sec:tails}
-There are a number of approaches to deriving the distribution of~\rr.
-Alternatively to the approach described in the main text~\citep{mantel}, one 
-could use the conditional distribution of \rr on 
-$Y_{t+1}$ and $Y_t$ as defined in equation A7 of Nishiura et 
-al.~\citep{nishiura}:
-\begin{equation}
-\label{seq:unorm}
-f_{R}(R_{t}) = (Y_tR_{t})^{Y_{t+1}} e^{-Y_tR_{t}}
-\end{equation}
-Noticing the kernel of (\ref{seq:unorm}) is that of a gamma distribution with 
-$a_2 = Y_{t+1}+1$ and $b_2 = Y_t$, we obtain a proper density from which to 
-construct $c_{\alpha}(R_t)$, simply by computing the appropriate quantiles of 
-said distribution.
- This density is
-\begin{equation}
-\label{seq:densityNishiura}
-f_N(R_t| a_2, b_2) =  \frac{b_2^{a_2}}{\Gamma(a_2)} R_t^{a_2-1} e^{-b_2 R_t}
-\end{equation}
-
-In order to decide which approach to take, it may be of use analysing the 
-tail behaviour of the derived distributions for \rr. 
-Consider the case of using a flat $\text{Uniform}(0, 1)$ prior for $\theta_t$.
-With $a_0 = b_0 = 1$, $a_1 = a_2$ and $b_1 = b_2 + 1$.
-The beta prime (inverse beta distribution) will have heavier tails compared to 
-the conditional distribution proposed by~\citep{nishiura}, thus providing more 
-conservative confidence/credibility intervals.
-To see that one needs simply take the ratio of the Beta prime and Gamma 
-(unnormalized) densities and evaluate the limit as $R_t$ goes to infinity:
-\begin{equation}
- \label{seq:densityratio}
- \lim_{R_t\to\infty}\frac{f_P(R_t| a_1, b_1)}{f_N(R_t| a_2, b_2)} =  
-\lim_{R_t\to\infty}\frac{e^{Y_{t}R_t}}{(1 +R_t)^{Y_{t} + Y_{t +1}+2}} = \infty
-\end{equation}
-Note also that we deliberately construct $c_{\alpha}(R_{t})$ as a 
-equal-tailed $100\alpha\%$ credible set, rather than a less conservative 
-highest posterior density (HPD) interval.
-Finally, we performed a simple simulation study to assess the result in equation
-~\ref{seq:densityratio}.
-The simulations were carried out as follows:
-\begin{enumerate}
- \item Create a two-dimensional grid of values for $\lambda_1$ and $\lambda_2$, 
-with rate values from $2$ to $1000$;
-\item For each point $(\lambda_1^{(j)}, \lambda_2^{(j)})$ in the grid,
-\begin{itemize}
- \item Generate $1000$ realisations of the random variables $Y_{1}^{(ij)} \sim 
-\text{Poisson}(\lambda_1^{(j)})$ and $Y_{2}^{(ij)} \sim 
-\text{Poisson}(\lambda_2^{(j)})$, with $i= 1, 2, \ldots, 1000$  and compute 
-$R_{t}^{(ij)} = Y_{2}^{(ij)}/
-Y_{1}^{(ij)}$;
-\item Compute the $\alpha\%$ credibility/confidence intervals using the method 
-proposed here $(\beta)$ and using the distribution proposed by Nishiura 
-(2010)~\cite{nishiura} ($\Gamma$), denoted by $c_{\beta}(R_{t}^{(ij)}; 
-\alpha)$ and 
-$c_{\Gamma}(R_{t}^{(ij)}; \alpha)$ respectively;
-\item Determine whether
-\begin{itemize}
- \item [(i)] $ c_{\Gamma}(R_{t}^{(ij)}; \alpha) \subseteq 
-c_{\beta}(R_{t}^{(ij)}; \alpha)$;
-\item [(ii)] $c_{\beta}(R_{t}^{(ij)}; \alpha) \subseteq 
-c_{\Gamma}(R_{t}^{(ij)}; \alpha)$ or;
-\item [(iii)] there is overlap
-\end{itemize}
-\item Calculate $Pr(R_{t}^{(ij)} > 1)$ using both distributions
-\end{itemize}
-\end{enumerate}
-The results show that the credibility intervals obtained using the distribution 
-proposed here were larger than (i.e. contained) those computed using the 
-distribution in~\ref{seq:densityNishiura} in $90.69\%$ of the simulations 
-(integrating over the grid), while the converse was observed in $1.14\%$ of the 
-cases.
-Moreover, our method yielded lower $Pr(R_{t} > 1)$  in $79.48\%$ of the 
-simulations, indicating it is indeed more robust to false alarm.
-An R script to perform the simulation study described above can be found 
-at~\url{
-https://github.com/fccoelho/paperLM1/blob/master/R/credibility_comparison.R}.
-
-As a side note, the Bayesian approach presented in this 
-paper will give similar results to orthodox confidence intervals~\citep{wilson} 
-and~\citep{clopper} for $Y_{t+1}$ and $Y_t >> 1$.
-Under the flat  uniform prior for $\theta_t$, the Bayesian posterior 
-credibility interval is nearly indistinguishable from the confidence interval 
-proposed by Clopper \& Pearson (1931)~\citep{clopper} for $Y_{t+1}, Y_t > 20$.
-Note also that the uniform prior ($Beta(1, 1)$) for $\theta_t$ constitutes a 
-poor prior choice mainly because the induced distribution for \rr is only 
-well-defined for $b_0 > 2$.
-
-An advantage of the Bayesian approach is that one can devise prior 
-distributions for $\theta_t$ taking advantage of the intuitive parametrization 
-and flexibility of the beta family of distributions.
-Prior elicitation can also be done for~\rr and the hyper-parameters directly 
-plugged into the prior for $\theta_t$. 
-One can, for example, choose prior mean and variance for~\rr and find $a_0$ 
-and $b_0$ that satisfy those conditions.
-Let $m_0$ and $v_0$ be the prior expectation and variance for $R_t$. 
-After some tedious algebra one finds
-\begin{align}
-\label{seq:elicitation}
-a_0 &= \frac{m_0v_0 + m_0^3 + m_0^2}{v_0} \\
-b_0 &= \frac{2v_0 + m_0^2 + m_0}{v_0}
-\end{align}
-If one wants only to specify $m_0$ and the coefficient of variation $c = 
-\sqrt{v_0}/ m_0$ for $R_t$ \textit{a priori}, some less boring 
-algebra gives:
-\begin{align}
-\label{seq:elicitationcv}
-a_0 &= \frac{m_0^3c^2 + m_0^3 + m_0^2}{m_0^2c^2} \\
-b_0 &= \frac{2m_0^2c^2 + m^2 + m}{m_0^2c^2}
-\end{align}
-
-This approach thus makes it possible to incorporate epidemiological knowledge 
-about disease Biology (e.g. the magnitude of $R_0$) into the computation of \rr.
-This may prove particularly important when disease counts are low and/or close 
-to the detection threshold.
-We provide an R script to perform the elicitation described above at 
-\url{https://github.com/fccoelho/paperLM1/blob/master/R/elicit_Rt_prior.R}.
-
-\section*{Estimating $S_0$ from simulated data}
-
-In order to determine whether the inference methodology proposed can 
-recover the true parameter values of the underlying simulation model, we have 
-devised a simple simulation experiment.
-Using the SIR model presented in the main paper we have simulated the incidence 
-curve of the  2013 epidemic (2012 in table~\ref{eq:AR2}), using 
-$S_0 = 0.0621$ and \rr, as estimated from actual incidence data, and $\tau=1$. 
-Figure~\ref{fig:sim_data} shows the posterior $S$ and $I$ alongside with 
-simulated incidence data.
-It is clear that the method can recover the correct value for $S_0$ 
-used to   simulate incidence.
-The script to generate the simulated data and run the inference is available at 
-the paper's Github repository 
-(\url{
-https://github.com/fccoelho/paperLM1/blob/master/python/fit_simulated_data.py}).
-
-\begin{figure}
- \centering
- \includegraphics[width=14cm]{./plots/Sim_DengueS2012_11_series.png}
- % Sim_DengueS2012_11_series.png: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=
- \caption{$S$ and $I$ series fitted to simulated incidence data (blue dots) 
-with $S_0=0.0621$, and \rr estimated from real incidence data.}
- \label{fig:sim_data}
-\end{figure}
-
-\subsection*{DiffeRential Evolution Adaptive Metropolis (DREAM)}
-
-Since the posterior distributions desired in this paper are not available in 
-closed-form, we need to resort to numerical methods to obtain approximations.
-We employ an adaptive Markov chain Monte Carlo (MCMC) algorithm, proposed
-by~\cite{vrugt2008}, called  DiffeRential Evolution Adaptive Metropolis (DREAM).
-
-DREAM draws on the basic structure of Differential Evolution Markov Chain 
-(DE-MC) which runs $N$ independent chains in parallel and accepts moves 
-proportional to the difference of two randomly sampled members (chains), thus 
-differentially evolving conditional on the proposal variances.
-An additional aspect of DREAM is the so-called delayed rejection.
-This procedure adapts the original Metropolis-Hastings algorithm by attempting 
-new moves whenever a proposal is rejected (instead of setting the current state 
-to the previously accepted one), thus delaying the rejection of proposals. 
-This potentially decreases autocorrelation and improves mixing 
-(see~\cite{mira2001} for details).
-
-A desirable property of DREAM is that one can scale the algorithm with model 
-complexity, meaning one can initialise as many parallel chains as there are 
-parameters in the model, thus exploiting the multi-core architecture of modern 
-computers to speed up convergence of the MCMC.
-
-A Python implementation of DREAM for dynamic models is available in the package 
-Bayesian inference with Python (BIP)~\citep{pone2011} available at 
-\url{http://code.google.com/p/bayesian-inference/}.
-The code to produce the results presented in this paper is available 
-at~\url{https://github.com/fccoelho/paperLM1/tree/master/python}.
-
 \end{document}

### 2015-08-03 scirep_version.tex
deleted appendix from draft_lm1 and updated scirep_version. closes #42

@@
 \newblock \bibinfo{title}{Temporal distribution of aedes aegypti in different
-  districts of rio de janeiro, brazil, measured by two types of traps}.
+  districts of rio de janeiro, {B}razil, measured by two types of traps}.
 \newblock \emph{\bibinfo{journal}{Journal of Medical Entomology}}
@@
   (\bibinfo{year}{2014}).
+
+\bibitem{nogueira1999dengue}
+\bibinfo{author}{Nogueira, R. M.~R.} \emph{et~al.}
+\newblock \bibinfo{title}{Dengue in the state of rio de janeiro, brazil,
+  1986-1998}.
+\newblock \emph{\bibinfo{journal}{Mem{\'o}rias do Instituto Oswaldo Cruz}}
+  \textbf{\bibinfo{volume}{94}}, \bibinfo{pages}{297--304}
+  (\bibinfo{year}{1999}).
+
+\bibitem{de2004dengue}
+\bibinfo{author}{De~Simone, T.} \emph{et~al.}
+\newblock \bibinfo{title}{Dengue virus surveillance: the co-circulation of
+  {DENV}-1, {DENV}-2 and {DENV}-3 in the state of {R}io de {J}aneiro,
+  {B}razil}.
+\newblock \emph{\bibinfo{journal}{Transactions of the Royal Society of Tropical
+  Medicine and Hygiene}} \textbf{\bibinfo{volume}{98}},
+  \bibinfo{pages}{553--562} (\bibinfo{year}{2004}).
+
+\bibitem{nogueira2011dengue}
+\bibinfo{author}{Nogueira, R.~M.} \& \bibinfo{author}{Eppinghaus, A.~L.}
+\newblock \bibinfo{title}{Dengue virus type 4 arrives in the state of {R}io de
+  {J}aneiro: a challenge for epidemiological surveillance and control}.
+\newblock \emph{\bibinfo{journal}{Mem{\'o}rias do Instituto Oswaldo Cruz}}
+  \textbf{\bibinfo{volume}{106}}, \bibinfo{pages}{255--256}
+  (\bibinfo{year}{2011}).
 
@@
 \newblock \bibinfo{title}{Virological surveillance for early warning of dengue
-  epidemics in the state of rio de janeiro, brazil}.
+  epidemics in the state of {R}io de {J}aneiro, {B}razil}.
 \newblock \emph{\bibinfo{journal}{Transactions of the Royal Society of Tropical
@@
 \newblock \bibinfo{title}{Disability adjusted life years lost to dengue in
-  brazil}.
+  {B}razil}.
 \newblock \emph{\bibinfo{journal}{Tropical Medicine \& International Health}}
@@
 \newblock \bibinfo{title}{Data quality of dengue epidemiological surveillance
-  in belo horizonte, southeastern brazil}.
+  in belo horizonte, southeastern {B}razil}.
 \newblock \emph{\bibinfo{journal}{Revista de Sa\'ude P\'ublica}}
@@
 
-\bibitem{mira2001}
-\bibinfo{author}{Mira, A.}
-\newblock \bibinfo{title}{On metropolis-hastings algorithms with delayed
-  rejection}.
-\newblock \emph{\bibinfo{journal}{Metron}} \textbf{\bibinfo{volume}{59}},
-  \bibinfo{pages}{231--241} (\bibinfo{year}{2001}).
+\bibitem{honorio2009spatial}
+\bibinfo{author}{Hon{\'o}rio, N.~A.} \emph{et~al.}
+\newblock \bibinfo{title}{Spatial evaluation and modeling of dengue
+  seroprevalence and vector density in {R}io de {J}aneiro, {B}razil}.
+\newblock \emph{\bibinfo{journal}{PLoS Negl Trop Dis}}
+  \textbf{\bibinfo{volume}{3}}, \bibinfo{pages}{e545} (\bibinfo{year}{2009}).
 

### 2015-08-31 maps_figure.tex
first stab at making the new maps figure

% \documentclass{standalone}
% \usepackage{capt-of}
% \usepackage{subfigure}
% \usepackage{graphicx}
% \begin{document}
% %  \begin{center}
% %  \includegraphics[width=8cm,height=4.82cm]{./plots/heatmap2010.jpg}
%   \subfigure[2010]{\includegraphics[width=8cm,height=4.82cm]{./plots/heatmap2010.jpg}}
%   \subfigure[2011]{\includegraphics[width=8cm]{./plots/heatmap2011.jpg}}
%   \subfigure[2012]{\includegraphics[width=8cm]{./plots/heatmap2012.jpg}}
%   \subfigure[2013]{\includegraphics[width=8cm]{./plots/heatmap2013.jpg}}
% %  \end{center}
% %    \captionof{figure}{}
% \end{document}

\documentclass[preview]{standalone}
\usepackage{mwe}
\usepackage[]{subfig}
\usepackage[margin=0.1cm]{geometry}
\begin{document}
% \def\FigSize{.23}
\begin{figure}[H]
\centering
\subfloat[\textbf{2010}]
  {
 \includegraphics[width=0.485\textwidth,height=4.0cm]{./plots/heatmap2010.jpg}
  }
\subfloat[\textbf{2011}]
  {
 \includegraphics[width=0.485\textwidth,height=4.0cm]{./plots/heatmap2011.jpg}
  }\\
\subfloat[\textbf{2012}]
  {
 \includegraphics[width=0.485\textwidth,height=4.0cm]{./plots/heatmap2012.jpg}
  }
  \subfloat[\textbf{2013}]
  {
 \includegraphics[width=0.485\textwidth,height=4.0cm]{./plots/heatmap2013.jpg}
  }
\end{figure}
\end{document}

### 2015-09-18 scirep_version.tex
first stab at making the new maps figure
latest modifications corresponding to the ressubmission to Scirep.

@@
 
+\begin{figure}
 \begin{center}
-\begin{figure}
- \subfigure[2010]{\includegraphics[width=8cm, 
-height=4.82cm]{./plots/heatmap2010.jpg}
-% heatmap2010.jpg: 1075x589 pixel, 96dpi, 28.44x15.58 cm, bb=0 0 806 442
-}
-  \subfigure[2011]{\includegraphics[width=8cm]{./plots/heatmap2011.jpg}}
-  \subfigure[2012]{\includegraphics[width=8cm]{./plots/heatmap2012.jpg}
-% heatmap2010.jpg: 1075x589 pixel, 96dpi, 28.44x15.58 cm, bb=0 0 806 442
-}
-  \subfigure[2013]{\includegraphics[width=8cm]{./plots/heatmap2013.jpg}}
+\includegraphics[width=\textwidth]{plots/4maps.png}
+\end{center}
   \caption{{\bf Maps showing the incidence of dengue in the city of Rio de 
-Janeiro from 2010 to 2013.} Circles indicate individual notified cases. A 
+Janeiro from 2010 to 2013 -- Clockwise from top left.} Circles indicate 
+individual notified cases. A 
 heatmap is overlayed on the maps showing absolute density of cases.
@@
 This suggests that although transmission risk varies spatially, there is 
-significant exposure over the entire city.}
+significant exposure over the entire city. Maps were generated with QGIS GIS 
+software, from incidence data.}
   \label{fig:mapas}
 \end{figure}
-\end{center}
 
@@
 
-\urlprefix\url{
-http://www.sciencedirect.com/science/article/pii/S0140673607616870}.
-
 \bibitem{Reich2013}
@@
   (\bibinfo{year}{2009}).
-\newblock \urlprefix\url{http://www.bioone.org/doi/abs/10.1603/033.046.0505}.
 
@@
 
-\urlprefix\url{
-http://www.sciencedirect.com/science/article/pii/S0048969708011509}.
-
 \bibitem{reiner_time-varying_2014}
@@
   \bibinfo{pages}{E2694--E2702} (\bibinfo{year}{2014}).
-\newblock \urlprefix\url{http://www.ncbi.nlm.nih.gov/pmc/articles/PMC4084484/}.
 
@@
   (\bibinfo{year}{2009}).
-\newblock
-  \urlprefix\url{http://link.springer.com/article/10.1007/s11538-009-9433-7}.
 
@@
   (\bibinfo{year}{2012}).
-\newblock
-
-\urlprefix\url{
-http://www.sciencedirect.com/science/article/pii/S0025556411001532}.
 
@@
 \bibinfo{author}{Stallybrass, C.~O.} \emph{et~al.}
-\newblock \bibinfo{title}{The principles of epidemiology and the process of
-  infection.}
-\newblock \emph{\bibinfo{journal}{The Principles of Epidemiology and the
-  Process of Infection.}}  (\bibinfo{year}{1931}).
+\newblock \emph{\bibinfo{title}{The Principles of Epidemiology and the Process
+  of Infection}}, vol.~\bibinfo{volume}{4} of \emph{\bibinfo{series}{10}}
+  (\bibinfo{publisher}{George Routledge \& Son, Ltd.}, \bibinfo{year}{1931}),
+  \bibinfo{edition}{1} edn.
 
@@
   in the case of the binomial}.
-\newblock \emph{\bibinfo{journal}{Biometrika}} \bibinfo{pages}{404--413}
-  (\bibinfo{year}{1934}).
+\newblock \emph{\bibinfo{journal}{Biometrika}} \textbf{\bibinfo{volume}{26}},
+  \bibinfo{pages}{404--413} (\bibinfo{year}{1934}).
 
@@
   (\bibinfo{year}{2009}).
-\newblock
-
-\urlprefix\url{
-http://onlinelibrary.wiley.com/doi/10.1111/j.1365-3156.2008.02203.x/abstract}.
 
@@
   (\bibinfo{year}{2006}).
-\newblock
-
-\urlprefix\url{
-http://www.scielosp.org/scielo.php?script=sci_abstract&pid=S0034-891020060001000
-  21&lng=en&nrm=iso&tlng=en}.
 
